### 2019-08-19 meta_info_collec/metafun/meta2preds.Rmd
meta info vs prediction accuracy fun

---
title: "meta2pred"
author: "Ali"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Meta data vs prediction accuracy

Investigating any relationship between the metadata going into genomes
and the prediction accuracy coming back out

```{r}
source('scripts/import.R')
```

## Reproducibility pred accuracy

comp `predacc4` (trained on Gmax,Creindhardtii,Mpolymorpha,Tcacao)

with `predacc9` (trained on Creinhardtii,Athaliana,Ptrichocarpa,Gmax,Sitalica,Bdistachyon,Slycopersicum,Mpolymorpha,Vvinifera)

```{r}
toplot <- predacc4
colnames(toplot) <- paste(colnames(toplot), '4', sep='.')
tmp <- predacc9
colnames(tmp) <- paste(colnames(tmp), '9', sep='.')
toplot <- merge(toplot, tmp, by.x = 'genome.4', by.y='genome.9')
```

```{r}
library(ggplot2)
bp <- ggplot(toplot, aes(color=factor(trained_on.4), fill=factor(trained_on.9))) +
  scale_fill_manual(values = c("grey55", "red2")) +
  scale_color_manual(values=c("white", "black")) +
  theme_bw() + 
  geom_abline(slope=1, intercept=0, color="grey20")
bp + geom_point(aes(x=acc_overall.4, y=acc_overall.9), shape=21, size=2.5) 
bp + geom_point(aes(x=f1_total_0.4, y=f1_total_0.9), shape=21, size=2.5) 
bp + geom_point(aes(x=f1_total_1.4, y=f1_total_1.9), shape=21, size=2.5) 
```
At a glance, yes the prediction accuracy is reproducible with different training sets,
yes being in the training set is a sort of bonus

## comparisons of high interest

```{r}
metadata$N50frac <- metadata$N50 / metadata$total_len
metadata$busco_C_loss_tran <- metadata$busco_C_tran - metadata$busco_C_geno
metadata$busco_C_loss_prot <- metadata$busco_C_prot - metadata$busco_C_geno

# stat for kmer diversity
k2mers <- metadata[, nchar(colnames(metadata)) == 2]
metadata$k2mer_sd <- apply(k2mers, 1, sd)

prepreg <- merge(metadata, predacc9, by.x='species', by.y='genome')
is_meta <- c(rep(TRUE, ncol(metadata)), rep(FALSE, ncol(predacc9)))

bp <- ggplot(prepreg, aes(color=factor(trained_on))) +
  scale_color_manual(values = c("grey55", "red2")) +
  theme_bw()

# genome size
bp + geom_point(aes(x=total_len, y=acc_overall))
bp + geom_point(aes(x=total_len, y=f1_total_1))
```

```{r}
# N50
bp + geom_point(aes(x=N50, y=acc_overall))
bp + geom_point(aes(x=N50, y=f1_total_1))
```

```{r}
# N50 as a fraction of genome size
bp + geom_point(aes(x=N50frac, y=acc_overall))
bp + geom_point(aes(x=N50frac, y=f1_total_1))
```

```{r}
# GC
bp + geom_point(aes(x=C, y=acc_overall))
bp + geom_point(aes(x=C, y=f1_total_1))
```
```{r}
# busco genomic completeness
bp + geom_point(aes(x=busco_C_geno, y=acc_overall))
bp + geom_point(aes(x=busco_C_geno, y=f1_total_1))
```

```{r}
# busco ratio transcriptome completeness - genome completeness
bp + geom_point(aes(x=busco_C_loss_tran, y=acc_overall))
bp + geom_point(aes(x=busco_C_loss_tran, y=f1_total_1))
```

```{r}
# busco ratio proteom completeness / genome completeness
bp + geom_point(aes(x=busco_C_loss_prot, y=acc_overall))
bp + geom_point(aes(x=busco_C_loss_prot, y=f1_total_1))
```

```{r}
# sd in rel abundance of k2 mers (wondering if we might be measuring,
# say, genome repetitivity / intergenic with the kmers)
bp + geom_point(aes(x=k2mer_sd, y=acc_overall))
bp + geom_point(aes(x=k2mer_sd, y=f1_total_1))
```

## single regression

```{r, results="hide", warning=FALSE}
col_is_na <- apply(apply(prepreg, 1:2, as.numeric), 2, function(x) is.na(x[1]))
to_scale <- (! col_is_na) & is_meta
prepreg_scale <- prepreg
prepreg_scale[, to_scale] <- scale(prepreg_scale[, to_scale])
```

### prep single binomial regression

```{r}
glm_binom <- function(y_frac, x=NULL, weights=prepreg$total_len){
  suc_fail <- round(cbind(y_frac, 1 - y_frac) * weights)
  if (is.null(x)){
    out <- glm(suc_fail ~ 1, family = binomial)
  }else{
    out <- glm(suc_fail ~ x, family = binomial)
  }
  return(out)
}

sigmoid_pred_line <- function(a_model){
  our_fit_fn <- function(x) predict(a_model, newdata = data.frame(x=x), type="response")
  stat_function(fun=our_fit_fn, color="grey30")
}

key_stats <- function(a_model){
  aic <- AIC(a_model)
  coeffs <- coef(a_model)
  p.value <- summary(a_model)$coefficients[2, 4]
  return(c(AIC=aic, coef=coeffs[2], p.value=p.value))
}

```

### vs accuracy
```{r}
xes <- colnames(prepreg_scale[to_scale])
indi_linregs_acc <- lapply(xes, 
                           function(x) glm_binom(y_frac=prepreg_scale$acc_overall, 
                                                 x=prepreg_scale[, x]))
names(indi_linregs_acc) <- xes
intercept_only <- glm_binom(y_frac=prepreg_scale$acc_overall)
indi_linregs_sum <- t(sapply(indi_linregs_acc, key_stats))
# sort by AIC score
indi_linregs_sum <- indi_linregs_sum[order(indi_linregs_sum[,1]), ]
```

#### most significant negative relationships (redundancy reduced)

```{r}
# fragmented buscos in the transcriptome (proteome was second)
bp <- ggplot(prepreg_scale, aes(y=acc_overall, color=factor(trained_on))) +
  scale_color_manual(values=c("black", "red2")) +
  theme_bw()
bp +
  geom_point(aes(x=busco_F_tran)) +
  sigmoid_pred_line(indi_linregs_acc[["busco_F_tran"]])
```

```{r}
# single copy buscos in the transcriptome (proteom was second); AKA lacks alternative splicing
bp + geom_point(aes(x=busco_S_tran)) +
  sigmoid_pred_line(indi_linregs_acc[["busco_S_tran"]])
```

#### most significant positive relationships (redundancy reduced)
```{r}
# duplicated buscos in the transcriptome (proteom was second); AKA has alternative splicing
bp + geom_point(aes(x=busco_D_tran)) +
  sigmoid_pred_line(indi_linregs_acc[["busco_D_tran"]])
```

```{r}
# count of CDS features from gff3 (maybe also related to alternative splicing?)
bp + geom_point(aes(x=CDS)) +
  sigmoid_pred_line(indi_linregs_acc[["CDS"]])
```

```{r}
# count of five_prime_UTR features from gff3 (maybe also related to alternative splicing?)
# that and probably the program that made the gff
bp + geom_point(aes(x=five_prime_UTR)) +
  sigmoid_pred_line(indi_linregs_acc[["five_prime_UTR"]])
# next would be more gff counts: 3', mRNA, exon
```

```{r}
# finally all the total_len_gt_X fall out with x in decreasing order
bp + geom_point(aes(x=total_len_gt_50k)) +
  sigmoid_pred_line(indi_linregs_acc[["total_len_gt_50k"]])
```

## multi regression (PLS after logit function)

```{r}
# linearizing transform for bounded 0-1 data (assuming it followed sigmoid to start with)
logit <- function(x){
  log(x) - log(1-x)
}

prepreg_scale$acc_overall_logit <- logit(prepreg_scale$acc_overall)
```

### OK, once just with lm
```{r}
non_redundant <- c("busco_F_tran","busco_D_tran","CDS","five_prime_UTR","mRNA","total_len_gt_50k","total_len","largest_contig","N75","gene","busco_C_tran","busco_C_loss_tran","N50frac","busco_F_geno","contigs_gt_5k","busco_M_tran","contigs","L50","CC","CA","busco_C_geno","CG","busco_M_geno","C","TA","AA","AC","N","AT","GA","GC","A","busco_D_geno","AG", "k2mer_sd")

x <- prepreg_scale[, non_redundant]
x$y <- prepreg_scale$acc_overall_logit
# bullshit bingo
lm.acc_overall.fit <- lm(y~., data=x)
summary(lm.acc_overall.fit)
```

### Now PLS

```{r}
library(pls)
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on = prepreg_scale$trained_on
x$y <- prepreg_scale$acc_overall_logit
pls.acc_overall.fit <- plsr(y~., data=x)
validationplot(pls.acc_overall.fit)
```

```{r}
# viz prediction accuracy
plot(x$y, predict(pls.acc_overall.fit, x, ncomp=3))
plot(x$y, predict(pls.acc_overall.fit, x, ncomp=12))
```

```{r}
library(reshape2)
metaclass <- c(rep("gff", 6), 
               rep("quast", 18),
               rep("buscoG", 5),
               rep("buscoP", 5),
               rep("buscoT", 5),
               rep("kmers", 13),
               rep("f.eng", 5))
metaclass <- factor(metaclass, levels=unique(metaclass))
# look at loadings of first components
plot_with_loadings <- function(amodel, metaclass){
  pls.loadings <- amodel$loadings
  toplot <- data.frame(pls.loadings[,1:3])
  toplot$meta <- factor(rownames(toplot), levels = rownames(toplot))
  toplot$metaclass <- metaclass
  toplot <- melt(toplot)
  #return(toplot)
  colnames(toplot)[3:4] <- c("component", "loading")

  bp <- ggplot(toplot, aes(y=loading, x=meta, fill=metaclass)) +
    scale_fill_brewer(palette = "Set3") +
    geom_bar(stat="identity") +
    facet_grid(rows=vars(component)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  bp
}
plot_with_loadings(pls.acc_overall.fit, metaclass)
```

#### looking at final coeffs with ncomp=2
```{r}

plot_with_ncomps <- function(ncomps, amodel, metaclass){
  toplot <- coef(amodel, ncomp=ncomps)
  toplot <- data.frame(coeffs3=toplot, 
                       meta=factor(rownames(toplot), 
                                   levels=rownames(toplot)),
                       metaclass=metaclass
                       )
  colnames(toplot)[1] <- 'coeffs'
  ggplot(toplot, aes(x=meta, y=coeffs, fill=metaclass)) +
    geom_bar(stat="identity") +
    scale_fill_brewer(palette = "Set3") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
}
plot_with_ncomps(2, pls.acc_overall.fit, metaclass)
```

#### looking at final coeffs with ncomp=3
```{r}
plot_with_ncomps(3, pls.acc_overall.fit, metaclass = metaclass)
```

#### looking at final coeffs with ncomp=9
```{r}
plot_with_ncomps(9, pls.acc_overall.fit, metaclass)
```

#### looking at final coeffs with ncomp=16
```{r}
plot_with_ncomps(16, pls.acc_overall.fit, metaclass)
```

## OK, and now that we have something resembling useful, PLS f1 1
```{r}
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg_scale$trained_on
x$y <- logit(prepreg_scale$f1_total_1)
pls.f1_total_f1.fit <- plsr(y~., data=x)
validationplot(pls.f1_total_f1.fit)
```

```{r}
# viz prediction accuracy
plot(x$y, predict(pls.f1_total_f1.fit, x, ncomp=3))
plot(x$y, predict(pls.f1_total_f1.fit, x, ncomp=12))
```

```{r}
# loadings
plot_with_loadings(pls.f1_total_f1.fit, metaclass)
```

#### looking at final coeffs with ncomp=2+
```{r}
plot_with_ncomps(2, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(3, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(9, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(16, pls.f1_total_f1.fit, metaclass)
```

## And further trained on 9, PLS f1_total_0
```{r}
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg_scale$trained_on
x$y <- logit(prepreg_scale$f1_total_0)
pls.f1_total_f0.fit <- plsr(y~., data=x)
validationplot(pls.f1_total_f0.fit)
```

```{r}
# viz prediction accuracy
plot(x$y, predict(pls.f1_total_f0.fit, x, ncomp=3))
plot(x$y, predict(pls.f1_total_f0.fit, x, ncomp=12))
```

```{r}
# loadings
plot_with_loadings(pls.f1_total_f0.fit, metaclass)
```

#### looking at final coeffs with ncomp=2+
```{r}
plot_with_ncomps(2, pls.f1_total_f0.fit, metaclass)
plot_with_ncomps(3, pls.f1_total_f0.fit, metaclass)
plot_with_ncomps(9, pls.f1_total_f0.fit, metaclass)
plot_with_ncomps(16, pls.f1_total_f0.fit, metaclass)
```

## kmers...

```{r}
kmers <- prepreg[, nchar(colnames(prepreg)) < 3]
mycols <- c('blue3', 'gold2', 'green3', 'grey40')
boxplot(kmers, col=mycols[c(1,2,4,1,3,3,1,3,2,2,3,2,1)], ylab='rel. freq.', main='kmers')
```

### model without kmers/kmers only (train 9, f1 1)

```{r}
# split input to kmers / no kmers
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg$trained_on
kmerish <- c("gc_content", "A", "C", "AA", "AC", "AG", "AT", "CA", "CC", "CG", "GA", "GC", "TA")
kmers_only <- x[, kmerish]
kmers_only$y <- logit(prepreg_scale$f1_total_1)

no_kmers <- x[, ! colnames(x) %in% kmerish]
no_kmers$y <- logit(prepreg_scale$f1_total_1)
```

#### kmers

```{r}
pls.kmers.f1_total_f1.fit <- plsr(y~., data=kmers_only)
validationplot(pls.kmers.f1_total_f1.fit)
plot(kmers_only$y, predict(pls.kmers.f1_total_f1.fit, kmers_only, ncomp=3))
mck <- metaclass[colnames(x) %in% kmerish]
plot_with_loadings(pls.kmers.f1_total_f1.fit, metaclass=mck)
plot_with_ncomps(2, pls.kmers.f1_total_f1.fit, metaclass=mck)
plot_with_ncomps(3, pls.kmers.f1_total_f1.fit, metaclass=mck)
```

#### no kmers

```{r}
pls.nokmers.f1_total_f1.fit <- plsr(y~., data=no_kmers)
validationplot(pls.nokmers.f1_total_f1.fit)
plot(no_kmers$y, predict(pls.nokmers.f1_total_f1.fit, no_kmers, ncomp=3))

mc2 <- metaclass[! colnames(x) %in% kmerish]
plot_with_loadings(pls.nokmers.f1_total_f1.fit, metaclass=mc2)
plot_with_ncomps(2, pls.nokmers.f1_total_f1.fit, metaclass=mc2)
plot_with_ncomps(3, pls.nokmers.f1_total_f1.fit, metaclass=mc2)
```

## Finally, PLS of f1_total_1 for trained on 4 
(to see if some or all is robust to exact training set)

```{r, results="hide", warning=FALSE}
prepreg <- merge(metadata, predacc4, by.x='species', by.y='genome')
col_is_na <- apply(apply(prepreg, 1:2, as.numeric), 2, function(x) is.na(x[1]))
to_scale <- (! col_is_na) & is_meta
prepreg_scale <- prepreg
prepreg_scale[, to_scale] <- scale(prepreg_scale[, to_scale])
```

```{r}
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg_scale$trained_on
x$y <- logit(prepreg_scale$f1_total_1)
pls4.f1_total_f1.fit <- plsr(y~., data=x)
validationplot(pls4.f1_total_f1.fit)
```
```{r}
# viz prediction accuracy
plot(x$y, predict(pls4.f1_total_f1.fit, x, ncomp=3))
plot(x$y, predict(pls4.f1_total_f1.fit, x, ncomp=12))
```

```{r}
# loadings
plot_with_loadings(pls4.f1_total_f1.fit, metaclass)
```

#### looking at final coeffs with ncomp=2+
```{r}
plot_with_ncomps(2, pls4.f1_total_f1.fit, metaclass)
plot_with_ncomps(3, pls4.f1_total_f1.fit, metaclass)
plot_with_ncomps(9, pls4.f1_total_f1.fit, metaclass)
plot_with_ncomps(16, pls4.f1_total_f1.fit, metaclass)
```

#### and finally side by side, 9 and 4
```{r}
plot_with_ncomps(3, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(3, pls4.f1_total_f1.fit, metaclass)
```

OK, on the side that just makes sense

* trained_on is positive
* fragmented buscos always hurt
* genomic buscos, sorts: complete, single copy, duplicated missing, fragmented

Makes some sense

* total length hurts

Confusing

* more contigs are good?

Maybe related to calling alternative splicing

* total counts of gff features help, particularly CDS, exon
* duplicated buscos are good, and single copy bad at the transcriptome/proteome level

Worrisome

* kmers

Adding the engineered feature from the kmers (2mer sd) certainly
helps a model without the kmers, but otherwise doesn't capture
that much of the kmer info / replace them.

```{r}
library(ggrepel)
coeffs9 <- coef(pls.f1_total_f1.fit, ncomp=3)
coeffs4 <- coef(pls4.f1_total_f1.fit, ncomp=3)

incl_name <- function(x, keep){
  if (x %in% keep){
    return(x)
  }else{
    return(NA)
  }
}
# prep labels

toorder <- abs(coeffs4 - coeffs9)
toorder <- data.frame(toorder)[order(toorder, decreasing = TRUE),, drop=FALSE]

toplot <- data.frame(coeffs4=coeffs4, coeffs9=coeffs9)
colnames(toplot) <- c("coeffs4", "coeffs9")
toplot$meta <- rownames(toplot)
toplot$meta_labs <- sapply(toplot$meta, incl_name, keep=rownames(toorder)[1:6])

ggplot(toplot, aes(x=coeffs4, y=coeffs9)) + 
  geom_abline(slope=1, intercept=0, color="grey") +
  geom_point() +
  geom_text_repel(aes(label=meta_labs), color="blue") +
  theme_bw()
```

### 2019-08-19 meta_info_collec/metafun/scripts/import.R

@@ -1,0 +1,4 @@
+metadata <- read.csv('metadata.csv')
+predacc9 <- read.csv('single_genome_results_10k_LSTM_O647Aaic_genomes.csv')
+predacc4 <- read.csv('single_genome_results_5k.csv')
+

### 2019-08-20 meta_info_collec/kmers/count_kmers.py
and now with cannonical kmers

@@ -1,6 +1,8 @@
 #!/usr/bin/env python
 import argparse
 from Bio import SeqIO
+from Bio.Seq import Seq
+from Bio.Alphabet import generic_dna
 from kpal.klib import Profile
 
 def count_seq(seq, k):
@@ -15,7 +17,20 @@ def main(fastain, min_k, max_k):
         seq = str(record.seq).upper()
         for k in range(min_k, max_k + 1, 1):
             counts, kmers = count_seq(seq, k)
+            count_dict = dict()
             for count, kmer in zip(counts, kmers):
+                count_dict[str(kmer)] = int(count)
+            # collapse to canonical kmers
+            to_del = []
+            for kmer in count_dict:
+                kmer_rc = str(Seq(kmer, generic_dna).reverse_complement())
+                if kmer > kmer_rc:
+                    count_dict[kmer_rc] += count_dict[kmer]
+                    to_del.append(kmer)
+            for kmer in to_del:
+                count_dict.pop(kmer)
+            for kmer in sorted(count_dict):
+                count = count_dict[kmer]
                 print('\t'.join([str(x) for x in [record.id, kmer, count, k]]))
         print('\t'.join([str(x) for x in [record.id, 'N', seq.count('N'), 1]]))
 

### 2019-10-04 meta_info_collec/metafun/meta2preds.Rmd
look @ meta-predictors for cds & aug cds

@@
 bp + geom_point(aes(x=N50, y=f1_total_1))
+bp + geom_point(aes(x=N50, y=f1_cds_1))
 ```
@@
 
+## now for Augustus (Marvin's internship results)
+
+### first: f1_cds_1 on the same subset of genomes but with pred 9
+```{r}
+# these are basically all the input vars
+aug <- read.csv('Augustus_results.csv')
+plus_aug <- merge(prepreg_scale, aug, by.x="species", by.y="genome")
+x <- plus_aug[, c(to_scale, FALSE)]  # new f1_cds_1_RT column is not scaled
+#x$trained_on <- plus_aug$trained_on
+x$y <- logit(plus_aug$f1_cds_1)
+pls.f1_cds_1.fit <- plsr(y~., data=x)
+validationplot(pls.f1_cds_1.fit)
+```
+
+```{r}
+# viz prediction accuracy
+plot(x$y, predict(pls.f1_cds_1.fit, x, ncomp=3))
+plot(x$y, predict(pls.f1_cds_1.fit, x, ncomp=12))
+```
+
+```{r}
+# loadings
+mc_aug <- metaclass[-length(metaclass)]
+plot_with_loadings(pls.f1_cds_1.fit, mc_aug)
+```
+
+#### looking at final coeffs with ncomp=2+
+```{r}
+plot_with_ncomps(2, pls.f1_cds_1.fit, mc_aug)
+plot_with_ncomps(3, pls.f1_cds_1.fit, mc_aug)
+plot_with_ncomps(9, pls.f1_cds_1.fit, mc_aug)
+plot_with_ncomps(16, pls.f1_cds_1.fit, mc_aug)
+```
+
+### now for Augustus
+
+```{r}
+# these are basically all the input vars
+x <- plus_aug[, c(to_scale, FALSE)]  # new f1_cds_1_RT column is 
+x$y <- logit(plus_aug$f1_cds_1_RT)
+pls.f1_cds_1_RT.fit <- plsr(y~., data=x)
+validationplot(pls.f1_cds_1_RT.fit)
+```
+```{r}
+# viz prediction accuracy
+plot(x$y, predict(pls.f1_cds_1_RT.fit, x, ncomp=3))
+plot(x$y, predict(pls.f1_cds_1_RT.fit, x, ncomp=12))
+```
+```{r}
+# loadings
+# skip trained_on for metaclass
+plot_with_loadings(pls.f1_cds_1_RT.fit, mc_aug)
+```
+#### looking at final coeffs with ncomp=2+
+```{r}
+plot_with_ncomps(2, pls.f1_cds_1_RT.fit, mc_aug)
+plot_with_ncomps(3, pls.f1_cds_1_RT.fit, mc_aug)
+plot_with_ncomps(9, pls.f1_cds_1_RT.fit, mc_aug)
+plot_with_ncomps(16, pls.f1_cds_1_RT.fit, mc_aug)
+```
+
 ## kmers...
@@
 ```
+
+Lucikly it looks like our training genomes are fairly well distributed in
+terms of kmers.
+```{r}
+source('scripts/PCA.R')
+```

### 2019-10-04 meta_info_collec/metafun/scripts/import.R

@@ -1,4 +1,5 @@
 metadata <- read.csv('metadata.csv')
 predacc9 <- read.csv('single_genome_results_10k_LSTM_O647Aaic_genomes.csv')
 predacc4 <- read.csv('single_genome_results_5k.csv')
-
+acc9withcds <- read.csv('single_genome_results_10k_LSTM_O647Aaic_CDS.csv')
+predacc9 <- merge(predacc9, acc9withcds[c("genome", "f1_cds_1")], by="genome")

### 2019-12-17 paper/main.tex
paper start

\documentclass{bioinfo}
\copyrightyear{2015} \pubyear{2015}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Subject Section}

\title[short Title]{General purpose gene annotation using deep learning}
\author[Sample \textit{et~al}.]{Corresponding Author\,$^{\text{\sfb 1,}*}$, Co-Author\,$^{\text{\sfb 2}}$ and Co-Author\,$^{\text{\sfb 2,}*}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
$^{\text{\sf 2}}$Department, Institution, City, Post Code,
Country.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:} Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text
Text Text Text Text Text.\\
\textbf{Results:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text\\
\textbf{Availability:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text  Text\\
\textbf{Contact:} \href{name@bio.com}{name@bio.com}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

\maketitle

\section{Introduction}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text. Figure~\ref{fig:01}
shows that the above method  Text Text Text Text  Text Text Text
Text Text Text  Text Text.  \citep{Bag01} wants to know about
{\ldots}{\ldots} text follows.
\begin{equation}
\sum \text{\it x}+ \text{\it y} =\text{\it Z}\label{eq:01}\vspace*{-10pt}
\end{equation}

%\enlargethispage{12pt}

\section{Approach}

Equation~(\ref{eq:01}) Text Text Text Text Text Text  Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
.....

\begin{methods}
\section{Methods}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text
text\vspace*{1pt}

\begin{itemize}
\item for bulleted list, use itemize
\item for bulleted list, use itemize
\item for bulleted list, use itemize\vspace*{1pt}
\end{itemize}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text Text Text Text Text Text Text Text Text Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text\vadjust{\newpage}.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text Text Text Text Text Text Text Text Text Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

\subsection{This is subheading}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text
text. Text Text Text Text Text Text  Text Text Text Text Text Text
Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

\subsubsection{This is subsubheading}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text\break
text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text
textText Text Text Text Text Text  Text Text Text Text Text Text
Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

\enlargethispage{6pt}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\begin{table}[!t]
\processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
head2 & head3 & head4\\\midrule
row1 & row1 & row1 & row1\\
row2 & row2 & row2 & row2\\
row3 & row3 & row3 & row3\\
row4 & row4 & row4 & row4\\\botrule
\end{tabular}}{This is a footnote}
\end{table}

\end{methods}

\begin{figure}[!tpb]%figure1
\fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
\end{minipage}}
%\centerline{\includegraphics{fig01.eps}}
\caption{Caption, caption.}\label{fig:01}
\end{figure}

%\begin{figure}[!tpb]%figure2
%%\centerline{\includegraphics{fig02.eps}}
%\caption{Caption, caption.}\label{fig:02}
%\end{figure}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\subsection{Test1}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text.

Table~\ref{Tab:01} shows that Text Text Text Text Text  Text Text
Text Text Text Text. Figure~2\vphantom{\ref{fig:02}} shows that
the above method Text Text. Text Text Text  Text Text Text Text
Text Text. Figure~2\vphantom{\ref{fig:02}} shows that the above
method Text Text. Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
%     as it may ignore the figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

(Table~\ref{Tab:01}) Text Text Text Text Text Text  Text Text Text
Text Text Text Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\begin{enumerate}
\item this is item, use enumerate
\item this is item, use enumerate
\item this is item, use enumerate
\end{enumerate}

Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows\vadjust{\pagebreak} that the
above method  Text Text Text Text Text Text Text Text Text Text
Text Text.  \citealp{Boffelli03} might want to know about text
text text text Text Text Text Text Text Text  Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text\break Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text\vspace*{-10pt}

\section*{Acknowledgements}

Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text
text text text\vspace*{-12pt}

\section*{Funding}

This work has been supported by the... Text Text  Text Text.\vspace*{-12pt}

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
%\bibliography{Document}

\begin{thebibliography}{}

\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.

\bibitem[Bag {\it et~al}., 2001]{Bag01}
Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.

\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
in nigral dopaminergic neurnol cell: correlation with the known
pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
Res.}, \textbf{110}(Suppl. 1), 76--84.

\bibitem[Lehmann, 1986]{Leh86}
Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.

\bibitem[Crenshaw and Jones, 2003]{Cre03}
Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
cancer management: one tumor, one chip. \textit{Bioinformatics},
doi:10.1093/bioinformatics/btn000.

\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
???--???.

\bibitem[Bardet, 1920]{Bar20}
Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
polydactylie et retinite pigmentaire (contribution a l'etude des
formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
institution, Paris, France.

\end{thebibliography}
\end{document}

### 2019-12-19 paper/main.tex
paper formalities

@@
 
-\subtitle{Subject Section}
+\subtitle{Really cool deep learning stuff}
 
 \title[short Title]{General purpose gene annotation using deep learning}
-\author[Sample \textit{et~al}.]{Corresponding Author\,$^{\text{\sfb 1,}*}$, Co-Author\,$^{\text{\sfb 2}}$ and Co-Author\,$^{\text{\sfb 2,}*}$}
-\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
-$^{\text{\sf 2}}$Department, Institution, City, Post Code,
-Country.}
+\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Andreas Weber\,$^{\text{\sfb 1,}*}$}
+\address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
 
-\corresp{$^\ast$To whom correspondence should be addressed.}
+\corresp{$^\dagger$Authors contributed equally. \newline $^\ast$To whom correspondence should be addressed.}
 
@@
 
-\abstract{\textbf{Motivation:} Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text
-Text Text Text Text Text.\\
-\textbf{Results:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text\\
+\abstract{\textbf{Motivation:} Current state of the art gene annotation programs for eukaryotic genomes have to be pretrained for each species and still often produce annotations that lack the required accuracy.
+\\
+\textbf{Results:} We set up a framework that can be used to build much more general gene callers that also outperform traditional methods on a variety of complex genomes.\\
 \textbf{Availability:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text
@@
 \section{Introduction}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text. Figure~\ref{fig:01}
-shows that the above method  Text Text Text Text  Text Text Text
-Text Text Text  Text Text.  \citep{Bag01} wants to know about
-{\ldots}{\ldots} text follows.
-\begin{equation}
-\sum \text{\it x}+ \text{\it y} =\text{\it Z}\label{eq:01}\vspace*{-10pt}
-\end{equation}
-
-%\enlargethispage{12pt}
+\begin{itemize}
+	\item Calling genes correctly is very important
+	\item but also very difficult in eukaryotes
+	\item All current state of the art gene callers work based on HMMs
+	\item The structure of the HMMs is finely tuned and engineered
+	\item Actual learning is very limited, and with that generalization across different genomes 
+	\item (Also have problems with genes that contain long introns, common in mammals)
+\end{itemize}
 
 \section{Approach}
+We seek to show a path for improving on current HMM-based methods both in performance and in the ability to generalize.
+To accomplish this, we use deep neural networks, which have been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods when detecting complex patterns in variety of fields.
+The genomic data was preprocessed with Geenuff \citealp{Denton19} abc
 
@@
 
-\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
-Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
-
-\bibitem[Bag {\it et~al}., 2001]{Bag01}
-Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.
-
-\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
-Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
-in nigral dopaminergic neurnol cell: correlation with the known
-pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
-Res.}, \textbf{110}(Suppl. 1), 76--84.
-
-\bibitem[Lehmann, 1986]{Leh86}
-Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.
+\bibitem[Denton {\it et~al}., 2019]{Denton19}
+Denton,A., Stiehler,F. (2019) Article title, {\it Journal Name}, {\bf 199}, 133-154.
 
@@
 
-\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
-Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
-(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
-???--???.
-
-\bibitem[Bardet, 1920]{Bar20}
-Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
-polydactylie et retinite pigmentaire (contribution a l'etude des
-formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
-institution, Paris, France.
-
 \end{thebibliography}

### 2019-12-20 paper/main.tex
stuff

@@
 \title[short Title]{General purpose gene annotation using deep learning}
-\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Andreas Weber\,$^{\text{\sfb 1,}*}$}
+\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Andreas Weber\,$^{\text{\sfb 1,}*}$}
 \address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
@@
 To accomplish this, we use deep neural networks, which have been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods when detecting complex patterns in variety of fields.
-The genomic data was preprocessed with Geenuff \citealp{Denton19} abc
+Our model uses the raw genomic sequence as input and produces base pair wise predictions for the present gene annotation.
+The genomic data was preprocessed with Geenuff \citealp{Denton19}
 

### 2019-12-21 backup_scripts/backup-home.sh
backup script

@@ -9,6 +9,7 @@ then
 	rsync -rzt --stats -e 'ssh -A -o "ProxyJump felix@134.99.224.62"' /home/felix/nni felix@134.99.224.58:/mnt/data/experiments_backup/nni_home_felix/
 	rsync -rzt --stats -e 'ssh -A -o "ProxyJump felix@134.99.224.62"' /home/felix/.local/nnictl felix@134.99.224.58:/mnt/data/experiments_backup/nni_home_felix/
 else
-	echo "Not connected to the vpn"
+	echo "Not connected to the vpn, writing data to the cluster"
+	rsync -rzt --stats --exclude='*model*.h5' /home/felix/nni festi100@hpc.rz.uni-duesseldorf.de:/gpfs/project/festi100/backup_home/
+	rsync -rzt --stats /home/felix/.local/nnictl festi100@hpc.rz.uni-duesseldorf.de:/gpfs/project/festi100/backup_home/
 fi
-

### 2019-12-23 paper/main.tex
paper

@@
 \section{Approach}
-We seek to show a path for improving on current HMM-based methods both in performance and in the ability to generalize.
-To accomplish this, we use deep neural networks, which have been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods when detecting complex patterns in variety of fields.
-Our model uses the raw genomic sequence as input and produces base pair wise predictions for the present gene annotation.
-The genomic data was preprocessed with Geenuff \citealp{Denton19}
-
-Equation~(\ref{eq:01}) Text Text Text Text Text Text  Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-.....
+Helixer is a set of tested guidelines for training and utilizing a general purpose gene annotation model for eukaryotic genomes.
+We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on more than 190 genomes in the animal case and 70 genomes for the plant model.
+We leverage modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields.
+Our model expects the raw genomic sequence as input and produces base pair wise predictions as output.
 
@@
 \section{Methods}
+The genomic data was preprocessed with Geenuff \citealp{Denton19}
 

### 2019-12-23 paper/main.tex
paper table

@@
 We leverage modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields.
-Our model expects the raw genomic sequence as input and produces base pair wise predictions as output.
+Our model expects the raw genomic sequence as main input and produces base pair wise predictions as output.
 
@@
 \section{Methods}
-The genomic data was preprocessed with Geenuff \citealp{Denton19}
+Helixer models work based on genomic data that was preprocessed and checked for errors by Geenuff \citealp{Denton19}, which stores all information in a SQL database.
+From there, training and evaluation data is generated.
+The generated data arrays can be found in Table
+
+\begin{table}[!t]
+\processtable{Data arrays generated and used by Helixer\label{Tab:01}} {
+\begin{tabular}{@{}llll@{}}
+\toprule Name & Information \\
+\midrule
+Input & Genomic sequence in a one hot encoding\\
+Output & Labels in a 4-dimensional one hot encoding \\
+Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
+Gene lengths & The length of a gene at a base or 0 if intergenic \\
+RNASeq score & A score in the interval $...$ to rank the coherence of the \\
+		     & annotation with RNASeq information \\
+\botrule
+\end{tabular}}{}
+\end{table}
 

### 2019-12-24 data_insight/animals/size_groups/make_search_space.sh
search space split

@@ -1,0 +1,15 @@
+#! /bin/bash
+
+echo -n '{
+    "test_data": {
+        "_type": "choice",
+        "_value": [
+'
+cat $1 | head -n -1 | cut -d " " -f1 | awk '{print "\"/gpfs/project/festi100/data/animals/'$2'/"$1"/test_data.h5\","}'
+cat $1 | tail -1 | cut -d " " -f1 | awk '{print "\"/gpfs/project/festi100/data/animals/'$2'/"$1"/test_data.h5\""}'
+
+echo -n '        ]
+    }
+}'
+
+

### 2020-01-04 cluster_eval/start_many.sh
start many

@@ -12,8 +12,8 @@ fi
 
 line_offset=$(<next_line)
 n_lines=$(cat datasets | wc -l)
-if [[ $(($line_offset+$1)) -gt $n_lines ]]; then
-	n_qsubs=$(($n_lines-$line_offset))
+if [[ $(($line_offset+$1)) -gt $(($n_lines+1)) ]]; then
+	n_qsubs=$(($n_lines-$line_offset+1))
 else
 	n_qsubs=$1
 fi

### 2020-01-05 backup_scripts/backup-experiments.sh
backup-experiments.sh

@@ -1,0 +1,21 @@
+#! /bin/bash
+unalias rsync &> /dev/null
+
+echo -e "\n----------------\n"
+date
+echo
+
+echo -e "\nnni local felix"
+rsync -rzt --stats --exclude="predictions.h5" /home/felix/nni nni_local_felix/
+rsync -rzt --stats --exclude="predictions.h5" /home/felix/.local/nnictl nni_local_felix/
+
+echo -e "\nnni clc server"
+rsync -rzt --stats  --exclude="predictions.h5" felix-stiehler@134.99.200.63:/home/felix-stiehler/nni nni_clc_server/
+rsync -rzt --stats  --exclude="predictions.h5" felix-stiehler@134.99.200.63:/home/felix-stiehler/.local/nnictl nni_clc_server/
+
+echo -e "\nnni cluster"
+rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/nni nni_cluster/
+rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/.local/nnictl nni_cluster/
+
+echo -e "\ncluster jobs"
+rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/git/HelixerPrep/jobs cluster_jobs/

### 2020-01-05 backup_scripts/backup_experiments.sh

@@ -1,21 +1,0 @@
-#! /bin/bash
-unalias rsync &> /dev/null
-
-echo -e "\n----------------\n"
-date
-echo
-
-echo -e "\nnni local felix"
-rsync -rzt --stats /home/felix/nni nni_local_felix/
-rsync -rzt --stats /home/felix/.local/nnictl nni_local_felix/
-
-echo -e "\nnni clc server"
-rsync -rzt --stats  felix-stiehler@134.99.200.63:/home/felix-stiehler/nni nni_clc_server/
-rsync -rzt --stats  felix-stiehler@134.99.200.63:/home/felix-stiehler/.local/nnictl nni_clc_server/
-
-echo -e "\nnni cluster"
-rsync -rzt --stats festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/nni nni_cluster/
-rsync -rzt --stats festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/.local/nnictl nni_cluster/
-
-echo -e "\ncluster jobs"
-rsync -rzt --stats festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/git/HelixerPrep/jobs cluster_jobs/

### 2020-01-07 paper/main.tex
paper

@@
 \section{Approach}
-Helixer is a set of tested guidelines for training and utilizing a general purpose gene annotation model for eukaryotic genomes.
-We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on more than 190 genomes in the animal case and 70 genomes for the plant model.
+Helixer is a prototype software for training and utilizing a general purpose gene annotation model for eukaryotic genomes.
+We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on 192 genomes in the animal case and 60 genomes for the plant model.
 We leverage modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields.
-Our model expects the raw genomic sequence as main input and produces base pair wise predictions as output.
+Our model expects the raw genomic sequence as main input and produces functional predictions on a base pair wise level.
 
@@
 From there, training and evaluation data is generated.
-The generated data arrays can be found in Table
+The generated data arrays can be found in Table \ref{Tab:01}.
+We worked with 192 animal and 60 plant genomes.
+The data is comprised of all animals genomes in ensembl Metazoa 45 (?) \citealp{Howe19} as well as all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
+Both groups were treated seperately both during training as well as evaluation.
+At the start of our work we split both genome groups into training and test sets, which contained 19 or 6 genomes, respectively.
+Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
 
@@
 
-\begin{table}[!t]
-\processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
-head2 & head3 & head4\\\midrule
-row1 & row1 & row1 & row1\\
-row2 & row2 & row2 & row2\\
-row3 & row3 & row3 & row3\\
-row4 & row4 & row4 & row4\\\botrule
-\end{tabular}}{This is a footnote}
-\end{table}
-
 \end{methods}
@@
 
-\bibitem[Crenshaw and Jones, 2003]{Cre03}
-Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
-cancer management: one tumor, one chip. \textit{Bioinformatics},
-doi:10.1093/bioinformatics/btn000.
+\bibitem[Howe KL {\it et~al}., 2019]{Howe19}
+Howe KL, Contreras-Moreira B, De Silva N, Maslen G, Akanni W, Allen J, Alvarez-Jarreta J, Barba M, Bolser DM, Cambell L, Carbajo M, Chakiachvili M, Christensen M, Cummins C, Cuzick A, Davis P, Fexova S, Gall A, George N, Gil L, Gupta P, Hammond-Kosack KE, Haskell E, Hunt SE, Jaiswal P, Janacek SH, Kersey PJ, Langridge N, Maheswari U, Maurel T, McDowall MD, Moore B, Muffato M, Naamati G, Naithani S, Olson A, Papatheodorou I, Patricio M, Paulini M, Pedro H, Perry E, Preece J, Rosello M, Russell M, Sitnik V, Staines DM, Stein J, Tello-Ruiz MK, Trevanion SJ, Urban M, Wei S, Ware D, Williams G, Yates AD, Flicek P.
+(2019) Ensembl Genomes 2020-enabling non-vertebrate genomic research.
+{\it Nucleic Acids Research 2019} https://doi.org/10.1093/nar/gkz890
 

### 2020-01-07 paper/main.tex
paper

@@
 \section{Methods}
-	Helixer models work based on genomic data that was preprocessed and checked for errors by Geenuff \citealp{Denton19}, which stores all information in a SQL database.
-From there, training and evaluation data is generated.
-The generated data arrays can be found in Table \ref{Tab:01}.
 We worked with 192 animal and 60 plant genomes.
 The data is comprised of all animals genomes in ensembl Metazoa 45 (?) \citealp{Howe19} as well as all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
-Both groups were treated seperately both during training as well as evaluation.
+Both data groups were used seperately during training as well as during evaluation.
 At the start of our work we split both genome groups into training and test sets, which contained 19 or 6 genomes, respectively.
 Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
+
+We preprocessed and stored the genomic information by using Geenuff \citealp{Denton19}.
+Geenuff is a datachecking and exploration tool, which stores all information inside a SQL database.
+Training and evaluation data was generated by querying this database and transforming the returned data into a numerical format suitable for machine learning.
+The generated data arrays can be found in Table \ref{Tab:01}.
+We only queried the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?).
+As Geenuff checks the genomic annotations for potential errors like missing UTR sections and marks areas as potentially erroneous, we were able to effectively blank out those bases during training.
+For the training, we divided each continous genomic sequence into 20.000 or 30.000 long subsequences, which are then used as input or label for the neural network.
+We appended zero padding if the subsequences are not long enough.
+If a subsequence is fully marked as erroneous, we do not include it in the training or evaluation datasets.
+
+The selection of training genomes required balancing multiple conflicting tradeoffs.
+On the one hand, we do want to train with as much data as possible but not all data has a good enough quality to enable a powerful generalization.
+It is also very desirable to have a broad spread in phylogenetic distance as well as genome size and average gene length as the model is increasingly unlikely to generalize well beyond the borders given by the training data.
+It was also important to us to get experimental results within a couple of days as we tested many different data and model configurations.
 

### 2020-01-07 paper/main.tex
paper

@@
 Our model expects the raw genomic sequence as main input and produces functional predictions on a base pair wise level.
+The source code and all data is publically available.
 
@@
 The selection of training genomes required balancing multiple conflicting tradeoffs.
-On the one hand, we do want to train with as much data as possible but not all data has a good enough quality to enable a powerful generalization.
+On the one hand, we do want to train with as much data as possible, but not all data has a good enough quality to enable a powerful generalization.
 It is also very desirable to have a broad spread in phylogenetic distance as well as genome size and average gene length as the model is increasingly unlikely to generalize well beyond the borders given by the training data.
-It was also important to us to get experimental results within a couple of days as we tested many different data and model configurations.
+It was also important to us to get experimental results within a couple of days as we tested many different data and model configurations, which limited the overall size of the training set and the model.
+
+\subsection{Model Architecture}
+	We use a 4-layer deep stacked LSTM network with 256 units per layer to produce the predictions.
+The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (citation).
+All relevant hyperparameters can be found in Table \ref{Tab:02} or in the Helixer source code repository.
+
+The model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
+This grouping enabled us to train with far longer sequences than normally done.
+
+\subsection{Evaluation}
 
@@
 \processtable{Data arrays generated and used by Helixer\label{Tab:01}} {
-\begin{tabular}{@{}llll@{}}
+\begin{tabular}{@{}ll@{}}
 \toprule Name & Information \\
@@
 		     & annotation with RNASeq information \\
+\botrule
+\end{tabular}}{}
+\end{table}
+
+\begin{table}[!t]
+\processtable{Hyperparameters\label{Tab:02}} {
+\begin{tabular}{@{}ll@{}}
+\toprule Name & Value \\
+\midrule
+Layers & 4\\
+Units per layer & 256\\
+Learning rate & 1e-4\\
+Optimizer & Adam\\
 \botrule

### 2020-01-07 paper/main.tex
paper

@@
 We worked with 192 animal and 60 plant genomes.
-The data is comprised of all animals genomes in ensembl Metazoa 45 (?) \citealp{Howe19} as well as all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
+The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format.
+We used all animals genomes in ensembl Metazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
 Both data groups were used seperately during training as well as during evaluation.
-At the start of our work we split both genome groups into training and test sets, which contained 19 or 6 genomes, respectively.
+At the start of our work we split both genome groups into training and test sets, which then contained 19 or 6 genomes, respectively.
 Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
 
-We preprocessed and stored the genomic information by using Geenuff \citealp{Denton19}.
-Geenuff is a datachecking and exploration tool, which stores all information inside a SQL database.
+We pre-processed and stored the genomic information by using Geenuff \citealp{Denton19}.
+Geenuff is a tool for checking and exploring genomic data, which stores all information inside a SQL database.
 Training and evaluation data was generated by querying this database and transforming the returned data into a numerical format suitable for machine learning.
-The generated data arrays can be found in Table \ref{Tab:01}.
-We only queried the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?).
+In the case of unambiguous input data, the genomic sequence was transformed into matrix that represents a one hot encoding with 4 rows.
+For the other case see Suplementary material ..
+().
+The annotations are also transformed into a 4-row one hot encoding matrix, where a one in each of the rows represents one of the possible classes {\it Intergenic}, {\it UTR}, {\it Exon} or {\it Intron}.
+An non-coding intron can be inferred from the prediction order.
+A full description of all generated data arrays can be found in table \ref{Tab:01}.
+
+During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?).
 As Geenuff checks the genomic annotations for potential errors like missing UTR sections and marks areas as potentially erroneous, we were able to effectively blank out those bases during training.
-For the training, we divided each continous genomic sequence into 20.000 or 30.000 long subsequences, which are then used as input or label for the neural network.
+For the training, we divided each continous genomic sequence into 20.000 or 30.000 bases long subsequences, which are then used as input or label for the neural network.
 We appended zero padding if the subsequences are not long enough.
@@
 
-The selection of training genomes required balancing multiple conflicting tradeoffs.
+The selection of training genomes requires balancing multiple conflicting tradeoffs.
 On the one hand, we do want to train with as much data as possible, but not all data has a good enough quality to enable a powerful generalization.
@@
 
-\subsection{Model Architecture}
-	We use a 4-layer deep stacked LSTM network with 256 units per layer to produce the predictions.
+\subsection{Model Architecture and Usage}
+We use a 4-layer deep stacked BLSTM network with 256 units per layer to produce the predictions in the form of a base pair wise classification.
 The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (citation).
@@
 
-\subsection{Evaluation}
+- how we train and validate the model
+- input of longer sequences
+- overlapping
+- ensembling
 
@@
 Input & Genomic sequence in a one hot encoding\\
+	  & for (A, T, C, G), otherwise different, see suplementary .. ()\\
 Output & Labels in a 4-dimensional one hot encoding \\
@@
 
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text
 text\vspace*{1pt}
@@
 Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
+
 \citealp{Boffelli03} might want to know about  text text text text
@@
 Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
-Text Text Text Text Text Text Text Text Text Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text\vadjust{\newpage}.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
-Text Text Text Text Text Text Text Text Text Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-\subsection{This is subheading}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text
-text. Text Text Text Text Text Text  Text Text Text Text Text Text
-Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-\subsubsection{This is subsubheading}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text\break
-text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text
-textText Text Text Text Text Text  Text Text Text Text Text Text
-Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-\enlargethispage{6pt}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
 
@@
 Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-\subsection{Test1}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
 \citealp{Boffelli03} might want to know about  text text text text
@@
 \section{Conclusion}
-
-(Table~\ref{Tab:01}) Text Text Text Text Text Text  Text Text Text
-Text Text Text Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-\begin{enumerate}
-\item this is item, use enumerate
-\item this is item, use enumerate
-\item this is item, use enumerate
-\end{enumerate}
-
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows\vadjust{\pagebreak} that the
-above method  Text Text Text Text Text Text Text Text Text Text
-Text Text.  \citealp{Boffelli03} might want to know about text
-text text text Text Text Text Text Text Text  Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text\break Text.
 

### 2020-01-07 cluster_eval/start_eval.sh
fixed job runtime cluster eval

@@ -40,16 +40,12 @@ cd $species
 declare -A values
 if [[ "$eval_len" == "20" ]]; then
 	values[__BATCH_SIZE__]="150"
-	values[__HOURS__]="3"
 elif [[ "$eval_len" == "50" ]]; then
 	values[__BATCH_SIZE__]="60"
-	values[__HOURS__]="47"
 elif [[ "$eval_len" == "100" ]]; then
 	values[__BATCH_SIZE__]="30"
-	values[__HOURS__]="47"
 elif [[ "$eval_len" == "200" ]]; then
 	values[__BATCH_SIZE__]="15"
-	values[__HOURS__]="47"
 else
 	echo "Unknown length $eval_len. exiting."
 	exit

### 2020-01-08 paper/main.tex
paper results table

@@
 The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format.
-We used all animals genomes in ensembl Metazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
+We used all animals genomes in EnsemblMetazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
 Both data groups were used seperately during training as well as during evaluation.
@@
 
+\begin{table}[!t]
+\processtable{Experimental Results and Comparison\label{Tab:03}} {
+\begin{tabular}{@{}lll@{}}
+\toprule & EnsemblMetazoa & JGI Phytosome\\
+\midrule
+Augustus & 0.62 (+-0.02) & 0.83 (+- 0.03)\\
+Best Helixer model & 0.62 (+-0.02) & 0.72 (+-0.02) \\
+Best model + varied input length & 0.62 (+-0.02) & 0.72 (+-0.02) \\
+Best model + varied input length & 0.62 (+-0.02) & 0.72 (+-0.02) \\
+Best model + varied input length + overlapping & 0.62 (+-0.02) & 0.72 (+-0.02) \\
+Varied input length + overlapping + ensembling & 0.62 (+-0.02) & 0.72 (+-0.02) \\
+\botrule
+\end{tabular}}{Values are averages in CDS F1 (?) across all validation and test genomes of the \\ respective group.}
+\end{table}
+
 Text Text Text Text Text Text  Text Text Text Text Text Text Text

### 2020-01-08 paper/main.tex
minor

@@
 Best model + varied input length & 0.62 (+-0.02) & 0.72 (+-0.02) \\
-Best model + varied input length & 0.62 (+-0.02) & 0.72 (+-0.02) \\
 Best model + varied input length + overlapping & 0.62 (+-0.02) & 0.72 (+-0.02) \\

### 2020-01-08 data_generation_insert/make-all-animal-single-genomes.sh
data_generation_insert folder renamed to data_scripts, some scripts added/updated

@@ -1,2 +1,0 @@
-#! /bin/bash
-cat ~/Desktop/data_insight/animals/genome_sizes_fragments | tail -n +3 | cut -f1 -d " " | shuf | xargs -P 8 -I % ~/git/HelixerPrep/export.py --db-path-in /mnt/data/felix/animal_geenuff.sqlite3 --out-dir /mnt/data/felix/datasets/animals/single_genomes/% --chunk-size 20000 --skip-meta-info --only-test-set --genomes %

### 2020-01-08 data_scripts/check-datasets.py

@@ -20,7 +20,12 @@ for genome_folder in os.listdir(args.folder):
         h5_files = glob.glob(os.path.join(args.folder, genome_folder, args.sub_folder, '*.h5'))
 
     for h5_file in h5_files:
-        f = h5py.File(h5_file, 'r')
+        try:
+            f = h5py.File(h5_file, 'r')
+        except OSError:
+            print('ERROR', h5_file)
+            continue
+
         if 'genomes' in f.attrs:
             print('OK', h5_file)
         else:

### 2020-01-08 data_scripts/eval-from-predictions.sh

@@ -1,0 +1,22 @@
+#! /bin/bash
+
+# generates predictions (what would have been in trial.log files when done with nni)
+# from a run that only generated predictions.h5 files in different folders
+# this is useful to save time as this step can be done concurrently
+# uses 16 jobs concurrently by default
+
+if [[ $# -lt 1 ]]; then
+	echo "Usage: ./eval-from-predictions.sh main_folder [output file name]"
+	exit
+fi
+
+main_folder=$1
+
+if [[ $# -lt 2 ]]; then
+	output_file_name="trials.log"
+else
+	output_file_name=$2
+fi
+
+
+

### 2020-01-08 data_scripts/make-all-animal-single-genomes.sh

@@ -1,0 +1,8 @@
+#! /bin/bash
+
+len_in_k=$1
+n_processes=$2
+
+# cat ~/Desktop/data_insight/animals/genome_sizes_fragments | tail -n +3 | cut -f1 -d " " | shuf | xargs -P 8 -I % ~/git/HelixerPrep/export.py --db-path-in /mnt/data/felix/animal_geenuff.sqlite3 --out-dir /mnt/data/felix/datasets/animals/single_genomes/% --chunk-size 20000 --only-test-set --genomes %
+
+cat ~/git/helixer_scratch/data_insight/animals/genome_sizes_fragments | tail -n +3 | cut -f1 -d " " | shuf | xargs -P $n_processes -I % ~/git/HelixerPrep/export.py --db-path-in ~/Desktop/animal_geenuff.sqlite3 --out-dir ~/Desktop/data/animals/single_genomes_"$len_in_k"k/% --chunk-size "$len_in_k"000 --only-test-set --genomes %

### 2020-01-08 data_scripts/eval-from-predictions-one-trial.sh
eval-from-predictions

@@ -1,0 +1,21 @@
+#! /bin/bash
+
+if [[ $# -lt 2 ]]; then
+	echo "Usage: ./eval-from-predictions-one-trial.sh trial_folder output file name"
+	exit
+fi
+
+trial_folder=$1
+output_file_name=$2
+
+# if the folder has a trial.log in it we assume nni, otherwise cluster
+if [[ -f "$trial_folder/trial.log" ]]; then
+	log_file="$trial_folder/trial.log"
+	dataset=$(grep "test_data.h5" "$log_file"| cut -d " " -f5 | tr -d "'")
+else
+	log_file="$trial_folder/*.sh.o*"
+	dataset=$(grep "test_data.h5" "$log_file"| cut -d " " -f3 | tr -d ",'")
+fi
+
+/home/felix-stiehler/git/HelixerPrep/scripts/basic_pred_cm.py -d $dataset -p "$trial_folder/predictions.h5" &>> "$trial_folder/$output_file_name"
+

### 2020-01-08 data_scripts/eval-from-predictions.sh

@@ -3,7 +3,8 @@
 # generates predictions (what would have been in trial.log files when done with nni)
 # from a run that only generated predictions.h5 files in different folders
 # this is useful to save time as this step can be done concurrently
-# uses 16 jobs concurrently by default
+# uses eval-from-predictions-one-trial.sh to work concurrently
+# concurrency is set to 16 currently
 
 if [[ $# -lt 1 ]]; then
 	echo "Usage: ./eval-from-predictions.sh main_folder [output file name]"
@@ -18,5 +19,4 @@ else
 	output_file_name=$2
 fi
 
-
-
+ls -1 -d "$main_folder"/*/ | xargs -P 16 -I % ./eval-from-predictions-one-trial.sh % $output_file_name

### 2020-01-09 paper/main.tex

@@
 ().
-The annotations are also transformed into a 4-row one hot encoding matrix, where a one in each of the rows represents one of the possible classes {\it Intergenic}, {\it UTR}, {\it Exon} or {\it Intron}.
+The annotations are also transformed into a 4-row one hot encoding matrix, where a one in each of the rows represents one of the possible classes {\it Intergenic}, {\it UTR}, {\it CDS} or {\it Intron}.
 An non-coding intron can be inferred from the prediction order.
 A full description of all generated data arrays can be found in table \ref{Tab:01}.
+
+The classes are not balanced equally and the tradeoff between the number of intergenic and intronic bases is quite different in both groups of genomes.
+Table \ref{Tab:statistics} shows the ratio of class occurences, together with other overall metrics.
+
+\begin{table}[!t]
+\processtable{Data group statistics\label{Tab:statistics}} {
+\begin{tabular}{@{}lll@{}}
+\toprule & EnsemblMetazoa & JGI Phytosome\\
+\midrule
+Fraction of classes  & 0.6, 0.02, 0.1, 0.2 & 0.6, 0.02, 0.1, 0.2 \\
+(Intergenic, UTR, CDS, Intron) & & \\
+Geenuff error rate & 0.1 (+- 0.02) & 0.3 (+- 0.02) \\
+Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
+Average gene length & 23000 (+- 5000) & 3370 (+- 50) \\
+\botrule
+\end{tabular}}{Values are averages across all genomes of a data group}
+\end{table}
 
@@
 
+Model evaluation during training was done with a metric that we call the {\it Genic F1}.
+This score is very similar to a classic F1 score calculated across multiple classes, except that it takes into account that the data is
+
 \subsection{Model Architecture and Usage}
-We use a 4-layer deep stacked BLSTM network with 256 units per layer to produce the predictions in the form of a base pair wise classification.
+	We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization (citation ?
+) between each BLSTM layer to produce the predictions in the form of a base pair wise classification.
 The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (citation).
@@
 - ensembling
+
+\subsection{Evaluation of Augustus}
+What Marvin did, why we don't use Genic F1 for comparison, etc
 

### 2020-01-09 paper/main.tex
data split and selection

@@
 It is also very desirable to have a broad spread in phylogenetic distance as well as genome size and average gene length as the model is increasingly unlikely to generalize well beyond the borders given by the training data.
+However, if the genomic patterns inside the data are too different from each other, it may be difficult for the model to generalize as well.
 It was also important to us to get experimental results within a couple of days as we tested many different data and model configurations, which limited the overall size of the training set and the model.
+
+We employed an iterative approach to effectively select a proper set of training genomes.
+We first started with 3-4 genomes that were almost certain to be of the highest quality.
+We then evaluated the performance of the resulting model on all genomes individually (including the training genomes) and looked for candidates to add to the training set or to remove from it.
+This process was repeated multiple times concurrently with the model search until we saw no more room for substantial improvement given our computational constraints.
+
+Once a set of training genomes was selected, we generated a validation set from it by splitting of 20\% of the FASTA sequences above and below the N90 of each genome, ensuring a proper distribution of large and short sequences in both the training and validation set.
+The validation set was only used to roughly judge the performance of our models during training.
+Evaluation of the annotations on all training and validation genomes was done regularly, as we ultimately only cared about the performance on fully unknown genomes.
 

### 2020-01-09 paper/main.tex
metrics

@@
 
-Model evaluation during training was done with a metric that we call the {\it Genic F1}.
-This score is very similar to a classic F1 score calculated across multiple classes, except that it takes into account that the data is
+	We use metrics two metrics to judge model performance.
+The {\it Genic F1} is our primary metric that provides the most comprehensive picture of annotation quality in one number.
+The {\it Subgenic F1} is similar to the Genic F1, except that it does not take UTR prediction quality into account.
+This is done for comparability with Augustus and further explained in chapter \ref{augustus}.
+Both metrics work by summing up True Positives (TP), False Positives (FP), etc of each considered class before calculating precision, recall and ultimatily a F1 score.
+Both metrics do also leave out the intergenic class, as this is by far the easiest to predict and could potentially skew the results too far away from the performance in the genic region.
+The Genic F1 thus essentially provides a weighted mean of the classes UTR, CDS and Intron while the Subgenic F1 sums up only CDS and Intron.
+All individual class performances are reported in the Suplementary Material (?).
 
@@
 
-\subsection{Evaluation of Augustus}
+\subsection{Evaluation of Augustus} \label{augustus}
 What Marvin did, why we don't use Genic F1 for comparison, etc

### 2020-01-09 cluster_eval/start_many.sh
start_many test for pwd

@@ -10,6 +10,11 @@ if [[ ! -f "next_line" ]]; then
 	exit
 fi
 
+if [[ $(basename $(pwd)) = "cluster_eval" ]]; then
+	echo "script seems to be run from the repository, which should be an error"
+	exit
+fi
+
 line_offset=$(<next_line)
 n_lines=$(cat datasets | wc -l)
 if [[ $(($line_offset+$1)) -gt $(($n_lines+1)) ]]; then

### 2020-01-09 paper/main.tex
datasets update to 50k default for non-invertebrates
post processing roughly

@@
 
-Once a set of training genomes was selected, we generated a validation set from it by splitting of 20\% of the FASTA sequences above and below the N90 of each genome, ensuring a proper distribution of large and short sequences in both the training and validation set.
+Once a set of training genomes was selected, we generated a validation set from it by splitting of 20\% of the FASTA sequences above and below the N90 of each genome, ensuring a proper distribution of large and short sequences in both the training and validation set and a split on the chromosome level.
 The validation set was only used to roughly judge the performance of our models during training.
-Evaluation of the annotations on all training and validation genomes was done regularly, as we ultimately only cared about the performance on fully unknown genomes.
+Evaluation of the annotations on all training and validation genomes was done regularly, as we ultimately only cared about the performance on for the model fully unknown genomes.
 
@@
 The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (citation).
+We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks and hybrid architectures.
+The hyperparameters were determined by a combination of manual and automatic hyperparameter optimization (talk about TPE?, quote nni?).
 All relevant hyperparameters can be found in Table \ref{Tab:02} or in the Helixer source code repository.
@@
 The model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
-This grouping enabled us to train with far longer sequences than normally done.
+This grouping enabled us to train with far longer sequences than usual.
+A class wise evaluation including the calculation of the Genic F1 score is performed for the validation set after every epoch.
+We use the model with the best Genic F1 score unless the improvement to the next best one was less than $0.01$ apart in order to prevent overfitting.
 
-- how we train and validate the model
-- input of longer sequences
-- overlapping
-- ensembling
+We also used multiple techniques to improve the prediction quality after the training was done.
+One very effective way was to input longer sequences during inference than during training.
+The sequences were up to 10 times larger, depending on species group and assembly quality.
+We also reconstructed the final predictions of a single model from overlapping predictions that were cut to a core sequence.
+This was done to drastically reduce a typical bias of the models towards the beginning and end of each sequence and also to provide the model with multiple different starting points.
+The different overlapping sequences were recombined by averaging the softmax values of each base.
+This was also done when we finally ensembled the whole output of different models on the softmax value level from (8?) different models.
+For more details on these post-processing techniques see Suplementary Material (?).
 

### 2020-01-09 paper/main.tex
paper figure

@@
 We also reconstructed the final predictions of a single model from overlapping predictions that were cut to a core sequence.
-This was done to drastically reduce a typical bias of the models towards the beginning and end of each sequence and also to provide the model with multiple different starting points.
+This was done to drastically reduce a typical bias of the models towards the beginning and end of each sequence (see figure \ref{fig:length_wise_bias}) and also to provide the model with multiple different starting points.
 The different overlapping sequences were recombined by averaging the softmax values of each base.
@@
 
-\begin{figure}[!tpb]%figure1
-\fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
-235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
-\end{minipage}}
-%\centerline{\includegraphics{fig01.eps}}
-\caption{Caption, caption.}\label{fig:01}
+\begin{figure}[!tpb]
+	\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
+\caption{Caption, caption.}\label{fig:length_wise_bias}
 \end{figure}

### 2020-01-09 backup_scripts/backup-experiments.sh
jobs dir on the cluster moved to gpfs dir

@@ -18,4 +18,4 @@ rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de
 rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/.local/nnictl nni_cluster/
 
 echo -e "\ncluster jobs"
-rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de:/home/festi100/git/HelixerPrep/jobs cluster_jobs/
+rsync -rzt --stats --exclude="predictions.h5" festi100@hpc.rz.uni-duesseldorf.de:/gpfs/project/festi100/jobs cluster_jobs/

### 2020-01-09 cluster_eval/start_many.sh
cluster eval script

@@ -20,7 +20,7 @@ main_data_folder=$2
 
 line_offset=$(<next_line)
 n_lines=$(cat datasets | wc -l)
-if [[ $(($line_offset+$1)) -gt $(($n_lines+1)) ]]; then
+if [[ $(($line_offset+$3)) -gt $(($n_lines+1)) ]]; then
 	n_qsubs=$(($n_lines-$line_offset+1))
 else
 	n_qsubs=$3

### 2020-01-10 data_scripts/eval-from-predictions-one-trial.sh
prediction eval fixed for cluster

@@ -11,11 +11,10 @@ output_file_name=$2
 # if the folder has a trial.log in it we assume nni, otherwise cluster
 if [[ -f "$trial_folder/trial.log" ]]; then
 	log_file="$trial_folder/trial.log"
-	dataset=$(grep "test_data.h5" "$log_file"| cut -d " " -f5 | tr -d "'")
+	dataset=$(grep "test_data.h5" $log_file| cut -d " " -f5 | tr -d "'")
 else
 	log_file="$trial_folder/*.sh.o*"
-	dataset=$(grep "test_data.h5" "$log_file"| cut -d " " -f3 | tr -d ",'")
+	dataset=$(grep "test_data.h5" $log_file| cut -d " " -f3 | tr -d ",'")
 fi
 
-/home/felix-stiehler/git/HelixerPrep/scripts/basic_pred_cm.py -d $dataset -p "$trial_folder/predictions.h5" &>> "$trial_folder/$output_file_name"
-
+~/git/HelixerPrep/scripts/basic_pred_cm.py -d $dataset -p "$trial_folder/predictions.h5" &>> "$trial_folder/$output_file_name"

### 2020-01-10 paper/main.tex
rewrite metric part paper

@@
 The {\it Genic F1} is our primary metric that provides the most comprehensive picture of annotation quality in one number.
-The {\it Subgenic F1} is similar to the Genic F1, except that it does not take UTR prediction quality into account.
-This is done for comparability with Augustus and further explained in chapter \ref{augustus}.
-Both metrics work by summing up True Positives (TP), False Positives (FP), etc of each considered class before calculating precision, recall and ultimatily a F1 score.
-Both metrics do also leave out the intergenic class, as this is by far the easiest to predict and could potentially skew the results too far away from the performance in the genic region.
-The Genic F1 thus essentially provides a weighted mean of the classes UTR, CDS and Intron while the Subgenic F1 sums up only CDS and Intron.
-All individual class performances are reported in the Suplementary Material (?).
+The {\it Subgenic F1} is similar to the Genic F1, except that it does not take UTR predictions into account.
+This is done for comparability with Augustus and further explained in Chapter \ref{augustus}.
+Both metrics work by summing up True Positives (TP), False Positives (FP) and False Negatives (FN) of each considered class before calculating precision and recall from these values and combining the two into a F1 score.
+Both metrics do not take the values of the intergenic class into account, as this class is very abundant and appears to be by far the easiest to predict.
+In the case of the Genic F1, this means that effectively only the TP of the intergenic class are disregarded.
+The metrics essentially provide a weighted mean of the performances in the considered classes.
+All individual class performances are reported in Suplementary Material (?).
 

### 2020-01-13 paper/main.tex
minor

@@
 \title[short Title]{General purpose gene annotation using deep learning}
-\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Andreas Weber\,$^{\text{\sfb 1,}*}$}
+\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
 \address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}

### 2020-01-13 data_scripts/eval-from-predictions.sh
eval predictions script

@@ -4,7 +4,7 @@
 # from a run that only generated predictions.h5 files in different folders
 # this is useful to save time as this step can be done concurrently
 # uses eval-from-predictions-one-trial.sh to work concurrently
-# concurrency is set to 16 currently
+# concurrency default is set to 16 currently
 
 if [[ $# -lt 1 ]]; then
 	echo "Usage: ./eval-from-predictions.sh main_folder [output file name]"
@@ -19,4 +19,22 @@ else
 	output_file_name=$2
 fi
 
-ls -1 -d "$main_folder"/*/ | xargs -P 16 -I % ./eval-from-predictions-one-trial.sh % $output_file_name
+for species_folder in $(ls -d "$main_folder"/*/); do
+	species=$(basename $species_folder)
+	cat <<- EOF > $species_folder/$species"_eval.sh"
+	#!/bin/bash
+	#PBS -l select=1:ncpus=1:mem=1gb
+	#PBS -l walltime=1:59:00
+	#PBS -A "HelixerOpt"
+
+	module load Python/3.6.5
+
+	/home/festi100/git/helixer_scratch/data_scripts/eval-from-predictions-one-trial.sh $species_folder $output_file_name
+	EOF
+
+	echo -n $species" "
+	cd $species_folder
+	qsub $species"_eval.sh"
+
+	sleep 3 # to check everything
+done

### 2020-01-13 data_scripts/eval-from-predictions.sh
minor

@@ -23,7 +23,7 @@ for species_folder in $(ls -d "$main_folder"/*/); do
 	species=$(basename $species_folder)
 	cat <<- EOF > $species_folder/$species"_eval.sh"
 	#!/bin/bash
-	#PBS -l select=1:ncpus=1:mem=1gb
+	#PBS -l select=1:ncpus=1:mem=500mb
 	#PBS -l walltime=1:59:00
 	#PBS -A "HelixerOpt"
 

### 2020-01-14 cluster_eval/start_eval.sh
overlap by default during cluster eval

@@ -39,13 +39,21 @@ cd $species
 # we could put this into a config file but for now this should work
 declare -A values
 if [[ "$eval_len" == "20" ]]; then
-	values[__BATCH_SIZE__]="150"
-elif [[ "$eval_len" == "50" ]]; then
-	values[__BATCH_SIZE__]="60"
-elif [[ "$eval_len" == "100" ]]; then
 	values[__BATCH_SIZE__]="30"
+	values[__OVERLAP_OFFSET__]="2500"
+	values[__CORE_LENGTH__]="10000"
+elif [[ "$eval_len" == "50" ]]; then
+	values[__BATCH_SIZE__]="12"
+	values[__OVERLAP_OFFSET__]="6250"
+	values[__CORE_LENGTH__]="25000"
+elif [[ "$eval_len" == "100" ]]; then
+	values[__BATCH_SIZE__]="6"
+	values[__OVERLAP_OFFSET__]="12500"
+	values[__CORE_LENGTH__]="10000"
 elif [[ "$eval_len" == "200" ]]; then
-	values[__BATCH_SIZE__]="15"
+	values[__BATCH_SIZE__]="3"
+	values[__OVERLAP_OFFSET__]="25000"
+	values[__CORE_LENGTH__]="100000"
 else
 	echo "Unknown length $eval_len. exiting."
 	exit

### 2020-01-14 cluster_eval/start_eval.sh
proper cluster batch sizes

@@ -39,19 +39,19 @@ cd $species
 # we could put this into a config file but for now this should work
 declare -A values
 if [[ "$eval_len" == "20" ]]; then
-	values[__BATCH_SIZE__]="30"
+	values[__BATCH_SIZE__]="400"
 	values[__OVERLAP_OFFSET__]="2500"
 	values[__CORE_LENGTH__]="10000"
 elif [[ "$eval_len" == "50" ]]; then
-	values[__BATCH_SIZE__]="12"
+	values[__BATCH_SIZE__]="160"
 	values[__OVERLAP_OFFSET__]="6250"
 	values[__CORE_LENGTH__]="25000"
 elif [[ "$eval_len" == "100" ]]; then
-	values[__BATCH_SIZE__]="6"
+	values[__BATCH_SIZE__]="80"
 	values[__OVERLAP_OFFSET__]="12500"
 	values[__CORE_LENGTH__]="10000"
 elif [[ "$eval_len" == "200" ]]; then
-	values[__BATCH_SIZE__]="3"
+	values[__BATCH_SIZE__]="40"
 	values[__OVERLAP_OFFSET__]="25000"
 	values[__CORE_LENGTH__]="100000"
 else

### 2020-01-15 cluster_eval/start_eval.sh
fix

@@ -49,7 +49,7 @@ elif [[ "$eval_len" == "50" ]]; then
 elif [[ "$eval_len" == "100" ]]; then
 	values[__BATCH_SIZE__]="80"
 	values[__OVERLAP_OFFSET__]="12500"
-	values[__CORE_LENGTH__]="10000"
+	values[__CORE_LENGTH__]="50000"
 elif [[ "$eval_len" == "200" ]]; then
 	values[__BATCH_SIZE__]="40"
 	values[__OVERLAP_OFFSET__]="25000"

### 2020-01-15 data_insight/plot_gene_lengths.py
plot_gene_lengths, gene length query

@@ -1,0 +1,28 @@
+#! /usr/bin/env python3
+
+# plots histograms of the gene lengths and the loss impact of each bin
+# for each genome in the given "gene_lengths" file which is expected to have 4 columns
+
+import argparse
+import numpy as np
+import matplotlib.pyplot as plt
+from collections import defaultdict
+
+parser = argparse.ArgumentParser()
+parser.add_argument('input_file', type=str)
+args = parser.parse_args()
+
+genome_values = defaultdict(list)
+for i, line in enumerate(open(args.input_file)):
+    if i > 1:
+        parts = [e for e in line.strip().split(" ") if len(e) > 0]
+        genome, value = parts[0], int(parts[2])
+        if value <= 100000:
+            genome_values[genome].append(value)
+
+for genome, values in genome_values.items():
+    plt.clf()
+    plt.title(genome)
+    plt.hist(values, bins=200)
+    plt.savefig(f'{genome}_lengths.png')
+    print(f'saved {genome}_lengths.png')

### 2020-01-21 eval_predictions/evaluations/length_wise/length-wise-eval-from-cluster-jobs.sh
renaming, --redo

@@ -1,0 +1,42 @@
+#! /bin/bash
+
+# initiates length wise prediction evaluations for all species in a folder
+# as would be typical for a job on the cluster
+# outputs a picture with the length wise performance and a lot of detailed info
+# to stdout, that should hopefully be enough to recover everything without having
+# to rerun it
+
+if [[ $# -lt 1 ]]; then
+	echo "Usage: ./length-wise-eval-from-cluster-jobs.sh predictions_main_folder [--redo]"
+	exit
+fi
+
+predictions_main_folder=$1
+
+for species_folder in $(ls -d "$predictions_main_folder"/*/); do
+	species=$(basename $species_folder)
+	output_file_path=$species_folder"/length_wise_eval.log"
+	if [[ $# -lt 2 || ($2 == "--redo" && (! -f $output_file_path || ! -s $output_file_path)) ]]; then
+		log_file="$species_folder/*.sh.o*"
+		dataset=$(grep "test_data.h5" $log_file| cut -d " " -f3 | tr -d ",'")
+		cat <<- EOF > $species_folder/$species"_length_wise_eval.sh"
+		#!/bin/bash
+		#PBS -l select=1:ncpus=1:mem=2gb
+		#PBS -l walltime=7:59:00
+		#PBS -A "HelixerOpt"
+
+		module load Python/3.6.5
+
+		/home/festi100/git/HelixerPrep/scripts/errors_position_wise.py -p $species_folder/predictions.h5 -d $dataset -g $species -o $species_folder &> $output_file_path
+		EOF
+
+		echo -n $species" "
+		cd $species_folder
+		qsub $species"_length_wise_eval.sh"
+
+		sleep 1
+	else
+		echo $species" skipped"
+	fi
+done
+

### 2020-01-21 predictions_eval/evaluations/length_wise/length-wise-eval-from-cluster-jobs.sh

@@ -1,37 +1,0 @@
-#! /bin/bash
-
-# initiates length wise prediction evaluations for all species in a folder
-# as would be typical for a job on the cluster
-# outputs a picture with the length wise performance and a lot of detailed info
-# to stdout, that should hopefully be enough to recover everything without having
-# to rerun it
-
-if [[ $# -lt 1 ]]; then
-	echo "Usage: ./length-wise-eval-from-cluster-jobs.sh predictions_main_folder"
-	exit
-fi
-
-predictions_main_folder=$1
-
-for species_folder in $(ls -d "$predictions_main_folder"/*/); do
-	species=$(basename $species_folder)
-	log_file="$species_folder/*.sh.o*"
-	dataset=$(grep "test_data.h5" $log_file| cut -d " " -f3 | tr -d ",'")
-	cat <<- EOF > $species_folder/$species"_length_wise_eval.sh"
-	#!/bin/bash
-	#PBS -l select=1:ncpus=1:mem=2gb
-	#PBS -l walltime=3:59:00
-	#PBS -A "HelixerOpt"
-
-	module load Python/3.6.5
-
-	/home/festi100/git/HelixerPrep/scripts/errors_position_wise.py -p $species_folder/predictions.h5 -d $dataset -g $species -o $species_folder &> $species_folder/length_wise_eval.log
-	EOF
-
-	echo -n $species" "
-	cd $species_folder
-	qsub $species"_length_wise_eval.sh"
-
-	sleep 1
-done
-

### 2020-01-22 paper/main.tex
stuff

@@
 \begin{table}[!t]
-\processtable{Data group statistics\label{Tab:statistics}} {
+\processtable{Data group statistics for all data\label{Tab:statistics_all}} {
+\begin{tabular}{@{}lll@{}}
+\toprule & EnsemblMetazoa & JGI Phytosome\\
+\midrule
+Fraction of classes  & 0.6, 0.02, 0.1, 0.2 & 0.6, 0.02, 0.1, 0.2 \\
+(Intergenic, UTR, CDS, Intron) & & \\
+Geenuff error rate & 0.1 (+- 0.02) & 0.3 (+- 0.02) \\
+Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
+Average gene length & 23000 (+- 5000) & 3370 (+- 50) \\
+\botrule
+\end{tabular}}{Values are averages across all genomes of a data group}
+\end{table}
+
+\begin{table}[!t]
+\processtable{Data group statistics for our training data\label{Tab:statistics_training}} {
 \begin{tabular}{@{}lll@{}}
@@
 As Geenuff checks the genomic annotations for potential errors like missing UTR sections and marks areas as potentially erroneous, we were able to effectively blank out those bases during training.
-For the training, we divided each continous genomic sequence into 20.000 or 30.000 bases long subsequences, which are then used as input or label for the neural network.
+For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network.
 We appended zero padding if the subsequences are not long enough.

### 2020-01-22 data_scripts/insert-full.sh

@@ -1,8 +1,8 @@
 #! /bin/bash
 
-db_path='/mnt/data/felix/animal_geenuff.sqlite3'
+# db_path='/mnt/data/felix/animal_geenuff.sqlite3'
+db_path='/home/felix/Desktop/animal_geenuff_test.sqlite3'
 
-rm -v $db_path;
-ls -1 /mnt/data/ali/share/ensembl_nosplit/ | xargs -I % ~/Desktop/data_scripts/insert-one.sh % $db_path
+ls -1 /mnt/data/ali/share/ensembl_test | xargs -I % /home/felix/git/helixer_scratch/data_scripts/insert-one.sh % $db_path
 
-# ge; ls -1 /mnt/data/ali/share/phytozome_organized/ready/train | xargs -I % ./import_genome.py --db-path ~/Desktop/geenuff_plants.sqlite3 --basedir /mnt/data/felix/rawdata/plants/%/ --species %
+# ge; ls -1 /mnt/data/ali/share/phytozome_organized/ready/test | xargs -I % ./import_genome.py --db-path ~/Desktop/geenuff_plants.sqlite3 --basedir /mnt/data/felix/rawdata/plants/%/ --species %

### 2020-01-22 data_scripts/insert-one.sh

@@ -1,13 +1,18 @@
 #! /bin/bash
 
 geenuff_basedir=~/git/GeenuFF
-data_basedir=/mnt/data/ali/share/ensembl_nosplit
+data_basedir=/mnt/data/ali/share/ensembl_test
+# data_basedir=/mnt/data/ali/share/ensembl_nosplit
+# data_basedir=/mnt/data/ali/share/phytozome_organized/ready/test
 genome=$1
 db_path=$2
 
 rm -v -r $geenuff_basedir/tmp
 mkdir -v -p $geenuff_basedir/tmp/input $geenuff_basedir/tmp/output
-gunzip -v -c $data_basedir/$genome/ensembl98/annotation/*.gff3.gz > $geenuff_basedir/tmp/input/$genome.gff3
-gunzip -v -c $data_basedir/$genome/ensembl98/assembly/*.fa.gz > $geenuff_basedir/tmp/input/$genome.fa
+
+# gunzip -v -c $data_basedir/$genome/ensembl98/annotation/*.gff3.gz > $geenuff_basedir/tmp/input/$genome.gff3
+# gunzip -v -c $data_basedir/$genome/ensembl98/assembly/*.fa.gz > $geenuff_basedir/tmp/input/$genome.fa
+
+cp -v $data_basedir/$genome/ensembl98/annotation/*.gff3 $data_basedir/$genome/ensembl98/assembly/*.fa $geenuff_basedir/tmp/input/
 
 $geenuff_basedir/import_genome.py --db-path $2 --basedir $geenuff_basedir/tmp/ --species $genome

### 2020-01-22 paper/main.tex
stuff

@@
 Both data groups were used seperately during training as well as during evaluation.
-At the start of our work we split both genome groups into training and test sets, which then contained 19 or 6 genomes, respectively.
+At the start of our work we split both genome groups into training and test sets.
+The test sets contained 19 or 6 genomes, respectively.
+The test genomes were chosen based on being frequently used model species while also representing a broad phylogenetic spread.
 Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
@@
 The classes are not balanced equally and the tradeoff between the number of intergenic and intronic bases is quite different in both groups of genomes.
-Table \ref{Tab:statistics} shows the ratio of class occurences, together with other overall metrics.
+Table \ref{Tab:statistics_all} shows the ratio of class occurences, together with other overall metrics.
 
@@
 	During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?).
-As Geenuff checks the genomic annotations for potential errors like missing UTR sections and marks areas as potentially erroneous, we were able to effectively blank out those bases during training.
+As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to effectively blank out those bases during training.
+The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it.
+Table \ref{Tab:statistics_all} and \ref{Tab:statistics_training} show statistics about the masking.
 For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network.

### 2020-01-23 eval_predictions/predictions/plants/mk-predictions-plants.sh
minor

@@ -12,6 +12,7 @@ job_folder=$1
 model_file=$2
 data_main_folder="/gpfs/project/festi100/data/plants/single_genomes"
 
+# for data_folder in Alyrata  Csativus  Macuminata  Ppatens  Taestivum  Vcarteri; do
 for data_folder in $(ls -d "$data_main_folder"/*/); do
 	species=$(basename $data_folder)
 	job_subfolder=$job_folder/$species
@@ -33,9 +34,9 @@ for data_folder in $(ls -d "$data_main_folder"/*/); do
 		echo -n $species" "
 		cd $job_subfolder
 		qsub $species".sh"
+
+		# sleep 60 # to check everything and spread out the jobs
 	else
 		echo "Folder $job_subfolder exists. skipping."
 	fi
-
-	sleep 1 # to check everything
 done

### 2020-01-24 cluster_eval_predictions/evaluations/length_wise/length-wise-eval-from-cluster-jobs.sh
minor

@@ -17,7 +17,7 @@ for species_folder in $(ls -d "$predictions_main_folder"/*/); do
 	species=$(basename $species_folder)
 	output_file_path=$species_folder"/length_wise_eval.log"
 	if [[ $# -lt 2 || ($2 == "--redo" && (! -f $output_file_path || ! -s $output_file_path)) ]]; then
-		log_file="$species_folder/*.sh.o*"
+		log_file="$species_folder/$species.sh.o*"
 		dataset=$(grep "test_data.h5" $log_file| cut -d " " -f3 | tr -d ",'")
 		cat <<- EOF > $species_folder/$species"_length_wise_eval.sh"
 		#!/bin/bash
@@ -34,7 +34,7 @@ for species_folder in $(ls -d "$predictions_main_folder"/*/); do
 		cd $species_folder
 		qsub $species"_length_wise_eval.sh"
 
-		sleep 1
+		sleep 10
 	else
 		echo $species" skipped"
 	fi

### 2020-01-24 method_comp/count_floatables.py
count numbers in a file

@@ -1,0 +1,27 @@
+import argparse
+import sys
+
+
+def main(input):
+    i = 0
+    with open(input) as f:
+        for line in f:
+            sline = line.split()  # splits at any white space... so hopefully mostly words
+            for word in sline:
+                try:
+                    num = float(word)
+                    print('found', num, file=sys.stderr)
+                    i += 1
+                except ValueError:
+                    pass
+    print('{}\ttotal floatable words'.format(i))
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--input', '-i', type=str, required=True,
+                        help='Path to input text file')
+
+    args = parser.parse_args()
+
+    main(args.input)

### 2020-01-28 cluster_eval_predictions/evaluations/overall/cp-log-files-for-ensemble-eval.sh
minor stuff

@@ -1,0 +1,16 @@
+#! /bin/bash
+
+# An evaluation dependes on a {species}.sh.o* file to be present so it knows where for which
+# data file the predictions were made. In case of an ensemble those files are not there
+# and need to be copied from one of the ensemble job folders.
+
+if [[ $# -lt 2 ]]; then
+	echo "Usage: ./cp-log-files-for-ensemble-eval ensemble_folder job_folder"
+	exit
+fi
+
+ensemble_folder=$1
+job_folder=$2
+
+cd $job_folder
+ls -1 -d */ | tr -d "/" | xargs -I % bash -c "cp -v $job_folder/%/%.sh.o* $ensemble_folder/%/"

### 2020-01-28 cluster_eval_predictions/predictions/ensembling/ensemble-predictions.sh

@@ -37,11 +37,11 @@ for species_folder in $(ls -d "$job_main_folder/$1"/*/); do
 
 		echo -n $species" "
 		cd $output_species_folder
-		qsub $species"_ensemble.sh"
+		# qsub $species"_ensemble.sh"
 
 		sleep 2
 	else
-		echo "Folder "$output_species_folder" already existing."
+		echo "Folder "$output_species_folder" already exists."
 	fi
 done
 

### 2020-01-28 method_comp/clean_UTRoff_gff3.py
one step insertion of exons and transcript -> mRNA for augustus gffs

@@ -1,0 +1,34 @@
+import argparse
+import copy
+
+
+def main(filein, fileout):
+    handleout = open(fileout, 'w')
+    with open(filein) as f:
+        for line in f:
+            line = line.rstrip()
+            sline = line.split('\t')
+            # insert exons for every CDS
+            if line.startswith('#'):
+                pass  # just copy over comment lines without further parsing
+            elif sline[2] == "CDS":
+                exonsline = copy.deepcopy(sline)
+                # change feature type
+                exonsline[2] = "exon"
+                # unique (exon) ID
+                exonsline[8] = exonsline[8].replace('cds', 'exon')
+                handleout.write('\t'.join(exonsline) + '\n')
+            # change transcript features to mRNA (bc idk why they're called transcript and have CDS here)
+            elif sline[2] == "transcript":
+                sline[2] = 'mRNA'
+            handleout.write('\t'.join(sline) + '\n')
+    handleout.close()
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser()
+    parser.add_argument('-i', '--gff_in', help='input gff3, produced from augustus called with --gff3=on --UTR=off',
+                        required=True)
+    parser.add_argument('-o', '--out', help='output gff3 filename', required=True)
+    args = parser.parse_args()
+    main(args.gff_in, args.out)

### 2020-01-28 meta_info_collec/coverage/cov_breakdown.R
cleaup labelling some and margins

@@ -52,16 +52,17 @@ paircov_plot <- function(cov,
 	  geom_bar(stat='identity', position=position) + 
           scale_fill_viridis_d(direction=1) +
 	  ylab(ylab_str) +
-          xlab("") +
-
-          #scale_x_discrete(breaks=1:4, labels=c('a   ig\na    ig','a\na','a\na','a\na')) +
-	  ggtitle(paste0(argmax0, " = ", decode[argmax0 + 1], 
-			 ", ", argmax1, " = ", decode[argmax1 + 1],
-			 "\n", covstring))
+          xlab("") 
+#
+#	  ggtitle(paste0(argmax0, " = ", decode[argmax0 + 1], 
+#			 ", ", argmax1, " = ", decode[argmax1 + 1],
+#			 "\n", covstring))
   ypos <- estimate_y_labs(bp)
   print(ypos)
   bp <- bp + annotation_custom(textGrob(" Ref.:\nPred.:"), 
-            xmin=0, xmax=0,ymin=ypos, ymax=ypos) #+
+            xmin=0, xmax=0,ymin=ypos, ymax=ypos) +
+        scale_x_discrete(labels=c('\n', '\n', '\n', '\n')) + 
+        theme(plot.margin=unit(c(8, 8, 24, 16), "pt"))
 #        theme(axis.text.x=element_blank())
   for (i in 1:4){
           bp <- bp + annotation_custom(textGrob(levels(x$ref_pred)[i]), 

### 2020-01-29 data_insight/mk_n75_by_genome_length.py
n75/genome_length csv output

@@ -1,0 +1,17 @@
+#! /usr/bin/env python3
+
+# reads in a metadata csv file and outputs a csv with just two columns: species,n75/genome_length
+
+import argparse
+import pandas as pd
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--metadata_file', type=str, default='../metadata_csvs/plant_metadata.csv')
+parser.add_argument('--output_file', type=str, default='output.csv')
+args = parser.parse_args()
+
+df = pd.read_csv(args.metadata_file)
+df['N75/total_len'] = df['N75'] / df['total_len']
+df = df.sort_values('N75/total_len', ascending=False)
+df.to_csv(args.output_file, columns=['species', 'N75', 'total_len', 'N75/total_len'], index=False)
+print(f'{args.output_file} written')

### 2020-01-29 data_insight/mk_n75_by_genome_length.py
n75 by genome length stuff

@@ -8,10 +8,11 @@ import pandas as pd
 parser = argparse.ArgumentParser()
 parser.add_argument('--metadata_file', type=str, default='../metadata_csvs/plant_metadata.csv')
 parser.add_argument('--output_file', type=str, default='output.csv')
+parser.add_argument('--sort-by', type=str, default='N75/total_len')
 args = parser.parse_args()
 
 df = pd.read_csv(args.metadata_file)
 df['N75/total_len'] = df['N75'] / df['total_len']
-df = df.sort_values('N75/total_len', ascending=False)
+df = df.sort_values(args.sort_by, ascending=False)
 df.to_csv(args.output_file, columns=['species', 'N75', 'total_len', 'N75/total_len'], index=False)
 print(f'{args.output_file} written')

### 2020-01-29 cluster_eval_predictions/predictions/animals/start_eval.sh
cluster eval

@@ -39,19 +39,19 @@ cd $species
 # we could put this into a config file but for now this should work
 declare -A values
 if [[ "$eval_len" == "20" ]]; then
-	values[__BATCH_SIZE__]="400"
+	values[__BATCH_SIZE__]="200"
 	values[__OVERLAP_OFFSET__]="2500"
 	values[__CORE_LENGTH__]="10000"
 elif [[ "$eval_len" == "50" ]]; then
-	values[__BATCH_SIZE__]="160"
+	values[__BATCH_SIZE__]="80"
 	values[__OVERLAP_OFFSET__]="6250"
 	values[__CORE_LENGTH__]="25000"
 elif [[ "$eval_len" == "100" ]]; then
-	values[__BATCH_SIZE__]="80"
+	values[__BATCH_SIZE__]="40"
 	values[__OVERLAP_OFFSET__]="12500"
 	values[__CORE_LENGTH__]="50000"
 elif [[ "$eval_len" == "200" ]]; then
-	values[__BATCH_SIZE__]="40"
+	values[__BATCH_SIZE__]="24"
 	values[__OVERLAP_OFFSET__]="25000"
 	values[__CORE_LENGTH__]="100000"
 else

### 2020-01-29 cluster_eval_predictions/predictions/animals/start_many.sh

@@ -10,7 +10,7 @@ if [[ ! -f "next_line" ]]; then
 	exit
 fi
 
-if [[ $(basename $(pwd)) = "cluster_eval" ]]; then
+if [[ $(basename $(pwd)) = "animals" ]]; then
 	echo "script seems to be run from the repository, which should be an error"
 	exit
 fi

### 2020-01-30 cluster_eval_predictions/predictions/animals/start_many.sh
start_many

@@ -15,6 +15,11 @@ if [[ $(basename $(pwd)) = "animals" ]]; then
 	exit
 fi
 
+if [[ $(basename -s ".h5" $model_file) != $(basename main_data_folder) ]]; then
+	echo "job folder does not match model file. could mean something is wrong, exiting"
+	exit
+fi
+
 model_file=$1
 main_data_folder=$2
 
@@ -28,5 +33,5 @@ fi
 
 for i in $(seq 1 $n_qsubs); do
 	./start_eval.sh $model_file $main_data_folder
-	sleep 30
+	sleep 20
 done

### 2020-01-30 paper/main.tex
paper

@@
 \section{Approach}
-Helixer is a prototype software for training and utilizing a general purpose gene annotation model for eukaryotic genomes.
+Helixer is a prototype software for training and utilizing a general purpose gene annotation model for large eukaryotic genomes.
 We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on 192 genomes in the animal case and 60 genomes for the plant model.
@@
 The test sets contained 19 or 6 genomes, respectively.
-The test genomes were chosen based on being frequently used model species while also representing a broad phylogenetic spread.
+The test genomes were chosen to be of the best possible quality (?) based on collected metadata while also representing a broad phylogenetic spread.
 Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
@@
 In the case of unambiguous input data, the genomic sequence was transformed into matrix that represents a one hot encoding with 4 rows.
-For the other case see Suplementary material ..
+For the other case see Supplementary material ..
 ().
@@
 The metrics essentially provide a weighted mean of the performances in the considered classes.
-All individual class performances are reported in Suplementary Material (?).
+All individual class performances are reported in Supplementary Material (?).
 
@@
 The model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
-This grouping enabled us to train with far longer sequences than usual.
-A class wise evaluation including the calculation of the Genic F1 score is performed for the validation set after every epoch.
-We use the model with the best Genic F1 score unless the improvement to the next best one was less than $0.01$ apart in order to prevent overfitting.
+This grouping enabled us to train effectively with far longer sequences than usual.
+A class wise evaluation including the calculation of the Genic F1 score is performed on the validation set after every epoch.
 
 We also used multiple techniques to improve the prediction quality after the training was done.
-One very effective way was to input longer sequences during inference than during training.
-The sequences were up to 10 times larger, depending on species group and assembly quality.
-We also reconstructed the final predictions of a single model from overlapping predictions that were cut to a core sequence.
-This was done to drastically reduce a typical bias of the models towards the beginning and end of each sequence (see figure \ref{fig:length_wise_bias}) and also to provide the model with multiple different starting points.
-The different overlapping sequences were recombined by averaging the softmax values of each base.
-This was also done when we finally ensembled the whole output of different models on the softmax value level from (8?) different models.
-For more details on these post-processing techniques see Suplementary Material (?).
+One very effective way for genomes with larger genes was to input longer sequences during inference than during training.
+This was done for all animal genomes except the invertebrates.
+The input sequences were up to 10 times larger, depending on species group and assembly quality.
+This also demonstrates the generalization capability of the model as it is able to make successful prediction on far longer sequences than it has ever seen.
+The sequence length during inference can be any of 20.000, 50.000, 100.000 or 200.000 and is determined by the species family(?).
+Families that tend to have longer genes get assigned longer sequence chunks.
+If the N75 of a species is less than twice as high as the supposed sequence length it is lowered until either criteria is met or a length of 50.000 is reached.
+For implementation details on this and the following techniques see Supplementary Material (?) or the source code..
+
+We also construct the final predictions of a single model by overlapping predictions that were made from a sliding window and then cut to a core sequence.
+This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see figure \ref{fig:length_wise_bias}) and also to improve model performance in general by providing the model with multiple different starting points.
+The different overlapping sequences were recombined by averaging the individual softmax values of each base.
+
+A model ensemble with 8 components was used to generate the final predictions for each species.
+For this, we trained 4 seperate animal and plant models and also used a model from an epoch that did not have the highest Genic F1 on the validation set.
+Depending on the results, we used a model from an epoch with either the highest genic recall or precision.
+The model with the best Genic F1 was always best in one of those.
+This was done to increase the diversity of the model ensemble.
+As with the overlapping, the fusion of the 8 individual predictions was done by a averaging the softmax values of each base pair prediction.
 

### 2020-01-31 cluster_eval_predictions/predictions/animals/start_eval.sh
animal queuing script

@@ -11,19 +11,19 @@ increase_line_number() {
 }
 
 if [[ $# -lt 2 ]]; then
-	echo "Usage: ./start_eval model_path datasets_basepath"
-	exit
+	echo "Usage: ./start_eval model_path datasets_basepath [sleep time in s]"
+	exit 1
 fi
 
 if [[ ! -f "next_line" ]]; then
 	echo "next_line file not found. Script has to be run from the root dir."
-	exit
+	exit 1
 fi
 
 line_offset=$(<next_line)
 if [[ $line_offset -gt $(cat datasets | wc -l) ]]; then
 	echo "Line offset $line_offset is too large. Maybe everything is done already."
-	exit
+	exit 1
 fi
 
 model=$1
@@ -35,7 +35,7 @@ eval_len=$(echo -n $line_content | cut -d "," -f2)
 if [[ -d $species ]]; then
 	echo "$species directory already existing. exiting."
 	increase_line_number
-	exit
+	exit 2
 fi
 
 mkdir $species
@@ -63,7 +63,7 @@ elif [[ "$eval_len" == "200" ]]; then
 	values[__CORE_LENGTH__]="100000"
 else
 	echo "Unknown length $eval_len. exiting."
-	exit
+	exit 1
 fi
 
 # general params
@@ -89,5 +89,9 @@ echo "line offset: $line_offset"
 qsub $species".sh"
 echo "job queued"
 cd ..
+
+if [[ $# -gt 2 ]]; then
+	sleep $3
+fi
 
 increase_line_number

### 2020-01-31 cluster_eval_predictions/predictions/animals/start_many.sh

@@ -33,5 +33,4 @@ fi
 
 for i in $(seq 1 $n_qsubs); do
 	./start_eval.sh $model_file $main_data_folder
-	sleep 20
 done

### 2020-01-31 cluster_eval_predictions/predictions/queue_animals.sh

@@ -1,0 +1,36 @@
+#! /bin/bash
+
+while true; do
+	for job in ZCPHo ZCPHo_6 duZo1 duZo1_7 lH8Hq lH8Hq_5 home home_5; do
+		echo "switching to $job"
+
+		# remove all failed jobs and reset next_line count
+		job_folder=/gpfs/project/festi100/jobs/$job
+		cd $job_folder
+		for species_folder in $(ls -d $job_folder/*/); do
+			if [[ ! -f $species_folder/predictions.h5 ]]; then
+				rm -rv $species_folder
+			fi
+		done;
+		echo -n "1" > next_line
+		echo "$job_folder/next_line reset"
+
+		while true; do
+			# queue jobs if theoretically possible
+			# 30 jobs seems to be the maximum for now
+			if [[ $(qstat -u festi100 | grep CUDA | grep " R " | egrep -c -v "00:00$") -lt 30 ]]; then
+				./start_eval.sh /gpfs/project/festi100/models/animals_final/$job".h5" /gpfs/project/festi100/data/animals 120
+				if [[ $? -eq 1 ]]; then
+					exit 1 # something went very wrong
+					# an exit code of 2 signals that the directory already existed, from which we just move on to try the next
+				fi
+				if [[ $(<next_line) -gt 192 ]]; then
+					break
+				fi
+			else
+				echo "found 30 jobs running"
+				sleep 120
+			fi
+		done
+	done
+done

### 2020-01-31 paper/main.tex

@@
 We also used multiple techniques to improve the prediction quality after the training was done.
-One very effective way for genomes with larger genes was to input longer sequences during inference than during training.
+One very effective way for genomes with larger genes was to input longer sequences during inference time than during training.
 This was done for all animal genomes except the invertebrates.
 The input sequences were up to 10 times larger, depending on species group and assembly quality.
-This also demonstrates the generalization capability of the model as it is able to make successful prediction on far longer sequences than it has ever seen.
+This also demonstrates ability of our model to generalize as it is able to make successful prediction on far longer sequences than it has ever seen.
 The sequence length during inference can be any of 20.000, 50.000, 100.000 or 200.000 and is determined by the species family(?).
-Families that tend to have longer genes get assigned longer sequence chunks.
-If the N75 of a species is less than twice as high as the supposed sequence length it is lowered until either criteria is met or a length of 50.000 is reached.
+Families that tend to have longer genes get assigned longer sequences.
+If the N75 of a species is less than twice as high as the supposed sequence length it is lowered until either this criteria is met or a length of 50.000 is reached.
 For implementation details on this and the following techniques see Supplementary Material (?) or the source code..
@@
 The different overlapping sequences were recombined by averaging the individual softmax values of each base.
+Supplementary Material (?) contains the influence on overlapping with respect to the position in a sequence for each genome, ordered by N75.
+We found that overlapping works best if the genomes are not very fragmented.
 
@@
 For this, we trained 4 seperate animal and plant models and also used a model from an epoch that did not have the highest Genic F1 on the validation set.
-Depending on the results, we used a model from an epoch with either the highest genic recall or precision.
+Depending on the results, we used the model iteration with either the highest genic recall or precision.
 The model with the best Genic F1 was always best in one of those.
 This was done to increase the diversity of the model ensemble.
-As with the overlapping, the fusion of the 8 individual predictions was done by a averaging the softmax values of each base pair prediction.
+As with the overlapping, the fusion of the 8 individual predictions was done by averaging the softmax values of each base pair prediction.
 

### 2020-01-31 cluster_eval_predictions/predictions/queue_animals.sh
minor

@@ -2,7 +2,7 @@
 
 while true; do
 	for job in ZCPHo ZCPHo_6 duZo1 duZo1_7 lH8Hq lH8Hq_5 home home_5; do
-		echo "switching to $job"
+		echo -e "\nswitching to $job"
 
 		# remove all failed jobs and reset next_line count
 		job_folder=/gpfs/project/festi100/jobs/$job
@@ -18,7 +18,7 @@ while true; do
 		while true; do
 			# queue jobs if theoretically possible
 			# 30 jobs seems to be the maximum for now
-			if [[ $(qstat -u festi100 | grep CUDA | grep " R " | egrep -c -v "00:00$") -lt 30 ]]; then
+			if [[ $(qstat -u festi100 | grep CUDA | grep -c " R ") -lt 30 ]]; then
 				./start_eval.sh /gpfs/project/festi100/models/animals_final/$job".h5" /gpfs/project/festi100/data/animals 120
 				if [[ $? -eq 1 ]]; then
 					exit 1 # something went very wrong

### 2020-01-31 cluster_eval_predictions/predictions/queue_animals.sh
minor

@@ -16,9 +16,9 @@ while true; do
 		echo "$job_folder/next_line reset"
 
 		while true; do
-			# queue jobs if theoretically possible
+			# queue jobs if theoretically possible and we have no more jobs than 10 queued atm
 			# 30 jobs seems to be the maximum for now
-			if [[ $(qstat -u festi100 | grep CUDA | grep -c " R ") -lt 30 ]]; then
+			if [[ $(qstat -u festi100 | grep CUDA | grep -c " Q ") -lt 11 && $(qstat -u festi100 | grep CUDA | grep -c " R ") -lt 30 ]]; then
 				./start_eval.sh /gpfs/project/festi100/models/animals_final/$job".h5" /gpfs/project/festi100/data/animals 120
 				if [[ $? -eq 1 ]]; then
 					exit 1 # something went very wrong

### 2020-01-31 data_scripts/parse.py
rm

@@ -1,18 +1,0 @@
-#! /usr/bin/env python3
-
-data = [['genome', 'percent_intergenic', 'percent_error']]
-genome = []
-for line in open('all_genomes_generation_output_truncated'):
-    if line.startswith('/home/'):
-        genome.append(line.split('/')[6])
-    elif line.startswith('Total intergenic'):
-        genome.append('{:.4f}'.format(float(line.split(' ')[2].split('%')[0]) / 100))
-    elif line.startswith('Total errors'):
-        genome.append('{:.4f}'.format(float(line.split(' ')[2].split('%')[0]) / 100))
-        data.append(genome)
-        genome = []
-
-for genome in data:
-    print('\t'.join(genome))
-
-

### 2020-01-31 cluster_eval_predictions/predictions/queue_animals.sh
minor

@@ -8,6 +8,7 @@ while true; do
 		job_folder=/gpfs/project/festi100/jobs/$job
 		cd $job_folder
 		for species_folder in $(ls -d $job_folder/*/); do
+			# if [[ -f $species_folder/*.sh.o* && ! -f $species_folder/predictions.h5 ]]; then
 			if [[ ! -f $species_folder/predictions.h5 ]]; then
 				rm -rv $species_folder
 			fi

### 2020-02-01 cluster_eval_predictions/predictions/animals/local/download_data_sets.sh
minor

@@ -1,0 +1,9 @@
+#! /bin/bash
+
+# download only the test_data.h5 files of the length per animal that is actually needed
+# for prediction according to the datasets file
+
+for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+
+
+done

### 2020-02-02 cluster_eval_predictions/check-datasets.py
minor

@@ -10,13 +10,16 @@ import argparse
 import shutil
 
 parser = argparse.ArgumentParser()
-parser.add_argument('folder', type=str)
+parser.add_argument('job_id', type=str)
+parser.add_argument('-jb', '--job-base-folder', type=str, default='/gpfs/project/festi100/jobs')
 parser.add_argument('-sf', '--sub-folder', type=str, default='')
 parser.add_argument('-del', '--delete', action='store_true', help='Remove folders with errors')
 args = parser.parse_args()
 
-for genome_folder in os.listdir(args.folder):
-    folder = os.path.join(args.folder, genome_folder)
+job_folder = os.path.join(args.job_base_folder, args.job_id)
+
+for genome_folder in os.listdir(job_folder):
+    folder = os.path.join(job_folder, genome_folder)
     if os.path.isdir(folder):
         if args.sub_folder == '':
             h5_files = glob.glob(os.path.join(folder, '*.h5'))

### 2020-02-02 cluster_eval_predictions/predictions/animals/local/download_data_sets.sh
local predictions

@@ -3,7 +3,16 @@
 # download only the test_data.h5 files of the length per animal that is actually needed
 # for prediction according to the datasets file
 
-for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+animals_datasets_base="/mnt/data/datasets/animals/"
 
+# for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mola_mola|lepisosteus_oculatus|erpetoichthys_calabaricus|saimiri_boliviensis_boliviensis" ~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+	species=$(echo -n $line | cut -d, -f1)
+	length=$(echo -n $line | cut -d, -f2)
 
+	species_dir=$animals_datasets_base"/single_genomes_"$length"k/"$species
+	mkdir -v -p $species_dir
+
+	# download from the clc
+	rsync --progress felix-stiehler@134.99.200.63:/home/felix-stiehler/Desktop/data/animals/single_genomes_$length"k/"$species/test_data.h5 $species_dir/
 done

### 2020-02-02 cluster_eval_predictions/predictions/animals/local/mk_predictions.sh

@@ -1,0 +1,51 @@
+#! /bin/bash
+
+# make local predictions on a set of animals
+
+animals_datasets_base="/mnt/data/datasets/animals/"
+job_id="duZo1"
+model="/home/felix/Desktop/models/animals_final/"$job_id".h5"
+predictions_folder_base="/mnt/data/predictions/"$job_id
+
+# for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mola_mola|lepisosteus_oculatus|erpetoichthys_calabaricus|saimiri_boliviensis_boliviensis" ~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+	species=$(echo -n $line | cut -d, -f1)
+	length=$(echo -n $line | cut -d, -f2)
+
+	data_dir=$animals_datasets_base"/single_genomes_"$length"k/"$species
+	predictions_dir=$predictions_folder_base/$species
+	mkdir -v -p $predictions_dir
+
+	case $length in
+		20)
+			batch_size=200
+			overlap_offset=2500
+			core_length=10000
+			;;
+		50)
+			batch_size=80
+			overlap_offset=6250
+			core_length=25000
+			;;
+		100)
+			batch_size=40
+			overlap_offset=12500
+			core_length=50000
+			;;
+		200)
+			batch_size=24
+			overlap_offset=25000
+			core_length=100000
+			;;
+		*)
+			echo "unknown length"
+			exit 1
+			;;
+	esac
+
+	# make predictions
+	/home/felix/git/HelixerPrep/helixerprep/prediction/LSTMModel.py -v -bs $batch_size -lm $model -td $data_dir/test_data.h5 -po $predictions_dir/predictions.h5 --overlap --overlap-offset $overlap_offset --core-length $core_length
+
+	# upload to the cluster in the background
+	rsync $predictions_dir/predictions.h5 festi100@hpc.rz.uni-duesseldorf.de:/gpfs/project/festi100/jobs/$job_id/$species/predictions.h5 &
+done

### 2020-02-03 cluster_eval_predictions/predictions/animals/local/mk_predictions.sh
minor

@@ -47,5 +47,5 @@ for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mol
 	/home/felix/git/HelixerPrep/helixerprep/prediction/LSTMModel.py -v -bs $batch_size -lm $model -td $data_dir/test_data.h5 -po $predictions_dir/predictions.h5 --overlap --overlap-offset $overlap_offset --core-length $core_length
 
 	# upload to the cluster in the background
-	rsync $predictions_dir/predictions.h5 festi100@hpc.rz.uni-duesseldorf.de:/gpfs/project/festi100/jobs/$job_id/$species/predictions.h5 &
+	rsync -r -v $predictions_dir festi100@hpc.rz.uni-duesseldorf.de:/gpfs/project/festi100/jobs/$job_id/ &
 done

### 2020-02-04 paper/main.tex
deleted old csv data, some paper

@@
 \botrule
-\end{tabular}}{Values are averages across all genomes of a data group}
+\end{tabular}}{Values are averages across all genomes of a data group. The genome size is excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.}
 \end{table}
@@
 Augustus & 0.62 (+-0.02) & 0.83 (+- 0.03)\\
+Dilated CNN Baseline & 0.62 (+-0.02) & 0.83 (+- 0.03)\\
 Best Helixer model & 0.62 (+-0.02) & 0.72 (+-0.02) \\
@@
 \citealp{Boffelli03} might want to know about  text text text text
+
+\section{Results}
+
+We propose a framework to develop a gene calling model that is both more general in the range genomes it can annotate and also outperforms traditional methods on a base pair wise evaluation.
+The overall results of Helixer, Augustus and a baseline build with dilated convolutional layers are shown in table \ref{Tab:03}.
 

### 2020-02-06 cluster_eval_predictions/predictions/animals/local/download_data_sets.sh
local predictions

@@ -2,13 +2,18 @@
 
 # download only the test_data.h5 files of the length per animal that is actually needed
 # for prediction according to the datasets file
+# only downloads single_genome files that are not 200k long to run those locally
 
 animals_datasets_base="/mnt/data/datasets/animals/"
 
-# for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
-for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mola_mola|lepisosteus_oculatus|erpetoichthys_calabaricus|saimiri_boliviensis_boliviensis" ~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+# for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mola_mola|lepisosteus_oculatus|erpetoichthys_calabaricus|saimiri_boliviensis_boliviensis" ~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
 	species=$(echo -n $line | cut -d, -f1)
 	length=$(echo -n $line | cut -d, -f2)
+
+	if [[ $length -eq 200 ]]; then
+		continue
+	fi
 
 	species_dir=$animals_datasets_base"/single_genomes_"$length"k/"$species
 	mkdir -v -p $species_dir

### 2020-02-06 cluster_eval_predictions/predictions/animals/local/mk_predictions.sh

@@ -1,16 +1,26 @@
 #! /bin/bash
 
 # make local predictions on a set of animals
+# skip the 200k long ones
 
+if [[ $# -lt 1 ]]; then
+	echo "Usage: ./mk_predictions.sh jobid"
+	exit
+fi
+
+job_id=$1
 animals_datasets_base="/mnt/data/datasets/animals/"
-job_id="duZo1"
 model="/home/felix/Desktop/models/animals_final/"$job_id".h5"
 predictions_folder_base="/mnt/data/predictions/"$job_id
 
-# for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
-for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mola_mola|lepisosteus_oculatus|erpetoichthys_calabaricus|saimiri_boliviensis_boliviensis" ~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+for line in $(<~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
+# for line in $(egrep "microcebus_murinus|dromaius_novaehollandiae|esox_lucius|mola_mola|lepisosteus_oculatus|erpetoichthys_calabaricus|saimiri_boliviensis_boliviensis" ~/git/helixer_scratch/cluster_eval_predictions/predictions/animals/cluster/datasets); do
 	species=$(echo -n $line | cut -d, -f1)
 	length=$(echo -n $line | cut -d, -f2)
+
+	if [[ $length -eq 200 ]]; then
+		continue
+	fi
 
 	data_dir=$animals_datasets_base"/single_genomes_"$length"k/"$species
 	predictions_dir=$predictions_folder_base/$species

### 2020-02-10 paper/main.tex
minor
added some stats

@@
 \begin{tabular}{@{}lll@{}}
-\toprule & EnsemblMetazoa & JGI Phytosome\\
+\toprule & Animals & Plants\\
 \midrule
-Fraction of classes  & 0.6, 0.02, 0.1, 0.2 & 0.6, 0.02, 0.1, 0.2 \\
-(Intergenic, UTR, CDS, Intron) & & \\
-Geenuff error rate & 0.1 (+- 0.02) & 0.3 (+- 0.02) \\
 Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
-Average gene length & 23000 (+- 5000) & 3370 (+- 50) \\
+Average gene length & tbd & tbd \\
+Geenuff error rate & tbd & tbd \\
+Fraction of class Intergenic  & 0.777 (+- 0.042) & 0.799 (+- 0.089) \\
+Fraction of class UTR & 0.006 (+- 0.006) & 0.017 (+- 0.016) \\
+Fraction of class CDS & 0.016 (+- 0.014) & 0.085 (+- 0.071) \\
+Fraction of class Intron & 0.201 (+- 0.035) & 0.099 (+- 0.057) \\
 \botrule
-\end{tabular}}{Values are averages across all genomes of a data group. The genome size is excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.}
+\end{tabular}}{Values are averages of the individual values of each genome in a data group.
+All statistics except the average gene lengths are excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.
+Each strand of DNA was counted seperately.
+The standart deviation is shown in the brackets.}
 \end{table}
@@
 \begin{tabular}{@{}lll@{}}
-\toprule & EnsemblMetazoa & JGI Phytosome\\
+\toprule & Animals & Plants\\
 \midrule
-Fraction of classes  & 0.6, 0.02, 0.1, 0.2 & 0.6, 0.02, 0.1, 0.2 \\
-(Intergenic, UTR, CDS, Intron) & & \\
-Geenuff error rate & 0.1 (+- 0.02) & 0.3 (+- 0.02) \\
-Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
-Average gene length & 23000 (+- 5000) & 3370 (+- 50) \\
+Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
+Average gene length & tbd & tbd \\
+Geenuff error rate & tbd & tbd \\
+Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
+Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
+Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
+Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
 \botrule
-\end{tabular}}{Values are averages across all genomes of a data group. The genome size is excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.}
+\end{tabular}}{Values are averages of the individual values of each genome in a data group.
+All statistics except the average gene lengths are excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.
+Each strand of DNA was counted seperately.
+The standart deviation is shown in the brackets.}
 \end{table}

### 2020-02-11 paper/main.tex
error stats

@@
 \midrule
-Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
+Average genome size in Gbp& 2.936 (+- 1.562) & 0.787 (+- 0.995) \\
 Average gene length & tbd & tbd \\
-Geenuff error rate & tbd & tbd \\
+Geenuff error rate & 0.311 (+- 0.129) & 0.351 (+- 0.249) \\
 Fraction of class Intergenic  & 0.777 (+- 0.042) & 0.799 (+- 0.089) \\
@@
 Average gene length & tbd & tbd \\
-Geenuff error rate & tbd & tbd \\
+Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
 Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\

### 2020-02-11 paper/main.tex
gene length stats

--- 2020-02-04 paper/main.tex
@@
 \begin{tabular}{@{}lll@{}}
-\toprule & EnsemblMetazoa & JGI Phytosome\\
+\toprule & Animals & Plants\\
 \midrule
-Fraction of classes  & 0.6, 0.02, 0.1, 0.2 & 0.6, 0.02, 0.1, 0.2 \\
-(Intergenic, UTR, CDS, Intron) & & \\
-Geenuff error rate & 0.1 (+- 0.02) & 0.3 (+- 0.02) \\
-Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
-Average gene length & 23000 (+- 5000) & 3370 (+- 50) \\
+Average genome size in Gbp& 2.936 (+- 1.562) & 0.787 (+- 0.995) \\
+Average gene length & 31,223 (+- 13,974)& 3,368 (+- 1,510)\\
+Geenuff error rate & 0.311 (+- 0.129) & 0.351 (+- 0.249) \\
+Fraction of class Intergenic  & 0.777 (+- 0.042) & 0.799 (+- 0.089) \\
+Fraction of class UTR & 0.006 (+- 0.006) & 0.017 (+- 0.016) \\
+Fraction of class CDS & 0.016 (+- 0.014) & 0.085 (+- 0.071) \\
+Fraction of class Intron & 0.201 (+- 0.035) & 0.099 (+- 0.057) \\
 \botrule
-\end{tabular}}{Values are averages across all genomes of a data group. The genome size is excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.}
+\end{tabular}}{Values are averages of the individual values of each genome in a data group.
+All statistics except the average gene lengths are excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.
+Each strand of DNA was counted seperately.
+The average gene length was determined by the average length of the longest transcript by cumulative exon length, which is how we query for genes during data generation.
+Brackets show the standart deviation.}
 \end{table}
@@
 \begin{tabular}{@{}lll@{}}
-\toprule & EnsemblMetazoa & JGI Phytosome\\
+\toprule & Animals & Plants\\
 \midrule
-Fraction of classes  & 0.6, 0.02, 0.1, 0.2 & 0.6, 0.02, 0.1, 0.2 \\
-(Intergenic, UTR, CDS, Intron) & & \\
-Geenuff error rate & 0.1 (+- 0.02) & 0.3 (+- 0.02) \\
-Average genome size in Gbp& 1.1 (+- 0.5) & 0.3 (+- 0.1) \\
-Average gene length & 23000 (+- 5000) & 3370 (+- 50) \\
+Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
+Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
+Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
+Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
+Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
+Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
+Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
 \botrule
-\end{tabular}}{Values are averages across all genomes of a data group. The genome size is excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.}
+\end{tabular}}{The description of Table \ref{Tab:statistics_all} applies here as well.}
 \end{table}

### 2020-03-26 paper/main.tex
some paper

@@
 \section{Approach}
-Helixer is a prototype software for training and utilizing a general purpose gene annotation model for large eukaryotic genomes.
-We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on 192 genomes in the animal case and 60 genomes for the plant model.
-We leverage modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields.
+Helixer is a prototype software for training and utilizing a general purpose model for the functional annotation of large eukaryotic genomes.
+We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on 192 genomes in the animal case and on 60 plant genomes.
+We use modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields.
 Our model expects the raw genomic sequence as main input and produces functional predictions on a base pair wise level.
@@
 Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
-Gene lengths & The length of a gene at a base or 0 if intergenic \\
 RNASeq score & A score in the interval $...$ to rank the coherence of the \\

### 2020-03-27 paper/main.tex
minor

@@
 \begin{tabular}{@{}lll@{}}
-\toprule & EnsemblMetazoa & JGI Phytosome\\
+\toprule & Animals & Plants\\
 \midrule
-Augustus & 0.62 (+-0.02) & 0.83 (+- 0.03)\\
-Dilated CNN Baseline & 0.62 (+-0.02) & 0.83 (+- 0.03)\\
-Best Helixer model & 0.62 (+-0.02) & 0.72 (+-0.02) \\
-Best model + varied input length & 0.62 (+-0.02) & 0.72 (+-0.02) \\
-Best model + varied input length + overlapping & 0.62 (+-0.02) & 0.72 (+-0.02) \\
-Varied input length + overlapping + ensembling & 0.62 (+-0.02) & 0.72 (+-0.02) \\
+Augustus & 0.632 (+-0.089) & - (+-)\\
+Dilated CNN Baseline & - (+-) & - (+-)\\
+Best Helixer model & - (+-) & - (+-) \\
+Best model + varied input length & - (+-) & - (+-) \\
+Best model + varied input length + overlapping & - (+-) & - (+-) \\
+Varied input length + overlapping + ensembling & - (+-) & - (+-) \\
 \botrule
-\end{tabular}}{Values are averages in CDS F1 (?) across all validation and test genomes of the \\ respective group.}
+\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group.}
 \end{table}

### 2020-03-28 paper/main.tex
numbers

@@
 \midrule
-Augustus & 0.632 (+-0.089) & - (+-)\\
+Augustus & 0.632 (+-0.089) & 0.749 (+-0.161)\\
 Dilated CNN Baseline & - (+-) & - (+-)\\
 Best Helixer model & - (+-) & - (+-) \\
-Best model + varied input length & - (+-) & - (+-) \\
-Best model + varied input length + overlapping & - (+-) & - (+-) \\
-Varied input length + overlapping + ensembling & - (+-) & - (+-) \\
+Best model + varied input length & - (+-) & - \\
+Best model (+ varied input length) + overlapping &  - (+-) & 0.843 (+-0.151) \\
+(Varied input length +) overlapping + ensembling & 0.868 (+-0.064) & 0.863 (+-0.148) \\
 \botrule
-\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group.}
+\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group. Varied input length was only used in the animal case.}
 \end{table}

### 2020-03-28 paper/main.tex
best animal model number in paper

@@
 Best model + varied input length & - (+-) & - \\
-Best model (+ varied input length) + overlapping &  - (+-) & 0.843 (+-0.151) \\
+Best model (+ varied input length) + overlapping &  0.85 (+-0.068) & 0.843 (+-0.151) \\
 (Varied input length +) overlapping + ensembling & 0.868 (+-0.064) & 0.863 (+-0.148) \\

### 2020-04-04 plots/phylogenetic_trees/ete3/plot.py
ete3

@@ -1,0 +1,20 @@
+#!/usr/bin/env python3
+import numpy as np
+import ete3
+from ete3 import ClusterTree, ProfileFace, ArrayTable, TreeStyle, AttrFace
+from ete3.treeview.faces import add_face_to_node
+
+# tree = ete3.PhyloTree("all_species.tre", sp_naming_function=lambda node: node.name)
+data = ArrayTable("results.array")
+ct = ete3.ClusterTree("all_species.tre", data)
+data_max = np.max(ct.arraytable.matrix)
+data_min = np.min(ct.arraytable.matrix)
+
+def mylayout(node):
+    if node.is_leaf():
+        profile_face = ProfileFace(data_max, data_min, 0.0, 100, 14, "heatmap")
+        ete3.treeview.faces.add_face_to_node(profile_face, node, 0, aligned=True)
+
+ts = TreeStyle()
+ts.layout_fn = mylayout
+ct.show(tree_style=ts)

### 2020-04-06 plots/phylogenetic_trees/ete3/plot.py
stuff

@@ -1,7 +1,7 @@
 #!/usr/bin/env python3
 import numpy as np
 import ete3
-from ete3 import ClusterTree, ProfileFace, ArrayTable, TreeStyle, AttrFace
+from ete3 import ClusterTree, ProfileFace, ArrayTable, TreeStyle, AttrFace, CircleFace, TextFace
 from ete3.treeview.faces import add_face_to_node
 
 # tree = ete3.PhyloTree("all_species.tre", sp_naming_function=lambda node: node.name)
@@ -16,5 +16,8 @@ def mylayout(node):
         ete3.treeview.faces.add_face_to_node(profile_face, node, 0, aligned=True)
 
 ts = TreeStyle()
+ts.show_scale = False
 ts.layout_fn = mylayout
+ts.legend.add_face(TextFace("0.5 support"), column=0)
+ts.legend_position = 4
 ct.show(tree_style=ts)

### 2020-04-06 plots/phylogenetic_trees/ggtree/plot.R

@@ -7,12 +7,13 @@ tree = read.nhx('all_species.tre')
 # tree_human = tree_subset(tree, node="Homosapiens", levels_back=5)
 
 
-f1_difference = read.csv('f1_difference.csv', header=TRUE)
+results = read.csv('f1_scores.csv', header=TRUE)
 sets = read.csv('sets.csv', header=TRUE)
 p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set)) 
 # p = p + geom_label(aes(label=node), fill='steelblue')
 p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
 png('animals.png', width=1200, height=2000)
-gheatmap(p, f1_difference, width=.2, high='green', low='red', colnames=F, offset=5) 
+gheatmap(p, results, width=.2, low="red", mid="white", high="blue", midpoint=0.8, 
+		 colnames=F, offset=5) 
 dev.off()

### 2020-04-06 plots/phylogenetic_trees/ete3/plot.py
stuff

@@ -5,15 +5,23 @@ from ete3 import ClusterTree, ProfileFace, ArrayTable, TreeStyle, AttrFace, Circ
 from ete3.treeview.faces import add_face_to_node
 
 # tree = ete3.PhyloTree("all_species.tre", sp_naming_function=lambda node: node.name)
-data = ArrayTable("results.array")
+data = ArrayTable("one_result.array")
 ct = ete3.ClusterTree("all_species.tre", data)
-data_max = np.max(ct.arraytable.matrix)
-data_min = np.min(ct.arraytable.matrix)
+data_max = np.max(ct.arraytable.matrix, axis=0)
+data_min = np.min(ct.arraytable.matrix, axis=0)
+data_median = np.median(ct.arraytable.matrix, axis=0)
 
 def mylayout(node):
     if node.is_leaf():
-        profile_face = ProfileFace(data_max, data_min, 0.0, 100, 14, "heatmap")
+        if node.name == "Danio_rerio":
+            # import pudb; pudb.set_trace()
+            pass
+        # profile_face = ProfileFace(data_max[1], data_min[1], 0.0, 100, 14, "heatmap")
+        profile_face = ProfileFace(data_max[0], data_min[0], 0.0, 100, 14, "heatmap")
         ete3.treeview.faces.add_face_to_node(profile_face, node, 0, aligned=True)
+
+        # profile_face = ProfileFace(data_max[0], data_min[0], data_median[0], 100, 14, "heatmap")
+        # ete3.treeview.faces.add_face_to_node(profile_face, node, 1, aligned=True)
 
 ts = TreeStyle()
 ts.show_scale = False

### 2020-04-07 paper/main.tex
minor
paper

@@
 Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
+We call the genomes that were ultimatily used during training the {\it training genomes} and the remaining non-test genomes the {\it validation genomes}.
 
@@
 
-Once a set of training genomes was selected, we generated a validation set from it by splitting of 20\% of the FASTA sequences above and below the N90 of each genome, ensuring a proper distribution of large and short sequences in both the training and validation set and a split on the chromosome level.
-The validation set was only used to roughly judge the performance of our models during training.
-Evaluation of the annotations on all training and validation genomes was done regularly, as we ultimately only cared about the performance on for the model fully unknown genomes.
+Once a set of training genomes was selected, we split it up again into a {\it training and validation set}.
+This split of the training genomes was done to get a quicker sense of the generalization capabilities and was used primarily during training.
+We split of the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level.
+Evaluation of the annotations on all training and validation genomes was done regularly after a promising model candidate was found based on its validation set performance.
+Ultimately we only cared about the cross-species performance of a model.
 
@@
 \botrule
-\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group. Varied input length was only used in the animal case.}
+\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group.
+Varied input length was only used in the animal case.
+The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.}
 \end{table}

### 2020-04-07 paper/main.tex
lots of reordering

@@
 
-\begin{table}[!t]
-\processtable{Data group statistics for our training data\label{Tab:statistics_training}} {
-\begin{tabular}{@{}lll@{}}
-\toprule & Animals & Plants\\
-\midrule
-Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
-Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
-Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
-Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
-Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
-Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
-Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
-\botrule
-\end{tabular}}{The description of Table \ref{Tab:statistics_all} applies here as well.}
-\end{table}
-
 During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?).
@@
 The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it.
-Table \ref{Tab:statistics_all} and \ref{Tab:statistics_training} show statistics about the masking.
+Table \ref{Tab:statistics_all} shows statistics about the masking.
+The corresponding statistics for the selected training genomes can be found in Table \ref{Tab:statistics_training}.
 For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network.
@@
 Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
-RNASeq score & A score in the interval $...$ to rank the coherence of the \\
-		     & annotation with RNASeq information \\
 \botrule
@@
 \end{table}
-
-\begin{table}[!t]
-\processtable{Hyperparameters\label{Tab:02}} {
-\begin{tabular}{@{}ll@{}}
-\toprule Name & Value \\
-\midrule
-Layers & 4\\
-Units per layer & 256\\
-Learning rate & 1e-4\\
-Optimizer & Adam\\
-\botrule
-\end{tabular}}{}
-\end{table}
-
-text\vspace*{1pt}
-
-\begin{itemize}
-\item for bulleted list, use itemize
-\item for bulleted list, use itemize
-\item for bulleted list, use itemize\vspace*{1pt}
-\end{itemize}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
 
@@
 \end{figure}
-
-%\begin{figure}[!tpb]%figure2
-%%\centerline{\includegraphics{fig02.eps}}
-%\caption{Caption, caption.}\label{fig:02}
-%\end{figure}
 
@@
 
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-\citealp{Boffelli03} might want to know about  text text text text
-
 \section{Results}
 
-We propose a framework to develop a gene calling model that is both more general in the range genomes it can annotate and also outperforms traditional methods on a base pair wise evaluation.
-The overall results of Helixer, Augustus and a baseline build with dilated convolutional layers are shown in table \ref{Tab:03}.
-
-\section{Discussion}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text.
-
-Table~\ref{Tab:01} shows that Text Text Text Text Text  Text Text
-Text Text Text Text. Figure~2\vphantom{\ref{fig:02}} shows that
-the above method Text Text. Text Text Text  Text Text Text Text
-Text Text. Figure~2\vphantom{\ref{fig:02}} shows that the above
-method Text Text. Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
-Text.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
-%     as it may ignore the figures.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We evaluate the performance of our final model base on ... The overall results are shown in table \ref{Tab:03}.
 
 \section{Conclusion}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text\vspace*{-10pt}
+We propose a gene calling model that is able to perform cross-species genomic annotiation for a wide variety of large eucaryotic genomes.
+Our model does this while outperforming AUGUSTUS on a base pair wise level and withstanding RNASeq data based scrutiny.
 
@@
 
-Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text
-text text text\vspace*{-12pt}
+We are very grateful for the support of...
 
@@
 \end{thebibliography}
+
+\newpage
+
+\section{Supplementary Material}
+\subsection{Dilated CNN Baseline Network Search}
+
+\subsection{Input Data Encoding}
+
+\subsection{Hyperparameters}
+
+\begin{table}[!t]
+\processtable{Hyperparameters\label{Tab:02}} {
+\begin{tabular}{@{}ll@{}}
+\toprule Name & Value \\
+\midrule
+Layers & 4\\
+Units per layer & 256\\
+Learning rate & 1e-4\\
+Optimizer & Adam\\
+\botrule
+\end{tabular}}{}
+\end{table}
+
+\subsection{Training Data statistics}
+\begin{table}[!t]
+\processtable{Data group statistics for our training data\label{Tab:statistics_training}} {
+\begin{tabular}{@{}lll@{}}
+\toprule & Animals & Plants\\
+\midrule
+Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
+Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
+Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
+Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
+Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
+Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
+Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
+\botrule
+\end{tabular}}{The description of Table \ref{Tab:statistics_all} applies here as well.}
+\end{table}
+
 \end{document}

### 2020-04-07 paper/main.tex
few things here and there

@@
 Table \ref{Tab:statistics_all} shows statistics about the masking.
-The corresponding statistics for the selected training genomes can be found in Table \ref{Tab:statistics_training}.
+The corresponding statistics for the selected training genomes only can be found in Section \ref{sec:training_data}.
+
 For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network.
 We appended zero padding if the subsequences are not long enough.
-If a subsequence is fully marked as erroneous, we do not include it in the training or evaluation datasets.
+If a subsequence is fully marked as erroneous, we do not include it anywhere.
 
@@
 \subsection{Model Architecture and Usage}
+\label{sec:model}
 We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization (citation ?
@@
 Families that tend to have longer genes get assigned longer sequences.
-If the N75 of a species is less than twice as high as the supposed sequence length it is lowered until either this criteria is met or a length of 50.000 is reached.
-For implementation details on this and the following techniques see Supplementary Material (?) or the source code..
+For more implementation details on this see Section \ref{sec:longer} or the source code.
 
-We also construct the final predictions of a single model by overlapping predictions that were made from a sliding window and then cut to a core sequence.
+The final predictions of a single model are constructed by overlapping predictions, that were made from a sliding window and then cut to a core sequence.
 This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see figure \ref{fig:length_wise_bias}) and also to improve model performance in general by providing the model with multiple different starting points.
@@
 Varied input length was only used in the animal case.
-The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.}
+The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.
+The Augustus result is the median score for all genomes.}
 \end{table}
@@
 \section{Results}
-
-We evaluate the performance of our final model base on ... The overall results are shown in table \ref{Tab:03}.
+We evaluate our performance based on the average cross-species Subgenic F1 for all genomes except the few used during training.
+The results are displayed in Table \ref{Tab:03} and show a clear improvement over the AUGUSTUS and dilated CNN baselines.
+We also demonstrate the effectiveness of two techniques used during inference with the usage of longer input lengths being enabled by our LSTM stack without any fully connected layers on top.
+To our knowledge, this technique was not used before with a model used for gene annotation.
 
@@
 \subsection{Dilated CNN Baseline Network Search}
+\label{sec:dcnn}
 
@@
 \subsection{Training Data statistics}
+\label{sec:training_data}
+
 \begin{table}[!t]
@@
 
+\subsection{Longer Sequence Input}
+\label{sec:longer}
+Implementation detail:
+If the N75 of a species is less than twice as high as the supposed sequence length it is lowered until either this criteria is met or a length of 50.000 is reached.
+
+Table of species with their length.
+
 \end{document}

### 2020-04-07 paper/main.tex
table naming

@@
 An non-coding intron can be inferred from the prediction order.
-A full description of all generated data arrays can be found in table \ref{Tab:01}.
+A full description of all generated data arrays can be found in table \ref{tab:arrays}.
 
 The classes are not balanced equally and the tradeoff between the number of intergenic and intronic bases is quite different in both groups of genomes.
-Table \ref{Tab:statistics_all} shows the ratio of class occurences, together with other overall metrics.
+Table \ref{tab:statistics_all} shows the ratio of class occurences, together with other overall metrics.
 
 \begin{table}[!t]
-\processtable{Data group statistics for all data\label{Tab:statistics_all}} {
+\processtable{Data group statistics for all data\label{tab:statistics_all}} {
 \begin{tabular}{@{}lll@{}}
@@
 The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it.
-Table \ref{Tab:statistics_all} shows statistics about the masking.
+Table \ref{tab:statistics_all} shows statistics about the masking.
 The corresponding statistics for the selected training genomes only can be found in Section \ref{sec:training_data}.
@@
 The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (citation).
-We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks and hybrid architectures.
+We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks (CNN) and hybrid architectures.
 The hyperparameters were determined by a combination of manual and automatic hyperparameter optimization (talk about TPE?, quote nni?).
-All relevant hyperparameters can be found in Table \ref{Tab:02} or in the Helixer source code repository.
+All relevant hyperparameters can be found in Table \ref{tab:params} or in the Helixer source code repository.
+
+We also implemented a dilated CNN baseline as this kind of architecture is widely used when working with DNA data as input.
+Details of the neural architecture search are given in Section \ref{sec:dcnn} and results are shown in Table \ref{tab:results}.
 
@@
 \begin{table}[!t]
-\processtable{Data arrays generated and used by Helixer\label{Tab:01}} {
+\processtable{Data arrays generated and used by Helixer\label{tab:arrays}} {
 \begin{tabular}{@{}ll@{}}
@@
 \begin{table}[!t]
-\processtable{Experimental Results and Comparison\label{Tab:03}} {
+\processtable{Experimental Results and Comparison\label{tab:results}} {
 \begin{tabular}{@{}lll@{}}
@@
 We evaluate our performance based on the average cross-species Subgenic F1 for all genomes except the few used during training.
-The results are displayed in Table \ref{Tab:03} and show a clear improvement over the AUGUSTUS and dilated CNN baselines.
+The results are displayed in Table \ref{tab:results} and show a clear improvement over the AUGUSTUS and dilated CNN baselines.
 We also demonstrate the effectiveness of two techniques used during inference with the usage of longer input lengths being enabled by our LSTM stack without any fully connected layers on top.
@@
 \begin{table}[!t]
-\processtable{Hyperparameters\label{Tab:02}} {
+\processtable{Hyperparameters\label{tab:params}} {
 \begin{tabular}{@{}ll@{}}
@@
 \begin{table}[!t]
-\processtable{Data group statistics for our training data\label{Tab:statistics_training}} {
+\processtable{Data group statistics for our training data\label{tab:statistics_training}} {
 \begin{tabular}{@{}lll@{}}
@@
 \botrule
-\end{tabular}}{The description of Table \ref{Tab:statistics_all} applies here as well.}
+\end{tabular}}{The description of Table \ref{tab:statistics_all} applies here as well.}
 \end{table}

### 2020-04-07 paper/main.tex
dCNN stuff started

@@
 
-We also implemented a dilated CNN baseline as this kind of architecture is widely used when working with DNA data as input.
-Details of the neural architecture search are given in Section \ref{sec:dcnn} and results are shown in Table \ref{tab:results}.
+We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
+Details of the neural architecture search are given in Section \ref{sec:dcnn}.
 
-The model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
+Our final model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
 This grouping enabled us to train effectively with far longer sequences than usual.
@@
 \section{Supplementary Material}
+
 \subsection{Dilated CNN Baseline Network Search}
 \label{sec:dcnn}
+\begin{table}[!t]
+\processtable{Parameter search space for the dilated CNN baseline\label{tab:dcnn}} {
+\begin{tabular}{@{}ll@{}}
+\toprule Name & Value \\
+\midrule
+Layers & 4\\
+Units per layer & 256\\
+Learning rate & 1e-4\\
+Optimizer & Adam\\
+\botrule
+\end{tabular}}{Above are all parameters used during neural architecture search for the dilated CNN baseline.
+The same overall space was used for plant and animal data.
+We did, however, run multiple distinct searches that sometimes only operated over a subset of the given parameters.
+This was done as those seemed to be the most promising.
+We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs.
+Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.}
+\end{table}
+
+\begin{table}[!t]
+\processtable{Final Parameters of dilated CNN baseline\label{tab:dcnn_best}} {
+\begin{tabular}{@{}ll@{}}
+\toprule Name & Animal Value & Plant Value\\
+\midrule
+Layers & 4 & \\
+Units per layer & 256 & \\
+Learning rate & 1e-4 & \\
+Optimizer & Adam & \\
+\botrule
+\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.}
+\end{table}
 
@@
 \subsection{Hyperparameters}
-
 \begin{table}[!t]
@@
 \label{sec:training_data}
-
 \begin{table}[!t]

### 2020-04-07 paper/main.tex
dcnn

@@
 \begin{table}[!t]
-\processtable{Parameter search space for the dilated CNN baseline\label{tab:dcnn}} {
+\processtable{Parameter search space for the dilated CNN baselines\label{tab:dcnn}} {
 \begin{tabular}{@{}ll@{}}
-\toprule Name & Value \\
+\toprule Parameter & Possible Values \\
 \midrule
-Layers & 4\\
-Units per layer & 256\\
-Learning rate & 1e-4\\
-Optimizer & Adam\\
+Kernel size & \{4, 8, 12, 16\}\\
+Initial Filter depth & \{32, 64, 96, 128\}\\
+Number of Layers before doubling filter count & \{1, 2\}\\
+Dilation Multiplier& \{2, 3\}\\
+
 \botrule
@@
 We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs.
-Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.}
+Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.
+Run with a Genic F1 below 0.5 after 10 epochs were stopped at that point and the overall maximum epoch was 15.
+The batch size used for almost all runs was 32 and the dilation was capped to 81.
+The performances seemed to be leveling of well before that.
+The implementation and parameter space is very roughly based on (paper we looked at?) and can be found in the Helixer source code repository.
+In total, we ran (?) runs for the plant and (?) runs with the animal data.}
 \end{table}
@@
 \begin{table}[!t]
-\processtable{Final Parameters of dilated CNN baseline\label{tab:dcnn_best}} {
-\begin{tabular}{@{}ll@{}}
-\toprule Name & Animal Value & Plant Value\\
+\processtable{Final Parameters of dilated CNN baselines\label{tab:dcnn_best}} {
+\begin{tabular}{@{}lll@{}}
+\toprule Parameter & Possible Animal Values & Possible Plant Values\\
 \midrule
@@
 \subsection{Hyperparameters}
+
 \begin{table}[!t]
@@
 \end{tabular}}{}
+\end{table}
+
+\begin{table}[!t]
+\processtable{Class weights\label{tab:params_cw}} {
+\begin{tabular}{@{}lll@{}}
+\toprule Class & Animal Value & Plant Value \\
+\midrule
+[0.7, 1.6, 1.2, 1.2] & 4 &\\
+\botrule
+\end{tabular}}{see data statistics tables..}
 \end{table}

### 2020-04-07 paper/main.tex
dcnn

@@
 \midrule
-Kernel size & \{4, 8, 12, 16\}\\
-Initial Filter depth & \{32, 64, 96, 128\}\\
-Number of Layers before doubling filter count & \{1, 2\}\\
-Dilation Multiplier& \{2, 3\}\\
+Kernel Size & \{4, 8, 12, 16\}\\
+Initial Filter Depth & \{32, 64, 96, 128\}\\
+Number of Layers Before Doubling Filter Count & \{1, 2\}\\
+Dilation Multiplier & \{2, 3\}\\
+Number of Convolutional Layers & \{2, 3, 4, 5, 6, 7, 8\}\\
+Number of Hidden Fully Connected (FC) Layers & \{0, 1, 2\}\\	
+Dropout Used on FC Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
+Learning Rate & \{1e-3, 1e-4\}\\
 
@@
 \botrule
-\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.}
+\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes. }
 \end{table}

### 2020-04-07 paper/main.tex
dcnn values

@@
 Number of Hidden Fully Connected (FC) Layers & \{0, 1, 2\}\\	
-Dropout Used on FC Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
+Dropout Used on FC and Final Conv Layer Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
 Learning Rate & \{1e-3, 1e-4\}\\
@@
 \begin{tabular}{@{}lll@{}}
-\toprule Parameter & Possible Animal Values & Possible Plant Values\\
+\toprule Parameter & Animals & Plants\\
 \midrule
-Layers & 4 & \\
-Units per layer & 256 & \\
-Learning rate & 1e-4 & \\
-Optimizer & Adam & \\
+Kernel Size & 16 & 16\\
+Initial Filter Depth & 96 & 32\\
+Number of Layers Before Doubling Filter Count & 2 & 2\\
+Dilation Multiplier & 3 & 3\\
+Number of Convolutional Layers & 6 & 8\\
+Number of Hidden Fully Connected (FC) Layers & 2 & 0\\
+Dropout Used on FC Output & 0.0 & 0.0\\
+Learning Rate & 1e-4 & 1e-3\\
 \botrule

### 2020-04-07 backup_scripts/backup-experiments-archaeopteryx.sh
forgotten bu

@@ -5,10 +5,16 @@ echo -e "\n----------------\n"
 date
 echo
 
-echo -e "\nnni denbi"
-rsync -rzt --stats --exclude="predictions.h5" triceratops:nni nni_denbi/
-rsync -rzt --stats --exclude="predictions.h5" triceratops:.local/nnictl nni_denbi/
+echo -e "\nnni denbi no longer active"
+#rsync -rzt --stats --exclude="predictions.h5" triceratops:nni nni_denbi/
+#rsync -rzt --stats --exclude="predictions.h5" triceratops:.local/nnictl nni_denbi/
 
 echo -e "\nnni mordred"
 rsync -rzt --stats  --exclude="predictions.h5" mordred:nni nni_mordred/
 rsync -rzt --stats  --exclude="predictions.h5" mordred:.local/nnictl nni_mordred/
+
+echo -e "\nnni aardonyx"
+rsync -rzt --stats  --exclude="predictions.h5" aardonyx:nni nni_aliclc/
+rsync -rzt --stats  --exclude="predictions.h5" aardonyx:.local/nnictl nni_aliclc/
+
+

### 2020-04-07 paper/main.tex
dcnn, still wip

@@
 Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.
-Run with a Genic F1 below 0.5 after 10 epochs were stopped at that point and the overall maximum epoch was 15.
+Runs with a Genic F1 below 0.5 after 10 epochs were stopped at that point and the overall maximum epoch was 15.
 The batch size used for almost all runs was 32 and the dilation was capped to 81.
 The performances seemed to be leveling of well before that.
+New parameters where chosen at random.
 The implementation and parameter space is very roughly based on (paper we looked at?) and can be found in the Helixer source code repository.
-In total, we ran (?) runs for the plant and (?) runs with the animal data.}
+In total, we ran (?) runs for the animals and (?) runs with the plant data.}
 \end{table}
@@
 \botrule
-\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes. }
+\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.
+The best animal model was determined by 7+ runs and has circa 4.6 million parameters; the best plant model has around 2.1 million and .}
 \end{table}
@@
 \midrule
-[0.7, 1.6, 1.2, 1.2] & 4 &\\
+[0.7, 1.6, 1.2, 1.2] & [0.3, 1.0, 0.9, 0.3] &\\
 \botrule

### 2020-04-07 paper/main.tex
minor

@@
 Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.
-Runs with a Genic F1 below 0.5 after 10 epochs were stopped at that point and the overall maximum epoch was 15.
-The batch size used for almost all runs was 32 and the dilation was capped to 81.
-The performances seemed to be leveling of well before that.
+Runs with a Genic F1 below 0.5 after 10 epochs were stopped and the overall maximum epoch was 15.
+The performances seemed to be leveling off well before that.
+The batch size used for almost all runs was 32, the dilation was capped to 81 and the size of each fully connected layer was fixed at 128.
 New parameters where chosen at random.

### 2020-04-11 cluster_eval_predictions/evaluations/overall/eval-from-predictions-one-trial_nothpc.sh
eval from preds on local machine

@@ -1,0 +1,20 @@
+#! /bin/bash
+
+if [[ $# -lt 2 ]]; then
+	echo "Usage: ./eval-from-predictions-one-trial.sh trial_folder output file name"
+	exit
+fi
+
+trial_folder=$1
+output_file_name=$2
+
+# if the folder has a trial.log in it we assume nni, otherwise cluster
+if [[ -f "$trial_folder/trial.log" ]]; then
+	log_file="$trial_folder/trial.log"
+	dataset=$(grep "test_data.h5" $log_file| cut -d " " -f5 | tr -d "'")
+else
+	log_file="$trial_folder/*.sh.o*"
+	dataset=$(grep "test_data.h5" $log_file| cut -d " " -f3 | tr -d ",'")
+fi
+
+$hppath/HelixerPrep/scripts/basic_pred_cm.py -d $dataset -p "$trial_folder/predictions.h5" &>> "$trial_folder/$output_file_name"

### 2020-04-11 cluster_eval_predictions/evaluations/overall/eval-from-predictions_nothpc.sh

@@ -1,0 +1,17 @@
+#! /bin/bash
+
+# evaluates predictions (what would have been in trial.log files when done with nni)
+# from a run that only generated predictions.h5 files in different folders
+# this is useful to save time as this step can be done concurrently
+# uses eval-from-predictions-one-trial.sh to work concurrently
+
+if [[ $# -lt 2 ]]; then
+	echo "Usage: ./eval-from-predictions.sh main_folder output_file_name [--redo]"
+	exit
+fi
+
+main_folder=$1
+output_file_name=$2
+source $hppath/venv/bin/activate
+ls -d "$main_folder"/*/| xargs -n1 -P8 -I% $hppath/../helixer_scratch/cluster_eval_predictions/evaluations/overall/eval-from-predictions-one-trial.sh % $output_file_name
+

### 2020-04-11 paper/main.tex
add first draft for introduction

@@
 \section{Introduction}
-\begin{itemize}
-	\item Calling genes correctly is very important
-	\item but also very difficult in eukaryotes
-	\item All current state of the art gene callers work based on HMMs
-	\item The structure of the HMMs is finely tuned and engineered
-	\item Actual learning is very limited, and with that generalization across different genomes 
-	\item (Also have problems with genes that contain long introns, common in mammals)
-\end{itemize}
+For many tasks, it is necessary to have DNA regions labeled according to their function.
+This process is called gene annotation and it can be performed in different levels of precision, from simple coding -- non-coding classification to detailed structural labeling, such as enhancer regions or long non-coding RNA sequences in eukaryotes.
+In this work on eucaryotic genomes, we differentiate between Intergenic regions, Untranslated regions (UTRS, parts at the beginning and end of gene, that are transcribed but not translated), Coding sequences (CDS) and Introns.
+Because of the sheer size of genomes alone, gene annotation is a very labour intensive process and genomic database providers like NCBI [\citealp{NCBI13}] and Ensembl [\citealp{Ensembl16}] offer gene annotation pipelines.
+These pipelines integrate experimental data (from e.g.
+RNA-seq or proteogenomics) with homolog sequences in the database and ab-initio gene predictions.
+The latter is an attractive approach, because it is cheap and fast.
+State of the art performance ab-initio gene prediction is achieved by higher order hidden markov models (HMMs), such as Genscan [\citealp{Genscan}] or Augustus [\citealp{Augustus}] but their accuracy leaves room for improvement.
+By choosing possible states and transitions, designers of HMMs assume structure in the sequence that may limit its predictive power.
+Indeed, in practice HMMs have problems with generalizing across species and it seems actual learning of underlying patterns is limited.
+In the last decade, deep neural networks have been applied with tremendous success in many areas of statistical modelling.
+For sequence data, such as DNA, recurrent neural networks (RNN) in the form of long-short term memory (LSTM) are an established architecture and here, we employ them to improve on HMMs for ab-initio genome annotation.
+
+RNNs have already shown promising results in related, but generally more limited investigations.
+[\citealp{Choudhary17}] carries out preliminary explorations on the potential of BLSTMs for cross species gene prediction, but trains his model on human genes only and tests it on two more species.
+DeepAnnotator [\citealp{DeepAnnotator}] uses bidirectional LSTMs (BLSTM) for gene finding in prokaryotes.
+Gene prediction in prokaryotes is considered more amenable than in eukaryotes, as genes in prokaryotes are proportionately more frequent in the genome, feature simpler control structures and do not use splicing [\citealp{Zhuo04}].
+DanQ [\citealp{DanQ}] proposes the use of a BLSTM after a concurrent neural net (CNN) to find detailed motifs in the human genome.
+DeePromotor[\citealp{DeePromotor}] trains a similar architecture for the recognition of promotor regions.
+Recently, several groups [\citealp{SpliceAi}, \citealp{SpliceFinder}] successfully used CNNs to find splicing sites.
 
@@
 
+\bibitem[Thibaud-Nissen {\it et~al}, 2013]{NCBI13}
+Thibaud-Nissen,F., Souvorov,A., Murphy,T., DiCuccio,M., Kitts,P. (2013) Eukaryotic Genome Annotation Pipeline{\it The NCBI Handbook, 2nd}
+
+\bibitem[Aken {\it et~al}, 2016]{Ensembl16}
+Aken,B., Ayling,S., Barrell,D., Clarke,L., {\it et~al.} (2016) The Ensembl gene annotation system, {\it Database}, {\bf 2016}, baw093
+
+\bibitem[Burge {\it et} Karlin, 1997]{Genscan}
+Burge,C., Karlin,S. (1997) Prediction of complete gene structures in human genomic DNA, {\it Journal of Molecular Biology}, {\bf 268-1}, 78-94
+
+\bibitem[Stanke {\it et} Waack, 2003]{Augustus}
+	Stanke,M., Waack,S., (2003) Gene prediction with a hidden Markov model and a new intron submodel, {\it Bioinformatics}, {\bf 19 Suppl.2}, ii215–ii225
+
+\bibitem[Zhuo {\it et~al}., 2004]{Zhuo04}
+Zhuo,W., Yazhu,C., Yixue,L. (2004) A Brief Review of Computational Gene Prediction Methods, {\it Genomics, Proteomics \& Bioinformatics}, 216-221
+
+\bibitem[Amin {\it et~al}., 2018]{DeepAnnotator}
+Amin,M., Yurovsky,A., Tian,Y., Skiena,S.
+(2018) DeepAnnotator: Genome Annotation with Deep Learning, {\it Proceedings of the 2018 ACM International Conference on Bioinformatics, Computational Biology, and Health Informatics}, 254-259.
+
+\bibitem[Choudhary, 2017]{Choudhary17}
+Choudhary,S. (2017) Predicting protein coding boundaries using RNNs, http://www.saket-choudhary.me/rnn-cds-prediction/predicting-protein-coding.pdf (accessed 2020-04-11)
+
+\bibitem[Daniel {\it et~al}., 2016]{DanQ}
+Daniel,Q., Xiaohui,X., (2016) DanQ: a hybrid convolutional and recurrent deep neural network for quantifying the function of DNA sequences, {\it Nucleic Acids Research}, {\bf 44-11}, e107
+
+\bibitem[Oubounyt {\it et~al}., 2019]{DeePromotor}
+Oubounyt,M., Louadi,Z., Tayara,H., Chong2,K. (2019) DeePromoter: Robust Promoter Predictor Using Deep Learning., {\it Frontiers in genetics}, {\bf 10}, 286
+
+\bibitem[Jaganathan {\it et~al}., 2019]{SpliceAi}
+Jaganathan,K., Panagiotopoulou,S., McRae,J., Darbandi,S., et.al., Predicting Splicing from Primary Sequence with Deep Learning, {\it Cell}, {\bf 176-3}, 535-548.e24
+
+\bibitem[Wang {\it et~al}., 2019]{SpliceFinder}
+Wang,R., Wang,Z., Wang,J. (2019) SpliceFinder: ab initio prediction of splice sites using convolutional neural network, {\it BMC Bioinformatics}, {\bf 20}, 652
+
 \bibitem[Denton {\it et~al}., 2019]{Denton19}

### 2020-04-12 paper/main.tex
minor

@@
 The implementation and parameter space is very roughly based on (paper we looked at?) and can be found in the Helixer source code repository.
-In total, we ran (?) runs for the animals and (?) runs with the plant data.}
+In total, we trained 30 models for the animals and (?) with the plant data.}
 \end{table}
@@
 \end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.
-The best animal model was determined by 7+ runs and has circa 4.6 million parameters; the best plant model has around 2.1 million and .}
+The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
 \end{table}

### 2020-04-13 paper/main.tex
run count
citation

@@
 
+\bibitem[Gupta {\it et~al}., 2017]{dCNNGupta}
+Gupta,A., Rush.,A. (2017) Dilated convolutions for modeling long-distance genomic dependencies, {\it arXiv preprint arXiv:1710.01278}
+
 \bibitem[Denton {\it et~al}., 2019]{Denton19}
@@
 New parameters where chosen at random.
-The implementation and parameter space is very roughly based on (paper we looked at?) and can be found in the Helixer source code repository.
-In total, we trained 30 models for the animals and (?) with the plant data.}
+The implementation and parameter space is very roughly based on [\citealp{dCNNGupta}] and can be found in the Helixer source code repository.
+In total, we trained 30 models for the animals and 48 with the plant data.}
 \end{table}

### 2020-04-13 paper/main.tex
boxplot images, text

@@
 
+\begin{figure}[!tpb]
+	\centerline{\includegraphics[width=0.45\textwidth]{images/overall_boxplot}}
+\caption{Caption, caption.}\label{fig:boxplot}
+\end{figure}
+
 \begin{table}[!t]
@@
 \section{Results}
-We evaluate our performance based on the average cross-species Subgenic F1 for all genomes except the few used during training.
-The results are displayed in Table \ref{tab:results} and show a clear improvement over the AUGUSTUS and dilated CNN baselines.
-We also demonstrate the effectiveness of two techniques used during inference with the usage of longer input lengths being enabled by our LSTM stack without any fully connected layers on top.
+Figure \ref{fig:boxplot} shows the side-by-side comparison of the distribution of performances on all non-training genomes in the animal and plant case by Subgenic F1.
+In the case of Helixer, these scores represent cross-species predictions.
+We also compare the average performances of different configurations of Helixer with Augustus and a dilated CNN baseline in Table \ref{tab:results}.
+The results show a clear improvement over the AUGUSTUS both in median performance and spread.
+We also significantly outperform the dilated CNN baseline and demonstrated the effectiveness of two techniques used during inference with the usage of longer input lengths being enabled by our LSTM stack without any fully connected layers on top.
 To our knowledge, this technique was not used before with a model used for gene annotation.

### 2020-04-13 paper/main.tex
caption

@@
 The {\it Subgenic F1} is similar to the Genic F1, except that it does not take UTR predictions into account.
-This is done for comparability with Augustus and further explained in Chapter \ref{augustus}.
+This is done for comparability with AUGUSTUS and further explained in Chapter \ref{augustus}.
 Both metrics work by summing up True Positives (TP), False Positives (FP) and False Negatives (FN) of each considered class before calculating precision and recall from these values and combining the two into a F1 score.
@@
 
-\subsection{Evaluation of Augustus} \label{augustus}
+\subsection{Evaluation of AUGUSTUS} \label{sec:augustus}
 What Marvin did, why we don't use Genic F1 for comparison, etc
@@
 	\centerline{\includegraphics[width=0.45\textwidth]{images/overall_boxplot}}
-\caption{Caption, caption.}\label{fig:boxplot}
+	\caption{Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
+Results are given for Helixer and AUGUSTUS with the median predictions scores being denoted by the red lines.
+Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}.}\label{fig:boxplot}
 \end{figure}
@@
 \midrule
-Augustus & 0.632 (+-0.089) & 0.749 (+-0.161)\\
+AUGUSTUS & 0.632 (+-0.089) & 0.749 (+-0.161)\\
 Dilated CNN Baseline & - (+-) & - (+-)\\
@@
 The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.
-The Augustus result is the median score for all genomes.}
+The AUGUSTUS result is the median score for all genomes.}
 \end{table}

### 2020-04-13 paper/main.tex
minor

@@
 \begin{figure}[!tpb]
-	\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
+\centerline{\includegraphics[width=0.45\textwidth]{images/overlapping/both_species_overlapping_aggregate_comparison}}
 \caption{Caption, caption.}\label{fig:length_wise_bias}
@@
 In the case of Helixer, these scores represent cross-species predictions.
-We also compare the average performances of different configurations of Helixer with Augustus and a dilated CNN baseline in Table \ref{tab:results}.
+We also compare the median performances of different configurations of Helixer with AUGUSTUS and a dilated CNN baseline in Table \ref{tab:results}.
 The results show a clear improvement over the AUGUSTUS both in median performance and spread.
-We also significantly outperform the dilated CNN baseline and demonstrated the effectiveness of two techniques used during inference with the usage of longer input lengths being enabled by our LSTM stack without any fully connected layers on top.
-To our knowledge, this technique was not used before with a model used for gene annotation.
+We also significantly outperform a baseline constructed of a dilated CNN architecture.
+
+Two techniques were used during inference to improve performance and limit model bias.
+The usage of longer input lengths helped especially for genomes that tend to have longer genes and was enabled by our model architecture being a relatively simple LSTM stack without any fully connected layers on top.
+We also constructed the final predictions out of a overlapping ones, which greatly helped to reduce prediction bias in most genomes.
+To our knowledge, both techniques were not used before in a model developed for gene annotation.
 
@@
 
+\subsection{Overlapping evaluation by species}
+\begin{figure}[!tpb]
+\centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
+\caption{Caption, caption.}\label{fig:animals_overlapping}
+\end{figure}
+
 \end{document}

### 2020-04-14 paper/main.tex
minor
class weights table

@@
 \midrule
-[0.7, 1.6, 1.2, 1.2] & [0.3, 1.0, 0.9, 0.3] &\\
+	Intergenic & 0.7 & 0.3 \\
+	UTR & 1.6 & 1.0 \\
+	CDS & 1.2 & 0.9 \\
+	Intron & 1.2 & 0.3 \\
 \botrule
-\end{tabular}}{see data statistics tables..}
+\end{tabular}}{The class weights are the only hyperparameters that are different for animals and plants.
+The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table \ref{tab:statistics_all}).
+All other hyperparameters are shown in Table \ref{tab:params}.}
 \end{table}
@@
 \begin{figure}[!tpb]
-\centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
+% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
 \caption{Caption, caption.}\label{fig:animals_overlapping}

### 2020-04-15 plots/phylogenetic_trees/ggtree/plot.R
trees

@@ -7,13 +7,18 @@ tree = read.nhx('all_species.tre')
 # tree_human = tree_subset(tree, node="Homosapiens", levels_back=5)
 
 
-results = read.csv('f1_scores.csv', header=TRUE)
+f1_scores = read.csv('f1_scores.csv', header=TRUE)
+differences = read.csv('differences.csv', header=TRUE)
 sets = read.csv('sets.csv', header=TRUE)
+
 p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set)) 
 # p = p + geom_label(aes(label=node), fill='steelblue')
-p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
+# p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
-png('animals.png', width=1200, height=2000)
-gheatmap(p, results, width=.2, low="red", mid="white", high="blue", midpoint=0.8, 
-		 colnames=F, offset=5) 
+png('animals_differences.png', width=1200, height=2000)
+gheatmap(p, differences, width=.2, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
+dev.off()
+
+png('animals_f1_scores.png', width=1200, height=2000)
+gheatmap(p, f1_scores, width=.2, low="red", high="blue", colnames=F, offset=5) 
 dev.off()

### 2020-04-16 plots/phylogenetic_trees/ggtree/plot.R
minor fixes

@@ -1,3 +1,4 @@
+# This code need the ggtree fork from https://github.com/soi/ggtree to run
 library(ggplot2)
 library(ggtree)
 library(treeio)
@@ -16,9 +17,9 @@ p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
 png('animals_differences.png', width=1200, height=2000)
-gheatmap(p, differences, width=.2, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
+gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
 dev.off()
 
 png('animals_f1_scores.png', width=1200, height=2000)
-gheatmap(p, f1_scores, width=.2, low="red", high="blue", colnames=F, offset=5) 
+gheatmap(p, f1_scores, width=.1, low="red", high="darkgreen", colnames=F, offset=5) 
 dev.off()

### 2020-04-16 plots/phylogenetic_trees/ggtree/plot.R
moved files around, plant tree data

@@ -4,15 +4,37 @@ library(ggtree)
 library(treeio)
 library(tidytree)
 
-tree = read.nhx('all_species.tre')
-# tree_human = tree_subset(tree, node="Homosapiens", levels_back=5)
+# animals
+tree = read.nhx('animals/all_species.tre')
+
+f1_scores = read.csv('animals/f1_scores.csv', header=TRUE)
+differences = read.csv('animals/differences.csv', header=TRUE)
+sets = read.csv('animals/sets.csv', header=TRUE)
+
+p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+# p = p + geom_label(aes(label=node), fill='steelblue')
+# p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
+
+png('animals_differences.png', width=1200, height=2000)
+gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
+dev.off()
+
+png('animals_f1_scores.png', width=1200, height=2000)
+gheatmap(p, f1_scores, width=.1, low="red", high="darkgreen", colnames=F, offset=5) 
+dev.off()
 
 
-f1_scores = read.csv('f1_scores.csv', header=TRUE)
-differences = read.csv('differences.csv', header=TRUE)
-sets = read.csv('sets.csv', header=TRUE)
 
-p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set)) 
+# plants
+tree = read.nhx('plants/all_species.tre')
+
+# f1_scores = read.csv('plants/f1_scores.csv', header=TRUE)
+# differences = read.csv('plants/differences.csv', header=TRUE)
+# sets = read.csv('plants/sets.csv', header=TRUE)
+
+# p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+ggtree(tree) # %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+quit()
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 

### 2020-04-16 plots/phylogenetic_trees/ggtree/plot.R
plant tree data

@@ -33,7 +33,7 @@ tree = read.nhx('plants/all_species.tre')
 # sets = read.csv('plants/sets.csv', header=TRUE)
 
 # p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
-ggtree(tree) # %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+ggtree(tree) + geom_tiplab(size=4, offset=.3) # + geom_tippoint(aes(color=set, shape=set, size=.5)) 
 quit()
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)

### 2020-04-16 data_scripts/rm_one_sp.py
thanks for bug hunting ChrisGuenther91

@@ -22,7 +22,7 @@ for grp in ['scores', 'data', 'evaluation']:
                             dtype=odat[key].dtype,
                             compression='lzf',
                             shuffle=odat[key].shuffle)
-        
+    by = 500
     for i in range(0, f['data/y'].shape[0], by):
         submask = mask[i:(i + by)]
         n_seqs = np.sum(submask)
@@ -38,6 +38,7 @@ for grp in ['scores', 'data', 'evaluation']:
     newf.flush()
 
 # copy over the metadata, start_end_i will be wrong, everything else should be OK
+newf.create_group('meta')
 for key in f['meta'].keys():
 
     bkey = key.encode('utf-8')

### 2020-04-16 paper/main.tex
trees
updated results with test plant genomes

@@
 \midrule
-AUGUSTUS & 0.632 (+-0.089) & 0.749 (+-0.161)\\
+AUGUSTUS & 0.632 (+-0.089) & 0.747 (+-0.188)\\
 Dilated CNN Baseline & - (+-) & - (+-)\\

### 2020-04-16 paper/main.tex
augustus results without our training genomes and with eptatretus_burgeri

@@
 \midrule
-AUGUSTUS & 0.632 (+-0.089) & 0.747 (+-0.188)\\
+AUGUSTUS & 0.632 (+-0.09) & 0.757 (+-0.182)\\
 Dilated CNN Baseline & - (+-) & - (+-)\\
@@
 Varied input length was only used in the animal case.
-The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.
-The AUGUSTUS result is the median score for all genomes.}
+The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.}
 \end{table}

### 2020-04-17 plots/phylogenetic_trees/ggtree/plot.R
plant plots

@@ -28,20 +28,19 @@ dev.off()
 # plants
 tree = read.nhx('plants/all_species.tre')
 
-# f1_scores = read.csv('plants/f1_scores.csv', header=TRUE)
-# differences = read.csv('plants/differences.csv', header=TRUE)
+f1_scores = read.csv('plants/f1_scores.csv', header=TRUE)
+differences = read.csv('plants/differences.csv', header=TRUE)
 # sets = read.csv('plants/sets.csv', header=TRUE)
 
 # p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
-ggtree(tree) + geom_tiplab(size=4, offset=.3) # + geom_tippoint(aes(color=set, shape=set, size=.5)) 
-quit()
+p = ggtree(tree) + geom_tiplab(size=4, offset=.3) # + geom_tippoint(aes(color=set, shape=set, size=.5)) 
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
-png('animals_differences.png', width=1200, height=2000)
+png('plants_differences.png', width=1200, height=2000)
 gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
 dev.off()
 
-png('animals_f1_scores.png', width=1200, height=2000)
+png('plants_f1_scores.png', width=1200, height=2000)
 gheatmap(p, f1_scores, width=.1, low="red", high="darkgreen", colnames=F, offset=5) 
 dev.off()

### 2020-04-17 plots/phylogenetic_trees/ggtree/plot.R
trees

@@ -15,11 +15,11 @@ p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
-png('animals_differences.png', width=1200, height=2000)
+png('animals_differences.png', width=1200, height=3000)
 gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
 dev.off()
 
-png('animals_f1_scores.png', width=1200, height=2000)
+png('animals_f1_scores.png', width=1200, height=3000)
 gheatmap(p, f1_scores, width=.1, low="red", high="darkgreen", colnames=F, offset=5) 
 dev.off()
 
@@ -30,17 +30,16 @@ tree = read.nhx('plants/all_species.tre')
 
 f1_scores = read.csv('plants/f1_scores.csv', header=TRUE)
 differences = read.csv('plants/differences.csv', header=TRUE)
-# sets = read.csv('plants/sets.csv', header=TRUE)
+sets = read.csv('plants/sets.csv', header=TRUE)
 
-# p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
-p = ggtree(tree) + geom_tiplab(size=4, offset=.3) # + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
-png('plants_differences.png', width=1200, height=2000)
-gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
+png('plants_differences.png', width=600, height=1000)
+gheatmap(p, differences, width=.15, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
 dev.off()
 
[…]

### 2020-04-17 paper/main.tex
length wise bias plot and caption

@@
 \begin{figure}[!tpb]
-\centerline{\includegraphics[width=0.45\textwidth]{images/overlapping/both_species_overlapping_aggregate_comparison}}
-\caption{Caption, caption.}\label{fig:length_wise_bias}
+\label{fig:length_wise_bias}
+\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
+\caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
+Each value is the average performance of a 200 basepair long chunk.
+The regular metrics show a significant negative bias towards the borders of each sequence.
+Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average.
+See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
 \end{figure}
@@
 \subsection{Overlapping evaluation by species}
+\label{sec:overlapping}
 \begin{figure}[!tpb]

### 2020-04-18 plots/phylogenetic_trees/ggtree/plot.R
plots

@@ -16,14 +16,12 @@ p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
 png('animals_differences.png', width=1200, height=3000)
-gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
+gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5, legend_title="Subgenic F1") 
 dev.off()
 
 png('animals_f1_scores.png', width=1200, height=3000)
-gheatmap(p, f1_scores, width=.1, low="red", high="darkgreen", colnames=F, offset=5) 
+gheatmap(p, f1_scores, width=.1, colnames=F, offset=5, use_scale_fill_viridis_d=T, legend_title="Subgenic F1") 
 dev.off()
-
-
 
 # plants
 tree = read.nhx('plants/all_species.tre')
@@ -37,9 +35,9 @@ p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
 png('plants_differences.png', width=600, height=1000)
-gheatmap(p, differences, width=.15, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5) 
+gheatmap(p, differences, width=.15, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5, legend_title="Subgenic F1") 
 dev.off()
 
 png('plants_f1_scores.png', width=600, height=1000)
-gheatmap(p, f1_scores, width=.15, low="red", high="darkgreen", colnames=F, offset=5) 
+gheatmap(p, f1_scores, width=.15, colnames=F, offset=5, use_scale_fill_viridis_d=T, legend_title="Subgenic F1") 
 dev.off()

### 2020-04-19 plots/phylogenetic_trees/ggtree/plot.R
tree plots

@@ -11,16 +11,16 @@ f1_scores = read.csv('animals/f1_scores.csv', header=TRUE)
 differences = read.csv('animals/differences.csv', header=TRUE)
 sets = read.csv('animals/sets.csv', header=TRUE)
 
-p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=Set, shape=Set, size=.5)) + guides(size=F) 
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
 png('animals_differences.png', width=1200, height=3000)
-gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5, legend_title="Subgenic F1") 
+gheatmap(p, differences, width=.1, low="red", mid="white", high="blue", midpoint=0.0, use_scale_fill_gradient2=T, colnames=F, offset=5, legend_title="Difference") 
 dev.off()
 
 png('animals_f1_scores.png', width=1200, height=3000)
-gheatmap(p, f1_scores, width=.1, colnames=F, offset=5, use_scale_fill_viridis_d=T, legend_title="Subgenic F1") 
+gheatmap(p, f1_scores, width=.1, colnames=F, offset=5, use_scale_fill_viridis_d=T, legend_title="Subgenic F1")
 dev.off()
 
 # plants
@@ -30,12 +30,12 @@ f1_scores = read.csv('plants/f1_scores.csv', header=TRUE)
 differences = read.csv('plants/differences.csv', header=TRUE)
 sets = read.csv('plants/sets.csv', header=TRUE)
 
-p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=set, shape=set, size=.5)) 
+p = ggtree(tree) %<+% sets + geom_tiplab(size=4, offset=.3) + geom_tippoint(aes(color=Set, shape=Set, size=.5)) + guides(size=F) 
 # p = p + geom_label(aes(label=node), fill='steelblue')
 # p = p + geom_cladelabel(node=201, label="Mammalia", align=T, geom='label', offset=12)
 
[…]

### 2020-04-20 paper/main.tex
minor
rewrite approach

@@
 \section{Approach}
-Helixer is a prototype software for training and utilizing a general purpose model for the functional annotation of large eukaryotic genomes.
-We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on 192 genomes in the animal case and on 60 plant genomes.
-We use modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields.
-Our model expects the raw genomic sequence as main input and produces functional predictions on a base pair wise level.
+Helixer is a prototype software for training and utilizing a general purpose model for the functional cross-species annotation of large eukaryotic genomes using only DNA data as input.
+We demonstrate the effectiveness of this approach by training a single model for the annotation of a large set of genomes from the kingdoms metazoa and viridiplantae, respectively, which we will call {\it animals} and {\it plants} from now on.
+In total we used 192 animal genomes and 60 plant genomes.
+In contrast to traditional approaches, we employ modelling techniques from the field of deep learning to improve both the annotation accuracy and the ability to generalize across genomes.
 The source code and all data is publically available.
@@
 \caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
-Each value is the average performance of a 200 basepair long chunk.
+Each value is the average performance of a 200 basepair long chunk by one of the final ensemble models.
 The regular metrics show a significant negative bias towards the borders of each sequence.

### 2020-04-20 paper/main.tex
text

@@
 The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format.
-We used all animals genomes in EnsemblMetazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation).
+We used all animals genomes in EnsemblMetazoa 45 (\citealp{Howe19}) as well as all non-embargoed plant genomes of the JGI Phytosome 13 database (\citealp{Phytozome}).
 Both data groups were used seperately during training as well as during evaluation.
-At the start of our work we split both genome groups into training and test sets.
-The test sets contained 19 or 6 genomes, respectively.
-The test genomes were chosen to be of the best possible quality (?) based on collected metadata while also representing a broad phylogenetic spread.
-Some of the remaining genomes were used for training purposes and all were used for evaluating our models.
-We call the genomes that were ultimatily used during training the {\it training genomes} and the remaining non-test genomes the {\it validation genomes}.
+At the start of our work we split both genome groups into {\it training genomes} and {\it test genomes}.
+The test genome group is made up of 19 or 6 genomes, respectively, and was chosen to be of the best possible quality based on collected metadata while also representing a broad phylogenetic spread.
+We then experimented which genomes to use for during training from the pool of training genomes while evaluating the performance of models on the remaining ones.
+We call the genomes that were ultimatily not used during training the {\it validation genomes}.
+While we did split off a group of test genomes at the beginning of our work, we ended up only reporting on the generalization performance of the combined set of validation and test genomes as the validation genomes are by far the largest set.
+We also do not find a noticible difference in performance between those two groups.
 
-We pre-processed and stored the genomic information by using Geenuff \citealp{Denton19}.
+We pre-processed and stored the genomic information by using Geenuff (\citealp{Denton19}).
 Geenuff is a tool for checking and exploring genomic data, which stores all information inside a SQL database.
@@
 
-\bibitem[Howe KL {\it et~al}., 2019]{Howe19}
+\bibitem[Howe {\it et~al}., 2019]{Howe19}
 Howe KL, Contreras-Moreira B, De Silva N, Maslen G, Akanni W, Allen J, Alvarez-Jarreta J, Barba M, Bolser DM, Cambell L, Carbajo M, Chakiachvili M, Christensen M, Cummins C, Cuzick A, Davis P, Fexova S, Gall A, George N, Gil L, Gupta P, Hammond-Kosack KE, Haskell E, Hunt SE, Jaiswal P, Janacek SH, Kersey PJ, Langridge N, Maheswari U, Maurel T, McDowall MD, Moore B, Muffato M, Naamati G, Naithani S, Olson A, Papatheodorou I, Patricio M, Paulini M, Pedro H, Perry E, Preece J, Rosello M, Russell M, Sitnik V, Staines DM, Stein J, Tello-Ruiz MK, Trevanion SJ, Urban M, Wei S, Ware D, Williams G, Yates AD, Flicek P.
@@
 {\it Nucleic Acids Research 2019} https://doi.org/10.1093/nar/gkz890
+
+\bibitem[Goodstein {\it et~al}., 2012]{Phytozome}
+	David M.
+Goodstein, Shengqiang Shu, Russell Howson, Rochak Neupane, Richard D.
+Hayes, Joni Fazo, Therese Mitros, William Dirks, Uffe Hellsten, Nicholas Putnam, Daniel S.
+Rokhsar (2012) Phytozome: a comparative platform for green plant genomics, {\it Nucleic Acids Research}, Volume 40, Issue D1, Pages D1178–D1186, https://doi.org/10.1093/nar/gkr944
 

### 2020-04-20 paper/main.tex
minor

@@
 At the start of our work we split both genome groups into {\it training genomes} and {\it test genomes}.
-The test genome group is made up of 19 or 6 genomes, respectively, and was chosen to be of the best possible quality based on collected metadata while also representing a broad phylogenetic spread.
-We then experimented which genomes to use for during training from the pool of training genomes while evaluating the performance of models on the remaining ones.
+The test genome group is made up of 19 and 6 genomes, respectively, and was chosen to be of the best possible quality based on collected metadata while also representing a broad phylogenetic spread.
+We then experimented which genomes to use for during training from the pool of non-test genomes while evaluating the performance of models on the remaining ones.
 We call the genomes that were ultimatily not used during training the {\it validation genomes}.

### 2020-04-21 paper/main.tex
many small things

@@
 While we did split off a group of test genomes at the beginning of our work, we ended up only reporting on the generalization performance of the combined set of validation and test genomes as the validation genomes are by far the largest set.
-We also do not find a noticible difference in performance between those two groups.
+We also did not find a noticible difference in performance between those two groups.
 
@@
 Training and evaluation data was generated by querying this database and transforming the returned data into a numerical format suitable for machine learning.
-In the case of unambiguous input data, the genomic sequence was transformed into matrix that represents a one hot encoding with 4 rows.
-For the other case see Supplementary material ..
-().
-The annotations are also transformed into a 4-row one hot encoding matrix, where a one in each of the rows represents one of the possible classes {\it Intergenic}, {\it UTR}, {\it CDS} or {\it Intron}.
-An non-coding intron can be inferred from the prediction order.
-A full description of all generated data arrays can be found in table \ref{tab:arrays}.
+The encoding of the genomic sequence is done in line with the IUPAC nucleic acid notation and the functional annotation used as labels during training is transformed to a one hot encoding with the four classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}.
+See also Table \ref{tab:arrays} for a more detailed description of the generated data formats.
 
@@
 
-During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?).
-As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to effectively blank out those bases during training.
+During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation.
+As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to blank out those bases during training by using the sample weights described in Table \ref{tab:arrays}.
 The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it.
@@
 This split of the training genomes was done to get a quicker sense of the generalization capabilities and was used primarily during training.
-We split of the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level.
+We split off the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level.
 Evaluation of the annotations on all training and validation genomes was done regularly after a promising model candidate was found based on its validation set performance.
-Ultimately we only cared about the cross-species performance of a model.
+Ultimately we only cared about the cross-species performance on all validation genomes of a model.
 
-We use metrics two metrics to judge model performance.
+We use two metrics to judge model performance.
 The {\it Genic F1} is our primary metric that provides the most comprehensive picture of annotation quality in one number.
 The {\it Subgenic F1} is similar to the Genic F1, except that it does not take UTR predictions into account.
-This is done for comparability with AUGUSTUS and further explained in Chapter \ref{augustus}.
+This is done for comparability with AUGUSTUS and further explained in Section \ref{sec:augustus}.
 Both metrics work by summing up True Positives (TP), False Positives (FP) and False Negatives (FN) of each considered class before calculating precision and recall from these values and combining the two into a F1 score.
 Both metrics do not take the values of the intergenic class into account, as this class is very abundant and appears to be by far the easiest to predict.
+(See Table \ref{tab:statistics_all} for the class distribution and Table \ref{tab:detailed_results} for a more in-depth report our model performance including the intergenic class).
 In the case of the Genic F1, this means that effectively only the TP of the intergenic class are disregarded.
 The metrics essentially provide a weighted mean of the performances in the considered classes.
-All individual class performances are reported in Supplementary Material (?).
 
@@
 
-\subsection{Evaluation of AUGUSTUS} \label{sec:augustus}
+\subsection{Evaluation of AUGUSTUS} 
+\label{sec:augustus}
 What Marvin did, why we don't use Genic F1 for comparison, etc
@@
 \midrule
-Input & Genomic sequence in a one hot encoding\\
-	  & for (A, T, C, G), otherwise different, see suplementary .. ()\\
-Output & Labels in a 4-dimensional one hot encoding \\
+Input & Genomic sequence in the 4-dimensional IUPAC encoding\\
+	  & (one hot encoding for non-ambiguous bases) \\
+Output & Labels in a 4-dimensional one hot encoding representing \\
+      & the classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}\\
 Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
 \botrule
-\end{tabular}}{}
+\end{tabular}}{An intron inside a UTR region is encoded as intron and can be inferred from the prediction order.}
 \end{table}
@@
 
-\subsection{Input Data Encoding}
-
 \subsection{Hyperparameters}
@@
 
+\subsection{Detailed model performance}
+
+\begin{table}[!t]
+\processtable{Class weights\label{tab:detailed_results}} {
+\begin{tabular}{@{}lll@{}}
+\toprule Class & Animal Value & Plant Value \\
+\midrule
+Intergenic & 0.7 & 0.3 \\
+UTR & 1.6 & 1.0 \\
+CDS & 1.2 & 0.9 \\
+Intron & 1.2 & 0.3 \\
+\botrule
+\end{tabular}}{}
+\end{table}
+
 \end{document}

### 2020-04-21 paper/main.tex
switch to bibtex

@@
 \appnotes{Manuscript Category}
+\usepackage{url}
 
@@
 In this work on eucaryotic genomes, we differentiate between Intergenic regions, Untranslated regions (UTRS, parts at the beginning and end of gene, that are transcribed but not translated), Coding sequences (CDS) and Introns.
-Because of the sheer size of genomes alone, gene annotation is a very labour intensive process and genomic database providers like NCBI [\citealp{NCBI13}] and Ensembl [\citealp{Ensembl16}] offer gene annotation pipelines.
+Because of the sheer size of genomes alone, gene annotation is a very labour intensive process and genomic database providers like NCBI \citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl} offer gene annotation pipelines.
 These pipelines integrate experimental data (from e.g.
@@
 The latter is an attractive approach, because it is cheap and fast.
-State of the art performance ab-initio gene prediction is achieved by higher order hidden markov models (HMMs), such as Genscan [\citealp{Genscan}] or Augustus [\citealp{Augustus}] but their accuracy leaves room for improvement.
+State of the art performance ab-initio gene prediction is achieved by higher order hidden markov models (HMMs), such as Genscan \citep{burge1997prediction} or Augustus \citep{stanke2003gene} but their accuracy leaves room for improvement.
 By choosing possible states and transitions, designers of HMMs assume structure in the sequence that may limit its predictive power.
@@
 RNNs have already shown promising results in related, but generally more limited investigations.
-[\citealp{Choudhary17}] carries out preliminary explorations on the potential of BLSTMs for cross species gene prediction, but trains his model on human genes only and tests it on two more species.
-DeepAnnotator [\citealp{DeepAnnotator}] uses bidirectional LSTMs (BLSTM) for gene finding in prokaryotes.
-Gene prediction in prokaryotes is considered more amenable than in eukaryotes, as genes in prokaryotes are proportionately more frequent in the genome, feature simpler control structures and do not use splicing [\citealp{Zhuo04}].
-DanQ [\citealp{DanQ}] proposes the use of a BLSTM after a concurrent neural net (CNN) to find detailed motifs in the human genome.
-DeePromotor[\citealp{DeePromotor}] trains a similar architecture for the recognition of promotor regions.
-Recently, several groups [\citealp{SpliceAi}, \citealp{SpliceFinder}] successfully used CNNs to find splicing sites.
+\citep{choudharypredicting} carries out preliminary explorations on the potential of BLSTMs for cross species gene prediction, but trains his model on human genes only and tests it on two more species.
+DeepAnnotator \citep{amin2018deepannotator} uses bidirectional LSTMs (BLSTM) for gene finding in prokaryotes.
+Gene prediction in prokaryotes is considered more amenable than in eukaryotes, as genes in prokaryotes are proportionately more frequent in the genome, feature simpler control structures and do not use splicing \citep{wang2004brief}.
+DanQ \citep{quang2016danq} proposes the use of a BLSTM after a concurrent neural net (CNN) to find detailed motifs in the human genome.
+DeePromotor \citep{oubounyt2019deepromoter} trains a similar architecture for the recognition of promotor regions.
+Recently, several groups \citep{jaganathan2019predicting, wang2019splicefinder} successfully used CNNs to find splicing sites.
 
@@
 \botrule
-\end{tabular}}{Values are averages of the individual values of each genome in a data group.
+\end{tabular}}{{\it Note:} Values are averages of the individual values of each genome in a data group.
 All statistics except the average gene lengths are excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.
@@
 
+\subsection{Metrics}
 We use two metrics to judge model performance.
@@
 \label{sec:model}
-We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization (citation ?
-) between each BLSTM layer to produce the predictions in the form of a base pair wise classification.
-The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (citation).
+We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization (\citealp{Layernorm}) between each BLSTM layer to produce the predictions in the form of a base pair wise classification.
+The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (\citealp{Tensorflow}).
 We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks (CNN) and hybrid architectures.
@@
 \botrule
-\end{tabular}}{An intron inside a UTR region is encoded as intron and can be inferred from the prediction order.}
+\end{tabular}}{{\it Note:} An intron inside a UTR region is encoded as intron and can be inferred from the prediction order.}
 \end{table}
@@
 	\centerline{\includegraphics[width=0.45\textwidth]{images/overall_boxplot}}
-	\caption{Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
+	\caption{Boxplots comparing the Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
 Results are given for Helixer and AUGUSTUS with the median predictions scores being denoted by the red lines.
@@
 \botrule
-\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group.
+\end{tabular}}{{\it Note:} Values are the median in Subgenic F1 across all validation and test genomes of the respective group.
 Varied input length was only used in the animal case.
@@
 
-\begin{thebibliography}{}
-
-\bibitem[Thibaud-Nissen {\it et~al}, 2013]{NCBI13}
-Thibaud-Nissen,F., Souvorov,A., Murphy,T., DiCuccio,M., Kitts,P. (2013) Eukaryotic Genome Annotation Pipeline{\it The NCBI Handbook, 2nd}
-
-\bibitem[Aken {\it et~al}, 2016]{Ensembl16}
-Aken,B., Ayling,S., Barrell,D., Clarke,L., {\it et~al.} (2016) The Ensembl gene annotation system, {\it Database}, {\bf 2016}, baw093
-
-\bibitem[Burge {\it et} Karlin, 1997]{Genscan}
-Burge,C., Karlin,S. (1997) Prediction of complete gene structures in human genomic DNA, {\it Journal of Molecular Biology}, {\bf 268-1}, 78-94
-
-\bibitem[Stanke {\it et} Waack, 2003]{Augustus}
-	Stanke,M., Waack,S., (2003) Gene prediction with a hidden Markov model and a new intron submodel, {\it Bioinformatics}, {\bf 19 Suppl.2}, ii215–ii225
-
-\bibitem[Zhuo {\it et~al}., 2004]{Zhuo04}
-Zhuo,W., Yazhu,C., Yixue,L. (2004) A Brief Review of Computational Gene Prediction Methods, {\it Genomics, Proteomics \& Bioinformatics}, 216-221
-
-\bibitem[Amin {\it et~al}., 2018]{DeepAnnotator}
-Amin,M., Yurovsky,A., Tian,Y., Skiena,S.
-(2018) DeepAnnotator: Genome Annotation with Deep Learning, {\it Proceedings of the 2018 ACM International Conference on Bioinformatics, Computational Biology, and Health Informatics}, 254-259.
-
-\bibitem[Choudhary, 2017]{Choudhary17}
-Choudhary,S. (2017) Predicting protein coding boundaries using RNNs, http://www.saket-choudhary.me/rnn-cds-prediction/predicting-protein-coding.pdf (accessed 2020-04-11)
-
-\bibitem[Daniel {\it et~al}., 2016]{DanQ}
-Daniel,Q., Xiaohui,X., (2016) DanQ: a hybrid convolutional and recurrent deep neural network for quantifying the function of DNA sequences, {\it Nucleic Acids Research}, {\bf 44-11}, e107
-
-\bibitem[Oubounyt {\it et~al}., 2019]{DeePromotor}
-Oubounyt,M., Louadi,Z., Tayara,H., Chong2,K. (2019) DeePromoter: Robust Promoter Predictor Using Deep Learning., {\it Frontiers in genetics}, {\bf 10}, 286
-
-\bibitem[Jaganathan {\it et~al}., 2019]{SpliceAi}
-Jaganathan,K., Panagiotopoulou,S., McRae,J., Darbandi,S., et.al., Predicting Splicing from Primary Sequence with Deep Learning, {\it Cell}, {\bf 176-3}, 535-548.e24
-
-\bibitem[Wang {\it et~al}., 2019]{SpliceFinder}
-Wang,R., Wang,Z., Wang,J. (2019) SpliceFinder: ab initio prediction of splice sites using convolutional neural network, {\it BMC Bioinformatics}, {\bf 20}, 652
-
-\bibitem[Gupta {\it et~al}., 2017]{dCNNGupta}
-Gupta,A., Rush.,A. (2017) Dilated convolutions for modeling long-distance genomic dependencies, {\it arXiv preprint arXiv:1710.01278}
-
-\bibitem[Denton {\it et~al}., 2019]{Denton19}
-Denton,A., Stiehler,F. (2019) Article title, {\it Journal Name}, {\bf 199}, 133-154.
-
-\bibitem[Howe {\it et~al}., 2019]{Howe19}
-Howe KL, Contreras-Moreira B, De Silva N, Maslen G, Akanni W, Allen J, Alvarez-Jarreta J, Barba M, Bolser DM, Cambell L, Carbajo M, Chakiachvili M, Christensen M, Cummins C, Cuzick A, Davis P, Fexova S, Gall A, George N, Gil L, Gupta P, Hammond-Kosack KE, Haskell E, Hunt SE, Jaiswal P, Janacek SH, Kersey PJ, Langridge N, Maheswari U, Maurel T, McDowall MD, Moore B, Muffato M, Naamati G, Naithani S, Olson A, Papatheodorou I, Patricio M, Paulini M, Pedro H, Perry E, Preece J, Rosello M, Russell M, Sitnik V, Staines DM, Stein J, Tello-Ruiz MK, Trevanion SJ, Urban M, Wei S, Ware D, Williams G, Yates AD, Flicek P.
-(2019) Ensembl Genomes 2020-enabling non-vertebrate genomic research.
-{\it Nucleic Acids Research 2019} https://doi.org/10.1093/nar/gkz890
-
-\bibitem[Goodstein {\it et~al}., 2012]{Phytozome}
-	David M.
-Goodstein, Shengqiang Shu, Russell Howson, Rochak Neupane, Richard D.
-Hayes, Joni Fazo, Therese Mitros, William Dirks, Uffe Hellsten, Nicholas Putnam, Daniel S.
-Rokhsar (2012) Phytozome: a comparative platform for green plant genomics, {\it Nucleic Acids Research}, Volume 40, Issue D1, Pages D1178–D1186, https://doi.org/10.1093/nar/gkr944
-
-\end{thebibliography}
+\bibliographystyle{natbib}
+\bibliography{literature}
 
@@
 \botrule
-\end{tabular}}{Above are all parameters used during neural architecture search for the dilated CNN baseline.
+\end{tabular}}{{\it Note:} Above are all parameters used during neural architecture search for the dilated CNN baseline.
 The same overall space was used for plant and animal data.
@@
 \botrule
-\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.
+\end{tabular}}{{\it Note:} The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.
 The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
@@
 \botrule
-\end{tabular}}{}
+\end{tabular}}{{\it Note:} }
 \end{table}
@@
 \botrule
-\end{tabular}}{The class weights are the only hyperparameters that are different for animals and plants.
+\end{tabular}}{{\it Note:} The class weights are the only hyperparameters that are different for animals and plants.
 The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table \ref{tab:statistics_all}).
@@
 \botrule
-\end{tabular}}{The description of Table \ref{tab:statistics_all} applies here as well.}
+\end{tabular}}{{\it Note:} The description of Table \ref{tab:statistics_all} applies here as well.}
 \end{table}

### 2020-04-21 paper/main.tex
literature

@@
 The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format.
-We used all animals genomes in EnsemblMetazoa 45 (\citealp{Howe19}) as well as all non-embargoed plant genomes of the JGI Phytosome 13 database (\citealp{Phytozome}).
+We used all animals genomes in EnsemblMetazoa 45 \citep{howe2020ensembl} as well as all non-embargoed plant genomes of the JGI Phytosome 13 database \citep{goodstein2012phytozome}.
 Both data groups were used seperately during training as well as during evaluation.
@@
 
-We pre-processed and stored the genomic information by using Geenuff (\citealp{Denton19}).
+We pre-processed and stored the genomic information by using Geenuff \citep{denton2019}.
 Geenuff is a tool for checking and exploring genomic data, which stores all information inside a SQL database.
@@
 \label{sec:model}
-We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization (\citealp{Layernorm}) between each BLSTM layer to produce the predictions in the form of a base pair wise classification.
-The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras (citation) on top of TensorFlow (\citealp{Tensorflow}).
+We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization \citep{ba2016layer} between each BLSTM layer to produce the predictions in the form of a base pair wise classification.
+The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras \citep{chollet2015keras} on top of TensorFlow \citep{abadi2016tensorflow}.
 We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks (CNN) and hybrid architectures.
-The hyperparameters were determined by a combination of manual and automatic hyperparameter optimization (talk about TPE?, quote nni?).
+The hyperparameters were optimized by a combination of manual and automatic optimization.
+Automatic optimization was carried out by using either the TPE algorithm \citep{bergstra2011algorithms}, random search or grid search depending on the situation.
+We used NNI \citep{nni2019} to facilitate the search.
 All relevant hyperparameters can be found in Table \ref{tab:params} or in the Helixer source code repository.

### 2020-04-21 paper/main.tex
longer input length

@@
 
-We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
-Details of the neural architecture search are given in Section \ref{sec:dcnn}.
-
 Our final model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
@@
 The input sequences were up to 10 times larger, depending on species group and assembly quality.
-This also demonstrates ability of our model to generalize as it is able to make successful prediction on far longer sequences than it has ever seen.
-The sequence length during inference can be any of 20.000, 50.000, 100.000 or 200.000 and is determined by the species family(?).
-Families that tend to have longer genes get assigned longer sequences.
+This also demonstrates the ability of our model to generalize as it is able to make successful prediction on far longer sequences than it has ever seen.
+The sequence length during inference can be any of 20.000, 50.000, 100.000 or 200.000 and is determined by the phylogenetic class.
+The assignment was chosen to keep the average gene length roughly proportional to the length of the sequence input.
 For more implementation details on this see Section \ref{sec:longer} or the source code.
@@
 As with the overlapping, the fusion of the 8 individual predictions was done by averaging the softmax values of each base pair prediction.
+
+We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
+Details of the neural architecture search are given in Section \ref{sec:dcnn}.
 
@@
 \label{sec:longer}
-Implementation detail:
-If the N75 of a species is less than twice as high as the supposed sequence length it is lowered until either this criteria is met or a length of 50.000 is reached.
-
-Table of species with their length.
+\begin{table}[!t]
+\processtable{Input Sequence Length by Animal Class\label{tab:input_lengths}} {
+\begin{tabular}{@{}lll@{}}
+\toprule Class & Input Length in bp\\
+\midrule
+Mammalia & 200.000 \\
+Reptilia (excluding Aves) & 200.000 \\
+Aves & 100.000 \\
+Amphibia & 100.000 \\
+Actinopterygii & 50.000 \\
+Chondrichthyes & 50.000 \\
+Actinistia & 50.000 \\
+Ascidiacea & 20.000 \\
+Insecta & 20.000 \\
+\botrule
+\end{tabular}}{{\it Note:} Input sequence length for each phylogenetic class present in our animal dataset.
+As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000].
+If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached.
+}
+\end{table}
 

### 2020-04-21 paper/main.tex
minor

@@
 This also demonstrates the ability of our model to generalize as it is able to make successful prediction on far longer sequences than it has ever seen.
-The sequence length during inference can be any of 20.000, 50.000, 100.000 or 200.000 and is determined by the phylogenetic class.
-The assignment was chosen to keep the average gene length roughly proportional to the length of the sequence input.
+The sequence length during inference can be up to 200.000 base pairs long and is determined by the phylogenetic class.
+The concrete lengths were chosen to keep the average gene length roughly proportional to the length of the sequence input.
 For more implementation details on this see Section \ref{sec:longer} or the source code.

### 2020-04-22 paper/main.tex
some rewrite

@@
 
-\subsection{Model Architecture and Usage}
+\subsection{Model Architecture}
 \label{sec:model}
@@
 
+We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
+Details of the neural architecture search are given in Section \ref{sec:dcnn}.
+
+\subsection{Inference Techniques}
 We also used multiple techniques to improve the prediction quality after the training was done.
@@
 The sequence length during inference can be up to 200.000 base pairs long and is determined by the phylogenetic class.
-The concrete lengths were chosen to keep the average gene length roughly proportional to the length of the sequence input.
+The concrete lengths were chosen to keep the typical average gene length roughly proportional to the length of the sequence input.
 For more implementation details on this see Section \ref{sec:longer} or the source code.
@@
 The final predictions of a single model are constructed by overlapping predictions, that were made from a sliding window and then cut to a core sequence.
-This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see figure \ref{fig:length_wise_bias}) and also to improve model performance in general by providing the model with multiple different starting points.
+This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see Figure \ref{fig:length_wise_bias}).
+It also improves the average model performance in general by providing the model with multiple different starting points.
 The different overlapping sequences were recombined by averaging the individual softmax values of each base.
-Supplementary Material (?) contains the influence on overlapping with respect to the position in a sequence for each genome, ordered by N75.
-We found that overlapping works best if the genomes are not very fragmented.
+The figures in Section \ref{sec:overlapping} show the effect of overlapping for each genome, ordered by N75.
+We found that overlapping tends to work best if the genomes are not very fragmented and we used it for both animals and plants.
 
@@
 As with the overlapping, the fusion of the 8 individual predictions was done by averaging the softmax values of each base pair prediction.
-
-We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
-Details of the neural architecture search are given in Section \ref{sec:dcnn}.
 

### 2020-04-22 paper/main.tex
minor
supplements started in extra file

@@
 
-\subtitle{Really cool deep learning stuff}
+\subtitle{Genome analysis}
 
-\title[short Title]{General purpose gene annotation using deep learning}
+\title[short Title]{Helixer: Cross-species gene annotation of large eucaryotic genomes using deep learning}
 \author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
@@
 
-\section{Conclusion}
+\section{Discussion}
 We propose a gene calling model that is able to perform cross-species genomic annotiation for a wide variety of large eucaryotic genomes.

### 2020-04-22 paper/supplements.tex
supplements started in extra file

\documentclass{article}
\usepackage{graphicx}
\usepackage{authblk}

\begin{document}

\title{Supplemental Material for \\ Helixer: Cross-species gene annotation of large eucaryotic genomes using deep learning}

\author[1]{Felix Stiehler}
\author[1]{Alisandra Denton}
\author[1]{Marvin Steinborn}
\author[ \hspace{-1ex}]{Stephan Scholz}
\author[1]{Andreas Weber}
\affil[1]{Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}

%\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
% \address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
% \author{Author's Name}

\date{}
\maketitle

\section{Introduction}
Here is the text of your introduction.

\subsection{Subsection Heading Here}
Write your subsection text here.

\section{Conclusion}
Write your conclusion here.

\end{document}

### 2020-04-22 paper/main.tex
lots of supplement stuff

@@
 
-%\bibliographystyle{natbib}
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
-%
-%\bibliographystyle{plain}
-%
-%\bibliography{Document}
-
 \bibliographystyle{natbib}
@@
 
-\newpage
-
-\section{Supplementary Material}
-
-\subsection{Dilated CNN Baseline Network Search}
-\label{sec:dcnn}
-\begin{table}[!t]
-\processtable{Parameter search space for the dilated CNN baselines\label{tab:dcnn}} {
-\begin{tabular}{@{}ll@{}}
-\toprule Parameter & Possible Values \\
-\midrule
-Kernel Size & \{4, 8, 12, 16\}\\
-Initial Filter Depth & \{32, 64, 96, 128\}\\
-Number of Layers Before Doubling Filter Count & \{1, 2\}\\
-Dilation Multiplier & \{2, 3\}\\
-Number of Convolutional Layers & \{2, 3, 4, 5, 6, 7, 8\}\\
-Number of Hidden Fully Connected (FC) Layers & \{0, 1, 2\}\\	
-Dropout Used on FC and Final Conv Layer Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
-Learning Rate & \{1e-3, 1e-4\}\\
-
-\botrule
-\end{tabular}}{{\it Note:} Above are all parameters used during neural architecture search for the dilated CNN baseline.
-The same overall space was used for plant and animal data.
-We did, however, run multiple distinct searches that sometimes only operated over a subset of the given parameters.
-This was done as those seemed to be the most promising.
-We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs.
-Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.
-Runs with a Genic F1 below 0.5 after 10 epochs were stopped and the overall maximum epoch was 15.
-The performances seemed to be leveling off well before that.
-The batch size used for almost all runs was 32, the dilation was capped to 81 and the size of each fully connected layer was fixed at 128.
-New parameters where chosen at random.
-The implementation and parameter space is very roughly based on [\citealp{dCNNGupta}] and can be found in the Helixer source code repository.
-In total, we trained 30 models for the animals and 48 with the plant data.}
-\end{table}
-
-\begin{table}[!t]
-\processtable{Final Parameters of dilated CNN baselines\label{tab:dcnn_best}} {
-\begin{tabular}{@{}lll@{}}
-\toprule Parameter & Animals & Plants\\
-\midrule
-Kernel Size & 16 & 16\\
-Initial Filter Depth & 96 & 32\\
-Number of Layers Before Doubling Filter Count & 2 & 2\\
-Dilation Multiplier & 3 & 3\\
-Number of Convolutional Layers & 6 & 8\\
-Number of Hidden Fully Connected (FC) Layers & 2 & 0\\
-Dropout Used on FC Output & 0.0 & 0.0\\
-Learning Rate & 1e-4 & 1e-3\\
-\botrule
-\end{tabular}}{{\it Note:} The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.
-The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
-\end{table}
-
-\subsection{Hyperparameters}
-
-\begin{table}[!t]
-\processtable{Hyperparameters\label{tab:params}} {
-\begin{tabular}{@{}ll@{}}
-\toprule Name & Value \\
-\midrule
-Layers & 4\\
-Units per layer & 256\\
-Learning rate & 1e-4\\
-Optimizer & Adam\\
-\botrule
-\end{tabular}}{{\it Note:} }
-\end{table}
-
-\begin{table}[!t]
-\processtable{Class weights\label{tab:params_cw}} {
-\begin{tabular}{@{}lll@{}}
-\toprule Class & Animal Value & Plant Value \\
-\midrule
-Intergenic & 0.7 & 0.3 \\
-UTR & 1.6 & 1.0 \\
-CDS & 1.2 & 0.9 \\
-Intron & 1.2 & 0.3 \\
-\botrule
-\end{tabular}}{{\it Note:} The class weights are the only hyperparameters that are different for animals and plants.
-The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table \ref{tab:statistics_all}).
-All other hyperparameters are shown in Table \ref{tab:params}.}
-\end{table}
-
-\subsection{Training Data statistics}
-\label{sec:training_data}
-\begin{table}[!t]
-\processtable{Data group statistics for our training data\label{tab:statistics_training}} {
-\begin{tabular}{@{}lll@{}}
-\toprule & Animals & Plants\\
-\midrule
-Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
-Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
-Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
-Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
-Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
-Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
-Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
-\botrule
-\end{tabular}}{{\it Note:} The description of Table \ref{tab:statistics_all} applies here as well.}
-\end{table}
-
-\subsection{Longer Sequence Input}
-\label{sec:longer}
-\begin{table}[!t]
-\processtable{Input Sequence Length by Animal Class\label{tab:input_lengths}} {
-\begin{tabular}{@{}lll@{}}
-\toprule Class & Input Length in bp\\
-\midrule
-Mammalia & 200.000 \\
-Reptilia (excluding Aves) & 200.000 \\
-Aves & 100.000 \\
-Amphibia & 100.000 \\
-Actinopterygii & 50.000 \\
-Chondrichthyes & 50.000 \\
-Actinistia & 50.000 \\
-Ascidiacea & 20.000 \\
-Insecta & 20.000 \\
-\botrule
-\end{tabular}}{{\it Note:} Input sequence length for each phylogenetic class present in our animal dataset.
-As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000].
-If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached.
-}
-\end{table}
-
-\subsection{Overlapping evaluation by species}
-\label{sec:overlapping}
-\begin{figure}[!tpb]
-% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
-\caption{Caption, caption.}\label{fig:animals_overlapping}
-\end{figure}
-
-\subsection{Detailed model performance}
-
-\begin{table}[!t]
-\processtable{Class weights\label{tab:detailed_results}} {
-\begin{tabular}{@{}lll@{}}
-\toprule Class & Animal Value & Plant Value \\
-\midrule
-Intergenic & 0.7 & 0.3 \\
-UTR & 1.6 & 1.0 \\
-CDS & 1.2 & 0.9 \\
-Intron & 1.2 & 0.3 \\
-\botrule
-\end{tabular}}{{\it Note:} }
-\end{table}
-
 \end{document}

### 2020-04-22 paper/supplements.tex
lots of supplement stuff

@@
 \usepackage{authblk}
+\usepackage{natbib}
 
@@
 
-\section{Introduction}
-Here is the text of your introduction.
+\newpage
+\section{Dilated CNN Baseline Network Search}
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}ll@{}}
+\hline
+Parameter & Search Space\\ [0.5ex]
+\hline
+Kernel Size & \{4, 8, 12, 16\}\\
+Initial Filter Depth & \{32, 64, 96, 128\}\\
+Number of Layers Before Doubling Filter Count & \{1, 2\}\\
+Dilation Multiplier & \{2, 3\}\\
+Number of Convolutional Layers & \{2, 3, 4, 5, 6, 7, 8\}\\
+Number of Hidden Fully Connected (FC) Layers & \{0, 1, 2\}\\	
+Dropout Used on FC and Final Conv Layer Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
+Learning Rate & \{1e-3, 1e-4\}\\
+\hline
+\end{tabular}
+\caption{Above are all parameters used during neural architecture search for the dilated CNN baseline.
+The same overall space was used for plant and animal data.
+We did, however, run multiple distinct searches that sometimes only operated over a subset of the given parameters.
+This was done as those seemed to be the most promising.
+We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs.
+Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.
+Runs with a Genic F1 below 0.5 after 10 epochs were stopped and the overall maximum epoch was 15.
+The performances seemed to be leveling off well before that.
+The batch size used for almost all runs was 32, the dilation was capped to 81 and the size of each fully connected layer was fixed at 128.
+New parameters where chosen at random.
+The implementation and parameter space is very roughly based on \citep{gupta2017dilated} and can be found in the Helixer source code repository.
+In total, we trained 30 models for the animals and 48 with the plant data.}
+\end{table}
 
-\subsection{Subsection Heading Here}
-Write your subsection text here.
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Parameter & Animals & Plants\\ [0.5ex]
+\hline
+Kernel Size & 16 & 16\\
+Initial Filter Depth & 96 & 32\\
+Number of Layers Before Doubling Filter Count & 2 & 2\\
+Dilation Multiplier & 3 & 3\\
+Number of Convolutional Layers & 6 & 8\\
+Number of Hidden Fully Connected (FC) Layers & 2 & 0\\
+Dropout Used on FC Output & 0.0 & 0.0\\
+Learning Rate & 1e-4 & 1e-3\\
+\hline
+\end{tabular}
+\caption{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.
+The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
+\end{table}
 
-\section{Conclusion}
-Write your conclusion here.
+\newpage
+\section{Hyperparameters}
+
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}ll@{}}
+\hline
+Parameter & Value \\ [0.5ex]
+\hline
+Layers & 4\\
+Units per layer & 256\\
+Learning rate & 1e-4\\
+Optimizer & Adam\\
+\hline
+\end{tabular}
+\caption{}
+\label{suptab:params}
+\end{table}
+
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Class & Animal Value & Plant Value \\ [0.5ex]
+\hline
+Intergenic & 0.7 & 0.3 \\
+UTR & 1.6 & 1.0 \\
+CDS & 1.2 & 0.9 \\
+Intron & 1.2 & 0.3 \\
+\hline
+\end{tabular}
+\caption{The class weights are the only hyperparameters that are different for animals and plants.
+The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}
+\end{table}
+
+\newpage
+\section{Training Data statistics}
+\label{sec:training_data}
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+& Animals & Plants\\ [0.5ex]
+\hline
+Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
+Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
+Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
+Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
+Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
+Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
+Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
+\hline
+\end{tabular}
+\caption{The description of Table [?] applies here as well.}
+\end{table}
+
+\newpage
+\section{Longer Sequence Input}
+\label{sec:longer}
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Class & Input Length in bp\\ [0.5ex]
+\hline
+Mammalia & 200.000 \\
+Reptilia (excluding Aves) & 200.000 \\
+Aves & 100.000 \\
+Amphibia & 100.000 \\
+Actinopterygii & 50.000 \\
+Chondrichthyes & 50.000 \\
+Actinistia & 50.000 \\
+Ascidiacea & 20.000 \\
+Insecta & 20.000 \\
+\hline
+\end{tabular}
+\caption{Input sequence length for each phylogenetic class present in our animal dataset.
+As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000].
+If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached.
+}
+\end{table}
+
+\newpage
+\section{Overlapping evaluation by species}
+\label{sec:overlapping}
+\begin{figure}[!h]
+% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
+\caption{Caption, caption.}\label{fig:animals_overlapping}
+\end{figure}
+
+\newpage
+\section{In-depth model performance}
+
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Class & Animal Value & Plant Value \\ [0.5ex]
+\hline
+Intergenic & 0.7 & 0.3 \\
+UTR & 1.6 & 1.0 \\
+CDS & 1.2 & 0.9 \\
+Intron & 1.2 & 0.3 \\
+\hline
+\end{tabular}
+\caption{}
+\end{table}
+
+\renewcommand\refname{Supplemental Reference}
+\bibliographystyle{natbib}
+\bibliography{literature}
 

### 2020-04-22 paper/supplements.tex
stuff

@@
 \usepackage{natbib}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks,
+    citecolor=black,
+    filecolor=black,
+    linkcolor=black,
+    urlcolor=black
+}
 
@@
 
-\title{Supplemental Material for \\ Helixer: Cross-species gene annotation of large eucaryotic genomes using deep learning}
+\title{Supplemental Material for \\ Helixer: Cross-species Gene Annotation of Large Eucaryotic Genomes Using Deep Learning}
 
@@
 \maketitle
+\tableofcontents
+
+\newpage
+\section{In-depth model performance}
+
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Class & Animal Value & Plant Value \\ [0.5ex]
+\hline
+Intergenic & 0.7 & 0.3 \\
+UTR & 1.6 & 1.0 \\
+CDS & 1.2 & 0.9 \\
+Intron & 1.2 & 0.3 \\
+\hline
+\end{tabular}
+\caption{}
+\end{table}
+
+\newpage
+\section{Hyperparameters}
+
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}ll@{}}
+\hline
+Parameter & Value \\ [0.5ex]
+\hline
+Layers & 4\\
+Units per layer & 256\\
+Learning rate & 1e-4\\
+Optimizer & Adam\\
+\hline
+\end{tabular}
+\caption{All hyperparameters used for the }
+\label{suptab:params}
+\end{table}
+
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Class & Animal Value & Plant Value \\ [0.5ex]
+\hline
+Intergenic & 0.7 & 0.3 \\
+UTR & 1.6 & 1.0 \\
+CDS & 1.2 & 0.9 \\
+Intron & 1.2 & 0.3 \\
+\hline
+\end{tabular}
+\caption{The class weights are the only hyperparameters that are different for animals and plants.
+The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}
+\end{table}
+
+\newpage
+\section{Longer Sequence Input}
+\label{sec:longer}
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Class & Input Length in bp\\ [0.5ex]
+\hline
+Mammalia & 200.000 \\
+Reptilia (excluding Aves) & 200.000 \\
+Aves & 100.000 \\
+Amphibia & 100.000 \\
+Actinopterygii & 50.000 \\
+Chondrichthyes & 50.000 \\
+Actinistia & 50.000 \\
+Ascidiacea & 20.000 \\
+Insecta & 20.000 \\
+\hline
+\end{tabular}
+\caption{Input sequence length for each phylogenetic class present in our animal dataset.
+As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000].
+If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached.
+}
+\end{table}
+
+\newpage
+\section{Overlapping Evaluation by Species}
+\label{sec:overlapping}
+\begin{figure}[!h]
+% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
+\caption{Caption, caption.}\label{fig:animals_overlapping}
+\end{figure}
+
+\newpage
+\section{Training Data statistics}
+\label{sec:training_data}
+\begin{table}[!h]
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+& Animals & Plants\\ [0.5ex]
+\hline
+Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
+Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
+Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
+Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
+Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
+Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
+Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
+\hline
+\end{tabular}
+\caption{The description of Table [?] applies here as well.}
+\end{table}
 
@@
 
-\newpage
-\section{Hyperparameters}
-
-\begin{table}[!h]
-\renewcommand\thetable{S1}
-\centering
-\begin{tabular}{@{}ll@{}}
-\hline
-Parameter & Value \\ [0.5ex]
-\hline
-Layers & 4\\
-Units per layer & 256\\
-Learning rate & 1e-4\\
-Optimizer & Adam\\
-\hline
-\end{tabular}
-\caption{}
-\label{suptab:params}
-\end{table}
-
-\begin{table}[!h]
-\renewcommand\thetable{S1}
-\centering
-\begin{tabular}{@{}lll@{}}
-\hline
-Class & Animal Value & Plant Value \\ [0.5ex]
-\hline
-Intergenic & 0.7 & 0.3 \\
-UTR & 1.6 & 1.0 \\
-CDS & 1.2 & 0.9 \\
-Intron & 1.2 & 0.3 \\
-\hline
-\end{tabular}
-\caption{The class weights are the only hyperparameters that are different for animals and plants.
-The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}
-\end{table}
-
-\newpage
-\section{Training Data statistics}
-\label{sec:training_data}
-\begin{table}[!h]
-\renewcommand\thetable{S1}
-\centering
-\begin{tabular}{@{}lll@{}}
-\hline
-& Animals & Plants\\ [0.5ex]
-\hline
-Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
-Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
-Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
-Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
-Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
-Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
-Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
-\hline
-\end{tabular}
-\caption{The description of Table [?] applies here as well.}
-\end{table}
-
-\newpage
-\section{Longer Sequence Input}
-\label{sec:longer}
-\begin{table}[!h]
-\renewcommand\thetable{S1}
-\centering
-\begin{tabular}{@{}lll@{}}
-\hline
-Class & Input Length in bp\\ [0.5ex]
-\hline
-Mammalia & 200.000 \\
-Reptilia (excluding Aves) & 200.000 \\
-Aves & 100.000 \\
-Amphibia & 100.000 \\
-Actinopterygii & 50.000 \\
-Chondrichthyes & 50.000 \\
-Actinistia & 50.000 \\
-Ascidiacea & 20.000 \\
-Insecta & 20.000 \\
-\hline
-\end{tabular}
-\caption{Input sequence length for each phylogenetic class present in our animal dataset.
-As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000].
-If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached.
-}
-\end{table}
-
-\newpage
-\section{Overlapping evaluation by species}
-\label{sec:overlapping}
-\begin{figure}[!h]
-% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
-\caption{Caption, caption.}\label{fig:animals_overlapping}
-\end{figure}
-
-\newpage
-\section{In-depth model performance}
-
-\begin{table}[!h]
-\renewcommand\thetable{S1}
-\centering
-\begin{tabular}{@{}lll@{}}
-\hline
-Class & Animal Value & Plant Value \\ [0.5ex]
-\hline
-Intergenic & 0.7 & 0.3 \\
-UTR & 1.6 & 1.0 \\
-CDS & 1.2 & 0.9 \\
-Intron & 1.2 & 0.3 \\
-\hline
-\end{tabular}
-\caption{}
-\end{table}
-
 \renewcommand\refname{Supplemental Reference}

### 2020-04-23 paper/images/overlapping/tile.sh
name files and tile script

@@ -1,0 +1,20 @@
+#! /bin/bash
+
+dataset="plants"
+folder="./$dataset"
+plots_per_page=15
+tile="3x"
+
+total_count=$(ls -1 $folder | grep -v "aggregate_comparison.png" | wc -l)
+count=1
+
+pwd_folder=$(pwd)
+cd $folder
+
+for offset in $(seq 1 $plots_per_page $total_count)
+do
+	files=$(cat ~/git/helixer_scratch/data_insight/$dataset/genomes_by_n75 | tail +$((offset+1)) | head -$plots_per_page | cut -d "," -f1 | sed -e 's/$/_comparison.png/')
+	montage -mode concatenate -tile $tile $files $pwd_folder/"montage_plants"$count".png"
+	((count++))
+done
+

### 2020-04-23 paper/images/overlapping/tile.sh
position wise eval pics

@@ -14,7 +14,7 @@ cd $folder
 for offset in $(seq 1 $plots_per_page $total_count)
 do
 	files=$(cat ~/git/helixer_scratch/data_insight/$dataset/genomes_by_n75 | tail +$((offset+1)) | head -$plots_per_page | cut -d "," -f1 | sed -e 's/$/_comparison.png/')
-	montage -mode concatenate -tile $tile $files $pwd_folder/"montage_plants"$count".png"
+	montage -mode concatenate -tile $tile $files $pwd_folder/"montage_"$dataset$count".png"
 	((count++))
 done
 

### 2020-04-24 paper/supplements.tex
position wise eval pics
supplement images

@@
 \newpage
+\section{Model Evaluation by Species}
+The following plots show the prediction performance in Subgenic F1 per species as well as the difference to AUGUSTUS for both animals and plants.
+
+\subsection{Animals}
+\newpage
+
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part1}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part2}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part1}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part2}}
+\end{figure}
+
+\clearpage
+\subsection{Plants}
+
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.0\textwidth]{images/plants_f1_scores}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.0\textwidth]{images/plants_differences}}
+\end{figure}
+
+\clearpage
 \section{Hyperparameters}
@@
 \newpage
-\section{Overlapping Evaluation by Species}
+\section{Effect of Overlapping by Species}
 \label{sec:overlapping}
+
+The following plots show the effect of overlapping for each species we worked with individually.
+The plots are ordered with descending N75, as we found that overlapping tends to produce less desirable patterns in the prediction quality for the very fragmented genomes.
+
+\subsection{Animals}
+\newpage
+
 \begin{figure}[!h]
-% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
-\caption{Caption, caption.}\label{fig:animals_overlapping}
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals1}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals2}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals3}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals4}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals5}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals6}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals7}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals8}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals9}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals10}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals11}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals12}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
 \end{figure}
 
-\newpage
+\clearpage
+\subsection{Plants}
+
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants1}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants2}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants3}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants4}}
+\end{figure}
+
+\clearpage
 \section{Training Data statistics}

### 2020-04-24 paper/supplements.tex
detailed plant performance added

--- 2020-04-22 paper/supplements.tex
@@
 \newpage
+\section{Model Evaluation by Species}
+The following tables and plots show the prediction performance of Hexlixer for each species we worked with.
+The plots display the Subgenic F1 as well as the difference to AUGUSTUS with respect to the phylogenetic placement by NCBI Taxonomy.
+
+\subsection{Animals}
+\newpage
+
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part1}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part2}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part1}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part2}}
+%\end{figure}
+
+\clearpage
+\subsection{Plants}
+
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\renewcommand\thetable{S1}
+\centering
+%\vspace*{-3.5cm}
+%\hspace*{-2.5cm}
+\begin{tabular}{@{}llllllll@{}}
+\hline
+Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1& Genic F1 \\ [0.5ex]
+\hline
+Ananas\_comosus & 0.8749 & 0.9224 & 0.5674 & 0.8591 & 0.7325 & 0.7691 & 0.753 \\
+Amaranthus\_hypochondriacus & 0.947 & 0.9728 & 0.5729 & 0.8535 & 0.8964 & 0.8856 & 0.864 \\
+Arabidopsis\_lyrata & 0.9438 & 0.9701 & 0.637 & 0.8957 & 0.8196 & 0.8649 & 0.8314 \\
+Asparagus\_officinalis & 0.6042 & 0.2635 & 0 & 0.8517 & 0.6979 & 0.7246 & 0.7206 \\
+Arabidopsis\_thaliana & 0.9542 & 0.9729 & 0.7294 & 0.953 & 0.9329 & 0.9456 & 0.9181 \\
+Amborella\_trichopoda & 0.6936 & 0.3451 & 0 & 0.8617 & 0.7816 & 0.7965 & 0.7933 \\
+Brachypodium\_distachyon & 0.9326 & 0.9625 & 0.6232 & 0.9027 & 0.8772 & 0.8879 & 0.8426 \\
+Brassica\_oleracea & 0.8648 & 0 & 0 & 0.9321 & 0.8785 & 0.911 & 0.9051 \\
+Cicer\_arietinum & 0.8583 & 0.7838 & 0 & 0.8872 & 0.8952 & 0.8921 & 0.8885 \\
+Citrus\_clementina & 0.9554 & 0.977 & 0.6277 & 0.9175 & 0.9027 & 0.9095 & 0.8771 \\
+Capsella\_grandiflora & 0.9465 & 0.9656 & 0.7117 & 0.9545 & 0.9139 & 0.9389 & 0.9145 \\
+Carica\_papaya & 0.9009 & 0.8926 & 0 & 0.9043 & 0.9156 & 0.9113 & 0.9057 \\
+Chenopodium\_quinoa & 0.944 & 0.9703 & 0.6057 & 0.855 & 0.8311 & 0.8382 & 0.8244 \\
+Chlamydomonas\_reinhardtii & 0.9484 & 0.9657 & 0.8039 & 0.9577 & 0.9546 & 0.9562 & 0.9254 \\
+Capsella\_rubella & 0.9455 & 0.9669 & 0.668 & 0.9485 & 0.9138 & 0.9352 & 0.9032 \\
+Cucumis\_sativus & 0.9515 & 0.9729 & 0.6174 & 0.9268 & 0.9237 & 0.9249 & 0.8962 \\
+Citrus\_sinensis & 0.9545 & 0.9766 & 0.5878 & 0.9147 & 0.9027 & 0.9082 & 0.8778 \\
+Coccomyxa\_subellipsoidea\_C-169 & 0.8467 & 0.8573 & 0.4164 & 0.8771 & 0.8896 & 0.8847 & 0.8409 \\
+Chromochloris\_zofingiensis & 0.8849 & 0.3095 & 0 & 0.9129 & 0.9299 & 0.9203 & 0.9027 \\
+Daucus\_carota & 0.8924 & 0.703 & 0 & 0.9284 & 0.9212 & 0.924 & 0.9175 \\
+Dunaliella\_salina & 0.7458 & 0.8473 & 0.1832 & 0.6152 & 0.4306 & 0.4585 & 0.4352 \\
+Eucalyptus\_grandis & 0.9447 & 0.9712 & 0.5327 & 0.8391 & 0.8245 & 0.8305 & 0.7956 \\
+Eutrema\_salsugineum & 0.9555 & 0.9755 & 0.6501 & 0.9301 & 0.8985 & 0.9179 & 0.8908 \\
+Fragaria\_vesca & 0.8776 & 0.9192 & 0.633 & 0.8666 & 0.7813 & 0.8218 & 0.7972 \\
+Glycine\_max & 0.9755 & 0.9886 & 0.6564 & 0.9265 & 0.9165 & 0.92 & 0.8825 \\
+Gossypium\_raimondii & 0.9784 & 0.9901 & 0.6836 & 0.9216 & 0.8835 & 0.9004 & 0.8701 \\
+Helianthus\_annuus & 0.9747 & 0.9875 & 0.4514 & 0.7758 & 0.6562 & 0.6975 & 0.6763 \\
+Hordeum\_vulgare & 0.9826 & 0.9917 & 0.4462 & 0.7774 & 0.624 & 0.6762 & 0.6503 \\
+Kalanchoe\_fedtschenkoi & 0.9415 & 0.9693 & 0.6067 & 0.9112 & 0.8569 & 0.8833 & 0.8372 \\
+Lactuca\_sativa & 0.98 & 0.9908 & 0.5771 & 0.867 & 0.7075 & 0.7848 & 0.7679 \\
+Linum\_usitatissimum & 0.8411 & 0.3605 & 0 & 0.9452 & 0.8435 & 0.8991 & 0.8912 \\
+Musa\_acuminata & 0.926 & 0.9598 & 0.4746 & 0.8856 & 0.8984 & 0.8946 & 0.8682 \\
+Malus\_domestica & 0.9629 & 0.9816 & 0.6214 & 0.9049 & 0.8661 & 0.8825 & 0.852 \\
+Manihot\_esculenta & 0.9748 & 0.9879 & 0.681 & 0.9365 & 0.9199 & 0.9261 & 0.8981 \\
+Mimulus\_guttatus & 0.9704 & 0.985 & 0.7101 & 0.9405 & 0.9234 & 0.9319 & 0.9062 \\
+Marchantia\_polymorpha & 0.9409 & 0.9708 & 0.6294 & 0.8786 & 0.8797 & 0.8792 & 0.8083 \\
+Micromonas\_pusilla & 0.6896 & 0.639 & 0.3364 & 0.7517 & 0.617 & 0.7358 & 0.7223 \\
+Micromonas\_sp.\_RCC299 & 0.592 & 0.5639 & 0.2706 & 0.637 & 0.5212 & 0.6301 & 0.6152 \\
+Medicago\_truncatula & 0.9181 & 0.9474 & 0.6813 & 0.8625 & 0.8639 & 0.8633 & 0.8475 \\
+Olea\_europaea & 0.7935 & 0.7695 & 0 & 0.794 & 0.8262 & 0.8142 & 0.8096 \\
+Ostreococcus\_sp.\_lucimarinus & 0.2174 & 0.3241 & 0.0451 & 0.0746 & 0.0441 & 0.0734 & 0.0725 \\
+Oryza\_sativa & 0.9061 & 0.9439 & 0.6234 & 0.8367 & 0.827 & 0.8312 & 0.8044 \\
+Oropetium\_thomaeum & 0.8682 & 0.9194 & 0.5482 & 0.8605 & 0.8045 & 0.8275 & 0.8058 \\
+Physcomitrella\_patens & 0.9589 & 0.9816 & 0.637 & 0.8701 & 0.8554 & 0.8626 & 0.8064 \\
+Prunus\_persica & 0.9434 & 0.9688 & 0.6422 & 0.9089 & 0.8944 & 0.9012 & 0.8605 \\
+Populus\_trichocarpa & 0.9609 & 0.9806 & 0.6771 & 0.9244 & 0.916 & 0.9196 & 0.882 \\
+Porphyra\_umbilicalis & 0.7648 & 0.8529 & 0.2087 & 0.6096 & 0.2986 & 0.565 & 0.5139 \\
+Ricinus\_communis & 0.9256 & 0.9442 & 0.4673 & 0.8984 & 0.9364 & 0.9192 & 0.9055 \\
+Sorghum\_bicolor & 0.9654 & 0.983 & 0.5792 & 0.9033 & 0.8608 & 0.8775 & 0.8292 \\
+Setaria\_italica & 0.9561 & 0.9775 & 0.6527 & 0.9166 & 0.8933 & 0.9037 & 0.8646 \\
+Solanum\_lycopersicum & 0.9476 & 0.9725 & 0.4948 & 0.8826 & 0.8367 & 0.8509 & 0.8127 \\
+Selaginella\_moellendorffii & 0.7166 & 0.7977 & 0.1825 & 0.7436 & 0.435 & 0.662 & 0.6388 \\
+Spirodela\_polyrhiza & 0.7784 & 0 & 0 & 0.903 & 0.8414 & 0.8624 & 0.8529 \\
+Solanum\_tuberosum & 0.7272 & 0.0994 & 0 & 0.8551 & 0.8199 & 0.8335 & 0.8299 \\
+Triticum\_aestivum & 0.9238 & 0.9601 & 0.5238 & 0.8455 & 0.8356 & 0.8399 & 0.7961 \\
+Theobroma\_cacao & 0.8743 & 0.9295 & 0.5257 & 0.8677 & 0.545 & 0.6421 & 0.627 \\
+Volvox\_carteri & 0.8546 & 0.9045 & 0.5535 & 0.831 & 0.8241 & 0.8269 & 0.7706 \\
+Vitis\_vinifera & 0.7918 & 0.0076 & 0 & 0.9229 & 0.8613 & 0.8771 & 0.8724 \\
+Zostera\_marina & 0.9148 & 0.9463 & 0.5865 & 0.9052 & 0.8261 & 0.8571 & 0.8387 \\
+Zea\_mays & 0.9843 & 0.9931 & 0.6073 & 0.886 & 0.854 & 0.8643 & 0.8406 \\
+\hline
+\end{tabular}}
+\end{table}
+
+\clearpage
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.0\textwidth]{images/plants_f1_scores}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.0\textwidth]{images/plants_differences}}
+\end{figure}
+
+\clearpage
 \section{Hyperparameters}
@@
 \newpage
-\section{Overlapping Evaluation by Species}
+\section{Effect of Overlapping by Species}
 \label{sec:overlapping}
-\begin{figure}[!h]
-% \centerline{\includegraphics[height=0.01\paperheight]{images/overlapping/animals_overlapping}}
-\caption{Caption, caption.}\label{fig:animals_overlapping}
-\end{figure}
 
+The following plots show the effect of overlapping for each species we worked with individually.
+The plots are ordered with descending N75, as we found that overlapping tends to produce less desirable patterns in the prediction quality for the very fragmented genomes.
+
+\subsection{Animals}
 \newpage
+
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals1}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals2}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals3}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals4}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals5}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals6}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals7}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals8}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals9}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals10}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals11}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals12}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
+%\end{figure}
+%
+%\clearpage
+%\subsection{Plants}
+%
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants1}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants2}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants3}}
+%\end{figure}
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants4}}
+%\end{figure}
+%
+\clearpage
 \section{Training Data statistics}

### 2020-04-24 paper/supplements.tex
detailed results added

@@
 \newpage
-\section{In-depth model performance}
-
-\begin{table}[!h]
-\renewcommand\thetable{S1}
-\centering
-\begin{tabular}{@{}lll@{}}
-\hline
-Class & Animal Value & Plant Value \\ [0.5ex]
-\hline
-Intergenic & 0.7 & 0.3 \\
-UTR & 1.6 & 1.0 \\
-CDS & 1.2 & 0.9 \\
-Intron & 1.2 & 0.3 \\
-\hline
-\end{tabular}
-\caption{}
-\end{table}
-
-\newpage
 \section{Model Evaluation by Species}
 The following tables and plots show the prediction performance of Hexlixer for each species we worked with.
-The plots display the Subgenic F1 as well as the difference to AUGUSTUS with respect to the phylogenetic placement by NCBI Taxonomy.
+An F1 score of zero can indicate that the reference did not include that class.
+The plots display the Subgenic F1 as well as the difference to AUGUSTUS with respect to the phylogenetic placement by the NCBI taxonomy database \citep{federhen2012ncbi}.
 
@@
 \newpage
+
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}llllllll@{}}
+\hline
+Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1 & Genic F1 \\ [0.5ex]
+\hline
+acanthochromis\_polyacanthus & 0.9036 & 0.9323 & 0.6 & 0.8663 & 0.8764 & 0.8753 & 0.8547 \\
+ailuropoda\_melanoleuca & 0.9343 & 0.9488 & 0 & 0.8757 & 0.9173 & 0.9153 & 0.913 \\
+amphilophus\_citrinellus & 0.8894 & 0.8501 & 0.0004 & 0.8609 & 0.9189 & 0.9123 & 0.9039 \\
+amphiprion\_ocellaris & 0.9001 & 0.9304 & 0.5473 & 0.8391 & 0.8763 & 0.8723 & 0.8472 \\
+amphiprion\_percula & 0.8942 & 0.923 & 0.5586 & 0.8375 & 0.8781 & 0.874 & 0.851 \\
+anabas\_testudineus & 0.9036 & 0.93 & 0.6076 & 0.8632 & 0.9037 & 0.8978 & 0.8709 \\
+anas\_platyrhynchos\_platyrhynchos & 0.9218 & 0.9498 & 0.3189 & 0.8432 & 0.8579 & 0.857 & 0.8478 \\
+anolis\_carolinensis & 0.9051 & 0.942 & 0.4548 & 0.7631 & 0.788 & 0.7867 & 0.7776 \\
+anser\_brachyrhynchus & 0.9286 & 0.9533 & 0.3737 & 0.8571 & 0.8766 & 0.8754 & 0.8675 \\
+aotus\_nancymaae & 0.944 & 0.9629 & 0.3618 & 0.8557 & 0.9012 & 0.8996 & 0.895 \\
+apteryx\_haastii & 0.9025 & 0.9207 & 0.2082 & 0.8287 & 0.89 & 0.886 & 0.8793 \\
+apteryx\_owenii & 0.9072 & 0.935 & 0.3271 & 0.8354 & 0.8641 & 0.8623 & 0.8534 \\
+apteryx\_rowi & 0.9102 & 0.9366 & 0.3352 & 0.8395 & 0.8716 & 0.8696 & 0.8608 \\
+astatotilapia\_calliptera & 0.8663 & 0.9014 & 0.5666 & 0.7731 & 0.8438 & 0.8352 & 0.8169 \\
+astyanax\_mexicanus & 0.8849 & 0.9124 & 0.5784 & 0.8246 & 0.8632 & 0.8601 & 0.8454 \\
+betta\_splendens & 0.8423 & 0.8834 & 0.3626 & 0.8065 & 0.8227 & 0.8197 & 0.7842 \\
+bison\_bison\_bison & 0.9021 & 0.9187 & 0 & 0.8575 & 0.8829 & 0.8819 & 0.8796 \\
+bos\_indicus\_hybrid & 0.9304 & 0.9569 & 0.5481 & 0.8079 & 0.8505 & 0.8489 & 0.8414 \\
+bos\_mutus & 0.9234 & 0.9498 & 0.3885 & 0.843 & 0.8593 & 0.8586 & 0.8525 \\
+bos\_taurus & 0.9385 & 0.9622 & 0.4702 & 0.8379 & 0.8637 & 0.8627 & 0.8549 \\
+caenorhabditis\_elegans & 0.7423 & 0.8288 & 0.1477 & 0.7385 & 0.5006 & 0.5965 & 0.5709 \\
+calidris\_pugnax & 0.9166 & 0.9452 & 0.4135 & 0.8282 & 0.8568 & 0.8552 & 0.846 \\
+calidris\_pygmaea & 0.9195 & 0.9461 & 0.4404 & 0.8485 & 0.8682 & 0.867 & 0.8576 \\
+callithrix\_jacchus & 0.9352 & 0.9584 & 0.2818 & 0.833 & 0.8692 & 0.8679 & 0.8642 \\
+callorhinchus\_milii & 0.8735 & 0.9123 & 0.3684 & 0.8019 & 0.8083 & 0.8079 & 0.7951 \\
+canis\_familiaris & 0.9271 & 0.9542 & 0.2754 & 0.8239 & 0.8481 & 0.8472 & 0.8388 \\
+canis\_lupus\_dingo & 0.937 & 0.9572 & 0.3463 & 0.8522 & 0.8976 & 0.8956 & 0.8906 \\
+capra\_hircus & 0.9395 & 0.962 & 0.4262 & 0.8521 & 0.8767 & 0.8757 & 0.8695 \\
+carlito\_syrichta & 0.9312 & 0.9554 & 0.4552 & 0.8477 & 0.8651 & 0.8644 & 0.8603 \\
+castor\_canadensis & 0.9177 & 0.9467 & 0.3725 & 0.8477 & 0.843 & 0.8433 & 0.8356 \\
+cavia\_aperea & 0.9405 & 0.9619 & 0 & 0.8462 & 0.8822 & 0.8805 & 0.8784 \\
+cavia\_porcellus & 0.9462 & 0.968 & 0.5285 & 0.8483 & 0.8652 & 0.8645 & 0.8563 \\
+cebus\_capucinus & 0.946 & 0.9659 & 0.5059 & 0.8455 & 0.8928 & 0.8911 & 0.885 \\
+cercocebus\_atys & 0.9457 & 0.9658 & 0.4896 & 0.8506 & 0.8917 & 0.8902 & 0.8834 \\
+chelonoidis\_abingdonii & 0.8808 & 0.9174 & 0.296 & 0.7582 & 0.8106 & 0.808 & 0.803 \\
+chinchilla\_lanigera & 0.9437 & 0.9664 & 0.507 & 0.8537 & 0.862 & 0.8617 & 0.8527 \\
+chlorocebus\_sabaeus & 0.9418 & 0.9645 & 0.4666 & 0.8569 & 0.8646 & 0.8643 & 0.8574 \\
+choloepus\_hoffmanni & 0.9036 & 0.9253 & 0 & 0.7481 & 0.8818 & 0.8744 & 0.8725 \\
+chrysemys\_picta\_bellii & 0.8882 & 0.9294 & 0.3391 & 0.7153 & 0.7733 & 0.7706 & 0.7631 \\
+ciona\_intestinalis & 0.7474 & 0.8196 & 0.1528 & 0.6708 & 0.6638 & 0.6655 & 0.6383 \\
+ciona\_savignyi & 0.7662 & 0.826 & 0.071 & 0.6471 & 0.6982 & 0.6878 & 0.6764 \\
+clupea\_harengus & 0.863 & 0.9009 & 0.3972 & 0.7854 & 0.8341 & 0.8282 & 0.8018 \\
+colobus\_angolensis\_palliatus & 0.9426 & 0.9631 & 0.5241 & 0.8557 & 0.8874 & 0.8863 & 0.8821 \\
+cottoperca\_gobio & 0.8889 & 0.9177 & 0.5953 & 0.8284 & 0.8756 & 0.8694 & 0.848 \\
+coturnix\_japonica & 0.9209 & 0.9461 & 0.3351 & 0.8604 & 0.8757 & 0.8747 & 0.8666 \\
+cricetulus\_griseus\_picr & 0.9315 & 0.9546 & 0.3524 & 0.8647 & 0.8778 & 0.8772 & 0.8712 \\
+crocodylus\_porosus & 0.917 & 0.9472 & 0.3471 & 0.8005 & 0.8337 & 0.8323 & 0.8262 \\
+cyanistes\_caeruleus & 0.9228 & 0.9518 & 0.3437 & 0.8501 & 0.8445 & 0.8449 & 0.8362 \\
+cynoglossus\_semilaevis & 0.8751 & 0.9027 & 0.5883 & 0.8718 & 0.8702 & 0.8704 & 0.8456 \\
+cyprinodon\_variegatus & 0.8978 & 0.9298 & 0.5196 & 0.8531 & 0.8577 & 0.8573 & 0.835 \\
+danio\_rerio & 0.873 & 0.9151 & 0.3666 & 0.7688 & 0.7974 & 0.7954 & 0.7734 \\
+dasypus\_novemcinctus & 0.9371 & 0.9628 & 0.4583 & 0.8148 & 0.8328 & 0.832 & 0.8245 \\
+denticeps\_clupeoides & 0.8766 & 0.9144 & 0.3753 & 0.8112 & 0.8603 & 0.8528 & 0.814 \\
+dipodomys\_ordii & 0.9385 & 0.9599 & 0.4335 & 0.8549 & 0.8877 & 0.8863 & 0.8798 \\
+dromaius\_novaehollandiae & 0.9118 & 0.9414 & 0.41 & 0.8481 & 0.8486 & 0.8486 & 0.8395 \\
+drosophila\_melanogaster & 0.8409 & 0.8902 & 0.4738 & 0.9278 & 0.6762 & 0.7623 & 0.7467 \\
+echinops\_telfairi & 0.7972 & 0.8476 & 0 & 0.741 & 0.7035 & 0.7048 & 0.7033 \\
+electrophorus\_electricus & 0.8849 & 0.9176 & 0.3795 & 0.837 & 0.882 & 0.8749 & 0.84 \\
+eptatretus\_burgeri & 0.7511 & 0.8463 & 0.1436 & 0.5056 & 0.3671 & 0.371 & 0.3674 \\
+equus\_asinus\_asinus & 0.9301 & 0.9476 & 0.2589 & 0.8651 & 0.9073 & 0.9055 & 0.9011 \\
+\hline
+\end{tabular}}
+\end{table}
+
+\clearpage
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}llllllll@{}}
+\hline
+Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1 & Genic F1 \\ [0.5ex]
+\hline
+equus\_caballus & 0.9327 & 0.9571 & 0.2984 & 0.8417 & 0.8686 & 0.8675 & 0.86 \\
+erinaceus\_europaeus & 0.8202 & 0.863 & 0 & 0.7519 & 0.7471 & 0.7472 & 0.7454 \\
+erpetoichthys\_calabaricus & 0.8659 & 0.9173 & 0.3119 & 0.6275 & 0.6816 & 0.6794 & 0.6738 \\
+esox\_lucius & 0.8749 & 0.912 & 0.5692 & 0.8331 & 0.8293 & 0.8297 & 0.8123 \\
+felis\_catus & 0.9301 & 0.9557 & 0.3452 & 0.85 & 0.8561 & 0.8559 & 0.8494 \\
+ficedula\_albicollis & 0.9287 & 0.9559 & 0.5561 & 0.8493 & 0.8587 & 0.8581 & 0.8473 \\
+fukomys\_damarensis & 0.9434 & 0.9664 & 0.4492 & 0.8454 & 0.857 & 0.8565 & 0.8479 \\
+fundulus\_heteroclitus & 0.8951 & 0.927 & 0.543 & 0.8397 & 0.863 & 0.8609 & 0.84 \\
+gadus\_morhua & 0.886 & 0.8931 & 0 & 0.8422 & 0.8924 & 0.887 & 0.8806 \\
+gallus\_gallus & 0.9325 & 0.9575 & 0.4856 & 0.8165 & 0.8784 & 0.8743 & 0.8637 \\
+gambusia\_affinis & 0.8873 & 0.9182 & 0.5344 & 0.8719 & 0.875 & 0.8747 & 0.8437 \\
+gasterosteus\_aculeatus & 0.8774 & 0.8987 & 0.1991 & 0.8583 & 0.8969 & 0.8896 & 0.8568 \\
+gopherus\_agassizii & 0.8679 & 0.9133 & 0.3725 & 0.7513 & 0.7523 & 0.7522 & 0.7456 \\
+gorilla\_gorilla & 0.9422 & 0.9635 & 0.5272 & 0.8459 & 0.8808 & 0.8796 & 0.8748 \\
+gouania\_willdenowi & 0.8533 & 0.8976 & 0.2835 & 0.7337 & 0.805 & 0.7979 & 0.7674 \\
+haplochromis\_burtoni & 0.9008 & 0.9309 & 0.6155 & 0.8732 & 0.8684 & 0.8689 & 0.8492 \\
+heterocephalus\_glaber\_female & 0.9416 & 0.9644 & 0.4958 & 0.828 & 0.8665 & 0.865 & 0.8565 \\
+hippocampus\_comes & 0.8695 & 0.9026 & 0.4316 & 0.8332 & 0.8622 & 0.8581 & 0.83 \\
+homo\_sapiens & 0.9492 & 0.9701 & 0.4823 & 0.8367 & 0.8648 & 0.8639 & 0.8553 \\
+hucho\_hucho & 0.874 & 0.9203 & 0.3735 & 0.7266 & 0.7813 & 0.7754 & 0.7537 \\
+ictalurus\_punctatus & 0.8791 & 0.9124 & 0.5555 & 0.8398 & 0.849 & 0.8481 & 0.8279 \\
+ictidomys\_tridecemlineatus & 0.944 & 0.9655 & 0.5465 & 0.8627 & 0.8771 & 0.8765 & 0.8686 \\
+jaculus\_jaculus & 0.9344 & 0.9578 & 0.2006 & 0.8484 & 0.8692 & 0.8685 & 0.8632 \\
+junco\_hyemalis & 0.9308 & 0.9538 & 0.4298 & 0.8626 & 0.893 & 0.8911 & 0.8797 \\
+kryptolebias\_marmoratus & 0.8961 & 0.9269 & 0.5683 & 0.8598 & 0.8748 & 0.873 & 0.8458 \\
+labrus\_bergylta & 0.894 & 0.9237 & 0.5812 & 0.8426 & 0.8788 & 0.8741 & 0.8513 \\
+larimichthys\_crocea & 0.8758 & 0.9079 & 0.3582 & 0.8572 & 0.8693 & 0.8677 & 0.8282 \\
+lates\_calcarifer & 0.8741 & 0.9062 & 0.3905 & 0.8387 & 0.867 & 0.8632 & 0.8276 \\
+latimeria\_chalumnae & 0.8817 & 0.9251 & 0.3917 & 0.7154 & 0.7628 & 0.7605 & 0.7529 \\
+lepidothrix\_coronata & 0.9261 & 0.9489 & 0.4337 & 0.8596 & 0.8919 & 0.8899 & 0.8805 \\
+lepisosteus\_oculatus & 0.9119 & 0.9401 & 0.5533 & 0.8676 & 0.8706 & 0.8703 & 0.8559 \\
+lonchura\_striata\_domestica & 0.9261 & 0.9503 & 0.43 & 0.8522 & 0.8859 & 0.8837 & 0.8731 \\
+loxodonta\_africana & 0.9311 & 0.948 & 0 & 0.8752 & 0.9058 & 0.9044 & 0.9024 \\
+macaca\_fascicularis & 0.9452 & 0.9654 & 0.4878 & 0.8549 & 0.8897 & 0.8885 & 0.8821 \\
+macaca\_mulatta & 0.9416 & 0.9637 & 0.3885 & 0.8124 & 0.876 & 0.8738 & 0.8664 \\
+macaca\_nemestrina & 0.946 & 0.966 & 0.4801 & 0.8495 & 0.8916 & 0.8901 & 0.8833 \\
+manacus\_vitellinus & 0.9125 & 0.9426 & 0.3877 & 0.8354 & 0.8483 & 0.8475 & 0.8378 \\
+mandrillus\_leucophaeus & 0.9414 & 0.9623 & 0.5191 & 0.8525 & 0.8856 & 0.8844 & 0.8802 \\
+marmota\_marmota\_marmota & 0.9238 & 0.9439 & 0.3195 & 0.8671 & 0.8922 & 0.8911 & 0.8869 \\
+mastacembelus\_armatus & 0.8885 & 0.9193 & 0.6051 & 0.8577 & 0.8714 & 0.8695 & 0.8464 \\
+maylandia\_zebra & 0.8784 & 0.9159 & 0.5813 & 0.762 & 0.8453 & 0.8352 & 0.8151 \\
+meleagris\_gallopavo & 0.912 & 0.9324 & 0.3154 & 0.8695 & 0.8911 & 0.8895 & 0.8826 \\
+melopsittacus\_undulatus & 0.9237 & 0.9493 & 0.4358 & 0.8516 & 0.8732 & 0.8719 & 0.8632 \\
+meriones\_unguiculatus & 0.9295 & 0.9535 & 0.4615 & 0.8735 & 0.8772 & 0.877 & 0.8682 \\
+mesocricetus\_auratus & 0.9448 & 0.9657 & 0.5247 & 0.8626 & 0.8837 & 0.8828 & 0.8759 \\
+microcebus\_murinus & 0.9429 & 0.9645 & 0.494 & 0.8372 & 0.8818 & 0.8801 & 0.8724 \\
+microtus\_ochrogaster & 0.9443 & 0.9646 & 0.5009 & 0.864 & 0.8923 & 0.8911 & 0.8831 \\
+mola\_mola & 0.8815 & 0.653 & 0 & 0.8648 & 0.9275 & 0.9197 & 0.9095 \\
+monodelphis\_domestica & 0.9257 & 0.9562 & 0.3206 & 0.7962 & 0.7846 & 0.785 & 0.7792 \\
+monopterus\_albus & 0.8819 & 0.9102 & 0.5546 & 0.8518 & 0.8671 & 0.8653 & 0.8461 \\
+mus\_caroli & 0.9372 & 0.9623 & 0.4256 & 0.8419 & 0.8348 & 0.8351 & 0.8305 \\
+mus\_musculus & 0.9542 & 0.9736 & 0.6151 & 0.8638 & 0.8659 & 0.8658 & 0.8569 \\
+mus\_pahari & 0.9352 & 0.9608 & 0.4229 & 0.8431 & 0.8344 & 0.8347 & 0.8301 \\
+mus\_spicilegus & 0.9339 & 0.9583 & 0.4436 & 0.8647 & 0.863 & 0.8631 & 0.8562 \\
+mus\_spretus & 0.9337 & 0.96 & 0.4093 & 0.8314 & 0.8267 & 0.8269 & 0.8224 \\
+mustela\_putorius\_furo & 0.9317 & 0.9578 & 0.4319 & 0.857 & 0.8494 & 0.8497 & 0.8413 \\
+myotis\_lucifugus & 0.942 & 0.9622 & 0.0703 & 0.8534 & 0.9013 & 0.8983 & 0.8933 \\
+nannospalax\_galili & 0.9486 & 0.9685 & 0.4998 & 0.8469 & 0.8844 & 0.883 & 0.8767 \\
+neolamprologus\_brichardi & 0.8895 & 0.9094 & 0.4797 & 0.8523 & 0.8911 & 0.8868 & 0.8696 \\
+neovison\_vison & 0.9362 & 0.9565 & 0.4219 & 0.8752 & 0.895 & 0.8942 & 0.8894 \\
+\hline
+\end{tabular}}
+\end{table}
+
+\clearpage
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}llllllll@{}}
+\hline
+Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1 & Genic F1 \\ [0.5ex]
+\hline
+nomascus\_leucogenys & 0.9433 & 0.9634 & 0.489 & 0.8497 & 0.8912 & 0.8897 & 0.8856 \\
+notamacropus\_eugenii & 0.8951 & 0.9203 & 0 & 0.7692 & 0.8619 & 0.8567 & 0.8552 \\
+notechis\_scutatus & 0.8491 & 0.89 & 0.2625 & 0.7682 & 0.7845 & 0.7834 & 0.776 \\
+nothoprocta\_perdicaria & 0.9062 & 0.8467 & 0 & 0.8533 & 0.9358 & 0.9294 & 0.9255 \\
+numida\_meleagris & 0.9276 & 0.952 & 0.3472 & 0.8565 & 0.8803 & 0.8788 & 0.871 \\
+ochotona\_princeps & 0.8374 & 0.8737 & 0 & 0.7899 & 0.7795 & 0.7799 & 0.7776 \\
+octodon\_degus & 0.9381 & 0.9602 & 0.198 & 0.8305 & 0.8797 & 0.8778 & 0.8721 \\
+oreochromis\_niloticus & 0.8967 & 0.9299 & 0.5269 & 0.8292 & 0.8495 & 0.8472 & 0.8284 \\
+ornithorhynchus\_anatinus & 0.9007 & 0.939 & 0.4056 & 0.7987 & 0.7723 & 0.7738 & 0.7665 \\
+oryctolagus\_cuniculus & 0.9025 & 0.9397 & 0.2709 & 0.7731 & 0.7786 & 0.7784 & 0.7715 \\
+oryzias\_latipes & 0.8864 & 0.9179 & 0.6008 & 0.8146 & 0.8671 & 0.8609 & 0.8401 \\
+oryzias\_melastigma & 0.8842 & 0.9189 & 0.5583 & 0.8437 & 0.8535 & 0.8524 & 0.8265 \\
+otolemur\_garnettii & 0.95 & 0.9681 & 0.1822 & 0.8767 & 0.9011 & 0.9001 & 0.8954 \\
+ovis\_aries & 0.9325 & 0.9569 & 0.2919 & 0.8441 & 0.8695 & 0.8684 & 0.8622 \\
+pan\_paniscus & 0.9532 & 0.972 & 0.5101 & 0.8553 & 0.8818 & 0.8809 & 0.8757 \\
+pan\_troglodytes & 0.9458 & 0.9665 & 0.499 & 0.8469 & 0.8822 & 0.881 & 0.8745 \\
+panthera\_pardus & 0.9366 & 0.9592 & 0.498 & 0.8428 & 0.8787 & 0.8774 & 0.8716 \\
+panthera\_tigris\_altaica & 0.9289 & 0.954 & 0.4484 & 0.835 & 0.8641 & 0.863 & 0.8574 \\
+papio\_anubis & 0.943 & 0.9608 & 0.3066 & 0.8673 & 0.9071 & 0.9057 & 0.9019 \\
+parambassis\_ranga & 0.8571 & 0.8962 & 0.3735 & 0.8001 & 0.845 & 0.8375 & 0.7963 \\
+paramormyrops\_kingsleyae & 0.8926 & 0.925 & 0.591 & 0.8542 & 0.8639 & 0.8628 & 0.8418 \\
+parus\_major & 0.9307 & 0.955 & 0.3955 & 0.8683 & 0.8818 & 0.881 & 0.8703 \\
+pelodiscus\_sinensis & 0.8755 & 0.92 & 0.348 & 0.6937 & 0.7558 & 0.7529 & 0.744 \\
+periophthalmus\_magnuspinnatus & 0.9109 & 0.9341 & 0 & 0.8391 & 0.9087 & 0.8999 & 0.889 \\
+peromyscus\_maniculatus\_bairdii & 0.9164 & 0.9274 & 0.002 & 0.8843 & 0.9059 & 0.905 & 0.903 \\
+petromyzon\_marinus & 0.9126 & 0.9496 & 0.094 & 0.714 & 0.8285 & 0.8155 & 0.8039 \\
+phascolarctos\_cinereus & 0.9254 & 0.953 & 0.3568 & 0.8302 & 0.8395 & 0.8392 & 0.8336 \\
+piliocolobus\_tephrosceles & 0.9374 & 0.9571 & 0.3872 & 0.8289 & 0.8967 & 0.8942 & 0.8908 \\
+poecilia\_formosa & 0.8969 & 0.9291 & 0.5553 & 0.8592 & 0.8762 & 0.8742 & 0.8445 \\
+poecilia\_latipinna & 0.8973 & 0.9281 & 0.4831 & 0.8562 & 0.879 & 0.8763 & 0.8486 \\
+poecilia\_mexicana & 0.8975 & 0.9298 & 0.5679 & 0.8584 & 0.8696 & 0.8683 & 0.8422 \\
+poecilia\_reticulata & 0.8654 & 0.8732 & 0.4516 & 0.8738 & 0.8772 & 0.8768 & 0.8591 \\
+pogona\_vitticeps & 0.8852 & 0.9268 & 0.2963 & 0.7753 & 0.7736 & 0.7737 & 0.7641 \\
+pongo\_abelii & 0.9393 & 0.9631 & 0.5004 & 0.8195 & 0.8536 & 0.8525 & 0.8472 \\
+procavia\_capensis & 0.8892 & 0.9093 & 0 & 0.7863 & 0.8673 & 0.8636 & 0.8619 \\
+prolemur\_simus & 0.9409 & 0.9534 & 0 & 0.885 & 0.9256 & 0.9239 & 0.9223 \\
+propithecus\_coquereli & 0.9408 & 0.9613 & 0.4612 & 0.854 & 0.8906 & 0.8892 & 0.8844 \\
+pteropus\_vampyrus & 0.8878 & 0.9056 & 0 & 0.8487 & 0.8679 & 0.8669 & 0.8647 \\
+pundamilia\_nyererei & 0.9026 & 0.9284 & 0.5454 & 0.8598 & 0.8918 & 0.8882 & 0.868 \\
+pygocentrus\_nattereri & 0.9061 & 0.9327 & 0.5784 & 0.8557 & 0.8799 & 0.878 & 0.8611 \\
+rattus\_norvegicus & 0.9436 & 0.9674 & 0.4773 & 0.8384 & 0.8336 & 0.8338 & 0.8247 \\
+rhinopithecus\_bieti & 0.9438 & 0.9645 & 0.4634 & 0.8464 & 0.8876 & 0.8861 & 0.8799 \\
+rhinopithecus\_roxellana & 0.9446 & 0.9647 & 0.5251 & 0.8489 & 0.8918 & 0.8901 & 0.8843 \\
+saimiri\_boliviensis\_boliviensis & 0.9412 & 0.9625 & 0.5019 & 0.8488 & 0.884 & 0.8828 & 0.8771 \\
+salvator\_merianae & 0.8947 & 0.9334 & 0.3899 & 0.7896 & 0.782 & 0.7824 & 0.774 \\
+sarcophilus\_harrisii & 0.9245 & 0.9523 & 0.4197 & 0.8504 & 0.8387 & 0.8392 & 0.8336 \\
+scleropages\_formosus & 0.9055 & 0.9342 & 0.5883 & 0.8616 & 0.8885 & 0.8855 & 0.8633 \\
+scophthalmus\_maximus & 0.8944 & 0.9223 & 0.6249 & 0.8864 & 0.8822 & 0.8828 & 0.8575 \\
+serinus\_canaria & 0.9198 & 0.9463 & 0.4193 & 0.8563 & 0.8723 & 0.8713 & 0.8605 \\
+seriola\_dumerili & 0.9002 & 0.9263 & 0.5769 & 0.8745 & 0.8923 & 0.8902 & 0.8646 \\
+seriola\_lalandi\_dorsalis & 0.903 & 0.9289 & 0.6162 & 0.8732 & 0.8925 & 0.89 & 0.8679 \\
+sorex\_araneus & 0.8512 & 0.8868 & 0 & 0.7659 & 0.7934 & 0.7923 & 0.7901 \\
+spermophilus\_dauricus & 0.9238 & 0.9418 & 0 & 0.8297 & 0.9005 & 0.8974 & 0.8951 \\
+sphenodon\_punctatus & 0.8151 & 0.884 & 0.2405 & 0.5038 & 0.5878 & 0.5838 & 0.58 \\
+stegastes\_partitus & 0.9075 & 0.9332 & 0.577 & 0.8697 & 0.8949 & 0.8922 & 0.87 \\
+sus\_scrofa & 0.9186 & 0.952 & 0.4099 & 0.731 & 0.8049 & 0.8018 & 0.7894 \\
+taeniopygia\_guttata & 0.8963 & 0.8963 & 0.2082 & 0.8642 & 0.9045 & 0.9017 & 0.8963 \\
+takifugu\_rubripes & 0.8759 & 0.9133 & 0.5619 & 0.8475 & 0.86 & 0.8575 & 0.8282 \\
+tetraodon\_nigroviridis & 0.8745 & 0.8944 & 0.1826 & 0.8695 & 0.8888 & 0.8838 & 0.8602 \\
+theropithecus\_gelada & 0.9554 & 0.9694 & 0.5335 & 0.8786 & 0.9283 & 0.9266 & 0.9233 \\
+\hline
+\end{tabular}}
+\end{table}
+
+\clearpage
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\renewcommand\thetable{S1}
+\centering
+\begin{tabular}{@{}llllllll@{}}
+\hline
+Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1 & Genic F1 \\ [0.5ex]
+\hline
+tupaia\_belangeri & 0.8092 & 0.8584 & 0 & 0.732 & 0.7162 & 0.7167 & 0.7147 \\
+tursiops\_truncatus & 0.8792 & 0.8984 & 0 & 0.8473 & 0.8556 & 0.8552 & 0.8535 \\
+urocitellus\_parryii & 0.9391 & 0.9593 & 0.4454 & 0.8788 & 0.8948 & 0.8941 & 0.888 \\
+ursus\_americanus & 0.9168 & 0.9455 & 0.3849 & 0.8449 & 0.8512 & 0.8508 & 0.8433 \\
+ursus\_maritimus & 0.9257 & 0.9528 & 0.3279 & 0.8313 & 0.8528 & 0.852 & 0.8453 \\
+vicugna\_pacos & 0.8516 & 0.886 & 0 & 0.7738 & 0.7945 & 0.7939 & 0.7923 \\
+vombatus\_ursinus & 0.9315 & 0.9572 & 0.4186 & 0.8413 & 0.8494 & 0.8491 & 0.8425 \\
+vulpes\_vulpes & 0.9349 & 0.9595 & 0.3522 & 0.8444 & 0.8616 & 0.8609 & 0.8541 \\
+xenopus\_tropicalis & 0.8699 & 0.9175 & 0.317 & 0.7277 & 0.7438 & 0.7427 & 0.7289 \\
+xiphophorus\_couchianus & 0.8575 & 0.5599 & 0 & 0.8585 & 0.9119 & 0.9052 & 0.8932 \\
+xiphophorus\_maculatus & 0.8854 & 0.9156 & 0.5964 & 0.8368 & 0.871 & 0.8671 & 0.8439 \\
+zonotrichia\_albicollis & 0.9216 & 0.9482 & 0.382 & 0.8601 & 0.8667 & 0.8663 & 0.857 \\
+\hline
+\end{tabular}}
+\end{table}
 

### 2020-04-24 paper/supplements.tex
removed comments
minor, images commented out again

@@
 \section{Model Evaluation by Species}
-The following tables and plots show the prediction performance of Hexlixer for each species we worked with.
+The following tables and plots show the prediction performance of Helixer for each species we worked with.
+The tables show a breakdown by basepair level accuracy, F1 score for each individual class and our two overall metrics, the Subgenic F1 and Genic F1.
 An F1 score of zero can indicate that the reference did not include that class.
-The plots display the Subgenic F1 as well as the difference to AUGUSTUS with respect to the phylogenetic placement by the NCBI taxonomy database \citep{federhen2012ncbi}.
+
+The plots display the raw Subgenic F1 score as well as the difference to AUGUSTUS in Subgenic F1 with respect to the phylogenetic placement by the NCBI taxonomy database \citep{federhen2012ncbi}.
 
@@
 
-\renewcommand\refname{Supplemental Reference}
+\renewcommand\refname{Supplemental References}
 \bibliographystyle{natbib}

### 2020-04-25 paper/supplements.tex
hyperparameters

@@
 \hline
-Layers & 4\\
-Units per layer & 256\\
-Learning rate & 1e-4\\
-Optimizer & Adam\\
+LSTM Layers & 4 \\
+LSTM Units per layer & 256 \\
+Learning rate & 1e-3 \\
+Batch Size & 52 \\
+Optimizer & Adam \\
+Clipnorm & 1.0 \\
+Loss & Categorical Crossentropy \\
+Float Precision & 32 bit \\
 \hline
 \end{tabular}
-\caption{All hyperparameters used for the }
+\caption{The hyperparameters that were used for the training of both the animal and plant models. For more implementation details please consult the source code.}
 \label{suptab:params}
@@
 \end{tabular}
-\caption{The class weights are the only hyperparameters that are different for animals and plants.
+\caption{The class weights used during training.
+We found this parameter to be highly impactful for establishing a more balanced performance across the different classes.
+These are the only hyperparameters that are different for animals and plants.
 The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}

### 2020-04-25 paper/main.tex
stuff
minor
results text

@@
 
-\title[short Title]{Helixer: Cross-species gene annotation of large eucaryotic genomes using deep learning}
+\title[short Title]{Helixer: Cross-species Gene Annotation Of Large Eucaryotic Genomes Using Deep Learning}
 \author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
@@
 
+Our models, however, tend to perform less well for the very smallest and largest genomes or species that are phylogenetically the furthest away from our training genomes.
+This is the case for both animals and plants and is visualized by the Supplemental Figures S?-?.
+We do not, for example, consistently predict very well on the algae as well as the non-avian reptiles.
+While one of the the algaes were included in our training genomes, the non-avian reptiles were not represented.
+
+AUGUSTUS does outperform us rather consistently on the smallest genomes, but struggles much more the larger the genomes get.
+The difference in prediction quality is especially strong for mammals and the largest plants.
+
 Two techniques were used during inference to improve performance and limit model bias.

### 2020-04-25 paper/supplements.tex
results text

@@
 
-\title{Supplemental Material for \\ Helixer: Cross-species Gene Annotation of Large Eucaryotic Genomes Using Deep Learning}
+\title{Supplemental Materials for \\ Helixer: Cross-species Gene Annotation of Large Eucaryotic Genomes Using Deep Learning}
 
@@
 Mammalia & 200.000 \\
-Reptilia (excluding Aves) & 200.000 \\
+Non-avian Reptilia & 200.000 \\
 Aves & 100.000 \\

### 2020-04-25 paper/main.tex
some results text

@@
 We do not, for example, consistently predict very well on the algae as well as the non-avian reptiles.
-While one of the the algaes were included in our training genomes, the non-avian reptiles were not represented.
+While one of the the algae were included in our training genomes, the non-avian reptiles were not.
+Our model also tends to show higher uncertainty around transitions of annotation classes (e.g.
+from UTR to CDS) due to the lack of further post-processing of the basepair wise predictions.
 
 AUGUSTUS does outperform us rather consistently on the smallest genomes, but struggles much more the larger the genomes get.
-The difference in prediction quality is especially strong for mammals and the largest plants.
+The difference in prediction quality is especially strong for mammals, which tend to have large genomes with very long genes, and the largest plants.
 
@@
 Our model does this while outperforming AUGUSTUS on a base pair wise level and withstanding RNASeq data based scrutiny.
+
+Different set of training genomes could lead to a shift in strengths and weaknesses of the model while a larger and more spread out set of high quality genomes could result in a wider range of genomes with decent predictions.
 

### 2020-04-25 paper/main.tex
discussion, citation

@@
 Indeed, in practice HMMs have problems with generalizing across species and it seems actual learning of underlying patterns is limited.
-In the last decade, deep neural networks have been applied with tremendous success in many areas of statistical modelling.
+In the last decade, deep neural networks have been applied with tremendous success in many areas of statistical modelling, including biology \citep{ching2018opportunities}.
 For sequence data, such as DNA, recurrent neural networks (RNN) in the form of long-short term memory (LSTM) are an established architecture and here, we employ them to improve on HMMs for ab-initio genome annotation.
@@
 
-AUGUSTUS does outperform us rather consistently on the smallest genomes, but struggles much more the larger the genomes get.
+AUGUSTUS does outperform us rather consistently on the smallest genomes, but falls off much more with growing genome length.
 The difference in prediction quality is especially strong for mammals, which tend to have large genomes with very long genes, and the largest plants.
@@
 
-Different set of training genomes could lead to a shift in strengths and weaknesses of the model while a larger and more spread out set of high quality genomes could result in a wider range of genomes with decent predictions.
+We found our models to be highly sensitive to the training genomes we chose.
+A different set could lead to a significant shift in strengths and weaknesses of the model and a larger and more spread out set of high quality genomes could also result in a wider range of genomes with decent predictions.
+For this, more computational resources would be required.
+
+An avenue for future research could be the addition of RNASeq data as additional input.
+This would bring Helixer on even footing with current tools and could lead to real world applicable performance improvements if the results of this work are any indication, as deep learning has been shown to excel in a multimodal settings \citep{ching2018opportunities}.
 

### 2020-04-25 paper/main.tex
some citations
changes introduction

@@
 \textbf{Results:} We set up a framework that can be used to build much more general gene callers that also outperform traditional methods on a variety of complex genomes.\\
-\textbf{Availability:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text Text  Text\\
+\textbf{Availability:} \\
 \textbf{Contact:} \href{name@bio.com}{name@bio.com}\\
@@
 \section{Introduction}
-For many tasks, it is necessary to have DNA regions labeled according to their function.
-This process is called gene annotation and it can be performed in different levels of precision, from simple coding -- non-coding classification to detailed structural labeling, such as enhancer regions or long non-coding RNA sequences in eukaryotes.
-In this work on eucaryotic genomes, we differentiate between Intergenic regions, Untranslated regions (UTRS, parts at the beginning and end of gene, that are transcribed but not translated), Coding sequences (CDS) and Introns.
+The annotation of genes in DNA is fundamental for many other tasks depending on it to be done as good and correct as possible.
+It can be performed at different levels of precision, from simple coding -- non-coding classification to detailed structural labeling.
 Because of the sheer size of genomes alone, gene annotation is a very labour intensive process and genomic database providers like NCBI \citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl} offer gene annotation pipelines.
@@
 The latter is an attractive approach, because it is cheap and fast.
-State of the art performance ab-initio gene prediction is achieved by higher order hidden markov models (HMMs), such as Genscan \citep{burge1997prediction} or Augustus \citep{stanke2003gene} but their accuracy leaves room for improvement.
-By choosing possible states and transitions, designers of HMMs assume structure in the sequence that may limit its predictive power.
-Indeed, in practice HMMs have problems with generalizing across species and it seems actual learning of underlying patterns is limited.
-In the last decade, deep neural networks have been applied with tremendous success in many areas of statistical modelling, including biology \citep{ching2018opportunities}.
-For sequence data, such as DNA, recurrent neural networks (RNN) in the form of long-short term memory (LSTM) are an established architecture and here, we employ them to improve on HMMs for ab-initio genome annotation.
+State of the art performance is achieved by higher order hidden markov models (HMMs), such as Genscan \citep{burge1997prediction} or AUGUSTUS \citep{stanke2003gene} but their accuracy leaves room for improvement.
+By encoding possible states and transitions in a probalistic model, designers of HMMs assume structure in the sequence that may limit its predictive power.
+Indeed, in practice HMMs have problems with generalizing across species and it seems actual learning of underlying patterns is limited despite the fundamental ability to do so being present.
 
-RNNs have already shown promising results in related, but generally more limited investigations.
-\citep{choudharypredicting} carries out preliminary explorations on the potential of BLSTMs for cross species gene prediction, but trains his model on human genes only and tests it on two more species.
+In the last decade, deep neural networks (DNN) have been applied with great success in many areas of statistical modelling, including biology \citep{ching2018opportunities}.
+For sequence data, such as DNA, speech or text, a special kind of recurrent neural network (RNN) called long-short term memory (LSTM) \citep{hochreiter1997long} is an established building block for many different architectures.
+LSTM units can also be used to process sequential input starting from both ends, forming a bi-directional LSTM (BLSTM).
+It has also been shown that HMMs can be successfully combined with DNNs \citep{liu2016novo, liu2016pedla}.
+
+For the purpose of gene annotation, RNNs have already shown promising results.
+\citep{choudharypredicting} carries out preliminary explorations on the potential of BLSTMs for cross species gene prediction and trains his model on human genes to test it later on two more species.
 DeepAnnotator \citep{amin2018deepannotator} uses bidirectional LSTMs (BLSTM) for gene finding in prokaryotes.
@@
 
-\section{Approach}
-Helixer is a prototype software for training and utilizing a general purpose model for the functional cross-species annotation of large eukaryotic genomes using only DNA data as input.
+In this work we present Helixer, a novel prototype software for training and utilizing a general purpose DNN for the ab-initio cross-species gene annotation of large eukaryotic genomes using only DNA data as input.
+Our model is trained to differentiate between 4 regions: Intergenic, Untranslated (UTR), Coding (CDS) and Intron.
 We demonstrate the effectiveness of this approach by training a single model for the annotation of a large set of genomes from the kingdoms metazoa and viridiplantae, respectively, which we will call {\it animals} and {\it plants} from now on.
-In total we used 192 animal genomes and 60 plant genomes.
-In contrast to traditional approaches, we employ modelling techniques from the field of deep learning to improve both the annotation accuracy and the ability to generalize across genomes.
-The source code and all data is publically available.
+We worked with the full data of 192 animal genomes and 60 plant genomes.
+Both the ability to generalize across genomes as well as the scope of the evaluation greatly surpasses all other work in this field.
+The source code and all input data is publically available.
 
@@
 \section{Methods}
-We worked with 192 animal and 60 plant genomes.
+The foundation of our work are 192 animal and 60 plant genomes.
 The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format.
@@
 
-This work has been supported by the... Text Text  Text Text.\vspace*{-12pt}
+This work has been supported by the... \vspace*{-12pt}
 

### 2020-04-27 paper/main.tex
supplemental references

@@
 Table \ref{tab:statistics_all} shows statistics about the masking.
-The corresponding statistics for the selected training genomes only can be found in Section \ref{sec:training_data}.
+The corresponding statistics for the selected training genomes only can be found in Figure S6.
 
@@
 Both metrics do not take the values of the intergenic class into account, as this class is very abundant and appears to be by far the easiest to predict.
-(See Table \ref{tab:statistics_all} for the class distribution and Table \ref{tab:detailed_results} for a more in-depth report our model performance including the intergenic class).
+(See Table \ref{tab:statistics_all} for the class distribution and Tables S1 and S2 for a more in-depth report our model performance including the intergenic class).
 In the case of the Genic F1, this means that effectively only the TP of the intergenic class are disregarded.
@@
 We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
-Details of the neural architecture search are given in Section \ref{sec:dcnn}.
+Details of the neural architecture search are given in Section 6 of the Supplements.
 
@@
 The concrete lengths were chosen to keep the typical average gene length roughly proportional to the length of the sequence input.
-For more implementation details on this see Section \ref{sec:longer} or the source code.
+For more implementation details on this see Section 3 of the Supplements or the source code.
 
@@
 The different overlapping sequences were recombined by averaging the individual softmax values of each base.
-The figures in Section \ref{sec:overlapping} show the effect of overlapping for each genome, ordered by N75.
+The figures in Section 4 of the Supplements show the effect of overlapping for each genome, ordered by N75.
 We found that overlapping tends to work best if the genomes are not very fragmented and we used it for both animals and plants.
@@
 Our models, however, tend to perform less well for the very smallest and largest genomes or species that are phylogenetically the furthest away from our training genomes.
-This is the case for both animals and plants and is visualized by the Supplemental Figures S?-?.
+This is the case for both animals and plants and is visualized by the Figures S1 and S3.
 We do not, for example, consistently predict very well on the algae as well as the non-avian reptiles.
@@
 
-AUGUSTUS does outperform us rather consistently on the smallest genomes, but falls off much more with growing genome length.
-The difference in prediction quality is especially strong for mammals, which tend to have large genomes with very long genes, and the largest plants.
+AUGUSTUS does outperform us rather consistently on the smallest genomes, but falls off much more the larger the genomes get.
+The difference in prediction quality is especially strong for mammals, which tend to have big genomes with very long genes as well as the largest plants.
+Figures S2 and S4 display the comparison to AUGUSTUS by phylogenetic position.
 

### 2020-04-27 paper/main.tex
some reordering, more headlines

@@
 \section{Methods}
+\subsection{Datasets}
 The foundation of our work are 192 animal and 60 plant genomes.
@@
 Both data groups were used seperately during training as well as during evaluation.
+
 At the start of our work we split both genome groups into {\it training genomes} and {\it test genomes}.
-The test genome group is made up of 19 and 6 genomes, respectively, and was chosen to be of the best possible quality based on collected metadata while also representing a broad phylogenetic spread.
-We then experimented which genomes to use for during training from the pool of non-test genomes while evaluating the performance of models on the remaining ones.
-We call the genomes that were ultimatily not used during training the {\it validation genomes}.
-While we did split off a group of test genomes at the beginning of our work, we ended up only reporting on the generalization performance of the combined set of validation and test genomes as the validation genomes are by far the largest set.
+The test genome groups are made up of 19 and 6 genomes, respectively, and were chosen to be of the best possible quality based on collected metadata while also representing a broad phylogenetic spread.
+
+The selection of training genomes also requires balancing multiple conflicting tradeoffs.
+On the one hand, we do want to train with as much data as possible, but not all data has a good enough quality to enable a powerful generalization.
+It is also very desirable to have a broad spread phylogenetically as well as in genome size and average gene length as the model is increasingly unlikely to generalize well beyond the borders given by the training data.
+However, if the genomic patterns inside the data are too different from each other, it may be difficult for the model to generalize, too.
+It was also important for us to get experimental results within a couple of days as we tested many different data and model configurations, which limited the overall size of the training set and the model.
+
+We employed an iterative approach to effectively select a proper set of training genomes.
+We first started with 3-4 genomes that were almost certain to be of the highest quality.
+We then evaluated the performance of the resulting model on all genomes individually (including the training genomes) and looked for candidates to add to the training set or to remove from it.
+This process was repeated multiple times concurrently with the model search until we saw no more room for substantial improvement given our computational constraints.
+
+We call the genomes that were ultimatily not used during training from the original split the {\it validation genomes}.
+While we did split off a group of test genomes at the beginning of our work, we ended up only reporting on the generalization performance on the combined set of validation and test genomes as the validation genomes are by far the largest set.
 We also did not find a noticible difference in performance between those two groups.
 
-We pre-processed and stored the genomic information by using Geenuff \citep{denton2019}.
-Geenuff is a tool for checking and exploring genomic data, which stores all information inside a SQL database.
-Training and evaluation data was generated by querying this database and transforming the returned data into a numerical format suitable for machine learning.
+Once a set of training genomes was selected, we split it up again into a {\it training and validation set}.
+This split of the training genomes was done to get a quicker sense of the generalization capabilities and was used primarily during training.
+We split off the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level.
+Evaluation of the annotations on all training and validation genomes was done regularly after a promising model candidate was found based on its validation set performance.
+Ultimately we only cared about the cross-species performance on all validation genomes of a model.
+
+\subsection{Data Pre-processing}
+We first pre-process and store the raw genomic information by using GeenuFF \citep{denton2019}.
+GeenuFF is a tool for checking and exploring genomic data, that stores all information inside a SQL database.
+Training and evaluation-ready data is generated by querying this database and then transforming the returned data into a numerical format suitable for machine learning.
 The encoding of the genomic sequence is done in line with the IUPAC nucleic acid notation and the functional annotation used as labels during training is transformed to a one hot encoding with the four classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}.
-See also Table \ref{tab:arrays} for a more detailed description of the generated data formats.
-
-The classes are not balanced equally and the tradeoff between the number of intergenic and intronic bases is quite different in both groups of genomes.
-Table \ref{tab:statistics_all} shows the ratio of class occurences, together with other overall metrics.
+See also Table \ref{tab:arrays} for a more detailed description of the generated data types.
 
@@
 
-During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation.
-As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to blank out those bases during training by using the sample weights described in Table \ref{tab:arrays}.
-The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it.
+During data generation, we query for the longest transcript of each gene and disregard FASTA sequences that have no functional annotation.
+GeenuFF checks the genomic annotations for potential errors during import and is able to mark those areas.
+We use this information to effectively blank out those bases during training by using the sample weights described in Table \ref{tab:arrays}.
+The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and GeenuFF marks a potentially large intergenic region for it.
 Table \ref{tab:statistics_all} shows statistics about the masking.
-The corresponding statistics for the selected training genomes only can be found in Figure S6.
 
-For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network.
-We appended zero padding if the subsequences are not long enough.
+For the training itself, we divide each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network together with the sample weights.
+We append zero padding if the subsequences are not long enough.
 If a subsequence is fully marked as erroneous, we do not include it anywhere.
-
-The selection of training genomes requires balancing multiple conflicting tradeoffs.
-On the one hand, we do want to train with as much data as possible, but not all data has a good enough quality to enable a powerful generalization.
-It is also very desirable to have a broad spread in phylogenetic distance as well as genome size and average gene length as the model is increasingly unlikely to generalize well beyond the borders given by the training data.
-However, if the genomic patterns inside the data are too different from each other, it may be difficult for the model to generalize as well.
-It was also important to us to get experimental results within a couple of days as we tested many different data and model configurations, which limited the overall size of the training set and the model.
-
-We employed an iterative approach to effectively select a proper set of training genomes.
-We first started with 3-4 genomes that were almost certain to be of the highest quality.
-We then evaluated the performance of the resulting model on all genomes individually (including the training genomes) and looked for candidates to add to the training set or to remove from it.
-This process was repeated multiple times concurrently with the model search until we saw no more room for substantial improvement given our computational constraints.
-
-Once a set of training genomes was selected, we split it up again into a {\it training and validation set}.
-This split of the training genomes was done to get a quicker sense of the generalization capabilities and was used primarily during training.
-We split off the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level.
-Evaluation of the annotations on all training and validation genomes was done regularly after a promising model candidate was found based on its validation set performance.
-Ultimately we only cared about the cross-species performance on all validation genomes of a model.
 
@@
 
-\subsection{Model Architecture}
+\subsection{Model Architecture and Training}
 \label{sec:model}
@@
 We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks (CNN) and hybrid architectures.
+We also use class weigths as the classes of the input data are not balanced equally enough, at least not for our network, and does also vary greatly between animals and plants.
+For example, the tradeoff between the number of intergenic and intronic bases is quite different in both groups (see Table \ref{tab:statistics_all}).
+
 The hyperparameters were optimized by a combination of manual and automatic optimization.
@@
 We used NNI \citep{nni2019} to facilitate the search.
-All relevant hyperparameters can be found in Table \ref{tab:params} or in the Helixer source code repository.
+All relevant hyperparameters can be found in Table S3 and S4 or in the Helixer source code repository.
 

### 2020-04-29 plots/coverage/scores_histo_frm_h5.py
take one, break up score by category and right/wrong

@@ -1,0 +1,69 @@
+import numpy as np
+import pandas as pd
+import h5py
+import argparse
+
+
+# overall score
+def score(f, other_y, n=20, by=1000):
+    """get counts for histogram of reference scores broken up by other_y
+
+    other_y should be a list of keys for h5 datasets to be compared to data/y, for each key
+    this will record and return 8 histograms under the same key:
+       [ig right, ig wrong, utr right, utr wrong, cds right, cds wrong, intron right, intron wrong]"""
+    breaks = [x / n for x in range(n + 1)]
+    histos = {}
+    for key in other_y:
+        histos[key] = [np.zeros((n,)) for _ in range(8)]
+
+    for i in range(0, f['data/X'].shape[0], by):
+        y = np.argmax(f['data/y'][i:(i + by)], axis=2).ravel()
+        preds = {}
+        for key in other_y:
+            preds[key] = np.argmax(f[key][i:(i + by)], axis=2).ravel()
+        scores = f['scores/by_bp'][i:(i + by)].ravel()
+
+        for col in range(4):
+            mask = y == col
+            ysm = y[mask]
+            scoressm = scores[mask]
+            for key in other_y:
+                predsm = preds[key][mask]
+                maskright = (predsm == ysm).astype('bool')
+                righthisto = np.histogram(scoressm[maskright], bins=breaks)[0]
+                histos[key][col * 2] += righthisto
+                histos[key][col * 2 + 1] += np.histogram(scoressm[np.logical_not(maskright)], bins=breaks)[0]
+    index = []
+    for cat in ['ig', 'utr', 'cds', 'exon']:
+        index += [cat + '_agree', cat + '_disagree']
+
+    pd_histos = {}
+    for key in histos:
+        pd_histos[key] = pd.DataFrame(histos[key])
+        pd_histos[key].set_index(index)
+
+    return pd_histos
+
+
+def main(h5_data, alternatives, out_dir):
+    f = h5py.File(h5_data, mode='r')
+    other_y = ['predictions'] + ['alternative/' + x for x in alternatives.split(',')]
+    pd_histos = score(f, other_y)
+
+    for key in pd_histos:
+        fkey = key.replace('/', '_')
+        pd_histos[key].to_csv('{}/{}_scores.csv'.format(out_dir, fkey))
+
+    f.close()
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser()
+    parser.add_argument('-d', '--h5-data', required=True,
[…]

### 2020-04-29 paper/main.tex
some files moved, minor

@@
 \section{Discussion}
-We propose a gene calling model that is able to perform cross-species genomic annotiation for a wide variety of large eucaryotic genomes.
-Our model does this while outperforming AUGUSTUS on a base pair wise level and withstanding RNASeq data based scrutiny.
+We propose a novel way to develop gene calling models that are able to perform cross-species genomic annotiation for a wide variety of large eucaryotic genomes.
+Helixer does this while outperforming AUGUSTUS on a base pair wise level and withstanding RNASeq data based scrutiny.
 

### 2020-04-29 plots/coverage/scores_histo_frm_h5.py
and split histograms of scores up and running

@@ -5,7 +5,7 @@ import argparse
 
 
 # overall score
-def score(f, other_y, n=20, by=1000):
+def score(f, other_y, predictions, n=20, by=1000):
     """get counts for histogram of reference scores broken up by other_y
 
     other_y should be a list of keys for h5 datasets to be compared to data/y, for each key
@@ -20,7 +20,10 @@ def score(f, other_y, n=20, by=1000):
         y = np.argmax(f['data/y'][i:(i + by)], axis=2).ravel()
         preds = {}
         for key in other_y:
-            preds[key] = np.argmax(f[key][i:(i + by)], axis=2).ravel()
+            of = f
+            if key == 'predictions':
+              of = predictions
+            preds[key] = np.argmax(of[key][i:(i + by)], axis=2).ravel()
         scores = f['scores/by_bp'][i:(i + by)].ravel()
 
         for col in range(4):
@@ -40,15 +43,20 @@ def score(f, other_y, n=20, by=1000):
     pd_histos = {}
     for key in histos:
         pd_histos[key] = pd.DataFrame(histos[key])
-        pd_histos[key].set_index(index)
+        pd_histos[key].index = index
+        pd_histos[key].columns = breaks[:-1]
 
     return pd_histos
 
 
-def main(h5_data, alternatives, out_dir):
+def main(h5_data, alternatives, out_dir, predictions):
     f = h5py.File(h5_data, mode='r')
-    other_y = ['predictions'] + ['alternative/' + x for x in alternatives.split(',')]
-    pd_histos = score(f, other_y)
+    if predictions is None:
+        pred_f = f
+    else:
+        pred_f = h5py.File(predictions, mode='r')
+    other_y = ['predictions'] + ['alternative/{}/y'.format(x) for x in alternatives.split(',')]
+    pd_histos = score(f, other_y, pred_f)
 
     for key in pd_histos:
         fkey = key.replace('/', '_')
@@ -65,5 +73,6 @@ if __name__ == "__main__":
                         help='comma sep list of h5 groups where we will use alternative/{alternative}/y,'
                              ' e.g. "augustus"')
     parser.add_argument('-o', '--out-dir', default='./', help='directory where output will be written to')
+    parser.add_argument('-p', '--predictions', help='point to h5 with /predictions if not the same')
     args = parser.parse_args()
-    main(args.h5_data, args.alternative, args.out_dir)
+    main(args.h5_data, args.alternatives, args.out_dir, args.predictions)

### 2020-04-29 plots/grid_plot/plot.py
plotting, wip

@@ -1,0 +1,89 @@
+#!/usr/bin/env python3
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+
+# load and transform animal results
+df_aug = pd.read_csv('final animal eval - Augustus_all.csv')
+df_nn = pd.read_csv('final animal eval - ensemble.csv')
+df_meta = pd.read_csv('metadata_animals.csv')
+
+genomes = set(df_aug['genome']).intersection(set(df_nn['genome']))
+
+df_aug = df_aug[df_aug['genome'].isin(genomes)].sort_values('genome').reset_index(drop=True)
+df_nn = df_nn[df_nn['genome'].isin(genomes)].sort_values('genome').reset_index(drop=True)
+df_meta = df_meta[df_meta['species'].isin(genomes)].sort_values('species').reset_index(drop=True)
+
+df_aug = df_aug.drop(columns=['acc_overall', 'f1_ig', 'f1_utr', 'f1_exon', 'f1_intron', 'legacy_f1_cds','f1_genic'])
+
+# add nn values and metadata to aug df
+df_aug['NN'] = df_nn['sub_genic']
+df_aug['total_len'] = df_meta['total_len']
+df_animals = df_aug.rename(columns={'sub_genic': 'AUG', 'genome': 'species'})
+
+# remove training genomes
+df_animals = df_animals[~df_animals['species'].isin(['mus_musculus','oryzias_latipes','gallus_gallus',
+                                                     'theropithecus_gelada','anabas_testudineus',
+                                                     'drosophila_melanogaster'])].reset_index(drop=True)
+
+# load and transform plant results
+df_aug = pd.read_csv('plant_LSTM_eval - Augustus_all.csv')
+df_nn = pd.read_csv('final plant eval - ensemble.csv')
+df_meta = pd.read_csv('metadata_plants.csv')
+
+genomes = set(df_aug['genome']).intersection(set(df_nn['genome']))
+
+df_aug = df_aug[df_aug['genome'].isin(genomes)].sort_values('genome').reset_index(drop=True)
+df_nn = df_nn[df_nn['genome'].isin(genomes)].sort_values('genome').reset_index(drop=True)
[…]

### 2020-04-30 plots/grid_plot/plot.py
grip plot boxplots

@@ -56,34 +56,34 @@ fig = plt.figure(figsize=(12, 8))
 outer = gridspec.GridSpec(1, 3, wspace=0.1, hspace=0.1)
 
 # boxplots
-boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 1,
+boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2,
                 subplot_spec=outer[0], wspace=0.1, hspace=0.1)
 
-ax_animals_box = plt.Subplot(fig, boxplot_grid[0])
+# ax = plt.Subplot(fig, boxplot_grid[0, 0])
+# ax = sns.boxplot(y=df_animals["AUG"], medianprops={'color':'red'},
+            # ax=ax, color="white", zorder=100)
+# for patch in ax.artists:
+    # r, g, b, a = patch.get_facecolor()
+    # patch.set_facecolor((r, g, b, 0.0))
 
-_, axes = plt.subplots(1, 2, sharey=True, figsize=(10,10))
-sns.boxplot(y=df_animals["AUG"], medianprops={'color':'red'},
-            ax=axes[0], color="white")
-sns.swarmplot(y="AUG", data=df_animals, ax=axes[0], color="chocolate")
-axes[0].set_title('AUGUSTUS')
-axes[0].set_ylabel("Subgenic F1 Score")
+def add_subplot(subplot_spec, data, swarm_color, title=''):
+    ax = plt.Subplot(fig, subplot_spec)
+    ax = sns.boxplot(y=data, medianprops={'color':'red'}, ax=ax, color='white',
+                     showfliers=False, showcaps=False, whiskerprops={'linewidth':0})
+    plt.setp(ax.lines, zorder=100)
+    ax = sns.swarmplot(y=data, ax=ax, color=swarm_color)
+    ax.set_ylim((0.0, 1.0))
+    ax.set_ylabel(None)
+    ax.set_xticks([], [])
+    if title:
+        ax.set_title(title)
+    fig.add_subplot(ax)
 
-sns.boxplot(y=df_animals["NN"], medianprops={'color':'red'},
-            ax=axes[1], color="white")
-sns.swarmplot(y="NN", data=df_animals, ax=axes[1], color="royalblue")
-axes[1].set_title('Helixer')
+add_subplot(boxplot_grid[0, 0], df_animals['AUG'], 'chocolate', 'AUGUSTUS')
[…]

### 2020-04-30 plots/grid_plot/plot.py
plotting

@@ -20,7 +20,7 @@ df_aug = df_aug.drop(columns=['acc_overall', 'f1_ig', 'f1_utr', 'f1_exon', 'f1_i
 
 # add nn values and metadata to aug df
 df_aug['NN'] = df_nn['sub_genic']
-df_aug['total_len'] = df_meta['total_len']
+df_aug['total_len'] = df_meta['total_len'] / 1e9
 df_animals = df_aug.rename(columns={'sub_genic': 'AUG', 'genome': 'species'})
 
 # remove training genomes
@@ -42,7 +42,7 @@ df_meta = df_meta[df_meta['species'].isin(genomes)].sort_values('species').reset
 df_aug = df_aug.drop(columns=['acc_overall', 'f1_ig', 'f1_utr', 'f1_exon', 'f1_intron', 'legacy_f1_cds', 'f1_genic'])
 
 df_aug['NN'] = df_nn['sub_genic']
-df_aug['total_len'] = df_meta['total_len']
+df_aug['total_len'] = df_meta['total_len'] / 1e9
 df_plants = df_aug.rename(columns={'sub_genic': 'AUG', 'genome': 'species'})
 df_plants = df_plants[~df_plants['species'].isin(['Athaliana','Bdistachyon','Creinhardtii','Gmax','Mguttatus','Mpolymorpha',
                                                   'Ptrichocarpa','Sitalica','Zmays'])].reset_index(drop=True)
@@ -51,13 +51,16 @@ df_plants = df_plants[~df_plants['species'].isin(['Athaliana','Bdistachyon','Cre
 df_ol_animals = pd.read_csv('animals_overlapping_aggregate.csv')
 df_ol_plants = pd.read_csv('plants_overlapping_aggregate.csv')
 
+# colors
+aug_color = 'chocolate'
+hel_color = 'royalblue'
+
 # make subplot grid
 fig = plt.figure(figsize=(12, 8))
 outer = gridspec.GridSpec(1, 3, wspace=0.1, hspace=0.1)
 
 # boxplots
-boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2,
-                subplot_spec=outer[0], wspace=0.1, hspace=0.1)
+boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=outer[0], wspace=0.1, hspace=0.1)
 
 # ax = plt.Subplot(fig, boxplot_grid[0, 0])
[…]

### 2020-04-30 plots/grid_plot/plot.py
plotting

@@ -102,6 +102,9 @@ def add_scatterplot(subplot_spec, data, xlabel=''):
     ax.set_ylabel(None)
     if xlabel:
         ax.set_xlabel(xlabel)
+    else:
+        ax.set_xticklabels([])
+        ax.set_xlabel(None)
     fig.add_subplot(ax)
 
 add_scatterplot(scatterplot_grid[0], df_animals)

### 2020-04-30 plots/grid_plot/plot.py
plotting

@@ -56,7 +56,7 @@ aug_color = 'chocolate'
 hel_color = 'royalblue'
 
 # make subplot grid
-fig = plt.figure(figsize=(12, 8))
+fig = plt.figure(figsize=(12, 6))
 outer = gridspec.GridSpec(1, 3, wspace=0.1, hspace=0.1)
 
 # boxplots
@@ -109,6 +109,33 @@ def add_scatterplot(subplot_spec, data, xlabel=''):
 
 add_scatterplot(scatterplot_grid[0], df_animals)
 add_scatterplot(scatterplot_grid[1], df_plants, 'Genome Size in Gbp')
+
+# line plots
+lineplot_grid = gridspec.GridSpecFromSubplotSpec(2, 1, subplot_spec=outer[2], wspace=0.1, hspace=0.1)
+
+def add_lineplot(subplot_spec, data, xlabel=''):
+    ax = plt.Subplot(fig, subplot_spec)
+    ax.plot(range(100), data['f1'], color='tab:red', linestyle='dashed')
+    ax.plot(range(100), data['f1_with_overlapping'], color='tab:red')
+    ax.plot(range(100), data['acc'], color='tab:purple', linestyle='dashed')
+    ax.plot(range(100), data['acc_with_overlapping'], color='tab:purple')
+    ax.set_ylim((0.0, 1.0))
+
+    ax.set_yticklabels([])
+    ax.set_ylabel(None)
+    ticks = [0, 50, 100]
+    ax.set_xticks(ticks)
+    if xlabel:
+        ax.set_xlabel(xlabel)
+        ax.set_xticklabels([t * 200 for t in ticks])
+    else:
+        ax.set_xlabel(None)
+        ax.set_xticklabels([])
+    fig.add_subplot(ax)
+
+add_lineplot(lineplot_grid[0], df_ol_animals)
+add_lineplot(lineplot_grid[1], df_ol_plants, xlabel='Basepair Offset in Sequence')
+
 
 plt.show()
 

### 2020-04-30 plots/grid_plot/plot.py
plotting

@@ -69,7 +69,7 @@ boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=outer[0], wsp
     # r, g, b, a = patch.get_facecolor()
     # patch.set_facecolor((r, g, b, 0.0))
 
-def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False):
+def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False, superscript_label=False):
     ax = plt.Subplot(fig, subplot_spec)
     ax = sns.boxplot(y=data, medianprops={'color':'red'}, ax=ax, color='white',
                      showfliers=False, showcaps=False, whiskerprops={'linewidth':0})
@@ -82,9 +82,11 @@ def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False):
         ax.set_yticklabels([])
     if title:
         ax.set_title(title)
+    if superscript_label:
+        ax.text(-0.2, 1.1, 'a)', transform=ax.transAxes, size=10, weight='bold')
     fig.add_subplot(ax)
 
-add_boxplot(boxplot_grid[0, 0], df_animals['AUG'], aug_color, tick_labels=True)
+add_boxplot(boxplot_grid[0, 0], df_animals['AUG'], aug_color, tick_labels=True, superscript_label=True)
 add_boxplot(boxplot_grid[0, 1], df_animals['NN'], hel_color)
 add_boxplot(boxplot_grid[1, 0], df_plants['AUG'], aug_color, tick_labels=True)
 add_boxplot(boxplot_grid[1, 1], df_plants['NN'], hel_color)
@@ -136,6 +138,7 @@ def add_lineplot(subplot_spec, data, xlabel=''):
 add_lineplot(lineplot_grid[0], df_ol_animals)
 add_lineplot(lineplot_grid[1], df_ol_plants, xlabel='Basepair Offset in Sequence')
 
+# add superscript labeling
 
 plt.show()
 

### 2020-04-30 plots/grid_plot/plot.py
plots

@@ -57,7 +57,7 @@ hel_color = 'royalblue'
 
 # make subplot grid
 fig = plt.figure(figsize=(12, 6))
-outer = gridspec.GridSpec(1, 3, wspace=0.1, hspace=0.1)
+outer = gridspec.GridSpec(1, 3, wspace=0.2, hspace=0.1)
 
 # boxplots
 boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=outer[0], wspace=0.1, hspace=0.1)
@@ -83,7 +83,7 @@ def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False, su
     if title:
         ax.set_title(title)
     if superscript_label:
-        ax.text(-0.2, 1.1, 'a)', transform=ax.transAxes, size=10, weight='bold')
+        ax.text(-0.45, 0.95, 'a)', transform=ax.transAxes, size=15, weight='bold')
     fig.add_subplot(ax)
 
 add_boxplot(boxplot_grid[0, 0], df_animals['AUG'], aug_color, tick_labels=True, superscript_label=True)
@@ -94,7 +94,7 @@ add_boxplot(boxplot_grid[1, 1], df_plants['NN'], hel_color)
 # scatterplots
 scatterplot_grid = gridspec.GridSpecFromSubplotSpec(2, 1, subplot_spec=outer[1], wspace=0.1, hspace=0.1)
 
-def add_scatterplot(subplot_spec, data, xlabel=''):
+def add_scatterplot(subplot_spec, data, xlabel='', superscript_label=False):
     ax = plt.Subplot(fig, subplot_spec)
     ax = sns.regplot(x=data['total_len'], y=data['AUG'], ax=ax, color=aug_color, scatter_kws={'s':8}, line_kws={'linewidth':1.5})
     ax = sns.regplot(x=data['total_len'], y=data['NN'], ax=ax, color=hel_color, scatter_kws={'s':8}, line_kws={'linewidth':1.5})
@@ -107,15 +107,17 @@ def add_scatterplot(subplot_spec, data, xlabel=''):
     else:
         ax.set_xticklabels([])
         ax.set_xlabel(None)
+    if superscript_label:
+        ax.text(-0.14, 0.95, 'b)', transform=ax.transAxes, size=15, weight='bold')
     fig.add_subplot(ax)
 
-add_scatterplot(scatterplot_grid[0], df_animals)
[…]

### 2020-05-01 plots/coverage/scores_histo_frm_h5.py
increase histo detail

@@ -2,48 +2,57 @@ import numpy as np
 import pandas as pd
 import h5py
 import argparse
+import itertools
+import copy
 
 
 # overall score
-def score(f, other_y, predictions, n=20, by=1000):
+def score(f, other_y, predictions, n=1000, by=2000):
     """get counts for histogram of reference scores broken up by other_y
 
     other_y should be a list of keys for h5 datasets to be compared to data/y, for each key
     this will record and return 8 histograms under the same key:
        [ig right, ig wrong, utr right, utr wrong, cds right, cds wrong, intron right, intron wrong]"""
     breaks = [x / n for x in range(n + 1)]
+
+    i_s = list(range(4))
+    i_s_by_y = [i_s for _ in range(len(other_y))]
+    argmax_sets = list(itertools.product(*i_s_by_y))
+
     histos = {}
-    for key in other_y:
-        histos[key] = [np.zeros((n,)) for _ in range(8)]
+    score_cats = ['ig', 'exon', 'intron']
+    for key in score_cats:
+        histos[key] = [np.zeros((n,)) for _ in range(len(argmax_sets))]
 
     for i in range(0, f['data/X'].shape[0], by):
-        y = np.argmax(f['data/y'][i:(i + by)], axis=2).ravel()
+        un_padded = np.sum(f['data/X'][i:(i + by)], axis=2).ravel().astype(bool)
         preds = {}
         for key in other_y:
             of = f
             if key == 'predictions':
               of = predictions
-            preds[key] = np.argmax(of[key][i:(i + by)], axis=2).ravel()
-        scores = f['scores/by_bp'][i:(i + by)].ravel()
+            preds[key] = np.argmax(of[key][i:(i + by)], axis=2).ravel()[un_padded]
+        scores = f['scores/ig_exon_intron_bp'][i:(i + by)].ravel()
 
-        for col in range(4):
-            mask = y == col
-            ysm = y[mask]
-            scoressm = scores[mask]
-            for key in other_y:
-                predsm = preds[key][mask]
-                maskright = (predsm == ysm).astype('bool')
-                righthisto = np.histogram(scoressm[maskright], bins=breaks)[0]
-                histos[key][col * 2] += righthisto
-                histos[key][col * 2 + 1] += np.histogram(scoressm[np.logical_not(maskright)], bins=breaks)[0]
-    index = []
-    for cat in ['ig', 'utr', 'cds', 'exon']:
[…]

### 2020-05-01 plots/coverage/scores_histo_frm_h5.py
ahh... actually running detailed score histos

@@ -24,7 +24,9 @@ def score(f, other_y, predictions, n=1000, by=2000):
     for key in score_cats:
         histos[key] = [np.zeros((n,)) for _ in range(len(argmax_sets))]
 
-    for i in range(0, f['data/X'].shape[0], by):
+    xshape = f['data/X'].shape[0]
+    for i in range(0, xshape, by):
+        print('at {} of {}'.format(i, xshape))
         un_padded = np.sum(f['data/X'][i:(i + by)], axis=2).ravel().astype(bool)
         preds = {}
         for key in other_y:
@@ -32,8 +34,7 @@ def score(f, other_y, predictions, n=1000, by=2000):
             if key == 'predictions':
               of = predictions
             preds[key] = np.argmax(of[key][i:(i + by)], axis=2).ravel()[un_padded]
-        scores = f['scores/ig_exon_intron_bp'][i:(i + by)].ravel()
-
+        scores = f['scores/ig_exon_intron_bp'][i:(i + by)].reshape([-1, 3])[un_padded]
         for ih, argmaxes in enumerate(argmax_sets):
             subpreds = copy.deepcopy(preds)
             scoressm = copy.deepcopy(scores)
@@ -43,7 +44,7 @@ def score(f, other_y, predictions, n=1000, by=2000):
                 mask = subpreds[other_y[iam]] == am
                 for key in subpreds:
                     subpreds[key] = subpreds[key][mask]
-                    scoressm = scoressm[mask]
+                scoressm = scoressm[mask, :]
 
             for isc, sc_key in enumerate(score_cats):
                 matchhisto = np.histogram(scoressm[:, isc], bins=breaks)[0]

### 2020-05-01 plots/grid_plot/plot.py
some plotting

@@ -57,17 +57,13 @@ hel_color = 'royalblue'
 
 # make subplot grid
 fig = plt.figure(figsize=(12, 6))
-outer = gridspec.GridSpec(1, 3, wspace=0.2, hspace=0.1)
+# nest gridspecs to have different spacing between column 2 and 3
+outer_main = gridspec.GridSpec(1, 2, width_ratios=[2, 1], wspace=0.19)
+outer_left = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=outer_main[0], wspace=0.2)
+outer_right = gridspec.GridSpecFromSubplotSpec(1, 1, subplot_spec=outer_main[1])
 
 # boxplots
-boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=outer[0], wspace=0.1, hspace=0.1)
-
-# ax = plt.Subplot(fig, boxplot_grid[0, 0])
-# ax = sns.boxplot(y=df_animals["AUG"], medianprops={'color':'red'},
-            # ax=ax, color="white", zorder=100)
-# for patch in ax.artists:
-    # r, g, b, a = patch.get_facecolor()
-    # patch.set_facecolor((r, g, b, 0.0))
+boxplot_grid = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=outer_left[0], wspace=0.1, hspace=0.1)
 
 def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False, superscript_label=False):
     ax = plt.Subplot(fig, subplot_spec)
@@ -76,14 +72,16 @@ def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False, su
     plt.setp(ax.lines, zorder=100)
     ax = sns.swarmplot(y=data, ax=ax, color=swarm_color, size=4)
     ax.set_ylim((0.0, 1.0))
-    ax.set_ylabel(None)
     ax.set_xticks([], [])
     if not tick_labels:
         ax.set_yticklabels([])
+        ax.set_ylabel(None)
+    else:
+        ax.set_ylabel('Subgenic F1')
     if title:
         ax.set_title(title)
     if superscript_label:
-        ax.text(-0.45, 0.95, 'a)', transform=ax.transAxes, size=15, weight='bold')
+        ax.text(-0.48, 0.95, 'a)', transform=ax.transAxes, size=15, weight='bold')
     fig.add_subplot(ax)
 
[…]

### 2020-05-01 paper/main.tex
main results plot added to latex

@@
 The latter is an attractive approach, because it is cheap and fast.
-State of the art performance is achieved by higher order hidden markov models (HMMs), such as Genscan \citep{burge1997prediction} or AUGUSTUS \citep{stanke2003gene} but their accuracy leaves room for improvement.
+State of the art performance is achieved by higher order hidden markov models (HMMs), such as Genscan \citep{burge1997prediction}, AUGUSTUS \citep{stanke2003gene} or SNAP \citep{johnson2008snap}.
+Their accuracy, however, leaves room for improvement.
 By encoding possible states and transitions in a probalistic model, designers of HMMs assume structure in the sequence that may limit its predictive power.
-Indeed, in practice HMMs have problems with generalizing across species and it seems actual learning of underlying patterns is limited despite the fundamental ability to do so being present.
+In practice HMMs have trouble to generalize across species and it seems actual learning of underlying patterns is limited despite the fundamental ability to do so being present.
 
@@
 
-\begin{figure}[!tpb]
-\label{fig:length_wise_bias}
-\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
-\caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
+%\begin{figure}[!tpb]
+%\label{fig:length_wise_bias}
+%\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
+%\caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
 Each value is the average performance of a 200 basepair long chunk by one of the final ensemble models.
@@
 See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
-\end{figure}
-
-\begin{figure}[!tpb]
-\centerline{\includegraphics[width=0.45\textwidth]{images/overall_boxplot}}
-\caption{Boxplots comparing the Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
+%\end{figure}
+%
+%\begin{figure}[!tpb]
+%\centerline{\includegraphics[width=0.45\textwidth]{images/overall_boxplot}}
+%\caption{Boxplots comparing the Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
 Results are given for Helixer and AUGUSTUS with the median predictions scores being denoted by the red lines.
 Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}.}\label{fig:boxplot}
-\end{figure}
+%\end{figure}
+
+\begin{figure*}[!tpb]
+\label{fig:results}
+\centerline{\includegraphics[width=\textwidth]{images/main_results}}
+\caption{Main overall results.
+{\bf a)} Boxplots comparing the Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
+Results are given for Helixer and AUGUSTUS with the median predictions scores being denoted by the red lines.
+Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}.
+{\bf c)} Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
+Each value is the average performance of a 200 basepair long chunk by one of the final ensemble models.
+The regular metrics show a significant negative bias towards the borders of each sequence.
+Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average.
+See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
+\end{figure*}
 

### 2020-05-01 plots/grid_plot/plot.py

@@ -81,7 +81,7 @@ def add_boxplot(subplot_spec, data, swarm_color, title='', tick_labels=False, su
     if title:
         ax.set_title(title)
     if superscript_label:
-        ax.text(-0.48, 0.95, 'a)', transform=ax.transAxes, size=15, weight='bold')
+        ax.text(-0.4, 0.95, 'a)', transform=ax.transAxes, size=15, weight='bold')
     fig.add_subplot(ax)
 
 add_boxplot(boxplot_grid[0, 0], df_animals['AUG'], aug_color, tick_labels=True, superscript_label=True)
@@ -107,7 +107,8 @@ def add_scatterplot(subplot_spec, data, xlabel='', superscript_label=False):
     else:
         ax.set_xticklabels([])
         ax.set_xlabel(None)
-        ax.legend(loc='lower left', fontsize='small')
+        # ax.legend(loc='lower left', fontsize='small')
+        ax.legend(loc='lower left')
     if superscript_label:
         ax.text(-0.14, 0.95, 'b)', transform=ax.transAxes, size=15, weight='bold')
     fig.add_subplot(ax)
@@ -135,12 +136,14 @@ def add_lineplot(subplot_spec, data, xlabel='', superscript_label=False):
     else:
         ax.set_xlabel(None)
         ax.set_xticklabels([])
-        ax.legend(loc='lower left', fontsize='small')
+        ax.legend(loc='lower left')
     if superscript_label:
-        ax.text(-0.21, 0.95, 'c)', transform=ax.transAxes, size=15, weight='bold')
+        ax.text(-0.17, 0.95, 'c)', transform=ax.transAxes, size=15, weight='bold')
     fig.add_subplot(ax)
 
 add_lineplot(lineplot_grid[0], df_ol_animals, superscript_label=True)
 add_lineplot(lineplot_grid[1], df_ol_plants, xlabel='Basepair Position in Input Sequence')
 
-plt.show()
+outer_main.tight_layout(fig)
+fig.savefig('main_results.png', dpi=350)
+# plt.show()

### 2020-05-02 paper/main.tex
minor
removed stddev from overall result table

@@
 
-%\begin{figure}[!tpb]
-%\label{fig:length_wise_bias}
-%\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
-%\caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
-Each value is the average performance of a 200 basepair long chunk by one of the final ensemble models.
-The regular metrics show a significant negative bias towards the borders of each sequence.
-Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average.
-See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
-%\end{figure}
-%
-%\begin{figure}[!tpb]
-%\centerline{\includegraphics[width=0.45\textwidth]{images/overall_boxplot}}
-%\caption{Boxplots comparing the Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
-Results are given for Helixer and AUGUSTUS with the median predictions scores being denoted by the red lines.
-Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}.}\label{fig:boxplot}
-%\end{figure}
-
 \begin{figure*}[!tpb]
@@
 \midrule
-AUGUSTUS & 0.632 (+-0.09) & 0.757 (+-0.182)\\
-Dilated CNN Baseline & - (+-) & - (+-)\\
-Best Helixer model & - (+-) & - (+-) \\
-Best model + varied input length & - (+-) & - \\
-Best model (+ varied input length) + overlapping &  0.85 (+-0.068) & 0.843 (+-0.151) \\
-(Varied input length +) overlapping + ensembling & 0.868 (+-0.064) & 0.863 (+-0.148) \\
+AUGUSTUS & 0.632 & 0.757 \\
+Dilated CNN Baseline & -  & - \\
+Best Helixer model & -  & -  \\
+Best model + varied input length & -  & - \\
+Best model (+ varied input length) + overlapping &  0.85  & 0.843  \\
+(Varied input length +) overlapping + ensembling & 0.868  & 0.863  \\
 \botrule

### 2020-05-02 paper/main.tex
caption, references

@@
 \subsection{Metrics}
-We use two metrics to judge model performance.
+We mainly use two metrics to judge model performance.
 The {\it Genic F1} is our primary metric that provides the most comprehensive picture of annotation quality in one number.
@@
 The metrics essentially provide a weighted mean of the performances in the considered classes.
+
+In Figure \ref{fig:main_results}c we also report on the overall base pair level accuracy besides the Genic F1.
 
@@
 The final predictions of a single model are constructed by overlapping predictions, that were made from a sliding window and then cut to a core sequence.
-This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see Figure \ref{fig:length_wise_bias}).
+This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see Figure \ref{fig:main_results}c).
 It also improves the average model performance in general by providing the model with multiple different starting points.
@@
 \begin{figure*}[!tpb]
-\label{fig:results}
+\label{fig:main_results}
 \centerline{\includegraphics[width=\textwidth]{images/main_results}}
-\caption{Main overall results.
-{\bf a)} Boxplots comparing the Subgenic F1 scores for all animal and plant genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
-Results are given for Helixer and AUGUSTUS with the median predictions scores being denoted by the red lines.
+\caption{Main overall results and other investigations.
+The top row shows the animal data; the bottom row is concerned with the plants.
+Yellow dots are for AUGUSTUS and the blue ones are for Helixer.
+{\bf a)} Swarm - and boxplots showing the Subgenic F1 scores for all genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
+Each dot represents the prediction performace for one genome.
+The median prediction scores are shown by the red line.
 Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}.
-{\bf c)} Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping.
-Each value is the average performance of a 200 basepair long chunk by one of the final ensemble models.
-The regular metrics show a significant negative bias towards the borders of each sequence.
+{\bf b)} Scatterplots showing the prediction performances measured in Subgenic F1 by genome length.
+The regression line is shown with a 95\% confidence interval.
+{\bf c)} The average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences.
+The dashed line shows the same without overlapping.
+Each value is the average performance of a 200 basepair long subsection by one of the models of the final ensemble.
+The usage of the model without overlapping results in a significant negative bias towards the borders of each sequence.
 Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average.
-See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
+See Section 4 for the Supplements for the effects of overlapping for each individual species.}
 \end{figure*}
@@
 \begin{table}[!t]
-\processtable{Experimental Results and Comparison\label{tab:results}} {
+\processtable{Summary of Experimental Results\label{tab:results}} {
 \begin{tabular}{@{}lll@{}}
@@
 \section{Results}
-Figure \ref{fig:boxplot} shows the side-by-side comparison of the distribution of performances on all non-training genomes in the animal and plant case by Subgenic F1.
+Figure \ref{fig:main_results}a shows the side-by-side comparison of the distribution of performances on all non-training genomes in the animal and plant case by Subgenic F1.
 In the case of Helixer, these scores represent cross-species predictions.

### 2020-05-02 plots/coverage/scores_histo_frm_h5.py
attempt update for normalized coverage

@@ -7,20 +7,23 @@ import copy
 
 
 # overall score
-def score(f, other_y, predictions, n=1000, by=2000):
-    """get counts for histogram of reference scores broken up by other_y
+def score(f, other_y, predictions, n=1000, by=2000, max_bin=4):
+    """get counts for histogram of normalized coverage broken up by other_y confusion matrix
 
-    other_y should be a list of keys for h5 datasets to be compared to data/y, for each key
-    this will record and return 8 histograms under the same key:
-       [ig right, ig wrong, utr right, utr wrong, cds right, cds wrong, intron right, intron wrong]"""
-    breaks = [x / n for x in range(n + 1)]
+    other_y should be a list of alternatives for y (ref or preds) as keys in the h5 file"""
+    # double check our 'max_bin' is actually high enough to summarize data
+    maxes_by_sp = f['meta/max_normalized_cov_sc'].attrs
+    for sp in maxes_by_sp:
+        assert max_bin > np.max(maxes_by_sp[sp])
+
+    breaks = [x / n * max_bin for x in range(n + 1)]
 
     i_s = list(range(4))
     i_s_by_y = [i_s for _ in range(len(other_y))]
     argmax_sets = list(itertools.product(*i_s_by_y))
 
     histos = {}
-    score_cats = ['ig', 'exon', 'intron']
+    score_cats = ['cov', 'sc']
     for key in score_cats:
         histos[key] = [np.zeros((n,)) for _ in range(len(argmax_sets))]
 
@@ -32,8 +35,8 @@ def score(f, other_y, predictions, n=1000, by=2000):
             if key == 'predictions':
               of = predictions
             preds[key] = np.argmax(of[key][i:(i + by)], axis=2).ravel()[un_padded]
-        scores = f['scores/ig_exon_intron_bp'][i:(i + by)].ravel()
-
+        scores = f['scores/norm_cov_by_bp'][i:(i + by)].reshape([-1, 2])
+        print(scores.shape)
         for ih, argmaxes in enumerate(argmax_sets):
             subpreds = copy.deepcopy(preds)
             scoressm = copy.deepcopy(scores)

### 2020-05-03 paper/main.tex
minor
abstract

@@
 
-\abstract{\textbf{Motivation:} Current state of the art gene annotation programs for eukaryotic genomes have to be pretrained for each species and still often produce annotations that lack the required accuracy.
-\\
-\textbf{Results:} We set up a framework that can be used to build much more general gene callers that also outperform traditional methods on a variety of complex genomes.\\
-\textbf{Availability:} \\
-\textbf{Contact:} \href{name@bio.com}{name@bio.com}\\
+\abstract{\textbf{Motivation:} Current state-of-the-art tools for the annotation of genes in eukaryotic genomes have to be pretrained for each species and still often produce annotations that can be improved much further.
+The fundamental algorithmic architecture for those tools also remains the same since about two decades, limiting learning capabilities.
+Here, we set out to improve the cross-species annotation of genes from DNA data alone with the help of deep learning.
+The goal is to eliminate the dependency on a closely related gene model while also improving the predictive quality in general with a fundamentally new architecture.\\
+\textbf{Results:} We present Helixer, a framework for the development and usage of a cross-species deep learning model that improves significantly on performance and generalizability when compared to more traditional methods.
+We evaluate our approach by building a single model for the annotation of 186 animal genomes as well as one for 51 plants.
+Our predictions are shown to be much less sensitive to the length of the genome than those of a current state-of-the-are tool.
+We also present two novel post-processing techniques that each worked to further strengthen our annotations and show in-depth results of an RNA-Seq based comparison of our predictions.\\
+\textbf{Availability:} The source code of this work is available at \url{https://github.com/weberlab-hhu/HelixerPrep} [change] under the GNU General Public License v3.0.\\
+\textbf{Contact:} \href{alisandra.denton@hhu.de}{alisandra.denton@hhu.de}\\
 \textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
@@
 The regression line is shown with a 95\% confidence interval.
-{\bf c)} The average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences.
+{\bf c)} The average Genic F1 (in red) and basepair wise accuracy (in purple) with respect to the position in the 20000 basepair long input sequences.
 The dashed line shows the same without overlapping.
@@
 
+Eliminating the need for a close genome annotation being present for pretraining.
+
 We found our models to be highly sensitive to the training genomes we chose.

### 2020-05-04 data_scripts/insert-one.sh
minor

@@ -1,9 +1,10 @@
 #! /bin/bash
 
 geenuff_basedir=~/git/GeenuFF
-data_basedir=/mnt/data/ali/share/ensembl_test
+# data_basedir=/mnt/data/ali/share/ensembl_test
 # data_basedir=/mnt/data/ali/share/ensembl_nosplit
 # data_basedir=/mnt/data/ali/share/phytozome_organized/ready/test
+data_basedir=/mnt/data/ali/share/phytozome_organized/ready/train
 genome=$1
 db_path=$2
 
@@ -13,6 +14,7 @@ mkdir -v -p $geenuff_basedir/tmp/input $geenuff_basedir/tmp/output
 # gunzip -v -c $data_basedir/$genome/ensembl98/annotation/*.gff3.gz > $geenuff_basedir/tmp/input/$genome.gff3
 # gunzip -v -c $data_basedir/$genome/ensembl98/assembly/*.fa.gz > $geenuff_basedir/tmp/input/$genome.fa
 
-cp -v $data_basedir/$genome/ensembl98/annotation/*.gff3 $data_basedir/$genome/ensembl98/assembly/*.fa $geenuff_basedir/tmp/input/
+# cp -v $data_basedir/$genome/ensembl98/annotation/*.gff3 $data_basedir/$genome/ensembl98/assembly/*.fa $geenuff_basedir/tmp/input/
+cp -v $data_basedir/$genome/v2.0/annotation/*.gff3 $data_basedir/$genome/v2.0/assembly/*.fa $geenuff_basedir/tmp/input/
 
 $geenuff_basedir/import_genome.py --db-path $2 --basedir $geenuff_basedir/tmp/ --species $genome

### 2020-05-04 paper/main.tex

@@
 Each value is the average performance of a 200 basepair long subsection by one of the models of the final ensemble.
-The usage of the model without overlapping results in a significant negative bias towards the borders of each sequence.
-Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average.
-See Section 4 for the Supplements for the effects of overlapping for each individual species.}
+See Section 4 for the Supplements shows the effects of overlapping for each individual species.}
 \end{figure*}

### 2020-05-04 paper/supplements.tex

@@
 \begin{table}[!h]
+\renewcommand\thetable{}
 \resizebox{\textwidth}{!}{%
-\renewcommand\thetable{S1}
 \centering
@@
 \begin{table}[!h]
+\renewcommand\thetable{S1}
 \resizebox{\textwidth}{!}{%
-\renewcommand\thetable{S1}
 \centering
@@
 \end{tabular}}
+\caption{Detailed results for all animal genomes.}
 \end{table}
 
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part1}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part2}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part1}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part2}}
-%\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part1}}
+\end{figure}
+\begin{figure}[!h]
+\renewcommand\thefigure{S1}
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part2}}
+\caption{Subgenic F1 results for all animal genomes shown by phylogenetic position.}
+\end{figure}
+
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part1}}
+\end{figure}
+\begin{figure}[!h]
+\renewcommand\thefigure{S2}
+\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part2}}
+\caption{Difference in Subgenic F1 to AUGUSTUS for all animal genomes shown by phylogenetic position.}
+\end{figure}
 
@@
 \begin{table}[!h]
+\renewcommand\thetable{S2}
 \resizebox{\textwidth}{!}{%
-\renewcommand\thetable{S1}
 \centering
@@
 \end{tabular}}
+\caption{Detailed results for all plant genomes.}
 \end{table}
@@
 \begin{figure}[!h]
+\renewcommand\thefigure{S3}
 \centerline{\includegraphics[width=1.0\textwidth]{images/plants_f1_scores}}
+\caption{Subgenic F1 results for all plant genomes shown by phylogenetic position.}
 \end{figure}
 \begin{figure}[!h]
+\renewcommand\thefigure{S4}
 \centerline{\includegraphics[width=1.0\textwidth]{images/plants_differences}}
+\caption{Difference in Subgenic F1 to AUGUSTUS for all plant genomes shown by phylogenetic position.}
 \end{figure}
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
+\renewcommand\thetable{S3}
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
+\renewcommand\thetable{S4}
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
+\renewcommand\thetable{S5}
 \centering
@@
 
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals1}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals2}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals3}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals4}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals5}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals6}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals7}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals8}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals9}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals10}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals11}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals12}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
-%\end{figure}
-%
-%\clearpage
-%\subsection{Plants}
-%
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants1}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants2}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants3}}
-%\end{figure}
-%\begin{figure}[!h]
-%\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants4}}
-%\end{figure}
-%
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals1}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals2}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals3}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals4}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals5}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals6}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals7}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals8}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals9}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals10}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals11}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals12}}
+\end{figure}
+\begin{figure}[!h]
+\renewcommand\thefigure{S5}
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
+\caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all animals genomes individually.}
+\end{figure}
+
+\clearpage
+\subsection{Plants}
+
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants1}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants2}}
+\end{figure}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants3}}
+\end{figure}
+\begin{figure}[!h]
+\renewcommand\thefigure{S6}
+\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants4}}
+\caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all plant genomes individually.}
+\end{figure}
+
 \clearpage
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
+\renewcommand\thetable{S6}
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
+\renewcommand\thetable{S7}
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
+\renewcommand\thetable{S8}
 \centering

### 2020-05-04 data_scripts/insert-one.sh
minor

@@ -14,7 +14,6 @@ mkdir -v -p $geenuff_basedir/tmp/input $geenuff_basedir/tmp/output
 # gunzip -v -c $data_basedir/$genome/ensembl98/annotation/*.gff3.gz > $geenuff_basedir/tmp/input/$genome.gff3
 # gunzip -v -c $data_basedir/$genome/ensembl98/assembly/*.fa.gz > $geenuff_basedir/tmp/input/$genome.fa
 
-# cp -v $data_basedir/$genome/ensembl98/annotation/*.gff3 $data_basedir/$genome/ensembl98/assembly/*.fa $geenuff_basedir/tmp/input/
-cp -v $data_basedir/$genome/v2.0/annotation/*.gff3 $data_basedir/$genome/v2.0/assembly/*.fa $geenuff_basedir/tmp/input/
+cp -v $data_basedir/$genome/*/annotation/*.gff3 $data_basedir/$genome/*/assembly/*.fa $geenuff_basedir/tmp/input/
 
 $geenuff_basedir/import_genome.py --db-path $2 --basedir $geenuff_basedir/tmp/ --species $genome

### 2020-05-05 paper/main.tex
results table finished

@@
 AUGUSTUS & 0.632 & 0.757 \\
-Dilated CNN Baseline & -  & - \\
-Best Helixer model & -  & -  \\
-Best model + varied input length & -  & - \\
-Best model (+ varied input length) + overlapping &  0.85  & 0.843  \\
+Dilated CNN Baseline & 0.666 &  0.802 \\
+Best Helixer model & 0.77  & 0.833  \\
+Best model + varied input length & 0.834  & - \\
+Best model (+ varied input length) + overlapping &  0.844  & 0.843  \\
 (Varied input length +) overlapping + ensembling & 0.868  & 0.863  \\

### 2020-05-06 meta_info_collec/coverage/actually_plot.R
maybe presentable coverage plots

@@ -1,0 +1,59 @@
+h=2.6
+widths = c(6, 4)
+cex=0.55
+
+targ_files <- c('/ref_vs_helixer.csv', '/aug_vs_helixer.csv')
+outdirs <- c('plots/vs_ref/', 'plots/vs_aug/')
+refs <- c('Ref', 'Aug')
+rows_sc <- list(c(1:4, 9:10, 13:16),
+                c(1, 3, 4, 9:10, 15:16))
+rows_cov <- list(c(1:8, 13:16),
+                 c(1, 3:4, 7:8, 15:16))
+mwidths <- c(4, 3)
+r_margin_cov <- c(2.1, 10.5) 
+
+for (j in 1:2){
+    targ_file <- targ_files[j]
+    outdir <- outdirs[j]
+    ref <- refs[j]
+    mw <- mwidths[j]
+    w <- widths[j]
+    source('import.R')
+    
+    postscript(file=paste0(outdir, 'averaged_spliced.eps'), height=h, width=w)
+      par(xpd=TRUE, mar=c(5.1, 4.1, 4.1, 10.5), cex=cex)
+      ticksat <- viribp(summed_demelted_sc[rows_sc[[j]], ], 
+                        ylab='average fraction of bp in\nspliced coverage bin',
+                        ref=ref, match_length=mw)
+      legend(ticksat[length(ticksat)] + 0.2, 1, colnames(summed_demelted_sc),
+             fill=viridis(6), bty='n', title='Bin')
+    dev.off()
+    
+    postscript(file=paste0(outdir, 'averaged_coverage.eps'), height=h, width=w)
+      par(xpd=TRUE, mar=c(5.1, 4.1, 4.1, r_margin_cov[j]), cex=cex)
+    
+      ticksat <- viribp(summed_demelted_cov[rows_cov[[j]],],
+                        ylab='average fraction of bp in\ncoverage bin',
+                        ref=ref, match_length=mw)
+    dev.off()
+    
+    for (sp in species){
+
+        postscript(file=paste0(outdir, sp, '_spliced.eps'), height=h, width=w)
+          par(xpd=TRUE, mar=c(5.1, 4.1, 4.1, 10.5), cex=cex)
+          ticksat <- viribp(sp_demelted_sc[[sp]][rows_sc[[j]], ],
+                            ylab='average fraction of bp in\nspliced coverage bin',
+                            ref=ref, match_length=mw)
[…]

### 2020-05-06 meta_info_collec/coverage/baseplot.R

@@ -1,0 +1,21 @@
+library(viridis)
+
+viribp <- function(demelt, ylab='fraction of bp with\ncoverage in bin', ref='Ref', match_length=4){
+
+    n_pairwise <- (nrow(demelt) - match_length) / 2
+    exon_widths <- c(0.0, 0.3)[c(rep(1, match_length), rep(c(2,1), n_pairwise))]
+
+    padj = 0.7
+    xat <- barplot(t(demelt), col=rev(viridis(6)), space=exon_widths, padj=padj)
+    title(ylab=ylab, line=1.7, cex=1.2)
+    ticksat <- unique(c(xat - 0.5, xat + 0.5))
+    ticksat <- sort(ticksat)
+    before_pairs <- match_length + 1
+    axis(1, at=ticksat[1:before_pairs], labels=rep('', before_pairs))
+    for (i in seq(1, n_pairwise * 3, 3)){
+        axis(1, at=ticksat[0:2 + i + before_pairs], labels=rep('', 3))
+    }
+    axis(1, at=-.5, labels=paste0('     ', ref, '.:   \nHelixer:   '), tick=F, padj=padj)
+    return(ticksat)
+}
+

### 2020-05-06 meta_info_collec/coverage/import.R

@@ -1,0 +1,55 @@
+# need a couple helper functions from this one
+source('cov_breakdown.R')
+
+sum_list_elements <- function(x){
+    out <- x[[1]]
+    for (i in 2:length(x)){
+        out <- out + x[[i]]
+    }
+    return(out)
+}
+
+mk_covbin <- function(x, covstring){
+    cov_bin <- factor(x[, covstring],
+                      levels=sort(unique(x[, covstring])))
+    levels(cov_bin) <- expl_cov(levels(cov_bin))
+    return(cov_bin)
+}
+
+# go from multi factor format to 2D table 
+sum_ref_pred <- function(x) {
+    sapply(split(x, x$ref_pred), function(w) sum(w$count))
+}
+
+unmelt <- function(x, cov_col){
+    # x has to be a table as created by coverage_counter.py, 
+    # but with the two 'argmax' columns combined to 'ref_pred'
+    demelt <- sapply(split(x, x[, cov_col]), sum_ref_pred)
+    demelt <- demelt[,ncol(demelt):1]
+    demelt <- demelt / rowSums(demelt)
+    return(demelt)
+}
+
+decode = c("IG", "UTR", "CDS", "Ntrn")
+
+## go from 
+species <- list.files('pairwise/')
+sp_list <- lapply(species, function(x) read.csv(paste0('pairwise/', x, targ_file)))
+# setup refernce vs prediction descriptive names
+for (i in 1:length(sp_list)){
+    sp_list[[i]]$ref_pred <- apply(sp_list[[i]][, c('argmax_y', 'argmax_pred')], 1, cm_labs, decode=decode)
+    sp_list[[i]]$ref_pred <- factor(sp_list[[i]]$ref_pred, levels=unique(sp_list[[i]]$ref_pred))
+    sp_list[[i]]$cov_bin  <- mk_covbin(sp_list[[i]], 'coverage_greater_eq')
+    sp_list[[i]]$sc_bin  <- mk_covbin(sp_list[[i]], 'spliced_coverage_greater_eq')
+}
+names(sp_list) <- species
+
+sp_demelted_cov <- lapply(sp_list, unmelt, cov_col='cov_bin')
+# convenience reorder
+sp_demelted_cov <- lapply(sp_demelted_cov, function(x) x[c(1,6,11,16, 2,5,3,9,4,13,7,10,8,14,12,15),])
[…]

### 2020-05-08 paper/main.tex
line breaks

@@
 
-\abstract{\textbf{Motivation:} Current state-of-the-art tools for the annotation of genes in eukaryotic genomes have to be pretrained for each species and still often produce annotations that can be improved much further.
-The fundamental algorithmic architecture for those tools also remains the same since about two decades, limiting learning capabilities.
-Here, we set out to improve the cross-species annotation of genes from DNA data alone with the help of deep learning.
-The goal is to eliminate the dependency on a closely related gene model while also improving the predictive quality in general with a fundamentally new architecture.\\
-\textbf{Results:} We present Helixer, a framework for the development and usage of a cross-species deep learning model that improves significantly on performance and generalizability when compared to more traditional methods.
-We evaluate our approach by building a single model for the annotation of 186 animal genomes as well as one for 51 plants.
-Our predictions are shown to be much less sensitive to the length of the genome than those of a current state-of-the-are tool.
-We also present two novel post-processing techniques that each worked to further strengthen our annotations and show in-depth results of an RNA-Seq based comparison of our predictions.\\
-\textbf{Availability:} The source code of this work is available at \url{https://github.com/weberlab-hhu/HelixerPrep} [change] under the GNU General Public License v3.0.\\
+\abstract{\textbf{Motivation:} Current state-of-the-art tools for the annotation of 
+genes in eukaryotic genomes have to be pretrained for each species and still often 
+produce annotations that can be improved much further. The fundamental algorithmic 
+architecture for those tools also remains the same since about two decades, limiting
+learning capabilities. Here, we set out to improve the cross-species annotation of
+genes from DNA data alone with the help of deep learning. The goal is to eliminate 
+the dependency on a closely related gene model while also improving the predictive
+quality in general with a fundamentally new architecture.\\
+\textbf{Results:} We present Helixer, a framework for the development and usage of 
+a cross-species deep learning model that improves significantly on performance and
+generalizability when compared to more traditional methods. We evaluate our 
+approach by building a single model for the annotation of 186 animal genomes as 
+well as one for 51 plants. Our predictions are shown to be much less sensitive to
+the length of the genome than those of a current state-of-the-are tool. We also 
+present two novel post-processing techniques that each worked to further strengthen 
+our annotations and show in-depth results of an RNA-Seq based comparison of our 
+predictions.\\
+\textbf{Availability:} The source code of this work is available at 
+\url{https://github.com/weberlab-hhu/HelixerPrep} [change] under the GNU General 
+Public License v3.0.\\
 \textbf{Contact:} \href{alisandra.denton@hhu.de}{alisandra.denton@hhu.de}\\
-\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
-online.}
+\textbf{Supplementary information:} Supplementary data are available at 
+\textit{Bioinformatics} online.}
 
@@
 \section{Introduction}
-The annotation of genes in DNA is fundamental for many other tasks depending on it to be done as good and correct as possible.
-It can be performed at different levels of precision, from simple coding -- non-coding classification to detailed structural labeling.
-Because of the sheer size of genomes alone, gene annotation is a very labour intensive process and genomic database providers like NCBI \citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl} offer gene annotation pipelines.
-These pipelines integrate experimental data (from e.g.
-RNA-seq or proteogenomics) with homolog sequences in the database and ab-initio gene predictions.
-The latter is an attractive approach, because it is cheap and fast.
-State of the art performance is achieved by higher order hidden markov models (HMMs), such as Genscan \citep{burge1997prediction}, AUGUSTUS \citep{stanke2003gene} or SNAP \citep{johnson2008snap}.
-Their accuracy, however, leaves room for improvement.
-By encoding possible states and transitions in a probalistic model, designers of HMMs assume structure in the sequence that may limit its predictive power.
-In practice HMMs have trouble to generalize across species and it seems actual learning of underlying patterns is limited despite the fundamental ability to do so being present.
+The annotation of genes in DNA is fundamental for many other tasks depending on it 
+to be done as good and correct as possible. It can be performed at different levels 
+of precision, from simple coding -- non-coding classification to detailed structural 
+labeling. Because of the sheer size of genomes alone, gene annotation is a very 
+labour intensive process and genomic database providers like NCBI 
+\citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl} offer gene 
+annotation pipelines. These pipelines integrate experimental data (from e.g. RNA-seq 
+or proteogenomics) with homolog sequences in the database and ab-initio gene 
+predictions. The latter is an attractive approach, because it is cheap and fast. 
+State of the art performance is achieved by higher order hidden markov models 
+(HMMs), such as Genscan \citep{burge1997prediction}, AUGUSTUS \citep{stanke2003gene} 
+or SNAP \citep{johnson2008snap}. Their accuracy, however, leaves room for 
+improvement. By encoding possible states and transitions in a probalistic model, 
+designers of HMMs assume structure in the sequence that may limit its predictive 
+power. In practice HMMs have trouble to generalize across species and it seems 
+actual learning of underlying patterns is limited despite the fundamental ability 
+to do so being present.
 
-In the last decade, deep neural networks (DNN) have been applied with great success in many areas of statistical modelling, including biology \citep{ching2018opportunities}.
-For sequence data, such as DNA, speech or text, a special kind of recurrent neural network (RNN) called long-short term memory (LSTM) \citep{hochreiter1997long} is an established building block for many different architectures.
-LSTM units can also be used to process sequential input starting from both ends, forming a bi-directional LSTM (BLSTM).
-It has also been shown that HMMs can be successfully combined with DNNs \citep{liu2016novo, liu2016pedla}.
+In the last decade, deep neural networks (DNN) have been applied with great success 
+in many areas of statistical modelling, including biology 
+\citep{ching2018opportunities}. For sequence data, such as DNA, speech or text, a 
+special kind of recurrent neural network (RNN) called long-short term memory (LSTM) 
+\citep{hochreiter1997long} is an established building block for many different 
+architectures. LSTM units can also be used to process sequential input starting 
+from both ends, forming a bi-directional LSTM (BLSTM). It has also been shown that 
+HMMs can be successfully combined with DNNs \citep{liu2016novo, liu2016pedla}.
 
 For the purpose of gene annotation, RNNs have already shown promising results. 
-\citep{choudharypredicting} carries out preliminary explorations on the potential of BLSTMs for cross species gene prediction and trains his model on human genes to test it later on two more species.
-DeepAnnotator \citep{amin2018deepannotator} uses bidirectional LSTMs (BLSTM) for gene finding in prokaryotes.
-Gene prediction in prokaryotes is considered more amenable than in eukaryotes, as genes in prokaryotes are proportionately more frequent in the genome, feature simpler control structures and do not use splicing \citep{wang2004brief}.
-DanQ \citep{quang2016danq} proposes the use of a BLSTM after a concurrent neural net (CNN) to find detailed motifs in the human genome.
-DeePromotor \citep{oubounyt2019deepromoter} trains a similar architecture for the recognition of promotor regions.
-Recently, several groups \citep{jaganathan2019predicting, wang2019splicefinder} successfully used CNNs to find splicing sites.
+\citep{choudharypredicting} carries out preliminary explorations on the potential 
+of BLSTMs for cross species gene prediction and trains his model on human genes to 
+test it later on two more species. DeepAnnotator \citep{amin2018deepannotator} uses 
+bidirectional LSTMs (BLSTM) for gene finding in prokaryotes. Gene prediction in 
+prokaryotes is considered more amenable than in eukaryotes, as genes in prokaryotes 
+are proportionately more frequent in the genome, feature simpler control structures 
+and do not use splicing \citep{wang2004brief}. DanQ \citep{quang2016danq} proposes 
+the use of a BLSTM after a concurrent neural net (CNN) to find detailed motifs in 
+the human genome. DeePromotor \citep{oubounyt2019deepromoter} trains a similar 
+architecture for the recognition of promotor regions. Recently, several groups 
+\citep{jaganathan2019predicting, wang2019splicefinder} successfully used CNNs to 
+find splicing sites.
 
-In this work we present Helixer, a novel prototype software for training and utilizing a general purpose DNN for the ab-initio cross-species gene annotation of large eukaryotic genomes using only DNA data as input.
-Our model is trained to differentiate between 4 regions: Intergenic, Untranslated (UTR), Coding (CDS) and Intron.
-We demonstrate the effectiveness of this approach by training a single model for the annotation of a large set of genomes from the kingdoms metazoa and viridiplantae, respectively, which we will call {\it animals} and {\it plants} from now on.
-We worked with the full data of 192 animal genomes and 60 plant genomes.
-Both the ability to generalize across genomes as well as the scope of the evaluation greatly surpasses all other work in this field.
-The source code and all input data is publically available.
+In this work we present Helixer, a novel prototype software for training and 
+utilizing a general purpose DNN for the ab-initio cross-species gene annotation of 
+large eukaryotic genomes using only DNA data as input. Our model is trained to 
+differentiate between 4 regions: Intergenic, Untranslated (UTR), Coding (CDS) and 
+Intron. We demonstrate the effectiveness of this approach by training a single model 
+for the annotation of a large set of genomes from the kingdoms metazoa and 
+viridiplantae, respectively, which we will call {\it animals} and {\it plants} from 
+now on. We worked with the full data of 192 animal genomes and 60 plant genomes. 
+Both the ability to generalize across genomes as well as the scope of the evaluation 
+greatly surpasses all other work in this field. The source code and all input data 
+is publically available.
 
@@
 \subsection{Datasets}
-The foundation of our work are 192 animal and 60 plant genomes.
-The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format.
-We used all animals genomes in EnsemblMetazoa 45 \citep{howe2020ensembl} as well as all non-embargoed plant genomes of the JGI Phytosome 13 database \citep{goodstein2012phytozome}.
-Both data groups were used seperately during training as well as during evaluation.
+The foundation of our work are 192 animal and 60 plant genomes. The data of each 
+genome consists of the latest publically available genomic assembly in form of a 
+FASTA file and the latest annotation in the GFF format. We used all animals genomes 
+in EnsemblMetazoa 45 \citep{howe2020ensembl} as well as all non-embargoed plant 
+genomes of the JGI Phytosome 13 database \citep{goodstein2012phytozome}. Both data 
+groups were used seperately during training as well as during evaluation.
 
-At the start of our work we split both genome groups into {\it training genomes} and {\it test genomes}.
-The test genome groups are made up of 19 and 6 genomes, respectively, and were chosen to be of the best possible quality based on collected metadata while also representing a broad phylogenetic spread.
+At the start of our work we split both genome groups into {\it training genomes} and 
+{\it test genomes}. The test genome groups are made up of 19 and 6 genomes, 
+respectively, and were chosen to be of the best possible quality based on collected 
+metadata while also representing a broad phylogenetic spread.
 
-The selection of training genomes also requires balancing multiple conflicting tradeoffs.
-On the one hand, we do want to train with as much data as possible, but not all data has a good enough quality to enable a powerful generalization.
-It is also very desirable to have a broad spread phylogenetically as well as in genome size and average gene length as the model is increasingly unlikely to generalize well beyond the borders given by the training data.
-However, if the genomic patterns inside the data are too different from each other, it may be difficult for the model to generalize, too.
-It was also important for us to get experimental results within a couple of days as we tested many different data and model configurations, which limited the overall size of the training set and the model.
+The selection of training genomes also requires balancing multiple conflicting 
+tradeoffs. On the one hand, we do want to train with as much data as possible, but 
+not all data has a good enough quality to enable a powerful generalization. It is 
+also very desirable to have a broad spread phylogenetically as well as in genome 
+size and average gene length as the model is increasingly unlikely to generalize 
+well beyond the borders given by the training data. However, if the genomic patterns 
+inside the data are too different from each other, it may be difficult for the model 
+to generalize, too. It was also important for us to get experimental results within 
+a couple of days as we tested many different data and model configurations, which 
+limited the overall size of the training set and the model.
 
-We employed an iterative approach to effectively select a proper set of training genomes.
-We first started with 3-4 genomes that were almost certain to be of the highest quality.
-We then evaluated the performance of the resulting model on all genomes individually (including the training genomes) and looked for candidates to add to the training set or to remove from it.
-This process was repeated multiple times concurrently with the model search until we saw no more room for substantial improvement given our computational constraints.
+We employed an iterative approach to effectively select a proper set of training 
+genomes. We first started with 3-4 genomes that were almost certain to be of the 
+highest quality. We then evaluated the performance of the resulting model on all 
+genomes individually (including the training genomes) and looked for candidates to 
+add to the training set or to remove from it. This process was repeated multiple 
+times concurrently with the model search until we saw no more room for substantial 
+improvement given our computational constraints.
 
-We call the genomes that were ultimatily not used during training from the original split the {\it validation genomes}.
-While we did split off a group of test genomes at the beginning of our work, we ended up only reporting on the generalization performance on the combined set of validation and test genomes as the validation genomes are by far the largest set.
-We also did not find a noticible difference in performance between those two groups.
+We call the genomes that were ultimatily not used during training from the original 
+split the {\it validation genomes}. While we did split off a group of test genomes 
+at the beginning of our work, we ended up only reporting on the generalization 
+performance on the combined set of validation and test genomes as the validation 
+genomes are by far the largest set. We also did not find a noticible difference in 
+performance between those two groups.
 
-Once a set of training genomes was selected, we split it up again into a {\it training and validation set}.
-This split of the training genomes was done to get a quicker sense of the generalization capabilities and was used primarily during training.
-We split off the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level.
-Evaluation of the annotations on all training and validation genomes was done regularly after a promising model candidate was found based on its validation set performance.
-Ultimately we only cared about the cross-species performance on all validation genomes of a model.
+Once a set of training genomes was selected, we split it up again into a {\it 
+training and validation set}. This split of the training genomes was done to get a 
+quicker sense of the generalization capabilities and was used primarily during 
+training. We split off the validation set by selecting 20\% of the FASTA sequences 
+above and below the N90 of each training genome, ensuring a proper distribution of 
+large and short sequences in both sets and a split on the chromosome level. 
+Evaluation of the annotations on all training and validation genomes was done 
+regularly after a promising model candidate was found based on its validation set 
+performance. Ultimately we only cared about the cross-species performance on all 
+validation genomes of a model.
 
 \subsection{Data Pre-processing}
-We first pre-process and store the raw genomic information by using GeenuFF \citep{denton2019}.
-GeenuFF is a tool for checking and exploring genomic data, that stores all information inside a SQL database.
-Training and evaluation-ready data is generated by querying this database and then transforming the returned data into a numerical format suitable for machine learning.
-The encoding of the genomic sequence is done in line with the IUPAC nucleic acid notation and the functional annotation used as labels during training is transformed to a one hot encoding with the four classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}.
-See also Table \ref{tab:arrays} for a more detailed description of the generated data types.
+We first pre-process and store the raw genomic information by using GeenuFF 
+\citep{denton2019}. GeenuFF is a tool for checking and exploring genomic data, that 
+stores all information inside a SQL database. Training and evaluation-ready data is 
+generated by querying this database and then transforming the returned data into a 
+numerical format suitable for machine learning. The encoding of the genomic sequence 
+is done in line with the IUPAC nucleic acid notation and the functional annotation 
+used as labels during training is transformed to a one hot encoding with the four 
+classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}. See also Table 
+\ref{tab:arrays} for a more detailed description of the generated data types.
 
@@
 \botrule
-\end{tabular}}{{\it Note:} Values are averages of the individual values of each genome in a data group.
-All statistics except the average gene lengths are excluding FASTA sequences without a gene and subsequences that were masked as completely erroneous.
-Each strand of DNA was counted seperately.
-The average gene length was determined by the average length of the longest transcript by cumulative exon length, which is how we query for genes during data generation.
-Brackets show the standart deviation.}
+\end{tabular}}{{\it Note:} Values are averages of the individual values of each 
+genome in a data group. All statistics except the average gene lengths are excluding 
+FASTA sequences without a gene and subsequences that were masked as completely 
+erroneous. Each strand of DNA was counted seperately. The average gene length was 
+determined by the average length of the longest transcript by cumulative exon 
+length, which is how we query for genes during data generation. Brackets show the 
+standart deviation.}
 \end{table}
 
-During data generation, we query for the longest transcript of each gene and disregard FASTA sequences that have no functional annotation.
-GeenuFF checks the genomic annotations for potential errors during import and is able to mark those areas.
-We use this information to effectively blank out those bases during training by using the sample weights described in Table \ref{tab:arrays}.
-The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and GeenuFF marks a potentially large intergenic region for it.
-Table \ref{tab:statistics_all} shows statistics about the masking.
+During data generation, we query for the longest transcript of each gene and 
+disregard FASTA sequences that have no functional annotation. GeenuFF checks the 
+genomic annotations for potential errors during import and is able to mark those 
+areas. We use this information to effectively blank out those bases during training 
+by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
+of bases masked lie in the intergenic region as the most prevelant error is a 
+missing UTR and GeenuFF marks a potentially large intergenic region for it. Table 
+\ref{tab:statistics_all} shows statistics about the masking.
 
-For the training itself, we divide each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network together with the sample weights.
-We append zero padding if the subsequences are not long enough.
-If a subsequence is fully marked as erroneous, we do not include it anywhere.
+For the training itself, we divide each continous genomic sequence into 20.000 bases 
+long subsequences, which are then used as input or label for the neural network 
+together with the sample weights. We append zero padding if the subsequences are 
+not long enough. If a subsequence is fully marked as erroneous, we do not include 
+it anywhere.
 
 \subsection{Metrics}
-We mainly use two metrics to judge model performance.
-The {\it Genic F1} is our primary metric that provides the most comprehensive picture of annotation quality in one number.
-The {\it Subgenic F1} is similar to the Genic F1, except that it does not take UTR predictions into account.
-This is done for comparability with AUGUSTUS and further explained in Section \ref{sec:augustus}.
-Both metrics work by summing up True Positives (TP), False Positives (FP) and False Negatives (FN) of each considered class before calculating precision and recall from these values and combining the two into a F1 score.
-Both metrics do not take the values of the intergenic class into account, as this class is very abundant and appears to be by far the easiest to predict.
-(See Table \ref{tab:statistics_all} for the class distribution and Tables S1 and S2 for a more in-depth report our model performance including the intergenic class).
-In the case of the Genic F1, this means that effectively only the TP of the intergenic class are disregarded.
-The metrics essentially provide a weighted mean of the performances in the considered classes.
+We mainly use two metrics to judge model performance. The {\it Genic F1} is our 
+primary metric that provides the most comprehensive picture of annotation quality 
+in one number. The {\it Subgenic F1} is similar to the Genic F1, except that it does 
+not take UTR predictions into account. This is done for comparability with AUGUSTUS 
+and further explained in Section \ref{sec:augustus}. Both metrics work by summing up 
+True Positives (TP), False Positives (FP) and False Negatives (FN) of each 
+considered class before calculating precision and recall from these values and 
+combining the two into a F1 score. Both metrics do not take the values of the 
+intergenic class into account, as this class is very abundant and appears to be by 
+far the easiest to predict. (See Table \ref{tab:statistics_all} for the class 
+distribution and Tables S1 and S2 for a more in-depth report our model performance 
+including the intergenic class). In the case of the Genic F1, this means that 
+effectively only the TP of the intergenic class are disregarded. The metrics 
+essentially provide a weighted mean of the performances in the considered classes.
 
-In Figure \ref{fig:main_results}c we also report on the overall base pair level accuracy besides the Genic F1.
+In Figure \ref{fig:main_results}c we also report on the overall base pair level 
+accuracy besides the Genic F1.
 
@@
 \label{sec:model}
-We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer normalization \citep{ba2016layer} between each BLSTM layer to produce the predictions in the form of a base pair wise classification.
-The model consists of circa 5.4 million parameters and was implemented with the deep learning library keras \citep{chollet2015keras} on top of TensorFlow \citep{abadi2016tensorflow}.
-We tested multiple different architectures before arriving at this model configuration, including convolutional neural networks (CNN) and hybrid architectures.
-We also use class weigths as the classes of the input data are not balanced equally enough, at least not for our network, and does also vary greatly between animals and plants.
-For example, the tradeoff between the number of intergenic and intronic bases is quite different in both groups (see Table \ref{tab:statistics_all}).
+We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer 
+normalization \citep{ba2016layer} between each BLSTM layer to produce the 
+predictions in the form of a base pair wise classification. The model consists of 
+circa 5.4 million parameters and was implemented with the deep learning library 
+keras \citep{chollet2015keras} on top of TensorFlow \citep{abadi2016tensorflow}. We 
+tested multiple different architectures before arriving at this model configuration, 
+including convolutional neural networks (CNN) and hybrid architectures. We also use 
+class weigths as the classes of the input data are not balanced equally enough, at 
+least not for our network, and does also vary greatly between animals and plants. 
+For example, the tradeoff between the number of intergenic and intronic bases is 
+quite different in both groups (see Table \ref{tab:statistics_all}).
 
-The hyperparameters were optimized by a combination of manual and automatic optimization.
-Automatic optimization was carried out by using either the TPE algorithm \citep{bergstra2011algorithms}, random search or grid search depending on the situation.
-We used NNI \citep{nni2019} to facilitate the search.
-All relevant hyperparameters can be found in Table S3 and S4 or in the Helixer source code repository.
+The hyperparameters were optimized by a combination of manual and automatic 
+optimization. Automatic optimization was carried out by using either the TPE 
+algorithm \citep{bergstra2011algorithms}, random search or grid search depending on 
+the situation. We used NNI \citep{nni2019} to facilitate the search. All relevant 
+hyperparameters can be found in Table S3 and S4 or in the Helixer source code 
+repository.
 
-Our final model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously.
-This grouping enabled us to train effectively with far longer sequences than usual.
-A class wise evaluation including the calculation of the Genic F1 score is performed on the validation set after every epoch.
+Our final model was trained with 10 bases of genomic sequence as input during each 
+time step and produces individual predictions for each of those 10 bases 
+simultaneously. This grouping enabled us to train effectively with far longer 
+sequences than usual. A class wise evaluation including the calculation of the Genic 
+F1 score is performed on the validation set after every epoch.
 
-We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input.
-Details of the neural architecture search are given in Section 6 of the Supplements.
+We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is 
+widely used when working with DNA data as input. Details of the neural architecture 
+search are given in Section 6 of the Supplements.
 
 \subsection{Inference Techniques}
-We also used multiple techniques to improve the prediction quality after the training was done.
-One very effective way for genomes with larger genes was to input longer sequences during inference time than during training.
-This was done for all animal genomes except the invertebrates.
-The input sequences were up to 10 times larger, depending on species group and assembly quality.
-This also demonstrates the ability of our model to generalize as it is able to make successful prediction on far longer sequences than it has ever seen.
-The sequence length during inference can be up to 200.000 base pairs long and is determined by the phylogenetic class.
-The concrete lengths were chosen to keep the typical average gene length roughly proportional to the length of the sequence input.
-For more implementation details on this see Section 3 of the Supplements or the source code.
+We also used multiple techniques to improve the prediction quality after the 
+training was done. One very effective way for genomes with larger genes was to input 
+longer sequences during inference time than during training. This was done for all 
+animal genomes except the invertebrates. The input sequences were up to 10 times 
+larger, depending on species group and assembly quality. This also demonstrates the 
+ability of our model to generalize as it is able to make successful prediction on 
+far longer sequences than it has ever seen. The sequence length during inference can 
+be up to 200.000 base pairs long and is determined by the phylogenetic class. The 
+concrete lengths were chosen to keep the typical average gene length roughly 
+proportional to the length of the sequence input. For more implementation details 
+on this see Section 3 of the Supplements or the source code.
 
-The final predictions of a single model are constructed by overlapping predictions, that were made from a sliding window and then cut to a core sequence.
-This was done to strongly reduce a typical drop in performance of the models towards the beginning and end of each sequence (see Figure \ref{fig:main_results}c).
-It also improves the average model performance in general by providing the model with multiple different starting points.
-The different overlapping sequences were recombined by averaging the individual softmax values of each base.
-The figures in Section 4 of the Supplements show the effect of overlapping for each genome, ordered by N75.
-We found that overlapping tends to work best if the genomes are not very fragmented and we used it for both animals and plants.
+The final predictions of a single model are constructed by overlapping predictions, 
+that were made from a sliding window and then cut to a core sequence. This was done 
+to strongly reduce a typical drop in performance of the models towards the beginning 
+and end of each sequence (see Figure \ref{fig:main_results}c). It also improves the 
+average model performance in general by providing the model with multiple different 
+starting points. The different overlapping sequences were recombined by averaging 
+the individual softmax values of each base. The figures in Section 4 of the 
+Supplements show the effect of overlapping for each genome, ordered by N75. We found 
+that overlapping tends to work best if the genomes are not very fragmented and we 
+used it for both animals and plants.
 
-A model ensemble with 8 components was used to generate the final predictions for each species.
-For this, we trained 4 seperate animal and plant models and also used a model from an epoch that did not have the highest Genic F1 on the validation set.
-Depending on the results, we used the model iteration with either the highest genic recall or precision.
-The model with the best Genic F1 was always best in one of those.
-This was done to increase the diversity of the model ensemble.
-As with the overlapping, the fusion of the 8 individual predictions was done by averaging the softmax values of each base pair prediction.
+A model ensemble with 8 components was used to generate the final predictions for 
+each species. For this, we trained 4 seperate animal and plant models and also used 
+a model from an epoch that did not have the highest Genic F1 on the validation set. 
+Depending on the results, we used the model iteration with either the highest genic 
+recall or precision. The model with the best Genic F1 was always best in one of 
+those. This was done to increase the diversity of the model ensemble. As with the 
+overlapping, the fusion of the 8 individual predictions was done by averaging the 
+softmax values of each base pair prediction.
 
@@
 \botrule
-\end{tabular}}{{\it Note:} An intron inside a UTR region is encoded as intron and can be inferred from the prediction order.}
+\end{tabular}}{{\it Note:} An intron inside a UTR region is encoded as intron and 
+can be inferred from the prediction order.}
 \end{table}
@@
 \centerline{\includegraphics[width=\textwidth]{images/main_results}}
-\caption{Main overall results and other investigations.
-The top row shows the animal data; the bottom row is concerned with the plants.
-Yellow dots are for AUGUSTUS and the blue ones are for Helixer.
-{\bf a)} Swarm - and boxplots showing the Subgenic F1 scores for all genomes that were not used during training of Helixer, which is equivalent to all validation and test genomes.
-Each dot represents the prediction performace for one genome.
-The median prediction scores are shown by the red line.
-Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}.
-{\bf b)} Scatterplots showing the prediction performances measured in Subgenic F1 by genome length.
-The regression line is shown with a 95\% confidence interval.
-{\bf c)} The average Genic F1 (in red) and basepair wise accuracy (in purple) with respect to the position in the 20000 basepair long input sequences.
-The dashed line shows the same without overlapping.
-Each value is the average performance of a 200 basepair long subsection by one of the models of the final ensemble.
-See Section 4 for the Supplements shows the effects of overlapping for each individual species.}
+\caption{Main overall results and other investigations. The top row shows the animal
+ data; the bottom row is concerned with the plants. Yellow dots are for AUGUSTUS and 
+the blue ones are for Helixer. {\bf a)} Swarm - and boxplots showing the Subgenic F1 
+scores for all genomes that were not used during training of Helixer, which is 
+equivalent to all validation and test genomes. Each dot represents the prediction 
+performace for one genome. The median prediction scores are shown by the red line. 
+Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}. {\bf b)} 
+Scatterplots showing the prediction performances measured in Subgenic F1 by genome 
+length. The regression line is shown with a 95\% confidence interval. {\bf c)} The 
+average Genic F1 (in red) and basepair wise accuracy (in purple) with respect to 
+the position in the 20000 basepair long input sequences. The dashed line shows the 
+same without overlapping. Each value is the average performance of a 200 basepair 
+long subsection by one of the models of the final ensemble. See Section 4 for the 
+Supplements shows the effects of overlapping for each individual species.}
 \end{figure*}
@@
 \botrule
-\end{tabular}}{{\it Note:} Values are the median in Subgenic F1 across all validation and test genomes of the respective group.
-Varied input length was only used in the animal case.
-The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.}
+\end{tabular}}{{\it Note:} Values are the median in Subgenic F1 across all 
+validation and test genomes of the respective group. Varied input length was only 
+used in the animal case. The best model was chosen out of the eight models of the 
+ensemble for having the best performance in the reported setup.}
 \end{table}
@@
 \section{Results}
-Figure \ref{fig:main_results}a shows the side-by-side comparison of the distribution of performances on all non-training genomes in the animal and plant case by Subgenic F1.
-In the case of Helixer, these scores represent cross-species predictions.
-We also compare the median performances of different configurations of Helixer with AUGUSTUS and a dilated CNN baseline in Table \ref{tab:results}.
-The results show a clear improvement over the AUGUSTUS both in median performance and spread.
-We also significantly outperform a baseline constructed of a dilated CNN architecture.
+Figure \ref{fig:main_results}a shows the side-by-side comparison of the distribution 
+of performances on all non-training genomes in the animal and plant case by Subgenic 
+F1. In the case of Helixer, these scores represent cross-species predictions. We 
+also compare the median performances of different configurations of Helixer with 
+AUGUSTUS and a dilated CNN baseline in Table \ref{tab:results}. The results show a 
+clear improvement over the AUGUSTUS both in median performance and spread. We also 
+significantly outperform a baseline constructed of a dilated CNN architecture.
 
-Our models, however, tend to perform less well for the very smallest and largest genomes or species that are phylogenetically the furthest away from our training genomes.
-This is the case for both animals and plants and is visualized by the Figures S1 and S3.
-We do not, for example, consistently predict very well on the algae as well as the non-avian reptiles.
-While one of the the algae were included in our training genomes, the non-avian reptiles were not.
-Our model also tends to show higher uncertainty around transitions of annotation classes (e.g.
-from UTR to CDS) due to the lack of further post-processing of the basepair wise predictions.
+Our models, however, tend to perform less well for the very smallest and largest 
+genomes or species that are phylogenetically the furthest away from our training 
+genomes. This is the case for both animals and plants and is visualized by the 
+Figures S1 and S3. We do not, for example, consistently predict very well on the 
+algae as well as the non-avian reptiles. While one of the the algae were included 
+in our training genomes, the non-avian reptiles were not. Our model also tends to 
+show higher uncertainty around transitions of annotation classes (e.g. from UTR to 
+CDS) due to the lack of further post-processing of the basepair wise predictions.
 
-AUGUSTUS does outperform us rather consistently on the smallest genomes, but falls off much more the larger the genomes get.
-The difference in prediction quality is especially strong for mammals, which tend to have big genomes with very long genes as well as the largest plants.
-Figures S2 and S4 display the comparison to AUGUSTUS by phylogenetic position.
+AUGUSTUS does outperform us rather consistently on the smallest genomes, but falls 
+off much more the larger the genomes get. The difference in prediction quality is 
+especially strong for mammals, which tend to have big genomes with very long genes 
+as well as the largest plants. Figures S2 and S4 display the comparison to AUGUSTUS 
+by phylogenetic position.
 
-Two techniques were used during inference to improve performance and limit model bias.
-The usage of longer input lengths helped especially for genomes that tend to have longer genes and was enabled by our model architecture being a relatively simple LSTM stack without any fully connected layers on top.
-We also constructed the final predictions out of a overlapping ones, which greatly helped to reduce prediction bias in most genomes.
-To our knowledge, both techniques were not used before in a model developed for gene annotation.
+Two techniques were used during inference to improve performance and limit model 
+bias. The usage of longer input lengths helped especially for genomes that tend to 
+have longer genes and was enabled by our model architecture being a relatively 
+simple LSTM stack without any fully connected layers on top. We also constructed the 
+final predictions out of a overlapping ones, which greatly helped to reduce 
+prediction bias in most genomes. To our knowledge, both techniques were not used 
+before in a model developed for gene annotation.
 
 \section{Discussion}
-We propose a novel way to develop gene calling models that are able to perform cross-species genomic annotiation for a wide variety of large eucaryotic genomes.
-Helixer does this while outperforming AUGUSTUS on a base pair wise level and withstanding RNASeq data based scrutiny.
+We propose a novel way to develop gene calling models that are able to perform 
+cross-species genomic annotiation for a wide variety of large eucaryotic genomes. 
+Helixer does this while outperforming AUGUSTUS on a base pair wise level and 
+withstanding RNASeq data based scrutiny.
 
@@
 
-We found our models to be highly sensitive to the training genomes we chose.
-A different set could lead to a significant shift in strengths and weaknesses of the model and a larger and more spread out set of high quality genomes could also result in a wider range of genomes with decent predictions.
-For this, more computational resources would be required.
+We found our models to be highly sensitive to the training genomes we chose. A 
+different set could lead to a significant shift in strengths and weaknesses of the 
+model and a larger and more spread out set of high quality genomes could also result 
+in a wider range of genomes with decent predictions. For this, more computational 
+resources would be required.
 
-An avenue for future research could be the addition of RNASeq data as additional input.
-This would bring Helixer on even footing with current tools and could lead to real world applicable performance improvements if the results of this work are any indication, as deep learning has been shown to excel in a multimodal settings \citep{ching2018opportunities}.
+An avenue for future research could be the addition of RNASeq data as additional 
+input. This would bring Helixer on even footing with current tools and could lead 
+to real world applicable performance improvements if the results of this work are 
+any indication, as deep learning has been shown to excel in a multimodal settings 
+\citep{ching2018opportunities}.
 

### 2020-05-08 paper/main.tex
some proofing, filling some methods gaps

@@
 
-\abstract{\textbf{Motivation:} Current state-of-the-art tools for the annotation of 
-genes in eukaryotic genomes have to be pretrained for each species and still often 
+\abstract{\textbf{Motivation:} Current state-of-the-art tools for the \textit{de novo} annotation of 
+genes in eukaryotic genomes have to be specifically fitted for each species and still often 
 produce annotations that can be improved much further. The fundamental algorithmic 
-architecture for those tools also remains the same since about two decades, limiting
+architecture for these tools has remained largely unchanged for about two decades, limiting
 learning capabilities. Here, we set out to improve the cross-species annotation of
-genes from DNA data alone with the help of deep learning. The goal is to eliminate 
+genes from DNA sequence alone with the help of deep learning. The goal is to eliminate 
 the dependency on a closely related gene model while also improving the predictive
@@
 well as one for 51 plants. Our predictions are shown to be much less sensitive to
-the length of the genome than those of a current state-of-the-are tool. We also 
+the length of the genome than those of a current state-of-the-art tool. We also 
 present two novel post-processing techniques that each worked to further strengthen 
@@
 \textbf{Contact:} \href{alisandra.denton@hhu.de}{alisandra.denton@hhu.de}\\  
+% normally contact is last author / person with permanent position (i.e. Andreas). 
+% Not sure how important that is, but if we're skipping that, then we can change it back to Felix
 \textbf{Supplementary information:} Supplementary data are available at 
@@
 \section{Introduction}
-The annotation of genes in DNA is fundamental for many other tasks depending on it 
-to be done as good and correct as possible. It can be performed at different levels 
+The annotation of genes in DNA is fundamental for many downstream tasks, which 
+depend on high quality annotations. Gene annotation can be performed at different levels 
 of precision, from simple coding -- non-coding classification to detailed structural 
-labeling. Because of the sheer size of genomes alone, gene annotation is a very 
-labour intensive process and genomic database providers like NCBI 
-\citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl} offer gene 
-annotation pipelines. These pipelines integrate experimental data (from e.g. RNA-seq 
-or proteogenomics) with homolog sequences in the database and ab-initio gene 
+labeling. Because of the sheer size of genomes alone, manual gene annotation is 
+generally intractable. Instead, researchers can use pipelines such as 
+Maker \citep{}, PASA \citep{} or those offered by genomic database providers like NCBI 
+\citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl}. 
+These pipelines integrate experimental data (from e.g. RNA-seq 
+or proteogenomics) with homologous sequences in the database and ab-initio gene 
 predictions. The latter is an attractive approach, because it is cheap and fast. 
-State of the art performance is achieved by higher order hidden markov models 
+State-of-the-art performance is achieved by higher order hidden markov models 
 (HMMs), such as Genscan \citep{burge1997prediction}, AUGUSTUS \citep{stanke2003gene} 
@@
 designers of HMMs assume structure in the sequence that may limit its predictive 
-power. In practice HMMs have trouble to generalize across species and it seems 
+power. In practice HMMs have trouble generalizing across species and it seems 
 actual learning of underlying patterns is limited despite the fundamental ability 
-to do so being present.
+to do so being present \citep{}.
 
@@
 and do not use splicing \citep{wang2004brief}. DanQ \citep{quang2016danq} proposes 
-the use of a BLSTM after a concurrent neural net (CNN) to find detailed motifs in 
+the use of a BLSTM after a convolutional neural net (CNN) to find detailed motifs in 
 the human genome. DeePromotor \citep{oubounyt2019deepromoter} trains a similar 
@@
 utilizing a general purpose DNN for the ab-initio cross-species gene annotation of 
-large eukaryotic genomes using only DNA data as input. Our model is trained to 
+large eukaryotic genomes using only DNA sequence as input. Our model is trained to 
 differentiate between 4 regions: Intergenic, Untranslated (UTR), Coding (CDS) and 
@@
 genome consists of the latest publically available genomic assembly in form of a 
-FASTA file and the latest annotation in the GFF format. We used all animals genomes 
+FASTA file and the latest annotation in the GFF format. One genome for each species of animal 
 in EnsemblMetazoa 45 \citep{howe2020ensembl} as well as all non-embargoed plant
-genomes of the JGI Phytosome 13 database \citep{goodstein2012phytozome}. Both data 
-groups were used seperately during training as well as during evaluation.
+species from the JGI Phytosome 13 database \citep{goodstein2012phytozome}; exact 
+genomes used are list here \ref{}. % add a supplemental table with names and versions
+Data was downloaded on 29/03/2019 and 15/10/2019 for plants and animals respectively. 
+Both data groups were used seperately during training as well as during evaluation.
+
+Both to divide genomes into training, validation and test sets, as well as to understand
+performance, we needed a way to evaluate genome and annotation quality.
+Metadata was generated for each genome, including: basic length and composition statistics \citep{}; % quast
+k-mer content \citep{}; % jellyfish
+conserved gene content of the genome, transcriptome, and proteome \citep{}; % BUSCO, files generated with gffread
+and finally simple counts of genic features from the gff files. In particular, 
+we considered the following features of the metadata when judging the quality of a genome and it's 
+annotation: assembly contiguity (N50 / genome size); retention in the proteome 
+of detected conserved orthogroups (BUSCOs) that were identified in the genome; and
+the presence of UTRs and alternative splicing as estimated by the gff feature-counts.
+Finally this metadata was combined with researcher knowledge of model species to guide genome
+selection as necessary.
 
@@
 {\it test genomes}. The test genome groups are made up of 19 and 6 genomes, 
-respectively, and were chosen to be of the best possible quality based on collected 
+respectively, and were chosen to be of decent quality and diversity based on collected 
 metadata while also representing a broad phylogenetic spread.
 
+For both feasibility and iterative cross-species generalization validattion, we
+selected only some of the remaining genomes for direct training.
 The selection of training genomes also requires balancing multiple conflicting 
@@
 We employed an iterative approach to effectively select a proper set of training 
-genomes. We first started with 3-4 genomes that were almost certain to be of the 
+genomes. We first started with 3-4 genomes that were expected to be of the 
 highest quality. We then evaluated the performance of the resulting model on all 

### 2020-05-08 meta_info_collec/coverage/actually_plot.R
more coverage stuff

@@ -25,7 +25,7 @@ for (j in 1:2){
       ticksat <- viribp(summed_demelted_sc[rows_sc[[j]], ], 
                         ylab='average fraction of bp in\nspliced coverage bin',
                         ref=ref, match_length=mw)
-      legend(ticksat[length(ticksat)] + 0.2, 1, colnames(summed_demelted_sc),
+      legend(ticksat[length(ticksat)] + 0.2, 1, rev(colnames(summed_demelted_sc)),
              fill=viridis(6), bty='n', title='Bin')
     dev.off()
     
@@ -44,7 +44,7 @@ for (j in 1:2){
           ticksat <- viribp(sp_demelted_sc[[sp]][rows_sc[[j]], ],
                             ylab='average fraction of bp in\nspliced coverage bin',
                             ref=ref, match_length=mw)
-          legend(ticksat[length(ticksat)] + 0.2, 1, colnames(summed_demelted_sc),
+          legend(ticksat[length(ticksat)] + 0.2, 1, rev(colnames(summed_demelted_sc)),
                  fill=viridis(6), bty='n', title='Bin')
         dev.off()
     

### 2020-05-11 paper/main.tex
some rephrasing and starting on Augustus stuff

@@
 size and average gene length as the model is increasingly unlikely to generalize 
-well beyond the borders given by the training data. However, if the genomic patterns 
-inside the data are too different from each other, it may be difficult for the model 
-to generalize, too. It was also important for us to get experimental results within 
-a couple of days as we tested many different data and model configurations, which 
-limited the overall size of the training set and the model.
+well beyond the borders given by the training data. However, it also may be difficult 
+for the model learn and generalize, if the genomic patterns inside the data are 
+too different from each other. It was also important for us to limit the size of
+our training genomes set to be small enough that enough that we could get experimental 
+results within a couple of days and thereby be able to test many different data and 
+model configurations.
 
@@
 
-Once a set of training genomes was selected, we split it up again into a {\it 
+Once a set of training genomes was selected, we further split the sequences therein into a {\it 
 training and validation set}. This split of the training genomes was done to get a 
@@
 regularly after a promising model candidate was found based on its validation set 
-performance. Ultimately we only cared about the cross-species performance on all 
-validation genomes of a model.
+performance. Ultimately we used the cross-species performance on all 
+validation genomes as the decisive measure of model quality.
 
@@
 \end{tabular}}{{\it Note:} Values are averages of the individual values of each 
-genome in a data group. All statistics except the average gene lengths are excluding 
-FASTA sequences without a gene and subsequences that were masked as completely 
-erroneous. Each strand of DNA was counted seperately. The average gene length was 
-determined by the average length of the longest transcript by cumulative exon 
-length, which is how we query for genes during data generation. Brackets show the 
-standart deviation.}
+genome in a data group. All statistics except the average gene lengths exclude
+FASTA sequences without a gene and any 20,000 bp subsequences that were masked as completely 
+erroneous. Each strand of DNA was counted seperately. The gene length was 
+determined by the length of the pre-mRNA of the longest protein at each
+loci. Brackets show the standard deviation.}
 \end{table}
 
-During data generation, we query for the longest transcript of each gene and 
-disregard FASTA sequences that have no functional annotation. GeenuFF checks the 
+During data generation, we query for the transcript with the longest protein of each gene and 
+disregard FASTA sequences that have no functional annotation as it is ambiguous
+whether such sequences contain no genes or were simply not annotated in the reference. GeenuFF checks the 
 genomic annotations for potential errors during import and is able to mark those 
@@
 by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
-of bases masked lie in the intergenic region as the most prevelant error is a 
+of masked bases lie in the intergenic region as the most prevelant error is a 
 missing UTR and GeenuFF marks a potentially large intergenic region for it. Table 
@@
 
-For the training itself, we divide each continous genomic sequence into 20.000 bases 
-long subsequences, which are then used as input or label for the neural network 
-together with the sample weights. We append zero padding if the subsequences are 
-not long enough. If a subsequence is fully marked as erroneous, we do not include 
-it anywhere.
+For the training itself, we divide each continous genomic sequence into 20,000 bp 
+long subsequences, for which one-hot vectors of the base pairs and annotations 
+are generated and respectively used as input and label for the neural network 
+together with the sample weights. We append zero padding if the sequences (or
+sequence ends) are shorter than 20,000 bp. 
+If a subsequence is fully marked as erroneous, we exclude it from all analyses.
 
@@
 considered class before calculating precision and recall from these values and 
-combining the two into a F1 score. Both metrics do not take the values of the 
+combining the two into a F1 score. Neither metric takes the values of the 
 intergenic class into account, as this class is very abundant and appears to be by 
@@
 effectively only the TP of the intergenic class are disregarded. The metrics 
-essentially provide a weighted mean of the performances in the considered classes.
+essentially provide a weighted mean of the performances in the considered classes. %?? hmmm, but this is not a weighted mean across classes is it? Do introns not weigh more heavily?
 
@@
 including convolutional neural networks (CNN) and hybrid architectures. We also use 
-class weigths as the classes of the input data are not balanced equally enough, at 
-least not for our network, and does also vary greatly between animals and plants. 
+class weigths as the classes of the input data are both unbalanced vary greatly 
+between animals and plants. 
 For example, the tradeoff between the number of intergenic and intronic bases is 
 quite different in both groups (see Table \ref{tab:statistics_all}).
+
+A class wise evaluation including the calculation of the Genic % still could use a better location...
+F1 score is performed on the validation set after every epoch. The best model
+from any training run was selected as the model with the highest Genic F1 after either
+maximum epochs were reached or Genic F1 stopped improving and training was interrupted. % check / what was patience set to?
 
@@
 simultaneously. This grouping enabled us to train effectively with far longer 
-sequences than usual. A class wise evaluation including the calculation of the Genic 
-F1 score is performed on the validation set after every epoch.
-
-We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is 
+sequences than usual. 
+We compare our final models to a dilated CNN (dCNN) baseline as this kind of architecture is 
 widely used when working with DNA data as input. Details of the neural architecture 
@@
 animal genomes except the invertebrates. The input sequences were up to 10 times 
-larger, depending on species group and assembly quality. This also demonstrates the 
+longer, depending on the phylogenetic group and assembly quality. This also demonstrates the 
 ability of our model to generalize as it is able to make successful prediction on 
-far longer sequences than it has ever seen. The sequence length during inference can 
-be up to 200.000 base pairs long and is determined by the phylogenetic class. The 
+far longer sequences than it has ever seen. The 
 concrete lengths were chosen to keep the typical average gene length roughly 
@@
 The final predictions of a single model are constructed by overlapping predictions, 
-that were made from a sliding window and then cut to a core sequence. This was done 
+that were made from a sliding window and then cropped to a core sequence. This was done 
 to strongly reduce a typical drop in performance of the models towards the beginning 
@@
 average model performance in general by providing the model with multiple different 
-starting points. The different overlapping sequences were recombined by averaging 
+starting points. The different overlapping sequences were combined by averaging 
 the individual softmax values of each base. The figures in Section 4 of the 
@@
 A model ensemble with 8 components was used to generate the final predictions for 
-each species. For this, we trained 4 seperate animal and plant models and also used 
-a model from an epoch that did not have the highest Genic F1 on the validation set. 
-Depending on the results, we used the model iteration with either the highest genic 
-recall or precision. The model with the best Genic F1 was always best in one of 
+each species. For this, we performed 4 separate animal and plant training runs and 
+selected two models each, specifically the model with the best precision and the  % please check
+other with the best recall. The model with the best Genic F1 was always best in one of 
 those. This was done to increase the diversity of the model ensemble. As with the 
@@
 \label{sec:augustus}
+Due to the practical time constraints for retraining and running, we decided to compare our Helixer
+models only to one existing {\it de novo} gene caller, the popular gene caller, Augustus.
+Typical usage requires retraining Augustus for each species, with the exception of a
+few lucky cases where a model for a close relative is already available. To scale this for 
+the large plant and animal datasets, we used protein-homology to create a training set and
+therefore could train and evaluate only models without UTRs.
+
+Orthologs of highly conserved single-copy genes were identified in each genome using
+BUSCO. The {\it viridiplantae} set was used for plants and the {\it metazoa} set for animals.
+The training genbank files were generated directly by BUSCO, by utilizing the ``--long"  % inline code?
+parameter. For plants, the entire retraining could be performed as above; however, for animals
+we randomly selected only half BUSCO-generated training set, which resulted in a training set 
+size and runtime comparable to the plants.
+
 What Marvin did, why we don't use Genic F1 for comparison, etc

### 2020-05-12 paper/main.tex
first draft of RNAseq methods

@@
 
-Orthologs of highly conserved single-copy genes were identified in each genome using
+Orthologs of highly conserved generally single-copy genes were identified in each genome using
 BUSCO. The {\it viridiplantae} set was used for plants and the {\it metazoa} set for animals.
@@
 parameter. For plants, the entire retraining could be performed as above; however, for animals
-we randomly selected only half BUSCO-generated training set, which resulted in a training set 
-size and runtime comparable to the plants.
+we randomly selected only half the BUSCO-generated training set, which resulted in a training set 
+size and runtime comparable to the plants (~400 genes and several days per species). For animals
+the training was carried out with the subsetted training file and using the following scripts provided
+by Augustus. An untrained model was setup with "new\_species.pl", and the model was fit by running
+"etraining" before and after the major hyper-parameter optimization with "optimize\_augustus.pl".
+Using the trained model for each species, we ran the main prediction ("augustus") with ``--UTR=off"
+and ``--gff3=on". The gffs produced by Augustus were imported into GeenuFF and exported as h5
+files in the same manor as the reference, allowing direct comparison with reference and 
+Helixer predictions. While tenable for some 237 species, this method allowed neither training
+nor prediction of UTR regions with Augustus, so the metric Sub Genic F1, which excludes UTR,
+was used for all comparisons between Augustus and Helixer.
 
-What Marvin did, why we don't use Genic F1 for comparison, etc
+\subsection{Evaluation against independent RNAseq data}
+\label{sec:rnaseq}
+The reference annotations were created with existing tools and largely with a pipeline incorporating
+{\it de novo} gene predictions with RNAseq and homology data. The references, like any
+data, are expected to contain errors; but here we were particularly concerned because
+the {\it de novo} gene caller used by the references was Augustus for many species (...)
+% todo list and cite examples (above and below)
+and a fundamentally similar HMM-based tool for many more (...), which would potentially bias 
+reference based comparisons of Helixer's and Augustus' performance.
+
+We therefore downloaded and processed public RNAseq data to obtain an independent option 
+for evaluating model performance. We selected three each of plant and animal genomes for
+detailed evaluation with RNAseq. These were selected to have relatively good 
+({\it Manihot esculenta} and {\it Papio anubis}), 
+typical ({\it Medicago truncatula} and {\it Equus caballus}), 
+and poor ({\it Theobroma cacao} and {\it Petromyzon marinus})
+performance compared to Augustus \ref{} % supp phylogenies maybe?
+within our generalizable range (i.e. excluding algae and
+invertebrates). Selections were further constrained by the availability of stranded RNAseq
+data.
+
+% todo, add versions to everything below
+For each of these six species the following search was performed on Sequence Read Archive
+'(("<species name>"[Organism] OR <species name>[All Fields]) AND stranded[All Fields]) 
+AND ("biomol rna"[Properties] AND "library layout paired"[Properties])'. In so far as 
+more than 50 samples were identified, every Nth sample was selected so that about 50
+samples were chosen for further processing, as listed here \ref{}. % todo make a supp table
+Each sample was quality controlled with FastQC \citep{}, before and after trimming
+with Trimmomatic \citep{}, and then mapped to the genome with Hisat2 \citep{}. 
+Samtools \citep{} was used for the sorting and indexing that was necessary both for
+RNAseq mapped metric collection with PicardTools \citep{} and custom downstream 
+processing below. MultiQC \citep{} was used to summarize the results. The above was 
+automated and the results visualized with the code here (... make available and link to RNAseq stuff).
+Samples were filtered to those that had {\it relatively} high mapping rates, particularly to
+exonic regions; that had low 3' bias, normal FastQC and Trimmomatic stats, and that
+were confirmed to be stranded with 2nd strand as sense strand. If more remained, seven
+of the high quality samples were selected randomly.
+
+Finally, the selected, quality, mapped RNAseq samples were quantified to
+get the coverage (number of reads matching; i.e. cigar =, M or X),
+and spliced coverage (number of reads with gap or splice; i.e. cigar N or D)
+for every base pair in the genomes as implemented in the "training\_rnaseq.py"
+script.
 

### 2020-05-12 paper/main.tex
draft

--- 2020-05-08 paper/main.tex
@@
 \usepackage{url}
+
+\usepackage{draftwatermark}
+\SetWatermarkText{DRAFT}
+\SetWatermarkScale{1}
 
@@
 \textbf{Availability:} The source code of this work is available at 
-\url{https://github.com/weberlab-hhu/HelixerPrep} [change] under the GNU General 
+\url{https://github.com/weberlab-hhu/HelixerPrep} under the GNU General 
 Public License v3.0.\\
@@
 
+\begin{table}[!t]
+\processtable{Data arrays generated and used by Helixer\label{tab:arrays}} {
+\begin{tabular}{@{}ll@{}}
+\toprule Name & Information \\
+\midrule
+Input & Genomic sequence in the 4-dimensional IUPAC encoding\\
+	  & (one hot encoding for non-ambiguous bases) \\
+Output & Labels in a 4-dimensional one hot encoding representing \\
+      & the classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}\\
+Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
+\botrule
+\end{tabular}}{{\it Note:} An intron inside a UTR region is encoded as intron and 
+can be inferred from the prediction order.}
+\end{table}
 \begin{methods}
@@
 
-\subsection{Evaluation of AUGUSTUS} 
-\label{sec:augustus}
-What Marvin did, why we don't use Genic F1 for comparison, etc
-
-\begin{table}[!t]
-\processtable{Data arrays generated and used by Helixer\label{tab:arrays}} {
-\begin{tabular}{@{}ll@{}}
-\toprule Name & Information \\
-\midrule
-Input & Genomic sequence in the 4-dimensional IUPAC encoding\\
-	  & (one hot encoding for non-ambiguous bases) \\
-Output & Labels in a 4-dimensional one hot encoding representing \\
-      & the classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}\\
-Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
-\botrule
-\end{tabular}}{{\it Note:} An intron inside a UTR region is encoded as intron and 
-can be inferred from the prediction order.}
-\end{table}
-
 \end{methods}
@@
 
-Eliminating the need for a close genome annotation being present for pretraining.
-
 We found our models to be highly sensitive to the training genomes we chose. A 
@@
 
-\section*{Acknowledgements}
-
-We are very grateful for the support of...
-
-\section*{Funding}
-
-This work has been supported by the... \vspace*{-12pt}
-
 \bibliographystyle{natbib}

### 2020-05-14 paper/main.tex
attempts to describe the rnaseq

--- 2020-05-12 paper/main.tex
@@
 
-A class wise evaluation including the calculation of the Genic % still could use a better location...
+A class wise evaluation including the calculation of the Genic 
 F1 score is performed on the validation set after every epoch. The best model
 from any training run was selected as the model with the highest Genic F1 after either
-maximum epochs were reached or Genic F1 stopped improving and training was interrupted. % check / what was patience set to?
+maximum epochs were reached or Genic F1 stopped improving and training was interrupted.
 
@@
 each species. For this, we performed 4 separate animal and plant training runs and 
-selected two models each, specifically the model with the best precision and the  % please check
+selected two models each, specifically the model with the best precision and the  % please check / epoch model / version something
 other with the best recall. The model with the best Genic F1 was always best in one of 
@@
 Helixer predictions. While tenable for some 237 species, this method allowed neither training
-nor prediction of UTR regions with Augustus, so the metric Sub Genic F1, which excludes UTR,
+nor prediction of UTR regions with Augustus, so the metric Subgenic F1, which excludes UTR,
 was used for all comparisons between Augustus and Helixer.
@@
 
-Finally, the selected, quality, mapped RNAseq samples were quantified to
+Finally, the selected, quality, mapped RNAseq samples were merged and quantified to
 get the coverage (number of reads matching; i.e. cigar =, M or X),
@@
 algae as well as the non-avian reptiles. While one of the the algae were included 
-in our training genomes, the non-avian reptiles were not. Our model also tends to 
-show higher uncertainty around transitions of annotation classes (e.g. from UTR to 
-CDS) due to the lack of further post-processing of the basepair wise predictions.
+in our training genomes it accounted for a tiny proportion of the total training 
+data and the non-avian reptiles were not included at all.
 
-AUGUSTUS does outperform us rather consistently on the smallest genomes, but falls 
-off much more the larger the genomes get. The difference in prediction quality is 
+AUGUSTUS outperforms us on some of the smallest genomes, but falls off much more 
+drastically as the genomes get larger. The difference in prediction quality is 
 especially strong for mammals, which tend to have big genomes with very long genes 
@@
 final predictions out of a overlapping ones, which greatly helped to reduce 
-prediction bias in most genomes. To our knowledge, both techniques were not used 
+prediction bias in most genomes. To our knowledge, neither technique has been used 
 before in a model developed for gene annotation.
+
+As an independent evaluation and to get a better qualitative understanding of
+performance we compared Helixer predictions, Augustus predictions and the
+reference to RNAseq coverage data. RNAseq coverage provides support for an 
+exonic annotation (UTR or CDS), spliced coverages provides support for an 
+annotation of intron, and neither coverage nor spliced coverage is expected 
+for intergenic annotations. Looking at selected subsequences \ref{fig:examples} % and supplemental figures.
+we identified cases where RNAseq supported both the annotations of the reference 
+and Helixer (a), neither (b), the reference, but not Helixer (c),
+and Helixer, but not the reference (d). Helixer models do not yet have postprocessing
+to make finalized single predictions, but instead output basewise probabilities.
+We see that the model exibits higher uncertainty around transitions between annotation `
+classes, for instance between UTR and CDS, or more dramatically between UTR and 
+intergenic even where Helixer predictions closely match the reference and RNAseq data (a).
+
+In some cases, where Helixer did not receive RNAseq support for its highest probability
+annotation, it assigned a low but non-trivial probability to the RNAseq-supported 
+exon/intron pattern (b and c,hSlA2 supp). However, in the extreme,
+there are cases where the Helixer model exibits substantial indecision or confusion 
+and shifts gradually between classes with no single class receiving a high probability 
+for extended stretches (hSlA2, hSlA3, lSlA3). In one case the 
+has not made a reasonably is not assigning a sufficiently 
+confident probability to any
 

### 2020-05-14 paper/main.tex
more rnaseq res

@@
 \title[short Title]{Helixer: Cross-species Gene Annotation Of Large Eucaryotic Genomes Using Deep Learning}
-\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
+\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Daniela Dey, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
 \address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
@@
 and Helixer (a), neither (b), the reference, but not Helixer (c),
-and Helixer, but not the reference (d). Helixer models do not yet have postprocessing
+and Helixer, but not the reference (d).
+
+Helixer models do not yet have postprocessing
 to make finalized single predictions, but instead output basewise probabilities.
-We see that the model exibits higher uncertainty around transitions between annotation `
+We see that the model exibits higher uncertainty around transitions between annotation 
 classes, for instance between UTR and CDS, or more dramatically between UTR and 
 intergenic even where Helixer predictions closely match the reference and RNAseq data (a). 
+Helixer's uncertainty around UTR <-> intergenic transitions may relate to a
+fundamentally harder problem (there is no conserved motif at the site as is observed
+for splice sites and start/stop codons), lack a single precise biological site \citep{},
+or noise in the reference, which we observed relative to the RNAseq data \ref{}. % supplemental UTR figures, 1/species
 
+Helixer models also sometimes showed uncertainty for larger regions.
 In some cases, where Helixer did not receive RNAseq support for its highest probability
@@
 and shifts gradually between classes with no single class receiving a high probability 
-for extended stretches (hSlA2, hSlA3, lSlA3). In one case the 
-has not made a reasonably is not assigning a sufficiently 
-confident probability to any
+for extended stretches (hSlA2, hSlA3, lSlA3). Notably, in one of the examples ()
+the RNAseq shows evidence of alternative splicing; and in another () Helixer's
+prediction falls between that of the reference and Augustus.
+
+Coverage and Spliced coverage were broken down by the confusion matrix (of both
+Helixer vs Augustus and Helixer vs the reference) for all genomes \ref{}. % H vs A main, H vs R supp
+Where tools were in agreement (far left), coverage and spliced coverage closely matched
+expectations. Specifically for CDS:CDS and UTR:UTR most base pairs had some, and many
+had moderate or high coverage. The same pattern was seen for intron:intron and spliced
+coverage. Finally, intergenic:intergenic showed only a small fraction of base pairs with 
+any of either coverage or spliced coverage.
+In cases of conflict, Helixer's predictions were substantially more consistent with
+the RNAseq data than those of Augustus. Specifically, for base pairs where one tool said CDS and
+the other intergenic or intron \ref{}, there was more coverage when 
+Helixer said CDS. Similarly, for base pairs where one tool said intron and the other 
+CDS or intergenic, there was more spliced coverage when Helixer said intron \ref{}. 
+These patterns were consistent in direction but varied in magnitude in the individual species. % double check
+RNAseq based comparison of Helixer and the reference was less clear cut.
+-Helixer is good at differentiating ig, ref better at introns
+-diffs between species. 
+-Helixer is substantially better for Tcacao, with a "poor Sub Genic F1"
+-taken together: competitive quality predictions, in some cases the mistakes really are in the reference / we wouldn't want 100\% accuracy against the reference
 
@@
 Helixer does this while outperforming AUGUSTUS on a base pair wise level and 
-withstanding RNASeq data based scrutiny.
+consistency with independent RNASeq data.
 

### 2020-05-14 paper/main.tex
fin brain dump, this probably got a little long...

@@
 Public License v3.0.\\
-\textbf{Contact:} \href{alisandra.denton@hhu.de}{alisandra.denton@hhu.de}\\  
-% normally contact is last author / person with permanent position (i.e. Andreas). 
-% Not sure how important that is, but if we're skipping that, then we can change it back to Felix
+\textbf{Contact:} \href{andreas.weber@hhu.de}{andreas.weber@hhu.de}\\
+
 \textbf{Supplementary information:} Supplementary data are available at 
@@
 performance we compared Helixer predictions, Augustus predictions and the
-reference to RNAseq coverage data. RNAseq coverage provides support for an 
+reference to RNAseq coverage data. RNAseq was processed for three each of plants
+and animals chosen because they had varied performance quality relative to Augustus. 
+RNAseq coverage provides support for an 
 exonic annotation (UTR or CDS), spliced coverages provides support for an 
@@
 coverage. Finally, intergenic:intergenic showed only a small fraction of base pairs with 
-any of either coverage or spliced coverage.
+any of either coverage or spliced coverage. In all conflicts the amount of RNAseq support
+fell between the cases where tools agreed, indicating that all options were at least
+capable of finding weak-spots in the other annotations.
 In cases of conflict, Helixer's predictions were substantially more consistent with
@@
 CDS or intergenic, there was more spliced coverage when Helixer said intron \ref{}. 
-These patterns were consistent in direction but varied in magnitude in the individual species. % double check
+These patterns were consistent in direction but varied in magnitude in the individual species,
+with the expception of {\it P. marinus} where Helixer and Augustus performed comparably
+at differentiating CDS and introns.
+
 RNAseq based comparison of Helixer and the reference was less clear cut.
--Helixer is good at differentiating ig, ref better at introns
--diffs between species. 
--Helixer is substantially better for Tcacao, with a "poor Sub Genic F1"
--taken together: competitive quality predictions, in some cases the mistakes really are in the reference / we wouldn't want 100\% accuracy against the reference
+Averaged across species, helixer's predictions received slightly more support
+than the reference when differentiating CDS, UTR and introns from intergenic;
+however helixer's predictions received slightly less support than the reference 
+when differentiating CDS and UTR from introns. Performance varies between individual
+species, from {\it P. marinus}, where the reference receives more support
+in every conflict, to {\it T. cacao}, where Helixer models receive equivalent
+or more support in every conflict. Interestingly, these are the two species that
+were selected because they had (relative to the usual difference) low Subgenic F1 compared to 
+Augustus; for the former the RNAseq confirms relatively weak performance for Helixer, while the 
+for the latter RNAseq rather indicates a sub-par reference.
 
 \section{Discussion}
-We propose a novel way to develop gene calling models that are able to perform 
-cross-species genomic annotiation for a wide variety of large eucaryotic genomes. 
-Helixer does this while outperforming AUGUSTUS on a base pair wise level and 
+We indtroduce Helixer a novel, deep-learning based tool for gene annotation
+prediction. Helixer outperforms AUGUSTUS on base pair wise metrics and on 
 consistency with independent RNASeq data.
 
-Eliminating the need for a close genome annotation being present for pretraining.
+We include trained models for plants and animals which achieve 
+high prediction accuracy on cross-species gene annotation for broad 
+phylogenetic groups (land plants and vertebrates, respectively).
+Within these groups, this eliminates the dependency on retraining and the
+expertise and data required therefore. Production of comparable, single-method
+annotations for broad groups has the potential to greatly facilitate downstream
+analyses which suffer when data is heterogeneous \citep{}.
 
@@
 in a wider range of genomes with decent predictions. For this, more computational 
-resources would be required.
+resources would be required. Simply training with a similar amount of data, 
+but for a different group (e.g. invertebrates of fungi) could also be used to 
+increase the functional predictive range.
 
@@
 
+Finally, developent of a post-processing method
+to go from base pair wise predictions to integrated predictions for whole 
+transcripts at each loci could both further improve performance and would
+greatly increase real world applications.
+
 \section*{Acknowledgements}
@@
 We are very grateful for the support of...
+
+HHU HPC-team, deNBI, Ebenhöh lab for computational (particularly GPU) resources.
 

### 2020-05-14 paper/main.tex
attempt to add figs (which themselves still need a touchup)

@@
 \usepackage{url}
-
+\usepackage{placeins}
 \begin{document}
@@
 
-\abstract{\textbf{Motivation:} Current state-of-the-art tools for the \textit{de novo} annotation of 
+\abstract{
+\textbf{Motivation:} Current state-of-the-art tools for the \textit{de novo} annotation of 
 genes in eukaryotic genomes have to be specifically fitted for each species and still often 
@@
 \textbf{Contact:} \href{andreas.weber@hhu.de}{andreas.weber@hhu.de}\\  
-
 \textbf{Supplementary information:} Supplementary data are available at 
-\textit{Bioinformatics} online.}
+\textit{Bioinformatics} online.
+}
 
@@
 
+\begin{figure*}[!h]
+\label{fig:cov_example_main}
+\centerline{\includegraphics[width=\textwidth]{images/cov_examples/cov_example_main}}
+\caption{high low accuracty by high low RNAseq support for ref}
+\end{figure*}
+
 As an independent evaluation and to get a better qualitative understanding of
@@
 annotation of intron, and neither coverage nor spliced coverage is expected 
-for intergenic annotations. Looking at selected subsequences \ref{fig:examples} % and supplemental figures.
+for intergenic annotations. Looking at selected subsequences (Figure \ref{fig:cov_example_main}) % and supplemental figures.
 we identified cases where RNAseq supported both the annotations of the reference 
@@
 
+\begin{figure*}[!h]
+\label{fig:average_vs_augustus}
+\includegraphics[width=5.6cm]{images/cov_examples/average_vs_augustus}  % should be 8.6 when floats are sorted out right
+\caption{A caption}
+\end{figure*}
+
 Coverage and Spliced coverage were broken down by the confusion matrix (of both
-Helixer vs Augustus and Helixer vs the reference) for all genomes \ref{}. % H vs A main, H vs R supp
+Helixer vs Augustus and Helixer vs the reference) for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
 Where tools were in agreement (far left), coverage and spliced coverage closely matched
@@
 greatly increase real world applications.
+\FloatBarrier
 
@@
 
-HHU HPC-team, deNBI, Ebenhöh lab for computational (particularly GPU) resources.
+HHU HPC-team, deNBI, Ebenh{\"o}h lab for computational (particularly GPU) resources.
 

### 2020-05-15 paper/main.tex
and now with figuresgit add main.tex :-)

@@
 \centerline{\includegraphics[width=\textwidth]{images/cov_examples/cov_example_main}}
-\caption{high low accuracty by high low RNAseq support for ref}
+\caption{Four example helixer predictions in the context of RNAseq data, the reference,
+and Augustus' prediction for {\it M. esculenta}. The examples were chosen so
+that a) Helixer had high accuracy against the reference and the reference
+was supported by the RNAseq data, b) Helixer had high accuracy but the 
+reference was not supported by the RNAseq data, c) Helixer had low accuracy
+and the reference was supported by the RNAseq data, and d) helixer had low
+accuracy but the reference was not supported by RNAseq data. Plotability was
+a major secondary consideration. Each subplot shows from top to
+bottom i) the natural log of the coverage ("cov", solid) and spliced coverage 
+("sc", dotted) + 1, ii) the reference annotation in matrix form, iii) 
+Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
+and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+Helixer emits a probability from 0-1 represented via gray-scale. "Ntrn" stand
+for intron, and "IG" stands for intergenic.
+}
 \end{figure*}
@@
 \label{fig:average_vs_augustus}
-\includegraphics[width=5.6cm]{images/cov_examples/average_vs_augustus}  % should be 8.6 when floats are sorted out right
-\caption{A caption}
+\includegraphics[width=8.6cm]{images/cov_examples/average_vs_augustus}  % should be 8.6 when floats are sorted out right
+\caption{Fraction of bp with color-indicated a) coverage and b) spliced coverage of genomic positions 
+broken down by the confusion matrix of Augustus and Helixer's predictions. Categories 
+are only displayed if they can be meaningfully compared by examining a) coverage or b) spliced
+coverage. The displayed fractions are the averages of the individual fractions for the 
+six RNAseq-evaluation species. The left-most bars show cases where the two tools agree, 
+while the remaining bars show paired conflicts. "Ntrn" stand
+for intron, and "IG" stands for intergenic.}
 \end{figure*}

### 2020-05-16 paper/main.tex
added supplemental RNAseq eval figs

@@
 a major secondary consideration. Each subplot shows from top to
-bottom i) the natural log of the coverage ("cov", solid) and spliced coverage 
-("sc", dotted) + 1, ii) the reference annotation in matrix form, iii) 
+bottom i) the natural log of the coverage (``cov", solid) and spliced coverage 
+(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii) 
 Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
 and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
-Helixer emits a probability from 0-1 represented via gray-scale. "Ntrn" stand
-for intron, and "IG" stands for intergenic.
+Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
 }
@@
 six RNAseq-evaluation species. The left-most bars show cases where the two tools agree, 
-while the remaining bars show paired conflicts. "Ntrn" stand
-for intron, and "IG" stands for intergenic.}
+while the remaining bars show paired conflicts. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.}
 \end{figure*}

### 2020-05-16 paper/supplements.tex
added supplemental RNAseq eval figs

@@
 
+\section{Evaluation against RNAseq}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S7}
+\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_002}}
+\caption{
+Four additional example helixer predictions in the context of RNAseq data, the reference,
+and Augustus' prediction for {\it M. esculenta}. The examples were chosen so
+that a) Helixer had high accuracy against the reference and the reference
+was supported by the RNAseq data, b) Helixer had high accuracy but the
+reference was not supported by the RNAseq data, c) Helixer had low accuracy
+and the reference was supported by the RNAseq data, and d) helixer had low
+accuracy but the reference was not supported by RNAseq data. Plotability was
+a major secondary consideration. Each subplot shows from top to
+bottom i) the natural log of the coverage (``cov", solid) and spliced coverage
+(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii)
+Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
+and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S8}
+\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_003}}
+\caption{
+Four additional example helixer predictions in the context of RNAseq data, the reference,
+and Augustus' prediction for {\it M. esculenta}. The examples were chosen so
+that a) Helixer had high accuracy against the reference and the reference
+was supported by the RNAseq data, b) Helixer had high accuracy but the
+reference was not supported by the RNAseq data, c) Helixer had low accuracy
+and the reference was supported by the RNAseq data, and d) helixer had low
+accuracy but the reference was not supported by RNAseq data. Plotability was
+a major secondary consideration. Each subplot shows from top to
+bottom i) the natural log of the coverage (``cov", solid) and spliced coverage
+(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii)
+Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
+and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S9}
+\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_UTRs_plants}}
+\caption{
+One example of helixer's gradual UTR to intergenic transition predictions 
+from each RNAseq evaluation plant species in the context of RNAseq data, the reference,
+and Augustus' prediction. The species are
+a) {\it M. esculenta}, b) {\it M. truncatula} and c) {\it T. cacao}. Each subplot shows from top to
+bottom i) the natural log of the coverage (``cov", solid) and spliced coverage
+(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii)
+Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
+and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S10}
+\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_UTRs_animals}}
+\caption{
+One example of helixer's gradual UTR to intergenic transition predictions 
+from each RNAseq evaluation animal species in the context of RNAseq data, the reference,
+and Augustus' prediction. The species are
+a) {\it P. anubis}, b) {\it E. caballus} and c) {\it P. marinus}. Each subplot shows from top to
+bottom i) the natural log of the coverage (``cov", solid) and spliced coverage
+(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii)
+Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
+and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S11}
+\centerline{\includegraphics[]{images/cov_examples/average_vs_ref}}
+\caption{
+Fraction of bp with color-indicated a) coverage and b) spliced coverage of genomic positions
+broken down by the confusion matrix of the reference and Helixer's predictions. Categories
+are only displayed if they can be meaningfully compared by examining a) coverage or b) spliced
+coverage. The displayed fractions are the averages of the individual fractions for the
+six RNAseq-evaluation species. The left-most bars show cases where the two tools agree,
+while the remaining bars show paired conflicts. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S12}
+\centerline{\includegraphics[]{images/cov_examples/each_sp_vs_augustus}}
+\caption{
+Fraction of bp with color-indicated a, c, e, g, i, k) coverage and b, d, f, h, j, l) spliced coverage of genomic positions
+broken down by the confusion matrix of Augustus' and Helixer's predictions. Categories
+are only displayed if they can be meaningfully compared by examining coverage or spliced
+coverage. The rows show the individual species: a, b) {\it M. esculenta}, c, d) {\it M. truncatula}, 
+e, f) {\it T. cacao}, g, h) {\it P. anubis}, i, j) {\it E. caballus} and k, l) {\it P. marinus} . 
+The left-most bars show cases where the two tools agree,
+while the remaining bars show paired conflicts. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+\renewcommand\thefigure{S13}
+\centerline{\includegraphics[]{images/cov_examples/each_sp_vs_ref}}
+\caption{
+Fraction of bp with color-indicated a, c, e, g, i, k) coverage and b, d, f, h, j, l) spliced coverage of genomic positions
+broken down by the confusion matrix of the reference and Helixer's predictions. Categories
+are only displayed if they can be meaningfully compared by examining coverage or spliced
+coverage. The rows show the individual species: a, b) {\it M. esculenta}, c, d) {\it M. truncatula}, 
+e, f) {\it T. cacao}, g, h) {\it P. anubis}, i, j) {\it E. caballus} and k, l) {\it P. marinus} . 
+The left-most bars show cases where the two tools agree,
+while the remaining bars show paired conflicts. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
 \renewcommand\refname{Supplemental References}

### 2020-05-19 paper/main.tex
some refs, some conventions, moving towards a hacked semi-dynamic ref for supplemental things

@@
 \usepackage{placeins}
+
+% extra counters to reference items in supplemental
+% note that thes are not directly linked, so the order of the environments 
+% should match where they occur in the supplement
+\newcounter{supfig}
+\newenvironment{supfig}[1][]{\refstepcounter{supfig}}
+
+\newcounter{suptab}
+\newenvironment{suptab}[1][]{\refstepcounter{suptab}}
+
+\newcounter{supsec}
+\newenvironment{supsec}[1][]{\refstepcounter{supsec}}
+
 \begin{document}
@@
 species from the JGI Phytosome 13 database \citep{goodstein2012phytozome}; exact 
-genomes used are list here \ref{}. % add a supplemental table with names and versions
+genomes used are list here \ref{suptab:source_data}. % add a supplemental table with names and versions
 Data was downloaded on 29/03/2019 and 15/10/2019 for plants and animals respectively. 
@@
 performance, we needed a way to evaluate genome and annotation quality.
-Metadata was generated for each genome, including: basic length and composition statistics \citep{}; % quast
-k-mer content \citep{}; % jellyfish
-conserved gene content of the genome, transcriptome, and proteome \citep{}; % BUSCO, files generated with gffread
+Metadata was generated for each genome, including: basic length and 
+composition statistics \citep{gurevich2013quast}; k-mer content \citep{marccais2011fast};
+conserved gene content of the genome, transcriptome, and proteome (\cite{simao2015busco};
+file conversion with gffread from \cite{trapnell2012differential}; % BUSCO, files generated with gffread
 and finally simple counts of genic features from the gff files. In particular, 
@@
 
-\begin{table}[!t]
+\begin{table}[!htb]
+\label{tab:arrays}
+\processtable{Data arrays generated and used by Helixer}{
+\begin{tabular}{@{}ll@{}}
+\toprule Name & Information \\
+\midrule
+Input & Genomic sequence in the 4-dimensional IUPAC encoding\\
+	  & (one hot encoding for non-ambiguous bases) \\
+Output & Labels in a 4-dimensional one hot encoding representing \\
+      & the classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}\\
+Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
+\botrule
+\end{tabular}}{{\it Note:} the encoding does not differentiate between introns
+in coding and non-coding regions.}
+\end{table}
+
+During data generation, we query for the transcript with the longest protein of each gene and 
+disregard FASTA sequences that have no functional annotation as it is ambiguous
+whether such sequences contain no genes or were simply not annotated in the reference. GeenuFF checks the 
+genomic annotations for potential errors during import and is able to mark those 
+areas. We use this information to effectively blank out those bases during training 
+by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
+of masked bases lie in the intergenic region as the most prevelant error is a 
+missing UTR and GeenuFF marks a potentially large intergenic region for it. Table 
+\ref{tab:statistics_all} shows statistics about the masking.
+
+For the training itself, we divide each continous genomic sequence into 20,000 bp 
+long subsequences, for which one-hot vectors of the base pairs and annotations 
+are generated and respectively used as input and label for the neural network 
+together with the sample weights. We append zero padding if the sequences (or
+sequence ends) are shorter than 20,000 bp. 
+If a subsequence is fully marked as erroneous, we exclude it from all analyses.
+
+\begin{table}[!htb]
 \processtable{Data group statistics for all data\label{tab:statistics_all}} {
@@
 \end{table}
-
-During data generation, we query for the transcript with the longest protein of each gene and 
-disregard FASTA sequences that have no functional annotation as it is ambiguous
-whether such sequences contain no genes or were simply not annotated in the reference. GeenuFF checks the 
-genomic annotations for potential errors during import and is able to mark those 
-areas. We use this information to effectively blank out those bases during training 
-by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
-of masked bases lie in the intergenic region as the most prevelant error is a 
-missing UTR and GeenuFF marks a potentially large intergenic region for it. Table 
-\ref{tab:statistics_all} shows statistics about the masking.
-
-For the training itself, we divide each continous genomic sequence into 20,000 bp 
-long subsequences, for which one-hot vectors of the base pairs and annotations 
-are generated and respectively used as input and label for the neural network 
-together with the sample weights. We append zero padding if the sequences (or
-sequence ends) are shorter than 20,000 bp. 
-If a subsequence is fully marked as erroneous, we exclude it from all analyses.
 
@@
 
-\begin{table}[!t]
-\processtable{Data arrays generated and used by Helixer\label{tab:arrays}} {
-\begin{tabular}{@{}ll@{}}
-\toprule Name & Information \\
-\midrule
-Input & Genomic sequence in the 4-dimensional IUPAC encoding\\
-	  & (one hot encoding for non-ambiguous bases) \\
-Output & Labels in a 4-dimensional one hot encoding representing \\
-      & the classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}\\
-Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
-\botrule
-\end{tabular}}{{\it Note:} An intron inside a UTR region is encoded as intron and 
-can be inferred from the prediction order.}
-\end{table}
-
 \end{methods}
@@
 \caption{Fraction of bp with color-indicated a) coverage and b) spliced coverage of genomic positions 
-broken down by the confusion matrix of Augustus and Helixer's predictions. Categories 
+broken down by the confusion matrix of Augustus' and Helixer's predictions. Categories 
 are only displayed if they can be meaningfully compared by examining a) coverage or b) spliced
@@
 \bibliography{literature}
+% place holders for supplemental things
+% ori version was grepped out of supplements, then
+% :'<,'>s/^/\\begin{supfig}/g
+% :'<,'>s/$/\\end{supfig}/g
+% and similar for tables/sections
+% tables
+\begin{suptab}\label{suptab:generalization_animals}\end{suptab}
+\begin{suptab}\label{suptab:generalization_plants}\end{suptab}
+\begin{suptab}\label{suptab:params}\end{suptab}
+\begin{suptab}\label{suptab:classweights}\end{suptab}
+\begin{suptab}\label{suptab:prediction_lengths}\end{suptab}
+\begin{suptab}\label{suptab:genome_stats}\end{suptab}
+\begin{suptab}\label{suptab:dcnn_xval}\end{suptab}
+\begin{suptab}\label{suptab:dcnn_params}\end{suptab}
+% figures
+\begin{supfig}\label{supfig:phylogeny_f1_animals}\end{supfig}
+\begin{supfig}\label{supfig:phylogeny_differences_animals}\end{supfig}
+\begin{supfig}\label{supfig:phylogeny_f1_plants}\end{supfig}
+\begin{supfig}\label{supfig:phylogeny_differences_plants}\end{supfig}
+\begin{supfig}\label{supfig:overlapping_animals}\end{supfig}
+\begin{supfig}\label{supfig:overlapping_plants}\end{supfig}
+\begin{supfig}\label{supfig:cov_example_02}\end{supfig}
+\begin{supfig}\label{supfig:cov_example_02}\end{supfig}
+\begin{supfig}\label{supfig:UTRs_animals}\end{supfig}
+\begin{supfig}\label{supfig:UTRs_plants}\end{supfig}
+\begin{supfig}\label{supfig:average_vs_ref}\end{supfig}
+\begin{supfig}\label{supfig:each_sp_vs_augustus}\end{supfig}
+\begin{supfig}\label{supfig:each_sp_vs_ref}\end{supfig}
+% sections
+\begin{supsec}\label{sec:longer}\end{supsec}
+\begin{supsec}\label{sec:overlapping}\end{supsec}
+\begin{supsec}\label{sec:training_data}\end{supsec}
 

### 2020-05-20 paper/main.tex
more semi-dynamic linking and refs

@@
 generally intractable. Instead, researchers can use pipelines such as 
-Maker \citep{}, PASA \citep{} or those offered by genomic database providers like NCBI 
+Maker \citep{cantarel2008maker}, PASA \citep{haas2003improving} or those 
+offered by genomic database providers like NCBI 
 \citep{thibaud2013eukaryotic} and Ensembl \citep{aken2016ensembl}. 
@@
 genomes used are list here \ref{suptab:source_data}. % add a supplemental table with names and versions
-Data was downloaded on 29/03/2019 and 15/10/2019 for plants and animals respectively. 
+Data was downloaded on 15/10/2019 and 29/03/2019 for animals and plants respectively. 
 Both data groups were used seperately during training as well as during evaluation.
@@
 The selection of training genomes also requires balancing multiple conflicting 
-tradeoffs. On the one hand, we do want to train with as much data as possible, but 
-not all data has a good enough quality to enable a powerful generalization. It is 
-also very desirable to have a broad spread phylogenetically as well as in genome 
-size and average gene length as the model is increasingly unlikely to generalize 
+tradeoffs. On the one hand, we want to train with as much data as possible, but 
+on the other
+not all data has sufficient quality to enable a powerful generalization. It is 
+also very desirable to have diverse training genomes with a broad phylogenetic spread, 
+and a variety of genome sizes and average gene length as the model is increasingly unlikely to generalize 
 well beyond the borders given by the training data. However, it also may be difficult 
 for the model learn and generalize, if the genomic patterns inside the data are 
-too different from each other. It was also important for us to limit the size of
+too different from each other. Practically, it was important for us to limit the size of
 our training genomes set to be small enough that enough that we could get experimental 
@@
 not take UTR predictions into account. This is done for comparability with AUGUSTUS 
-and further explained in Section \ref{sec:augustus}. Both metrics work by summing up 
+as further explained in Section \ref{sec:augustus}. Both metrics work by summing up 
 True Positives (TP), False Positives (FP) and False Negatives (FN) of each 
@@
 far the easiest to predict. (See Table \ref{tab:statistics_all} for the class 
-distribution and Tables S1 and S2 for a more in-depth report our model performance 
+distribution and Tables S\ref{suptab:generalization_animals} 
+and S\ref{suptab:generalization_plants} for a more in-depth report our model performance 
 including the intergenic class). In the case of the Genic F1, this means that 
 effectively only the TP of the intergenic class are disregarded. The metrics 
-essentially provide a weighted mean of the performances in the considered classes. %?? hmmm, but this is not a weighted mean across classes is it? Do introns not weigh more heavily?
+essentially provide a weighted mean of the performances in the considered classes,
+with weights proportional to class frequency.
 
@@
 the situation. We used NNI \citep{nni2019} to facilitate the search. All relevant 
-hyperparameters can be found in Table S3 and S4 or in the Helixer source code 
-repository.
+hyperparameters can be found in Table S\ref{suptab:params} and 
+S\ref{suptab:classweights} or in the Helixer source code repository.
 
@@
 widely used when working with DNA data as input. Details of the neural architecture 
-search are given in Section 6 of the Supplements.
+search are given in Section \ref{sec:dcnn} of the Supplements.
 
@@
 proportional to the length of the sequence input. For more implementation details 
-on this see Section 3 of the Supplements or the source code.
+on this see Section \ref{sec:longer} of the Supplements or the source code.
 
@@
 starting points. The different overlapping sequences were combined by averaging 
-the individual softmax values of each base. The figures in Section 4 of the 
+the individual softmax values of each base. The figures in Section \ref{sec:overlapping} of the 
 Supplements show the effect of overlapping for each genome, ordered by N75. We found 
@@
 each species. For this, we performed 4 separate animal and plant training runs and 
-selected two models each, specifically the model with the best precision and the  % please check / epoch model / version something
-other with the best recall. The model with the best Genic F1 was always best in one of 
+selected two checkpoints each, specifically the checkpoint from the epoch with the best precision and the  % please check / epoch model / version something
+other from the epoch with the best recall. The model checkpoint with the best Genic F1 was always best in one of 
 those. This was done to increase the diversity of the model ensemble. As with the 
@@
 Due to the practical time constraints for retraining and running, we decided to compare our Helixer
-models only to one existing {\it de novo} gene caller, the popular gene caller, Augustus.
+models only to one existing {\it de novo} gene caller, the popular gene caller, Augustus \citep{stanke2003gene}.
 Typical usage requires retraining Augustus for each species, with the exception of a
-few lucky cases where a model for a close relative is already available. To scale this for 
+few lucky cases where a model for a sufficiently close relative is already available. To scale this for 
 the large plant and animal datasets, we used protein-homology to create a training set and
@@
 data, are expected to contain errors; but here we were particularly concerned because
-the {\it de novo} gene caller used by the references was Augustus for many species (...)
-% todo list and cite examples (above and below)
-and a fundamentally similar HMM-based tool for many more (...), which would potentially bias 
-reference based comparisons of Helixer's and Augustus' performance.
+the {\it de novo} gene caller used by the references was Augustus for many species 
+(e.g. {\it Arabidopsis lyrata}, \cite{rawat2015improving}; {\it Asparagus officinalis}, 
+\cite{harkess2017asparagus}; and {\it Chenopodium quinoa}, \cite{jarvis2017genome})  
+and a different HMM-based tool for many more (e.g. {\it Citrus clementina}, \cite{wu2018genomics}; 
+{\it Capsella grandiflora}, \cite{slotte2013capsella};,...), which would potentially bias 
+reference based comparisons of Helixer's and Augustus' performance. % examples chosen alphabetically and skipping ons where I wasn't sure from the methods
 
@@
 Each sample was quality controlled with FastQC \citep{}, before and after trimming
-with Trimmomatic \citep{}, and then mapped to the genome with Hisat2 \citep{}. 
-Samtools \citep{} was used for the sorting and indexing that was necessary both for
-RNAseq mapped metric collection with PicardTools \citep{} and custom downstream 
-processing below. MultiQC \citep{} was used to summarize the results. The above was 
+with Trimmomatic \citep{bolger2014trimmomatic}, and then mapped to the genome with Hisat2 \citep{kim2019graph}. 
+Samtools \citep{li2009sequence} was used for the sorting and indexing that was necessary both for
+RNAseq mapped metric collection with PicardTools \citep{Picard2019toolkit} and custom downstream 
+processing below. MultiQC \citep{ewels2016multiqc} was used to summarize the results. The above was 
 automated and the results visualized with the code here (... make available and link to RNAseq stuff).
@@
 % sections
+\begin{supsec}\label{sec:generalization}\end{supsec}
+\begin{supsec}\label{sec:hyperparameters}\end{supsec}
 \begin{supsec}\label{sec:longer}\end{supsec}
@@
 \begin{supsec}\label{sec:training_data}\end{supsec}
+\begin{supsec}\label{sec:dcnn}\end{supsec}
+\begin{supsec}\label{sec:rnaseq_eval}\end{supsec}
 

### 2020-05-22 paper/main.tex
trying to explain sp selection

@@
 {\it de novo} gene predictions with RNAseq and homology data. The references, like any
-data, are expected to contain errors; but here we were particularly concerned because
-the {\it de novo} gene caller used by the references was Augustus for many species 
-(e.g. {\it Arabidopsis lyrata}, \cite{rawat2015improving}; {\it Asparagus officinalis}, 
+data, are expected to contain errors so we chose to use RNAseq data for independent evaluation.
+In plants there is the additional concern that the references may share biases with
+Augustus or HMMs as Augustus was used as the {\it de novo} gene caller for the reference of
+many plant species (e.g. {\it Arabidopsis lyrata}, \cite{rawat2015improving}; 
+{\it Asparagus officinalis}, 
 \cite{harkess2017asparagus}; and {\it Chenopodium quinoa}, \cite{jarvis2017genome})  
-and a different HMM-based tool for many more (e.g. {\it Citrus clementina}, \cite{wu2018genomics}; 
-{\it Capsella grandiflora}, \cite{slotte2013capsella};,...), which would potentially bias 
-reference based comparisons of Helixer's and Augustus' performance. % examples chosen alphabetically and skipping ons where I wasn't sure from the methods
+and other HMM-based tools for many more (e.g. {\it Citrus clementina}, \cite{wu2018genomics}; and
+{\it Capsella grandiflora}, \cite{slotte2013capsella}). The Ensemble animal dataset by and large
+used the Ensembl annotation pipeline, which includes no HMM-based tool.
 
-We therefore downloaded and processed public RNAseq data to obtain an independent option 
+We downloaded and processed public RNAseq data to obtain an independent option 
 for evaluating model performance. We selected three each of plant and animal genomes for
@@
 automated and the results visualized with the code here (... make available and link to RNAseq stuff).
-Samples were filtered to those that had {\it relatively} high mapping rates, particularly to
-exonic regions; that had low 3' bias, normal FastQC and Trimmomatic stats, and that
+Samples were filtered to those that had {\it relatively} high mapping rates, high mapping to
+exonic regions relative to other regions; that had low 3' bias, normal FastQC and 
+Trimmomatic stats, and that
 were confirmed to be stranded with 2nd strand as sense strand. If more remained, seven

### 2020-05-23 paper/main.tex
refs n cites

@@
 conserved gene content of the genome, transcriptome, and proteome (\cite{simao2015busco};
-file conversion with gffread from \cite{trapnell2012differential}; % BUSCO, files generated with gffread
+file conversion with gffread from \cite{trapnell2012differential}; 
 and finally simple counts of genic features from the gff files. In particular, 
@@
 and poor ({\it Theobroma cacao} and {\it Petromyzon marinus})
-performance compared to Augustus \ref{} % supp phylogenies maybe?
-within our generalizable range (i.e. excluding algae and
+performance compared to Augustus \ref{supfig:rna_select} % subgenic F1 sorted by Helixer new plots
+within our generalizable range (i.e. excluding the outgroups algae and
 invertebrates). Selections were further constrained by the availability of stranded RNAseq
@@
 samples were chosen for further processing, as listed here \ref{}. % todo make a supp table
-Each sample was quality controlled with FastQC \citep{}, before and after trimming
+Each sample was quality controlled with FastQC (www.bioinformatics.babraham.ac.uk/projects/fastqc/), before and after trimming
 with Trimmomatic \citep{bolger2014trimmomatic}, and then mapped to the genome with Hisat2 \citep{kim2019graph}. 
@@
 \caption{Main overall results and other investigations. The top row shows the animal
- data; the bottom row is concerned with the plants. Yellow dots are for AUGUSTUS and 
+ data; the bottom row is concerned with the plants. Orange dots are for AUGUSTUS and 
 the blue ones are for Helixer. {\bf a)} Swarm - and boxplots showing the Subgenic F1 
@@
 same without overlapping. Each value is the average performance of a 200 basepair 
-long subsection by one of the models of the final ensemble. See Section 4 for the 
-Supplements shows the effects of overlapping for each individual species.}
+long subsection by one of the models of the final ensemble. 
+See Section \ref{supsec:overlapping} of the 
+Supplements for the effects of overlapping in individual species.}
 \end{figure*}
@@
 genomes. This is the case for both animals and plants and is visualized by the 
-Figures S1 and S3. We do not, for example, consistently predict very well on the 
-algae as well as the non-avian reptiles. While one of the the algae were included 
+Figures S\ref{supfig:phylogeny_f1_animals} and S\ref{supfig:phylogeny_f1_plants}. 
+We do not, for example, consistently predict very well on the 
+algae nor on the non-avian reptiles. While one of the the algae were included 
 in our training genomes it accounted for a tiny proportion of the total training 
@@
 especially strong for mammals, which tend to have big genomes with very long genes 
-as well as the largest plants. Figures S2 and S4 display the comparison to AUGUSTUS 
+as well as the largest plants. Figures S\ref{supfig:phylogeny_differences_animals} 
+and S\ref{supfig:phylogeny_differences_plants} display the comparison to AUGUSTUS 
 by phylogenetic position.
@@
 we identified cases where RNAseq supported both the annotations of the reference 
-and Helixer (a), neither (b), the reference, but not Helixer (c),
-and Helixer, but not the reference (d).
+and Helixer a), neither b), the reference, but not Helixer c),
+and Helixer, but not the reference d).
 
@@
 fundamentally harder problem (there is no conserved motif at the site as is observed
-for splice sites and start/stop codons), lack a single precise biological site \citep{},
-or noise in the reference, which we observed relative to the RNAseq data \ref{}. % supplemental UTR figures, 1/species
+for splice sites and start/stop codons), lack of a precise one 
+base pair biological site \citep{carninci2006genome},  % todo, citation for termination site as well
+or noise in the reference, which we observed relative to the RNAseq data 
+(Figures S\ref{supfig:UTRs_animals}-\ref{supfig:UTRs_plants}). % supplemental UTR figures, 1/species
 
@@
 annotation, it assigned a low but non-trivial probability to the RNAseq-supported 
-exon/intron pattern (b and c,hSlA2 supp). However, in the extreme,
+exon/intron pattern (Figures \ref{fig:cov_example_main}b-c,S\ref{supfig:cov_example_02}c). However, in the extreme,
 there are cases where the Helixer model exibits substantial indecision or confusion 
 and shifts gradually between classes with no single class receiving a high probability 
-for extended stretches (hSlA2, hSlA3, lSlA3). Notably, in one of the examples ()
-the RNAseq shows evidence of alternative splicing; and in another () Helixer's
+for extended stretches (Figures S\ref{supfig:cov_example_02}c, S\ref{supfig:cov_example_02}c-d). 
+Notably, in one of the examples (S\ref{supfig:cov_example_02}d))
+the RNAseq shows evidence of alternative splicing; 
+and in another (S\ref{supfig:cov_example_02}c) Helixer's
 prediction falls between that of the reference and Augustus.
@@
 \begin{supfig}\label{supfig:cov_example_02}\end{supfig}
-\begin{supfig}\label{supfig:cov_example_02}\end{supfig}
+\begin{supfig}\label{supfig:cov_example_03}\end{supfig}
 \begin{supfig}\label{supfig:UTRs_animals}\end{supfig}

### 2020-05-24 paper/main.tex
ahha, the key term is microheterogeneity in poly(A) sites... also more fig refs

@@
 for splice sites and start/stop codons), lack of a precise one 
-base pair biological site \citep{carninci2006genome},  % todo, citation for termination site as well
+base pair biological site \citep{carninci2006genome,hon2013quantification}
 or noise in the reference, which we observed relative to the RNAseq data 
@@
 the RNAseq data than those of Augustus. Specifically, for base pairs where one tool said CDS and
-the other intergenic or intron \ref{}, there was more coverage when 
-Helixer said CDS. Similarly, for base pairs where one tool said intron and the other 
-CDS or intergenic, there was more spliced coverage when Helixer said intron \ref{}. 
+the other intergenic or intron, there was more coverage when 
+Helixer said CDS (Figure \ref{fig:average_vs_augustus}a). Similarly, for base pairs where one tool said intron and the other 
+CDS or intergenic, there was more spliced coverage when Helixer said intron (Figure \ref{fig:average_vs_augustus}b). 
 These patterns were consistent in direction but varied in magnitude in the individual species,
 with the expception of {\it P. marinus} where Helixer and Augustus performed comparably
-at differentiating CDS and introns.
+at differentiating CDS and introns (Figure S\ref{supfig:each_sp_vs_augustus}).
 
@@
 however helixer's predictions received slightly less support than the reference 
-when differentiating CDS and UTR from introns. Performance varies between individual
+when differentiating CDS and UTR from introns (Figure S\ref{supfig:average_vs_ref}). Performance varies between individual
 species, from {\it P. marinus}, where the reference receives more support
 in every conflict, to {\it T. cacao}, where Helixer models receive equivalent
-or more support in every conflict. Interestingly, these are the two species that
+or more support in every conflict (Figure S\ref{supfig:each_sp_vs_ref}). Interestingly, these are the two species that
 were selected because they had (relative to the usual difference) low Subgenic F1 compared to 

### 2020-05-25 paper/main.tex
dynamic supplement and stuff

@@
 \usepackage{url}
-\usepackage{placeins}
 
@@
 \subsection{Data Pre-processing}
-We first pre-process and store the raw genomic information by using GeenuFF 
-\citep{denton2019}. GeenuFF is a tool for checking and exploring genomic data, that 
+We first pre-process and store the raw genomic information by using GeenuFF (https://https:
+//github.com/weberlab-hhu/GeenuFF). GeenuFF is a tool for checking and exploring genomic data, that 
 stores all information inside a SQL database. Training and evaluation-ready data is 
@@
 widely used when working with DNA data as input. Details of the neural architecture 
-search are given in Section \ref{sec:dcnn} of the Supplements.
+search are given in Section \ref{ssec:dcnn} of the Supplements.
 
@@
 proportional to the length of the sequence input. For more implementation details 
-on this see Section \ref{sec:longer} of the Supplements or the source code.
+on this see Section \ref{ssec:longer} of the Supplements or the source code.
 
@@
 starting points. The different overlapping sequences were combined by averaging 
-the individual softmax values of each base. The figures in Section \ref{sec:overlapping} of the 
+the individual softmax values of each base. The figures in Section \ref{ssec:overlapping} of the 
 Supplements show the effect of overlapping for each genome, ordered by N75. We found 
@@
 Due to the practical time constraints for retraining and running, we decided to compare our Helixer
-models only to one existing {\it de novo} gene caller, the popular gene caller, Augustus \citep{stanke2003gene}.
-Typical usage requires retraining Augustus for each species, with the exception of a
+models only to one existing {\it de novo} gene caller, the popular gene caller, AUGUSTUS \citep{stanke2003gene}.
+Typical usage requires retraining AUGUSTUS for each species, with the exception of a
 few lucky cases where a model for a sufficiently close relative is already available. To scale this for 
@@
 the training was carried out with the subsetted training file and using the following scripts provided
-by Augustus. An untrained model was setup with "new\_species.pl", and the model was fit by running
+by AUGUSTUS. An untrained model was setup with "new\_species.pl", and the model was fit by running
 "etraining" before and after the major hyper-parameter optimization with "optimize\_augustus.pl".
 Using the trained model for each species, we ran the main prediction ("augustus") with ``--UTR=off"
-and ``--gff3=on". The gffs produced by Augustus were imported into GeenuFF and exported as h5
+and ``--gff3=on". The gffs produced by AUGUSTUS were imported into GeenuFF and exported as h5
 files in the same manor as the reference, allowing direct comparison with reference and 
 Helixer predictions. While tenable for some 237 species, this method allowed neither training
-nor prediction of UTR regions with Augustus, so the metric Subgenic F1, which excludes UTR,
-was used for all comparisons between Augustus and Helixer.
+nor prediction of UTR regions with AUGUSTUS, so the metric Subgenic F1, which excludes UTR,
+was used for all comparisons between AUGUSTUS and Helixer.
 
@@
 In plants there is the additional concern that the references may share biases with
-Augustus or HMMs as Augustus was used as the {\it de novo} gene caller for the reference of
-many plant species (e.g. {\it Arabidopsis lyrata}, \cite{rawat2015improving}; 
-{\it Asparagus officinalis}, 
-\cite{harkess2017asparagus}; and {\it Chenopodium quinoa}, \cite{jarvis2017genome})  
-and other HMM-based tools for many more (e.g. {\it Citrus clementina}, \cite{wu2018genomics}; and
-{\it Capsella grandiflora}, \cite{slotte2013capsella}). The Ensemble animal dataset by and large
+AUGUSTUS or HMMs as AUGUSTUS was used as the {\it de novo} gene caller for the reference of
+many plant species % (e.g. {\it Arabidopsis lyrata}, \cite{rawat2015improving}; 
+%{\it Asparagus officinalis}, 
+%\cite{harkess2017asparagus}; and {\it Chenopodium quinoa}, \cite{jarvis2017genome})  
+and other HMM-based tools for many more %(e.g. {\it Citrus clementina}, \cite{wu2018genomics}; and
+%{\it Capsella grandiflora}, \cite{slotte2013capsella}). 
+The Ensemble animal dataset by and large
 used the Ensembl annotation pipeline, which includes no HMM-based tool.
@@
 and poor ({\it Theobroma cacao} and {\it Petromyzon marinus})
-performance compared to Augustus \ref{supfig:rna_select} % subgenic F1 sorted by Helixer new plots
+performance compared to AUGUSTUS (Figure S\ref{supfig:rna_select_animals}-\ref{supfig:rna_select_plants}) % subgenic F1 sorted by Helixer new plots
 within our generalizable range (i.e. excluding the outgroups algae and
@@
 For each of these six species the following search was performed on Sequence Read Archive
-'(("<species name>"[Organism] OR <species name>[All Fields]) AND stranded[All Fields]) 
+`(("<species name>"[Organism] OR <species name>[All Fields]) AND stranded[All Fields]) 
 AND ("biomol rna"[Properties] AND "library layout paired"[Properties])'. In so far as 
-more than 50 samples were identified, every Nth sample was selected so that about 50
-samples were chosen for further processing, as listed here \ref{}. % todo make a supp table
+more than 50 samples were identified, every Nth sample was selected so that under 50
+samples were chosen for further processing (Table S\ref{suptab:rnaseq_samples}).
 Each sample was quality controlled with FastQC (www.bioinformatics.babraham.ac.uk/projects/fastqc/), before and after trimming
@@
 Samtools \citep{li2009sequence} was used for the sorting and indexing that was necessary both for
-RNAseq mapped metric collection with PicardTools \citep{Picard2019toolkit} and custom downstream 
+RNAseq mapped metric collection with PicardTools (http://broadinstitute.github.io/picard/) and custom downstream 
 processing below. MultiQC \citep{ewels2016multiqc} was used to summarize the results. The above was 
@@
 long subsection by one of the models of the final ensemble. 
-See Section \ref{supsec:overlapping} of the 
+See Section \ref{ssec:overlapping} of the 
 Supplements for the effects of overlapping in individual species.}
@@
 \caption{Four example helixer predictions in the context of RNAseq data, the reference,
-and Augustus' prediction for {\it M. esculenta}. The examples were chosen so
+and AUGUSTUS' prediction for {\it M. esculenta}. The examples were chosen so
 that a) Helixer had high accuracy against the reference and the reference
@@
 (``sc", dotted) + 1, ii) the reference annotation in matrix form, iii) 
-Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
-and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+AUGUSTUS' predictions in matrix form, and iv) Helixer's predictions. The reference
+and AUGUSTUS have either 0 (white) or 1 (black) for each base pair and category, while
 Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
@@
 As an independent evaluation and to get a better qualitative understanding of
-performance we compared Helixer predictions, Augustus predictions and the
+performance we compared Helixer predictions, AUGUSTUS predictions and the
 reference to RNAseq coverage data. RNAseq was processed for three each of plants
-and animals chosen because they had varied performance quality relative to Augustus. 
+and animals chosen because they had varied performance quality relative to AUGUSTUS. 
 RNAseq coverage provides support for an 
@@
 and in another (S\ref{supfig:cov_example_02}c) Helixer's
-prediction falls between that of the reference and Augustus.
+prediction falls between that of the reference and AUGUSTUS.
 
-\begin{figure*}[!h]
+\begin{figure}[hbt]
 \label{fig:average_vs_augustus}
@@
 \caption{Fraction of bp with color-indicated a) coverage and b) spliced coverage of genomic positions 
-broken down by the confusion matrix of Augustus' and Helixer's predictions. Categories 
+broken down by the confusion matrix of AUGUSTUS' and Helixer's predictions. Categories 
 are only displayed if they can be meaningfully compared by examining a) coverage or b) spliced
@@
 for intron, and ``IG" stands for intergenic.}
-\end{figure*}
+\end{figure}
 
 Coverage and Spliced coverage were broken down by the confusion matrix (of both
-Helixer vs Augustus and Helixer vs the reference) for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
+Helixer vs AUGUSTUS and Helixer vs the reference) for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
 Where tools were in agreement (far left), coverage and spliced coverage closely matched
@@
 In cases of conflict, Helixer's predictions were substantially more consistent with
-the RNAseq data than those of Augustus. Specifically, for base pairs where one tool said CDS and
+the RNAseq data than those of AUGUSTUS. Specifically, for base pairs where one tool said CDS and
 the other intergenic or intron, there was more coverage when 
@@
 These patterns were consistent in direction but varied in magnitude in the individual species,
-with the expception of {\it P. marinus} where Helixer and Augustus performed comparably
+with the expception of {\it P. marinus} where Helixer and AUGUSTUS performed comparably
 at differentiating CDS and introns (Figure S\ref{supfig:each_sp_vs_augustus}).
@@
 were selected because they had (relative to the usual difference) low Subgenic F1 compared to 
-Augustus; for the former the RNAseq confirms relatively weak performance for Helixer, while the 
+AUGUSTUS; for the former the RNAseq confirms relatively weak performance for Helixer, while the 
 for the latter RNAseq rather indicates a sub-par reference.
@@
 greatly increase real world applications.
-\FloatBarrier
 
@@
 % tables
+\begin{suptab}\label{suptab:rnaseq_samples}\end{suptab}
 \begin{suptab}\label{suptab:generalization_animals}\end{suptab}
@@
 % figures
+\begin{supfig}\label{supfig:rna_select_animals}\end{supfig}
+\begin{supfig}\label{supfig:rna_select_plants}\end{supfig}
 \begin{supfig}\label{supfig:phylogeny_f1_animals}\end{supfig}
@@
 % sections
-\begin{supsec}\label{sec:generalization}\end{supsec}
-\begin{supsec}\label{sec:hyperparameters}\end{supsec}
-\begin{supsec}\label{sec:longer}\end{supsec}
-\begin{supsec}\label{sec:overlapping}\end{supsec}
-\begin{supsec}\label{sec:training_data}\end{supsec}
-\begin{supsec}\label{sec:dcnn}\end{supsec}
-\begin{supsec}\label{sec:rnaseq_eval}\end{supsec}
+\begin{supsec}\label{ssec:data_prep}\end{supsec}
+\begin{supsec}\label{ssec:generalization}\end{supsec}
+\begin{supsec}\label{ssec:hyperparameters}\end{supsec}
+\begin{supsec}\label{ssec:longer}\end{supsec}
+\begin{supsec}\label{ssec:overlapping}\end{supsec}
+\begin{supsec}\label{ssec:training_data}\end{supsec}
+\begin{supsec}\label{ssec:dcnn}\end{supsec}
+\begin{supsec}\label{ssec:rnaseq_eval}\end{supsec}
 

### 2020-05-25 paper/supplements.tex
some refs, some conventions, moving towards a hacked semi-dynamic ref for supplemental things
more semi-dynamic linking and refs
refs n cites
dynamic supplement and stuff

@@
 \usepackage{natbib}
+\usepackage{geometry}
 \usepackage{hyperref}
+\usepackage{amssymb}
 \hypersetup{
@@
 }
+\newgeometry{vmargin={20mm}, hmargin={25mm}}
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+\def \tablescale{0.88}
 \begin{document}
@@
 \author[1]{Marvin Steinborn}
+\author[ \hspace{-1ex}]{Daniela Dey}
 \author[ \hspace{-1ex}]{Stephan Scholz}
@@
 \newpage
+\section{Data collection and preparation}
+\label{ssec:data_prep}
+% species and versions table
+% meta data table?
+% selecting species for RNAseq
+\begin{figure}[!h]
+\label{supfig:rna_select_animals}
+\centerline{\includegraphics[width=1.0\textwidth]{images/RNAseq_animals_fit}}
+\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
+was used to select three animal species (larger points) for detailed RNAseq analysis. The 
+species were chosen to have good, medium and poor performance; invertebrates (diamonds) were
+not considered as the generalizability and recommended use of the Helixer animal model
+currently only extends to vertebrates; candidate species were checked for the availibility of
+stranded, paired-end RNAseq data and skipped if none was available.}
+\end{figure}
+
+\clearpage
+\begin{figure}[!hbt]
+\label{supfig:rna_select_plants}
+\centerline{\includegraphics[width=0.5\textwidth]{images/RNAseq_plants}}
+\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
+was used to select three plant species (larger points) for detailed RNAseq analysis. The 
+species were chosen to have good, medium and poor performance; algae (diamonds) were
+not considered as the generalizability and recommended use of the Helixer animal model
+currently only extends to higher plants; candidate species were checked for the availibility of
+stranded, paired-end RNAseq data and skipped if none was available.}
+\end{figure}
+\clearpage
+% RNAseq tested samples table
+\begin{table}[!h]
+\resizebox{0.8\textwidth}{!}{%
+\centering
+\begin{tabular}{@{}lll|lll|lll|lll@{}}
+\hline
+Samples&HQ&S&Samples&HQ&S&Samples&HQ&S&Samples&HQ&S\\
+\hline
+\\
+\textit{P. anubis}\\
+\hline
+SRS3122053&&&SRS3122094&\checkmark&\checkmark&SRS3270781&&&SRS3270847&&\\
+SRS3122057&\checkmark&\checkmark&SRS3122111&&&SRS3270785&&&SRS3270853&&\\
+SRS3122063&\checkmark&\checkmark&SRS3270734&&&SRS3270808&&&SRS3270858&&\\
+SRS3122070&&&SRS3270735&&&SRS3270811&&&SRS3270862&&\\
+SRS3122074&&&SRS3270737&&&SRS3270812&&&SRS3270863&&\\
+SRS3122079&\checkmark&\checkmark&SRS3270738&&&SRS3270821&&&SRS3270870&&\\
+SRS3122081&&&SRS3270746&&&SRS3270824&&&SRS3270872&&\\
+SRS3122088&&&SRS3270751&&&SRS3270839&&&SRS3270880&&\\
+SRS3122092&\checkmark&\checkmark&SRS3270763&&&SRS3270840&&&SRS819311&\checkmark&\checkmark\\
+SRS3122093&&&SRS3270773&&&SRS3270841&&&SRS819317&\checkmark&\checkmark\\
+\hline
+\\
+\textit{E. caballus}\\
+\hline
+ERS1517592&&&ERS2382444&&&ERS2681771&&&SRS2552997&\checkmark&\checkmark\\
+ERS1517600&&&ERS2681720&\checkmark&\checkmark&ERS2681772&\checkmark&\checkmark&SRS2553004&\checkmark&\checkmark\\
+ERS1517608&&&ERS2681730&&&ERS2681775&&&SRS2770963&&\\
+ERS1517617&&&ERS2681733&&&ERS3563898&&&SRS2770972&&\\
+ERS1517625&&&ERS2681734&&&SRS1102326&&&SRS3347177&&\\
+ERS1517633&&&ERS2681740&&&SRS1102328&&&SRS3347182&&\\
+ERS1517641&&&ERS2681742&&&SRS1102336&&&SRS4826859&&\\
+ERS2382439&&&ERS2681750&&&SRS2158774&&&SRS4826867&&\\
+ERS2382440&&&ERS2681754&&&SRS2158777&&&SRS4890460&\checkmark&\checkmark\\
+ERS2382442&&&ERS2681758&&&SRS2158788&&&SRS5330332&\checkmark&\checkmark\\
+ERS2382443&&&ERS2681770&&&SRS2552988&\checkmark&\checkmark&SRS5330340&&\\
+\hline
+\\
+\textit{P. marinus}\\
+\hline
+SRS5315360&\checkmark&\checkmark&SRS5315363&\checkmark&&SRS5315366&\checkmark&&SRS5315369&\checkmark&\\
+SRS5315361&\checkmark&\checkmark&SRS5315364&\checkmark&\checkmark&SRS5315367&\checkmark&\checkmark&&&\\
+SRS5315362&\checkmark&\checkmark&SRS5315365&\checkmark&\checkmark&SRS5315368&\checkmark&\checkmark&&&\\
+\hline
+\\
+\textit{M. esculenta}\\
+\hline
+SRS1608507&&&SRS1610089&\checkmark&\checkmark&SRS1610201&&&SRS946172&&\\
+SRS1608510&\checkmark&\checkmark&SRS1610128&\checkmark&\checkmark&SRS1610205&\checkmark&\checkmark&SRS946227&&\\
+SRS1610036&&&SRS1610150&&&SRS1610207&&&SRS946633&&\\
+SRS1610041&&&SRS1610177&&&SRS1610208&&&SRS946650&&\\
+SRS1610076&\checkmark&\checkmark&SRS1610181&&&SRS945853&&&SRS946703&&\\
+SRS1610079&\checkmark&\checkmark&SRS1610187&&&SRS946078&&&SRS947897&&\\
+SRS1610084&\checkmark&\checkmark&SRS1610197&&&SRS946138&&&&&\\
+\hline
+\\
+\textit{M. truncatula}\\
+\hline
+ERS1556003&&&SRS1948297&&&SRS3700721&\checkmark&&SRS4425365&\checkmark&\\
+ERS1556006&&&SRS1948300&&&SRS3700725&\checkmark&\checkmark&SRS4538246&&\\
+ERS1556009&&&SRS1948304&\checkmark&&SRS4425341&&&SRS4538249&&\\
+ERS1556012&&&SRS1948306&&&SRS4425343&\checkmark&&SRS5045319&&\\
+ERS1556015&&&SRS1948309&&&SRS4425346&\checkmark&\checkmark&SRS5045322&&\\
+ERS1556018&\checkmark&&SRS1948313&\checkmark&\checkmark&SRS4425347&\checkmark&\checkmark&SRS5045325&&\\
+ERS1556021&&&SRS1948315&&&SRS4425350&\checkmark&&SRS5045328&&\\
+ERS1556024&&&SRS3700712&\checkmark&&SRS4425353&&&SRS5045330&&\\
+SRS1736395&\checkmark&\checkmark&SRS3700715&\checkmark&&SRS4425356&\checkmark&\checkmark&SRS5045334&&\\
+SRS1736398&\checkmark&&SRS3700717&&&SRS4425357&&&SRS5045337&&\\
+SRS1948293&&&SRS3700718&\checkmark&\checkmark&SRS4425364&\checkmark&&&&\\
+\hline
+\\
+\textit{T. cacao}\\
+\hline
+SRS5244538&\checkmark&&SRS5244546&\checkmark&\checkmark&SRS5244554&&&SRS5244562&\checkmark&\checkmark\\
+SRS5244539&&&SRS5244547&&&SRS5244555&&&SRS5244563&&\\
+SRS5244540&&&SRS5244548&&&SRS5244556&\checkmark&\checkmark&SRS5244564&&\\
+SRS5244541&&&SRS5244549&&&SRS5244557&\checkmark&&SRS5244565&\checkmark&\checkmark\\
+SRS5244542&&&SRS5244550&&&SRS5244558&&&SRS5244566&&\\
+SRS5244543&\checkmark&\checkmark&SRS5244551&&&SRS5244559&&&SRS5244567&&\\
+SRS5244544&\checkmark&\checkmark&SRS5244552&&&SRS5244560&&&SRS5244568&&\\
+SRS5244545&&&SRS5244553&\checkmark&\checkmark&SRS5244561&&&&&\\
+\hline
+\end{tabular}}
+\label{suptab:rnaseq_samples}
+\caption{List of RNAseq samples that were selected for processing, whether they were of high quality (HQ), and 
+finally, whether they were ultimately selected (S) for merging and use in coverage calculation.}
+\end{table}
+\clearpage
+%
+
 \section{Model Evaluation by Species}
+\label{ssec:generalization}
 The following tables and plots show the prediction performance of Helixer for each species we worked with.
@@
 \begin{table}[!h]
-\renewcommand\thetable{}
-\resizebox{\textwidth}{!}{%
+%\renewcommand\thetable{}
+\resizebox{\tablescale\textwidth}{!}{%
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S1}
-\resizebox{\textwidth}{!}{%
+%\renewcommand\thetable{S1}
+\resizebox{\tablescale\textwidth}{!}{%
 \centering
@@
 \end{tabular}}
+\label{suptab:generalization_animals}
 \caption{Detailed results for all animal genomes.}
@@
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part1}}
+\label{supfig:phylogeny_f1_animals}
+\centerline{\includegraphics[width=\textwidth]{images/animals_f1_scores_part1}}
 \end{figure}
 \begin{figure}[!h]
-\renewcommand\thefigure{S1}
-\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part2}}
+%\renewcommand\thefigure{S1}  % just commenting these out for now, incase I missed or miss-ordered anything in the text while making dynamic
+\centerline{\includegraphics[width=\textwidth]{images/animals_f1_scores_part2}}
 \caption{Subgenic F1 results for all animal genomes shown by phylogenetic position.}
@@
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part1}}
+\label{supfig:phylogeny_differences_animals}
+\centerline{\includegraphics[width=\textwidth]{images/animals_differences_part1}}
 \end{figure}
 \begin{figure}[!h]
-\renewcommand\thefigure{S2}
-\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part2}}
+%\renewcommand\thefigure{S2}
+\centerline{\includegraphics[width=\textwidth]{images/animals_differences_part2}}
 \caption{Difference in Subgenic F1 to AUGUSTUS for all animal genomes shown by phylogenetic position.}
@@
 \begin{table}[!h]
-\renewcommand\thetable{S2}
-\resizebox{\textwidth}{!}{%
+%\renewcommand\thetable{S2}
+\resizebox{\tablescale\textwidth}{!}{%
 \centering
@@
 \end{tabular}}
+\label{suptab:generalization_plants}
 \caption{Detailed results for all plant genomes.}
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S3}
-\centerline{\includegraphics[width=1.0\textwidth]{images/plants_f1_scores}}
+\label{supfig:phylogeny_f1_plants}
+%\renewcommand\thefigure{S3}
+\centerline{\includegraphics[width=\tablescale\textwidth]{images/plants_f1_scores}}
 \caption{Subgenic F1 results for all plant genomes shown by phylogenetic position.}
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S4}
-\centerline{\includegraphics[width=1.0\textwidth]{images/plants_differences}}
+\label{supfig:phylogeny_differences_plants}
+%\renewcommand\thefigure{S4}
+\centerline{\includegraphics[width=\tablescale\textwidth]{images/plants_differences}}
 \caption{Difference in Subgenic F1 to AUGUSTUS for all plant genomes shown by phylogenetic position.}
@@
 \section{Hyperparameters}
-
+\label{sec:hyperparameters}
 \begin{table}[!h]
-\renewcommand\thetable{S3}
+%\renewcommand\thetable{S3}
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S4}
+%\renewcommand\thetable{S4}
 \centering
@@
 The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}
+\label{suptab:classweights}
 \end{table}
@@
 \begin{table}[!h]
-\renewcommand\thetable{S5}
+\label{suptab:prediction_lengths}
+%\renewcommand\thetable{S5}
 \centering
@@
 \newpage
-
+\def \overlapscale{1.07}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals1}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals1}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals2}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals2}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals3}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals3}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals4}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals4}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals5}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals5}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals6}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals6}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals7}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals7}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals8}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals8}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals9}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals9}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals10}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals10}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals11}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals11}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals12}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals12}}
 \end{figure}
 \begin{figure}[!h]
-\renewcommand\thefigure{S5}
+%\renewcommand\thefigure{S5}
 \centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
 \caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all animals genomes individually.}
+\label{supfig:overlapping_animals}
 \end{figure}
@@
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants1}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants1}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants2}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants2}}
 \end{figure}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants3}}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants3}}
 \end{figure}
 \begin{figure}[!h]
-\renewcommand\thefigure{S6}
-\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants4}}
+%\renewcommand\thefigure{S6}
+\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants4}}
 \caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all plant genomes individually.}
+\label{supfig:overlapping_plants}
 \end{figure}
@@
 \begin{table}[!h]
-\renewcommand\thetable{S6}
+\label{suptab:genome_stats}
+%\renewcommand\thetable{S6}
 \centering
@@
 \end{tabular}
-\caption{The description of Table [?] applies here as well.}
+\caption{The description of Table [?] applies here as well.}  % 2
 \end{table}
@@
 \section{Dilated CNN Baseline Network Search}
+\label{sec:dcnn}
 \begin{table}[!h]
-\renewcommand\thetable{S7}
+\label{suptab:dcnn_xval}
+%\renewcommand\thetable{S7}
 \centering
@@
 \begin{table}[!h]
-\renewcommand\thetable{S8}
+\label{suptab:dcnn_params}
+%\renewcommand\thetable{S8}
 \centering
@@
 
+\newpage
 \section{Evaluation against RNAseq}
+\label{sec:rnaseq_eval}
 
 \begin{figure}[!h]
-\renewcommand\thefigure{S7}
-\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_002}}
+\label{supfig:cov_example_02}
+%\renewcommand\thefigure{S7}
+\centerline{\includegraphics[width=\textwidth]{images/cov_examples/cov_example_002}}
 \caption{
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S8}
-\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_003}}
+\label{supfig:cov_example_03}
+%\renewcommand\thefigure{S8}
+\centerline{\includegraphics[width=\textwidth]{images/cov_examples/cov_example_003}}
 \caption{
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S9}
-\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_UTRs_plants}}
+%\renewcommand\thefigure{S9}
+\label{supfig:UTRs_animals}
+\centerline{\includegraphics[width=\textwidth]{images/cov_examples/cov_example_UTRs_animals}}
+\caption{
+One example of helixer's gradual UTR to intergenic transition predictions 
+from each RNAseq evaluation animal species in the context of RNAseq data, the reference,
+and Augustus' prediction. The species are
+a) {\it P. anubis}, b) {\it E. caballus} and c) {\it P. marinus}. Each subplot shows from top to
+bottom i) the natural log of the coverage (``cov", solid) and spliced coverage
+(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii)
+Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
+and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
+Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
+for intron, and ``IG" stands for intergenic.
+}
+\end{figure}
+
+\begin{figure}[!h]
+%\renewcommand\thefigure{S10}
+\label{supfig:UTRs_plants}
+\centerline{\includegraphics[width=\textwidth]{images/cov_examples/cov_example_UTRs_plants}}
 \caption{
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S10}
-\centerline{\includegraphics[width=17.8cm]{images/cov_examples/cov_example_UTRs_animals}}
-\caption{
-One example of helixer's gradual UTR to intergenic transition predictions 
-from each RNAseq evaluation animal species in the context of RNAseq data, the reference,
-and Augustus' prediction. The species are
-a) {\it P. anubis}, b) {\it E. caballus} and c) {\it P. marinus}. Each subplot shows from top to
-bottom i) the natural log of the coverage (``cov", solid) and spliced coverage
-(``sc", dotted) + 1, ii) the reference annotation in matrix form, iii)
-Augustus' predictions in matrix form, and iv) Helixer's predictions. The reference
-and Augustus have either 0 (white) or 1 (black) for each base pair and category, while
-Helixer emits a probability from 0-1 represented via gray-scale. ``Ntrn" stand
-for intron, and ``IG" stands for intergenic.
-}
-\end{figure}
-
-\begin{figure}[!h]
-\renewcommand\thefigure{S11}
+%\renewcommand\thefigure{S11}
+\label{supfig:average_vs_ref}
 \centerline{\includegraphics[]{images/cov_examples/average_vs_ref}}
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S12}
+%\renewcommand\thefigure{S12}
+\label{supfig:each_sp_vs_augustus}
 \centerline{\includegraphics[]{images/cov_examples/each_sp_vs_augustus}}
@@
 \begin{figure}[!h]
-\renewcommand\thefigure{S13}
+%\renewcommand\thefigure{S13}
+\label{supfig:each_sp_vs_ref}
 \centerline{\includegraphics[]{images/cov_examples/each_sp_vs_ref}}

### 2020-05-25 meta_info_collec/coverage/sumsamples.py

@@ -1,0 +1,24 @@
+samplepath = 'sample_ids.txt'
+qualitypath = 'good_stranded.txt'
+chosenpath = 'good_stranded_files.txt'
+outpath = 'selected.csv'
+
+# import files for included checks
+with open(qualitypath, 'r') as f:
+    qualities = f.readlines()
+    qualities = [x.rstrip() for x in qualities]
+
+with open(chosenpath, 'r') as f:
+    chosens = f.readlines()
+    chosens = [x.rstrip() for x in chosens]
+
+chosens = [x.replace('mapped/', '') for x in chosens]
+chosens = [x.replace('.bam', '') for x in chosens]
+
+handleout = open(outpath, 'w')
+with open(samplepath, 'r') as f:
+    for line in f:
+        sample = line.rstrip()
+        outline = '{},{},{}\n'.format(sample, sample in qualities, sample in chosens)
+        handleout.write(outline)
+handleout.close()

### 2020-05-26 paper/main.tex
content presentable-ish

@@
 \newenvironment{supsec}[1][]{\refstepcounter{supsec}}
+
+\newcounter{supdata}
+\newenvironment{supdata}[1][]{\refstepcounter{supdata}}
 
@@
 FASTA file and the latest annotation in the GFF format. One genome for each species of animal 
-in EnsemblMetazoa 45 \citep{howe2020ensembl} as well as all non-embargoed plant
-species from the JGI Phytosome 13 database \citep{goodstein2012phytozome}; exact 
-genomes used are list here \ref{suptab:source_data}. % add a supplemental table with names and versions
+in EnsemblMetazoa 45 (\cite{howe2020ensembl}; Table S\ref{suptab:downloads_animals}) 
+as well as all non-embargoed plant species from the JGI Phytosome 13 database 
+(\cite{goodstein2012phytozome}; Table S\ref{suptab:downloads_plants}); exact 
+genomes used are list here . % add a supplemental table with names and versions
 Data was downloaded on 15/10/2019 and 29/03/2019 for animals and plants respectively. 
 Both data groups were used seperately during training as well as during evaluation.
-
-Both to divide genomes into training, validation and test sets, as well as to understand
-performance, we needed a way to evaluate genome and annotation quality.
-Metadata was generated for each genome, including: basic length and 
-composition statistics \citep{gurevich2013quast}; k-mer content \citep{marccais2011fast};
-conserved gene content of the genome, transcriptome, and proteome (\cite{simao2015busco};
-file conversion with gffread from \cite{trapnell2012differential}; 
-and finally simple counts of genic features from the gff files. In particular, 
-we considered the following features of the metadata when judging the quality of a genome and it's 
-annotation: assembly contiguity (N50 / genome size); retention in the proteome 
-of detected conserved orthogroups (BUSCOs) that were identified in the genome; and
-the presence of UTRs and alternative splicing as estimated by the gff feature-counts.
-Finally this metadata was combined with researcher knowledge of model species to guide genome
-selection as necessary.
 
@@
 respectively, and were chosen to be of decent quality and diversity based on collected 
-metadata while also representing a broad phylogenetic spread.
+metadata (see section S\ref{ssec:data_prep} and Dataset S\ref{supdata:metadata}) 
+while also representing a broad phylogenetic spread.
 
@@
 the training was carried out with the subsetted training file and using the following scripts provided
-by AUGUSTUS. An untrained model was setup with "new\_species.pl", and the model was fit by running
-"etraining" before and after the major hyper-parameter optimization with "optimize\_augustus.pl".
+by AUGUSTUS. An untrained model was setup with ``new\_species.pl", and the model was fit by running
+"etraining" before and after the major hyper-parameter optimization with ``optimize\_augustus.pl".
 Using the trained model for each species, we ran the main prediction ("augustus") with ``--UTR=off"
@@
 %\cite{harkess2017asparagus}; and {\it Chenopodium quinoa}, \cite{jarvis2017genome})  
-and other HMM-based tools for many more %(e.g. {\it Citrus clementina}, \cite{wu2018genomics}; and
+and other HMM-based tools for many more. %(e.g. {\it Citrus clementina}, \cite{wu2018genomics}; and
 %{\it Capsella grandiflora}, \cite{slotte2013capsella}). 
@@
 annotations for broad groups has the potential to greatly facilitate downstream
-analyses which suffer when data is heterogeneous \citep{}.
+analyses; for instance it could avoid some of the inconsistency and errors that
+are otherwise seen in RNAseq analyses when different annotations are used 
+\citep{zhao2015comprehensive,torres2016robust}.
 
@@
 % tables
-\begin{suptab}\label{suptab:rnaseq_samples}\end{suptab}
+
+\begin{suptab}\label{suptab:downloads_animals}\end{suptab}
+\begin{suptab}\label{suptab:downloads_plants}\end{suptab}
 \begin{suptab}\label{suptab:generalization_animals}\end{suptab}
@@
 \begin{suptab}\label{suptab:classweights}\end{suptab}
+\begin{suptab}\label{suptab:rnaseq_samples}\end{suptab}
 \begin{suptab}\label{suptab:prediction_lengths}\end{suptab}
@@
 % figures
-\begin{supfig}\label{supfig:rna_select_animals}\end{supfig}
-\begin{supfig}\label{supfig:rna_select_plants}\end{supfig}
 \begin{supfig}\label{supfig:phylogeny_f1_animals}\end{supfig}
@@
 \begin{supfig}\label{supfig:phylogeny_differences_plants}\end{supfig}
+\begin{supfig}\label{supfig:rna_select_animals}\end{supfig}
+\begin{supfig}\label{supfig:rna_select_plants}\end{supfig}
 \begin{supfig}\label{supfig:overlapping_animals}\end{supfig}
@@
 \begin{supsec}\label{ssec:hyperparameters}\end{supsec}
+\begin{supsec}\label{ssec:rnaseq_setup}\end{supsec}
 \begin{supsec}\label{ssec:longer}\end{supsec}
@@
 \begin{supsec}\label{ssec:rnaseq_eval}\end{supsec}
+% datasets
+\begin{supdata}\label{supdata:metadata}\end{supdata}  % just a placeholder for the CSV and where it's referenced
 

### 2020-05-26 paper/supplements.tex
content presentable-ish

@@
 \def \tablescale{0.88}
+
+\newcounter{supdata}
+\newenvironment{supdata}[1][]{\refstepcounter{supdata}}
+
 \begin{document}
@@
 \label{ssec:data_prep}
+This section describes exactly which animal and plant genomes were downloaded
+from Ensemble (Table \ref{suptab:downloads_animals}) and 
+Phytozome13 (Table \ref{suptab:downloads_plants}), respectively.
+Further we describe assesment of genome quality based on metadata.
+
+For each acquired genome we assembled a set of metadata for assessing 
+genome and annotation quality. 
+This metadata includes basic length and 
+composition statistics with Quast \citep{gurevich2013quast}; 
+bp content, 2-mer content with Jellyfish \citep{marccais2011fast};
+conserved gene content of the genome, transcriptome, and proteome (\cite{simao2015busco};
+the transcriptome and proteome used were generated from the genome and gff3 file 
+using gffread from \cite{trapnell2012differential}); 
+and finally simple counts of genic features from the gff3 files.
+
+While the metadata ultimately does not provide a way to quantify ``genome quality",
+there were several features that were, all the same, particularly useful to consider
+when evaulating, e.g. whether a particular genome was of sufficient quality to be
+used for training. Often these features did not directly measure quality of
+the annotation, but still reflect the time, resources and skill invested in
+creating the genome and annotation.
+
+\begin{itemize}
+    \item Assembly Quality
+    \begin{itemize}
+        \item Assembly contiguity--as measured by N50 / total assembly size
+    \end{itemize}
+    \item Annotation Quality
+    \begin{itemize}
+        \item Gene detection consistency--we looked at the difference between complete BUSCOs that could be
+          identified in the proteom and genome. An inability to identify BUSCOs in the proteome
+          that could alredy be identified in the genome without employing an extensive annotation
+          pipeline is a sign of poor annotation quality.
+        \item Alternative splicing--we looked at two measures of alternative splicing annotation: the 
+          ratio mRNA to gene gff3 features, and the ratio of `single' to `duplicated' BUSCOs identified
+          in the transcriptome. For species where extensive alternative splicing was expected, 
+          its absence indicates an absence of effort or extrinsic information inclusion in gene annotation.
+        \item UTRs--The average of 3' and 5' UTR features per mRNA feature. A handful of genomes included no, or very few, 
+          UTR annotations, and were thus innapropriate for training a model to predict UTRs. 
+    \end{itemize}
+\end{itemize}
+
+All of the raw metadata and the four derived features mentioned above can be found in 
+Dataset S\ref{supdata:metadata}).
+
+\begin{supdata}\label{supdata:metadata}\end{supdata}  % just a placeholder for the CSV and where it's referenced
+
 % species and versions table
+\def \versionscale{0.95}
+\begin{table}[!h]
+\resizebox*{!}{\versionscale\textheight}{%
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Ensembl98 name&Version info&Test\\
+\hline
+acanthochromis\_polyacanthus&ASM210954v1&\\
+ailuropoda\_melanoleuca&ailMel1&\\
+amphilophus\_citrinellus&Midas\_v5&\\
+amphiprion\_ocellaris&AmpOce10&\\
+amphiprion\_percula&Nemo\_v1&\\
+anabas\_testudineus&fAnaTes11&\\
+anas\_platyrhynchos\_platyrhynchos&CAU\_duck10&\\
+anolis\_carolinensis&AnoCar20&\\
+anser\_brachyrhynchus&ASM259213v1&\\
+aotus\_nancymaae&Anan\_20&\\
+apteryx\_haastii&aptHaa1&\\
+apteryx\_owenii&aptOwe1&\\
+astatotilapia\_calliptera&fAstCal12&\\
+astyanax\_mexicanus&Astyanax\_mexicanus-20&\\
+betta\_splendens&fBetSpl52&\\
+bison\_bison\_bison&Bison\_UMD10&\\
+bos\_indicus\_hybrid&UOA\_Brahman\_1&\\
+bos\_mutus&BosGru\_v20&\\
+bos\_taurus&ARS-UCD12&\\
+caenorhabditis\_elegans&WBcel235&\\
+calidris\_pygmaea&ASM369795v1&\\
+callithrix\_jacchus&ASM275486v1&\\
+callorhinchus\_milii&Callorhinchus\_milii-613&\\
+canis\_familiaris&CanFam31&\\
+canis\_lupus\_dingo&ASM325472v1&\\
+capra\_hircus&ARS1&\\
+carlito\_syrichta&Tarsius\_syrichta-201&\\
+castor\_canadensis&can\_genome\_v10&\\
+cavia\_aperea&CavAp10&\\
+cavia\_porcellus&Cavpor30&\\
+cebus\_capucinus&Cebus\_imitator-10&\\
+cercocebus\_atys&Caty\_10&\\
+chelonoidis\_abingdonii&ASM359739v1&\\
+chinchilla\_lanigera&ChiLan10&\\
+chlorocebus\_sabaeus&ChlSab11&\\
+chrysemys\_picta\_bellii&Chrysemys\_picta\_bellii-303&\\
+ciona\_intestinalis&KH&\\
+ciona\_savignyi&CSAV20&\\
+clupea\_harengus&Ch\_v202&\\
+colobus\_angolensis\_palliatus&pa\_10&\\
+cottoperca\_gobio&fCotGob31&\\
+coturnix\_japonica&Coturnix\_japonica\_20&\\
+cricetulus\_griseus\_picr&CriGri-PICR&\\
+crocodylus\_porosus&CroPor\_comp1&\\
+cynoglossus\_semilaevis&Cse\_v10&\\
+cyprinodon\_variegatus&C\_variegatus-10&\\
+dasypus\_novemcinctus&Dasnov30&\\
+denticeps\_clupeoides&fDenClu11&\\
+dipodomys\_ordii&Dord\_20&\\
+dromaius\_novaehollandiae&droNov1&\\
+drosophila\_melanogaster&BDGP622&\\
+echinops\_telfairi&TENREC&\\
+electrophorus\_electricus&Ee\_SOAP\_WITH\_SSPACE&\\
+eptatretus\_burgeri&Eburgeri\_32&\\
+equus\_asinus\_asinus&ASM303372v1&\\
+equus\_caballus&EquCab30&\\
+erinaceus\_europaeus&HEDGEHOG&\\
+erpetoichthys\_calabaricus&fErpCal11&\\
+esox\_lucius&Eluc\_V3&\\
+felis\_catus&Felis\_catus\_90&\\
+ficedula\_albicollis&FicAlb\_14&\\
+fukomys\_damarensis&DMR\_v10&\\
+fundulus\_heteroclitus&Fundulus\_heteroclitus-302&\\
+gallus\_gallus&GRCg6a&\\
+\hline
+\end{tabular}}
+\end{table}
+\clearpage
+
+\begin{table}[!h]
+\resizebox*{!}{\versionscale\textheight}{%
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Ensembl98 name&Version info&Test\\
+\hline
+gambusia\_affinis&ASM309773v1&\\
+gasterosteus\_aculeatus&BROADS1&\\
+gopherus\_agassizii&ASM289641v1&\\
+gorilla\_gorilla&gorGor4&\\
+gouania\_willdenowi&fGouWil21&\\
+haplochromis\_burtoni&AstBur10&\\
+heterocephalus\_glaber\_female&HetGla\_female\_10&\\
+hippocampus\_comes&H\_comes\_QL1\_v1&\\
+homo\_sapiens&GRCh38&\\
+hucho\_hucho&ASM331708v1&\\
+ictalurus\_punctatus&IpCoco\_12&\\
+ictidomys\_tridecemlineatus&SpeTri20&\\
+jaculus\_jaculus&JacJac10&\\
+kryptolebias\_marmoratus&ASM164957v1&\\
+labrus\_bergylta&BallGen\_V1&\\
+larimichthys\_crocea&L\_crocea\_20&\\
+lates\_calcarifer&ASB\_HGAPassembly\_v1&\\
+latimeria\_chalumnae&LatCha1&\\
+lepidothrix\_coronata&Lepidothrix\_coronata-10&\\
+lepisosteus\_oculatus&LepOcu1&\\
+lonchura\_striata\_domestica&LonStrDom1&\\
+loxodonta\_africana&loxAfr3&\\
+macaca\_fascicularis&Macaca\_fascicularis\_50&\\
+macaca\_nemestrina&Mnem\_10&\\
+manacus\_vitellinus&ASM171598v2&\\
+mandrillus\_leucophaeus&le\_10&\\
+marmota\_marmota\_marmota&marMar21&\\
+mastacembelus\_armatus&fMasArm11&\\
+meleagris\_gallopavo&Turkey\_201&\\
+melopsittacus\_undulatus&Melopsittacus\_undulatus\_63&\\
+meriones\_unguiculatus&MunDraft-v10&\\
+mesocricetus\_auratus&MesAur10&\\
+microcebus\_murinus&Mmur\_30&\\
+microtus\_ochrogaster&MicOch10&\\
+mola\_mola&ASM169857v1&\\
+monopterus\_albus&M\_albus\_10&\\
+mus\_caroli&CAROLI\_EIJ\_v11&\\
+mus\_musculus&GRCm38&\\
+mus\_pahari&PAHARI\_EIJ\_v11&\\
+mus\_spicilegus&MUSP714&\\
+mus\_spretus&SPRET\_EiJ\_v1&\\
+mustela\_putorius\_furo&MusPutFur10&\\
+myotis\_lucifugus&Myoluc20&\\
+neolamprologus\_brichardi&NeoBri10&\\
+neovison\_vison&v01&\\
+nomascus\_leucogenys&Nleu\_30&\\
+notamacropus\_eugenii&Meug\_10&\\
+notechis\_scutatus&TS10Xv2-PRI&\\
+ochotona\_princeps&OchPri20-Ens&\\
+octodon\_degus&OctDeg10&\\
+oreochromis\_niloticus&Orenil10&\\
+ornithorhynchus\_anatinus&OANA5&\\
+oryctolagus\_cuniculus&OryCun20&\\
+oryzias\_latipes&ASM223467v1&\\
+otolemur\_garnettii&OtoGar3&\\
+ovis\_aries&Oar\_v31&\\
+pan\_paniscus&panpan11&\\
+panthera\_pardus&PanPar10&\\
+panthera\_tigris\_altaica&PanTig10&\\
+pan\_troglodytes&Pan\_tro\_30&\\
+papio\_anubis&Panu\_30&\\
+parambassis\_ranga&fParRan21&\\
+paramormyrops\_kingsleyae&PKINGS\_01&\\
+parus\_major&Parus\_major11&\\
+\hline
+\end{tabular}}
+\end{table}
+\clearpage
+
+\begin{table}[!h]
+\resizebox*{!}{\versionscale\textheight}{%
+\centering
+\begin{tabular}{@{}lll@{}}
+\hline
+Ensembl98 name&Version info&Test\\
+\hline
+pelodiscus\_sinensis&PelSin\_10&\\
+periophthalmus\_magnuspinnatus&fa&\\
+peromyscus\_maniculatus\_bairdii&HU\_Pman\_21&\\
+petromyzon\_marinus&Pmarinus\_70&\\
+phascolarctos\_cinereus&phaCin\_unsw\_v41&\\
+piliocolobus\_tephrosceles&ASM277652v2&\\
+poecilia\_formosa&PoeFor\_512&\\
+poecilia\_latipinna&P\_latipinna-10&\\
+poecilia\_mexicana&P\_mexicana-10&\\
+poecilia\_reticulata&Guppy\_female\_10\_MT&\\
+pogona\_vitticeps&pvi11&\\
+pongo\_abelii&PPYG2&\\
+procavia\_capensis&proCap1&\\
+prolemur\_simus&Prosim\_10&\\
+propithecus\_coquereli&Pcoq\_10&\\
+pteropus\_vampyrus&pteVam1&\\
+pygocentrus\_nattereri&Pygocentrus\_nattereri-102&\\
+rhinopithecus\_bieti&ASM169854v1&\\
+rhinopithecus\_roxellana&Rrox\_v1&\\
+saimiri\_boliviensis\_boliviensis&SaiBol10&\\
+salvator\_merianae&HLtupMer3&\\
+sarcophilus\_harrisii&DEVIL70&\\
+scleropages\_formosus&ASM162426v1&\\
+seriola\_dumerili&Sdu\_10&\\
+seriola\_lalandi\_dorsalis&Sedor1&\\
+sorex\_araneus&COMMON\_SHREW1&\\
+spermophilus\_dauricus&ASM240643v1&\\
+sphenodon\_punctatus&ASM311381v1&\\
+stegastes\_partitus&Stegastes\_partitus-102&\\
+sus\_scrofa&Sscrofa111&\\
+taeniopygia\_guttata&taeGut324&\\
+takifugu\_rubripes&FUGU5&\\
+tetraodon\_nigroviridis&TETRAODON8&\\
+theropithecus\_gelada&Tgel\_10&\\
+tupaia\_belangeri&TREESHREW&\\
+tursiops\_truncatus&turTru1&\\
+urocitellus\_parryii&ASM342692v1&\\
+ursus\_americanus&ASM334442v1&\\
+ursus\_maritimus&UrsMar\_10&\\
+vicugna\_pacos&vicPac1&\\
+vombatus\_ursinus&bare-nosed\_wombat\_genome\_assembly&\\
+vulpes\_vulpes&VulVul22&\\
+xenopus\_tropicalis&Xenopus\_tropicalis\_v91&\\
+xiphophorus\_couchianus&Xiphophorus\_couchianus-401&\\
+xiphophorus\_maculatus&X\_maculatus-50-male&\\
+apteryx\_rowi&aptRow1&\checkmark\\
+calidris\_pugnax&ASM143184v1&\checkmark\\
+choloepus\_hoffmanni&choHof1&\checkmark\\
+cyanistes\_caeruleus&cyaCae2&\checkmark\\
+danio\_rerio&GRCz11&\checkmark\\
+gadus\_morhua&gadMor1&\checkmark\\
+junco\_hyemalis&ASM382977v1&\checkmark\\
+macaca\_mulatta&Mmul\_10&\checkmark\\
+maylandia\_zebra&M\_zebra\_UMD2a&\checkmark\\
+monodelphis\_domestica&ASM229v1&\checkmark\\
+nannospalax\_galili&galili\_v10&\checkmark\\
+nothoprocta\_perdicaria&notPer1&\checkmark\\
+numida\_meleagris&NumMel10&\checkmark\\
+oryzias\_melastigma&Om\_v07RACA&\checkmark\\
+pundamilia\_nyererei&PunNye10&\checkmark\\
+rattus\_norvegicus&Rnor\_60&\checkmark\\
+scophthalmus\_maximus&ASM318616v1&\checkmark\\
+serinus\_canaria&SCA1&\checkmark\\
+zonotrichia\_albicollis&Zonotrichia\_albicollis-101&\checkmark\\
+\hline
+\end{tabular}}
+\caption{Exact animal genome and annotation versions as download from Ensembl98
+Genomes not included were redundant (same species,
+different accession, assembly or version) with one of the above genomes.}
+\label{suptab:downloads_animals}
+\end{table}
+\clearpage
+
+\begin{table}[!h]
+\resizebox{0.7\textwidth}{!}{%
+\centering
+\begin{tabular}{@{}lllll@{}}
+\hline
+Species Name&Phytozome name&Phytozome ID&Version&Test\\
+\hline
+Ananas comosus&Acomosus&321&v3&\\
+Amaranthus hypochondriacus&Ahypochondriacus&459&v2.1&\\
+Asparagus officinalis&Aofficinalis&498&V1.1&\\
+Arabidopsis thaliana&Athaliana&167&TAIR10&\\
+Amborella trichopoda&Atrichopoda&291&v1.0&\\
+Brachypodium distachyon&Bdistachyon&314&v3.1&\\
+Brassica oleracea&Boleraceacapitata&446&v1.0&\\
+Cicer arietinum&Carietinum&492&v1.0&\\
+Citrus clementina&Cclementina&182&v1.0&\\
+Capsella grandiflora&Cgrandiflora&266&v1.1&\\
+Carica papaya&Cpapaya&113&ASGPBv0.4&\\
+Chenopodium quinoa&Cquinoa&392&v1.0&\\
+Chlamydomonas reinhardtii&Creinhardtii&281&v5.6&\\
+Capsella rubella&Crubella&474&v1.1&\\
+Citrus sinensis&Csinensis&154&v1.1&\\
+Coccomyxa subellipsoidea C-169&CsubellipsoideaC169&227&v2.0&\\
+Chromochloris zofingiensis&Czofingiensis&461&v5.2.3.2&\\
+Daucus carota&Dcarota&388&v2.0&\\
+Dunaliella salina&Dsalina&325&v1.0&\\
+Eucalyptus grandis&Egrandis&297&v2.0&\\
+Eutrema salsugineum&Esalsugineum&173&v1.0&\\
+Fragaria vesca&Fvesca&501&v2.0.a2&\\
+Glycine max&Gmax&508&Wm82.a4.v1&\\
+Gossypium raimondii&Graimondii&221&v2.1&\\
+Helianthus annuus&Hannuus&494&r1.2&\\
+Hordeum vulgare&Hvulgare&462&r1&\\
+Kalanchoe fedtschenkoi&Kfedtschenkoi&382&v1.1&\\
+Lactuca sativa&Lsativa&467&v5&\\
+Linum usitatissimum&Lusitatissimum&200&v1.0&\\
+Malus domestica&Mdomestica&491&v1.1&\\
+Manihot esculenta&Mesculenta&305&v6.1&\\
+Mimulus guttatus&Mguttatus&256&v2.0&\\
+Marchantia polymorpha&Mpolymorpha&320&v3.1&\\
+Micromonas pusilla&MpusillaCCMP1545&228&v3.0&\\
+Micromonas sp. RCC299&MspRCC299&229&v3.0&\\
+Medicago truncatula&Mtruncatula&285&Mt4.0v1&\\
+Olea europaea&Oeuropaea&451&v1.0&\\
+Ostreococcus lucimarinus&Olucimarinus&231&v2.0&\\
+Oryza sativa&Osativa&323&v7.0&\\
+Oropetium thomaeum&Othomaeum&386&v1.0&\\
+Prunus persica&Ppersica&298&v2.1&\\
+Populus trichocarpa&Ptrichocarpa&444&v3.1&\\
+Porphyra umbilicalis&Pumbilicalis&456&v1.5&\\
+Ricinus communis&Rcommunis&119&v0.1&\\
+Sorghum bicolor&Sbicolor&454&v3.1.1&\\
+Setaria italica&Sitalica&312&v2.2&\\
+Solanum lycopersicum&Slycopersicum&514&ITAG3.2&\\
+Selaginella moellendorffii&Smoellendorffii&91&v1.0&\\
+Spirodela polyrhiza&Spolyrhiza&290&v2&\\
+Solanum tuberosum&Stuberosum&448&v4.03&\\
+Theobroma cacao&Tcacao&233&v1.1&\\
+Vitis vinifera&Vvinifera&457&v2.1&\\
+Zostera marina&Zmarina&324&v2.2&\\
+Zea mays&Zmays&493&RefGen\_V4&\\
+Volvox carteri&Vcarteri&317&v2.1&\checkmark\\
+Cucumis sativus&Csativus&122&v1.0&\checkmark\\
+Musa acuminata&Ppatens&318&v3.3&\checkmark\\
+Physcomitrella patens&Macuminata&304&v1&\checkmark\\
+Triticum aestivum&Taestivum&296&v2.2&\checkmark\\
+Arabidopsis lyrata&Alyrata&384&v2.1&\checkmark\\
+\hline
+\end{tabular}}
+\caption{Exact plant genome and annotation versions as download from Phytozome13.
+Genomes not included were either still under embargo or redundant (same species,
+different accession, assembly or version) with one of the above genomes.}
+\label{suptab:downloads_plants}
+\end{table}
+\clearpage
 % meta data table?
-% selecting species for RNAseq
-\begin{figure}[!h]
-\label{supfig:rna_select_animals}
-\centerline{\includegraphics[width=1.0\textwidth]{images/RNAseq_animals_fit}}
-\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
-was used to select three animal species (larger points) for detailed RNAseq analysis. The 
-species were chosen to have good, medium and poor performance; invertebrates (diamonds) were
-not considered as the generalizability and recommended use of the Helixer animal model
-currently only extends to vertebrates; candidate species were checked for the availibility of
-stranded, paired-end RNAseq data and skipped if none was available.}
-\end{figure}
-
-\clearpage
-\begin{figure}[!hbt]
-\label{supfig:rna_select_plants}
-\centerline{\includegraphics[width=0.5\textwidth]{images/RNAseq_plants}}
-\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
-was used to select three plant species (larger points) for detailed RNAseq analysis. The 
-species were chosen to have good, medium and poor performance; algae (diamonds) were
-not considered as the generalizability and recommended use of the Helixer animal model
-currently only extends to higher plants; candidate species were checked for the availibility of
-stranded, paired-end RNAseq data and skipped if none was available.}
-\end{figure}
-\clearpage
-% RNAseq tested samples table
-\begin{table}[!h]
-\resizebox{0.8\textwidth}{!}{%
-\centering
-\begin{tabular}{@{}lll|lll|lll|lll@{}}
-\hline
-Samples&HQ&S&Samples&HQ&S&Samples&HQ&S&Samples&HQ&S\\
-\hline
-\\
-\textit{P. anubis}\\
-\hline
-SRS3122053&&&SRS3122094&\checkmark&\checkmark&SRS3270781&&&SRS3270847&&\\
-SRS3122057&\checkmark&\checkmark&SRS3122111&&&SRS3270785&&&SRS3270853&&\\
-SRS3122063&\checkmark&\checkmark&SRS3270734&&&SRS3270808&&&SRS3270858&&\\
-SRS3122070&&&SRS3270735&&&SRS3270811&&&SRS3270862&&\\
-SRS3122074&&&SRS3270737&&&SRS3270812&&&SRS3270863&&\\
-SRS3122079&\checkmark&\checkmark&SRS3270738&&&SRS3270821&&&SRS3270870&&\\
-SRS3122081&&&SRS3270746&&&SRS3270824&&&SRS3270872&&\\
-SRS3122088&&&SRS3270751&&&SRS3270839&&&SRS3270880&&\\
-SRS3122092&\checkmark&\checkmark&SRS3270763&&&SRS3270840&&&SRS819311&\checkmark&\checkmark\\
-SRS3122093&&&SRS3270773&&&SRS3270841&&&SRS819317&\checkmark&\checkmark\\
-\hline
-\\
-\textit{E. caballus}\\
-\hline
-ERS1517592&&&ERS2382444&&&ERS2681771&&&SRS2552997&\checkmark&\checkmark\\
-ERS1517600&&&ERS2681720&\checkmark&\checkmark&ERS2681772&\checkmark&\checkmark&SRS2553004&\checkmark&\checkmark\\
-ERS1517608&&&ERS2681730&&&ERS2681775&&&SRS2770963&&\\
-ERS1517617&&&ERS2681733&&&ERS3563898&&&SRS2770972&&\\
-ERS1517625&&&ERS2681734&&&SRS1102326&&&SRS3347177&&\\
-ERS1517633&&&ERS2681740&&&SRS1102328&&&SRS3347182&&\\
-ERS1517641&&&ERS2681742&&&SRS1102336&&&SRS4826859&&\\
-ERS2382439&&&ERS2681750&&&SRS2158774&&&SRS4826867&&\\
-ERS2382440&&&ERS2681754&&&SRS2158777&&&SRS4890460&\checkmark&\checkmark\\
-ERS2382442&&&ERS2681758&&&SRS2158788&&&SRS5330332&\checkmark&\checkmark\\
-ERS2382443&&&ERS2681770&&&SRS2552988&\checkmark&\checkmark&SRS5330340&&\\
-\hline
-\\
-\textit{P. marinus}\\
-\hline
-SRS5315360&\checkmark&\checkmark&SRS5315363&\checkmark&&SRS5315366&\checkmark&&SRS5315369&\checkmark&\\
-SRS5315361&\checkmark&\checkmark&SRS5315364&\checkmark&\checkmark&SRS5315367&\checkmark&\checkmark&&&\\
-SRS5315362&\checkmark&\checkmark&SRS5315365&\checkmark&\checkmark&SRS5315368&\checkmark&\checkmark&&&\\
-\hline
-\\
-\textit{M. esculenta}\\
-\hline
-SRS1608507&&&SRS1610089&\checkmark&\checkmark&SRS1610201&&&SRS946172&&\\
-SRS1608510&\checkmark&\checkmark&SRS1610128&\checkmark&\checkmark&SRS1610205&\checkmark&\checkmark&SRS946227&&\\
-SRS1610036&&&SRS1610150&&&SRS1610207&&&SRS946633&&\\
-SRS1610041&&&SRS1610177&&&SRS1610208&&&SRS946650&&\\
-SRS1610076&\checkmark&\checkmark&SRS1610181&&&SRS945853&&&SRS946703&&\\
-SRS1610079&\checkmark&\checkmark&SRS1610187&&&SRS946078&&&SRS947897&&\\
-SRS1610084&\checkmark&\checkmark&SRS1610197&&&SRS946138&&&&&\\
-\hline
-\\
-\textit{M. truncatula}\\
-\hline
-ERS1556003&&&SRS1948297&&&SRS3700721&\checkmark&&SRS4425365&\checkmark&\\
-ERS1556006&&&SRS1948300&&&SRS3700725&\checkmark&\checkmark&SRS4538246&&\\
-ERS1556009&&&SRS1948304&\checkmark&&SRS4425341&&&SRS4538249&&\\
-ERS1556012&&&SRS1948306&&&SRS4425343&\checkmark&&SRS5045319&&\\
-ERS1556015&&&SRS1948309&&&SRS4425346&\checkmark&\checkmark&SRS5045322&&\\
-ERS1556018&\checkmark&&SRS1948313&\checkmark&\checkmark&SRS4425347&\checkmark&\checkmark&SRS5045325&&\\
-ERS1556021&&&SRS1948315&&&SRS4425350&\checkmark&&SRS5045328&&\\
-ERS1556024&&&SRS3700712&\checkmark&&SRS4425353&&&SRS5045330&&\\
-SRS1736395&\checkmark&\checkmark&SRS3700715&\checkmark&&SRS4425356&\checkmark&\checkmark&SRS5045334&&\\
-SRS1736398&\checkmark&&SRS3700717&&&SRS4425357&&&SRS5045337&&\\
-SRS1948293&&&SRS3700718&\checkmark&\checkmark&SRS4425364&\checkmark&&&&\\
-\hline
-\\
-\textit{T. cacao}\\
-\hline
-SRS5244538&\checkmark&&SRS5244546&\checkmark&\checkmark&SRS5244554&&&SRS5244562&\checkmark&\checkmark\\
-SRS5244539&&&SRS5244547&&&SRS5244555&&&SRS5244563&&\\
-SRS5244540&&&SRS5244548&&&SRS5244556&\checkmark&\checkmark&SRS5244564&&\\
-SRS5244541&&&SRS5244549&&&SRS5244557&\checkmark&&SRS5244565&\checkmark&\checkmark\\
-SRS5244542&&&SRS5244550&&&SRS5244558&&&SRS5244566&&\\
-SRS5244543&\checkmark&\checkmark&SRS5244551&&&SRS5244559&&&SRS5244567&&\\
-SRS5244544&\checkmark&\checkmark&SRS5244552&&&SRS5244560&&&SRS5244568&&\\
-SRS5244545&&&SRS5244553&\checkmark&\checkmark&SRS5244561&&&&&\\
-\hline
-\end{tabular}}
-\label{suptab:rnaseq_samples}
-\caption{List of RNAseq samples that were selected for processing, whether they were of high quality (HQ), and 
-finally, whether they were ultimately selected (S) for merging and use in coverage calculation.}
-\end{table}
-\clearpage
-%
 
@@
 \end{tabular}}
+\caption{Detailed results for all animal genomes.}
 \label{suptab:generalization_animals}
-\caption{Detailed results for all animal genomes.}
 \end{table}
@@
 \end{tabular}}
+\caption{Detailed results for all plant genomes.}
 \label{suptab:generalization_plants}
-\caption{Detailed results for all plant genomes.}
 \end{table}
@@
 \end{table}
+\newpage
+
+\section{RNAseq Evaluation Setup}
+\label{sec:rnaseq_setup}
+% selecting species for RNAseq
+\begin{figure}[!h]
+\label{supfig:rna_select_animals}
+\centerline{\includegraphics[width=1.0\textwidth]{images/RNAseq_animals_fit}}
+\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
+was used to select three animal species (larger points) for detailed RNAseq analysis. The 
+species were chosen to have good, medium and poor performance; invertebrates (diamonds) were
+not considered as the generalizability and recommended use of the Helixer animal model
+currently only extends to vertebrates; candidate species were checked for the availibility of
+stranded, paired-end RNAseq data and skipped if none was available.}
+\end{figure}
+
+\clearpage
+\begin{figure}[!hbt]
+\label{supfig:rna_select_plants}
+\centerline{\includegraphics[width=0.5\textwidth]{images/RNAseq_plants}}
+\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
+was used to select three plant species (larger points) for detailed RNAseq analysis. The 
+species were chosen to have good, medium and poor performance; algae (diamonds) were
+not considered as the generalizability and recommended use of the Helixer animal model
+currently only extends to higher plants; candidate species were checked for the availibility of
+stranded, paired-end RNAseq data and skipped if none was available.}
+\end{figure}
+\clearpage
+
+% RNAseq tested samples table
+\begin{table}[!h]
+\resizebox{0.8\textwidth}{!}{%
+\centering
+\begin{tabular}{@{}lll|lll|lll|lll@{}}
+\hline
+Samples&HQ&S&Samples&HQ&S&Samples&HQ&S&Samples&HQ&S\\
+\hline
+\\
+\textit{P. anubis}\\
+\hline
+SRS3122053&&&SRS3122094&\checkmark&\checkmark&SRS3270781&&&SRS3270847&&\\
+SRS3122057&\checkmark&\checkmark&SRS3122111&&&SRS3270785&&&SRS3270853&&\\
+SRS3122063&\checkmark&\checkmark&SRS3270734&&&SRS3270808&&&SRS3270858&&\\
+SRS3122070&&&SRS3270735&&&SRS3270811&&&SRS3270862&&\\
+SRS3122074&&&SRS3270737&&&SRS3270812&&&SRS3270863&&\\
+SRS3122079&\checkmark&\checkmark&SRS3270738&&&SRS3270821&&&SRS3270870&&\\
+SRS3122081&&&SRS3270746&&&SRS3270824&&&SRS3270872&&\\
+SRS3122088&&&SRS3270751&&&SRS3270839&&&SRS3270880&&\\
+SRS3122092&\checkmark&\checkmark&SRS3270763&&&SRS3270840&&&SRS819311&\checkmark&\checkmark\\
+SRS3122093&&&SRS3270773&&&SRS3270841&&&SRS819317&\checkmark&\checkmark\\
+\hline
+\\
+\textit{E. caballus}\\
+\hline
+ERS1517592&&&ERS2382444&&&ERS2681771&&&SRS2552997&\checkmark&\checkmark\\
+ERS1517600&&&ERS2681720&\checkmark&\checkmark&ERS2681772&\checkmark&\checkmark&SRS2553004&\checkmark&\checkmark\\
+ERS1517608&&&ERS2681730&&&ERS2681775&&&SRS2770963&&\\
+ERS1517617&&&ERS2681733&&&ERS3563898&&&SRS2770972&&\\
+ERS1517625&&&ERS2681734&&&SRS1102326&&&SRS3347177&&\\
+ERS1517633&&&ERS2681740&&&SRS1102328&&&SRS3347182&&\\
+ERS1517641&&&ERS2681742&&&SRS1102336&&&SRS4826859&&\\
+ERS2382439&&&ERS2681750&&&SRS2158774&&&SRS4826867&&\\
+ERS2382440&&&ERS2681754&&&SRS2158777&&&SRS4890460&\checkmark&\checkmark\\
+ERS2382442&&&ERS2681758&&&SRS2158788&&&SRS5330332&\checkmark&\checkmark\\
+ERS2382443&&&ERS2681770&&&SRS2552988&\checkmark&\checkmark&SRS5330340&&\\
+\hline
+\\
+\textit{P. marinus}\\
+\hline
+SRS5315360&\checkmark&\checkmark&SRS5315363&\checkmark&&SRS5315366&\checkmark&&SRS5315369&\checkmark&\\
+SRS5315361&\checkmark&\checkmark&SRS5315364&\checkmark&\checkmark&SRS5315367&\checkmark&\checkmark&&&\\
+SRS5315362&\checkmark&\checkmark&SRS5315365&\checkmark&\checkmark&SRS5315368&\checkmark&\checkmark&&&\\
+\hline
+\\
+\textit{M. esculenta}\\
+\hline
+SRS1608507&&&SRS1610089&\checkmark&\checkmark&SRS1610201&&&SRS946172&&\\
+SRS1608510&\checkmark&\checkmark&SRS1610128&\checkmark&\checkmark&SRS1610205&\checkmark&\checkmark&SRS946227&&\\
+SRS1610036&&&SRS1610150&&&SRS1610207&&&SRS946633&&\\
+SRS1610041&&&SRS1610177&&&SRS1610208&&&SRS946650&&\\
+SRS1610076&\checkmark&\checkmark&SRS1610181&&&SRS945853&&&SRS946703&&\\
+SRS1610079&\checkmark&\checkmark&SRS1610187&&&SRS946078&&&SRS947897&&\\
+SRS1610084&\checkmark&\checkmark&SRS1610197&&&SRS946138&&&&&\\
+\hline
+\\
+\textit{M. truncatula}\\
+\hline
+ERS1556003&&&SRS1948297&&&SRS3700721&\checkmark&&SRS4425365&\checkmark&\\
+ERS1556006&&&SRS1948300&&&SRS3700725&\checkmark&\checkmark&SRS4538246&&\\
+ERS1556009&&&SRS1948304&\checkmark&&SRS4425341&&&SRS4538249&&\\
+ERS1556012&&&SRS1948306&&&SRS4425343&\checkmark&&SRS5045319&&\\
+ERS1556015&&&SRS1948309&&&SRS4425346&\checkmark&\checkmark&SRS5045322&&\\
+ERS1556018&\checkmark&&SRS1948313&\checkmark&\checkmark&SRS4425347&\checkmark&\checkmark&SRS5045325&&\\
+ERS1556021&&&SRS1948315&&&SRS4425350&\checkmark&&SRS5045328&&\\
+ERS1556024&&&SRS3700712&\checkmark&&SRS4425353&&&SRS5045330&&\\
+SRS1736395&\checkmark&\checkmark&SRS3700715&\checkmark&&SRS4425356&\checkmark&\checkmark&SRS5045334&&\\
+SRS1736398&\checkmark&&SRS3700717&&&SRS4425357&&&SRS5045337&&\\
+SRS1948293&&&SRS3700718&\checkmark&\checkmark&SRS4425364&\checkmark&&&&\\
+\hline
+\\
+\textit{T. cacao}\\
+\hline
+SRS5244538&\checkmark&&SRS5244546&\checkmark&\checkmark&SRS5244554&&&SRS5244562&\checkmark&\checkmark\\
+SRS5244539&&&SRS5244547&&&SRS5244555&&&SRS5244563&&\\
+SRS5244540&&&SRS5244548&&&SRS5244556&\checkmark&\checkmark&SRS5244564&&\\
+SRS5244541&&&SRS5244549&&&SRS5244557&\checkmark&&SRS5244565&\checkmark&\checkmark\\
+SRS5244542&&&SRS5244550&&&SRS5244558&&&SRS5244566&&\\
+SRS5244543&\checkmark&\checkmark&SRS5244551&&&SRS5244559&&&SRS5244567&&\\
+SRS5244544&\checkmark&\checkmark&SRS5244552&&&SRS5244560&&&SRS5244568&&\\
+SRS5244545&&&SRS5244553&\checkmark&\checkmark&SRS5244561&&&&&\\
+\hline
+\end{tabular}}
+\caption{List of RNAseq samples that were selected for processing, whether they were of high quality (HQ), and 
+finally, whether they were ultimately selected (S) for merging and use in coverage calculation.}
+\label{suptab:rnaseq_samples}
+\end{table}
+\clearpage
+%
 
@@
 \begin{table}[!h]
-\label{suptab:prediction_lengths}
 %\renewcommand\thetable{S5}
@@
 }
+\label{suptab:prediction_lengths}
 \end{table}
@@
 \begin{table}[!h]
-\label{suptab:genome_stats}
 %\renewcommand\thetable{S6}
@@
 \caption{The description of Table [?] applies here as well.}  % 2
+\label{suptab:genome_stats}
 \end{table}
@@
 \begin{table}[!h]
-\label{suptab:dcnn_xval}
 %\renewcommand\thetable{S7}
@@
 In total, we trained 30 models for the animals and 48 with the plant data.}
-\end{table}
-
-\begin{table}[!h]
-\label{suptab:dcnn_params}
+\label{suptab:dcnn_xval}
+\end{table}
+
+\begin{table}[!h]
 %\renewcommand\thetable{S8}
@@
 The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
+\label{suptab:dcnn_params}
 \end{table}
@@
 \label{supfig:cov_example_02}
+
 %\renewcommand\thefigure{S7}
@@
 
+\clearpage
 \renewcommand\refname{Supplemental References}
@@
 \bibliography{literature}
-
+\clearpage
 \end{document}

### 2020-05-28 paper/main.tex
minor things

@@
 as well as all non-embargoed plant species from the JGI Phytosome 13 database 
-(\cite{goodstein2012phytozome}; Table S\ref{suptab:downloads_plants}); exact 
-genomes used are list here . % add a supplemental table with names and versions
-Data was downloaded on 15/10/2019 and 29/03/2019 for animals and plants respectively. 
+(\cite{goodstein2012phytozome}; Table S\ref{suptab:downloads_plants}); The exact 
+genomes are listed in the aformentioned tables.
+Data was downloaded on the 15/10/2019 and 29/03/2019 for animals and plants respectively. 
 Both data groups were used seperately during training as well as during evaluation.
@@
 
-For both feasibility and iterative cross-species generalization validattion, we
+For both feasibility and the frequent validation of cross-species generalization during
+development, we
 selected only some of the remaining genomes for direct training.
 The selection of training genomes also requires balancing multiple conflicting 
-tradeoffs. On the one hand, we want to train with as much data as possible, but 
+tradeoffs in general. On the one hand, we want to train with as much data as possible, but 
 on the other
@@
 also very desirable to have diverse training genomes with a broad phylogenetic spread, 
-and a variety of genome sizes and average gene length as the model is increasingly unlikely to generalize 
+a variety of genome sizes and average gene lengths as the model is increasingly unlikely to generalize 
 well beyond the borders given by the training data. However, it also may be difficult 
-for the model learn and generalize, if the genomic patterns inside the data are 
+for the model learn and generalize if the genomic patterns inside the data are 
 too different from each other. Practically, it was important for us to limit the size of
@@
 training. We split off the validation set by selecting 20\% of the FASTA sequences 
-above and below the N90 of each training genome, ensuring a proper distribution of 
+above and below the N90 of each training genome at random, ensuring a proper distribution of 
 large and short sequences in both sets and a split on the chromosome level. 
@@
 \subsection{Data Pre-processing}
-We first pre-process and store the raw genomic information by using GeenuFF (https://https:
-//github.com/weberlab-hhu/GeenuFF). GeenuFF is a tool for checking and exploring genomic data, that 
+We first pre-process and store the raw genomic information by using GeenuFF 
+(https://github.com/weberlab-hhu/GeenuFF). 
+GeenuFF is a tool for checking and exploring genomic data, that 
 stores all information inside a SQL database. Training and evaluation-ready data is 
@@
 \botrule
-\end{tabular}}{{\it Note:} the encoding does not differentiate between introns
+\end{tabular}}{{\it Note:} The encoding does not differentiate between introns
 in coding and non-coding regions.}
@@
 circa 5.4 million parameters and was implemented with the deep learning library 
-keras \citep{chollet2015keras} on top of TensorFlow \citep{abadi2016tensorflow}. We 
+Keras \citep{chollet2015keras} on top of TensorFlow \citep{abadi2016tensorflow}. We 
 tested multiple different architectures before arriving at this model configuration, 
 including convolutional neural networks (CNN) and hybrid architectures. We also use 
-class weigths as the classes of the input data are both unbalanced vary greatly 
-between animals and plants. 
+class weigths as the class frequencies are both unbalanced and 
+vary greatly between animals and plants. 
 For example, the tradeoff between the number of intergenic and intronic bases is 
@@
 sequences than usual.
-We compare our final models to a dilated CNN (dCNN) baseline as this kind of architecture is 
+
+We also compare our final models to a dilated CNN (dCNN) baseline as this kind of architecture is 
 widely used when working with DNA data as input. Details of the neural architecture 

### 2020-05-29 paper/main.tex
minor changes

@@
 \subsection{Metrics}
+\label{sec:metrics}
 We mainly use two metrics to judge model performance. The {\it Genic F1} is our 
@@
 we randomly selected only half the BUSCO-generated training set, which resulted in a training set 
-size and runtime comparable to the plants (~400 genes and several days per species). For animals
+size and runtime comparable to the plants (about 400 genes and several days per species). For animals
 the training was carried out with the subsetted training file and using the following scripts provided
@@
 Using the trained model for each species, we ran the main prediction ("augustus") with ``--UTR=off"
-and ``--gff3=on". The gffs produced by AUGUSTUS were imported into GeenuFF and exported as h5
-files in the same manor as the reference, allowing direct comparison with reference and 
+and ``--gff3=on". The gffs produced by AUGUSTUS were imported into GeenuFF and exported as HDF5
+files in the same manor as the reference, allowing for the direct comparison with both the reference and 
 Helixer predictions. While tenable for some 237 species, this method allowed neither training
-nor prediction of UTR regions with AUGUSTUS, so the metric Subgenic F1, which excludes UTR,
-was used for all comparisons between AUGUSTUS and Helixer.
+nor prediction of UTR regions with AUGUSTUS, so the metric Subgenic F1, which disregards the class UTR,
+was used for all comparisons between AUGUSTUS and Helixer (see Section \ref{sec:metrics} for more details) .
 
@@
 The reference annotations were created with existing tools and largely with a pipeline incorporating
-{\it de novo} gene predictions with RNAseq and homology data. The references, like any
-data, are expected to contain errors so we chose to use RNAseq data for independent evaluation.
+{\it de novo} gene predictions with RNAseq and homology data. As the references, like any
+data, are expected to contain errors we chose to use RNAseq data for an independent evaluation.
 In plants there is the additional concern that the references may share biases with
@@
 AND ("biomol rna"[Properties] AND "library layout paired"[Properties])'. In so far as 
-more than 50 samples were identified, every Nth sample was selected so that under 50
-samples were chosen for further processing (Table S\ref{suptab:rnaseq_samples}).
+more than 50 samples were identified and every Nth sample was selected so that under 50
+samples were chosen for further processing (see Table S\ref{suptab:rnaseq_samples}).
 Each sample was quality controlled with FastQC (www.bioinformatics.babraham.ac.uk/projects/fastqc/), before and after trimming
@@
 automated and the results visualized with the code here (... make available and link to RNAseq stuff).
-Samples were filtered to those that had {\it relatively} high mapping rates, high mapping to
-exonic regions relative to other regions; that had low 3' bias, normal FastQC and 
-Trimmomatic stats, and that
-were confirmed to be stranded with 2nd strand as sense strand. If more remained, seven
+The samples were filtered based on many different metrics. Relatively high mapping rates, high mapping to
+exonic regions relative to other regions, low 3' bias, normal FastQC, Trimmomatic stats and a confirmation
+to be stranded with 2nd strand as sense strand were all taken into account. If more remained, seven
 of the high quality samples were selected randomly.
 
-Finally, the selected, quality, mapped RNAseq samples were merged and quantified to
+Finally, the selected and mapped RNAseq samples were merged and quantified to
 get the coverage (number of reads matching; i.e. cigar =, M or X),

### 2020-05-29 paper/main.tex
minor

@@
 
-For both feasibility and the frequent validation of cross-species generalization during
-development, we
-selected only some of the remaining genomes for direct training.
-The selection of training genomes also requires balancing multiple conflicting 
-tradeoffs in general. On the one hand, we want to train with as much data as possible, but 
+For the selection of training genomes we had to balance multiple conflicting 
+tradeoffs. On the one hand, we want to train with as much data as possible, but 
 on the other
@@
 are generated and respectively used as input and label for the neural network 
-together with the sample weights. We append zero padding if the sequences (or
-sequence ends) are shorter than 20,000 bp. 
+together with the sample weights. We append zero padding if the subsequences  
+are shorter than 20,000 bp. 
 If a subsequence is fully marked as erroneous, we exclude it from all analyses.
@@
 files in the same manor as the reference, allowing for the direct comparison with both the reference and 
-Helixer predictions. While tenable for some 237 species, this method allowed neither training
+Helixer predictions. While still feasible for some 237 species, this method allowed neither training
 nor prediction of UTR regions with AUGUSTUS, so the metric Subgenic F1, which disregards the class UTR,
@@
 intergenic even where Helixer predictions closely match the reference and RNAseq data (a). 
-Helixer's uncertainty around UTR <-> intergenic transitions may relate to a
+Helixer's uncertainty around transitions from UTR to intergenic regions may relate to a
 fundamentally harder problem (there is no conserved motif at the site as is observed
@@
 and shifts gradually between classes with no single class receiving a high probability 
-for extended stretches (Figures S\ref{supfig:cov_example_02}c, S\ref{supfig:cov_example_02}c-d). 
-Notably, in one of the examples (S\ref{supfig:cov_example_02}d))
+for extended stretches (Figures S\ref{supfig:cov_example_02}c-d). 
+Notably, in one of the examples (Figure S\ref{supfig:cov_example_02}d)
 the RNAseq shows evidence of alternative splicing; 
-and in another (S\ref{supfig:cov_example_02}c) Helixer's
-prediction falls between that of the reference and AUGUSTUS.
+and in another (Figure S\ref{supfig:cov_example_02}c) Helixer's
+prediction falls between that of the reference and of AUGUSTUS.
 
@@
 
-Coverage and Spliced coverage were broken down by the confusion matrix (of both
-Helixer vs AUGUSTUS and Helixer vs the reference) for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
-Where tools were in agreement (far left), coverage and spliced coverage closely matched
-expectations. Specifically for CDS:CDS and UTR:UTR most base pairs had some, and many
+Coverage and Spliced coverage were broken down by the confusion matrix of both
+Helixer vs AUGUSTUS as well as Helixer vs the reference for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
+Where both tools agreee on far left in the figure, coverage and spliced coverage closely matched expectations. 
+Specifically for CDS:CDS and UTR:UTR most base pairs had some, and many
 had moderate or high coverage. The same pattern was seen for intron:intron and spliced
@@
 capable of finding weak-spots in the other annotations.
+
 In cases of conflict, Helixer's predictions were substantially more consistent with
-the RNAseq data than those of AUGUSTUS. Specifically, for base pairs where one tool said CDS and
+the RNAseq data than those of AUGUSTUS. Specifically, for base pairs where one tool predicted CDS and
 the other intergenic or intron, there was more coverage when 
-Helixer said CDS (Figure \ref{fig:average_vs_augustus}a). Similarly, for base pairs where one tool said intron and the other 
-CDS or intergenic, there was more spliced coverage when Helixer said intron (Figure \ref{fig:average_vs_augustus}b). 
+Helixer predicted CDS (Figure \ref{fig:average_vs_augustus}a). Similarly, for base pairs where one tool predicted intron and the other 
+CDS or intergenic, there was more spliced coverage when Helixer predicted intron (Figure \ref{fig:average_vs_augustus}b). 
 These patterns were consistent in direction but varied in magnitude in the individual species,
 with the expception of {\it P. marinus} where Helixer and AUGUSTUS performed comparably
-at differentiating CDS and introns (Figure S\ref{supfig:each_sp_vs_augustus}).
+at differentiating between CDS and intron regions (Figure S\ref{supfig:each_sp_vs_augustus}).
 
@@
 \section{Discussion}
-We indtroduce Helixer a novel, deep-learning based tool for gene annotation
+With Helixer, we introduce a novel, deep-learning based tool for gene annotation
 prediction. Helixer outperforms AUGUSTUS on base pair wise metrics and on 

### 2020-06-02 paper/main.tex
paper

@@
 \textbf{Availability:} The source code of this work is available at 
-\url{https://github.com/weberlab-hhu/HelixerPrep} [change] under the GNU General 
+\url{https://github.com/weberlab-hhu/Helixer} under the GNU General 
 Public License v3.0.\\
@@
 designers of HMMs assume structure in the sequence that may limit its predictive 
-power. In practice HMMs have trouble generalizing across species and it seems 
-actual learning of underlying patterns is limited despite the fundamental ability 
-to do so being present \citep{}.
+power. In practice HMMs have trouble generalizing across species and any actual learning 
+related to sequence motifs is limited to very short sequences that indicate state transitions.
 
@@
 circa 5.4 million parameters and was implemented with the deep learning library 
-Keras \citep{chollet2015keras} on top of TensorFlow \citep{abadi2016tensorflow}. We 
+Keras (https://keras.io) on top of TensorFlow \citep{abadi2016tensorflow}. We 
 tested multiple different architectures before arriving at this model configuration, 
@@
 algorithm \citep{bergstra2011algorithms}, random search or grid search depending on 
-the situation. We used NNI \citep{nni2019} to facilitate the search. All relevant 
+the situation. We used NNI (https://github.com/microsoft/nni) to facilitate the search. All relevant 
 hyperparameters can be found in Table S\ref{suptab:params} and 
@@
 used in the animal case. The best model was chosen out of the eight models of the 
-ensemble for having the best performance in the reported setup.}
+ensemble for having the best performance on the validation set of the training genomes.}
 \end{table}
@@
 and AUGUSTUS' prediction for {\it M. esculenta}. The examples were chosen so
-that a) Helixer had high accuracy against the reference and the reference
-was supported by the RNAseq data, b) Helixer had high accuracy but the 
-reference was not supported by the RNAseq data, c) Helixer had low accuracy
-and the reference was supported by the RNAseq data, and d) helixer had low
+that {\bf a)} Helixer had high accuracy against the reference and the reference
+was supported by the RNAseq data, {\bf b)} Helixer had high accuracy but the 
+reference was not supported by the RNAseq data, {\bf c)} Helixer had low accuracy
+and the reference was supported by the RNAseq data, and {\bf d)} helixer had low
 accuracy but the reference was not supported by RNAseq data. Plotability was
@@
 for intergenic annotations. Looking at selected subsequences (Figure \ref{fig:cov_example_main}) % and supplemental figures.
-we identified cases where RNAseq supported both the annotations of the reference 
-and Helixer a), neither b), the reference, but not Helixer c),
-and Helixer, but not the reference d).
+we identified cases where RNAseq supported a) both the annotations of the reference 
+and Helixer, b) neither, c) the reference, but not Helixer,
+and d) Helixer, but not the reference.
 
@@
 annotation, it assigned a low but non-trivial probability to the RNAseq-supported 
-exon/intron pattern (Figures \ref{fig:cov_example_main}b-c,S\ref{supfig:cov_example_02}c). However, in the extreme,
+exon/intron pattern (Figures \ref{fig:cov_example_main}b-c, S\ref{supfig:cov_example_02}c). However, in the extreme,
 there are cases where the Helixer model exibits substantial indecision or confusion 
@@
 \includegraphics[width=8.6cm]{images/cov_examples/average_vs_augustus}  % should be 8.6 when floats are sorted out right
-\caption{Fraction of bp with color-indicated a) coverage and b) spliced coverage of genomic positions 
+\caption{Fraction of bp with color-indicated {\bf a)} coverage and {\bf b)} spliced coverage of genomic positions 
 broken down by the confusion matrix of AUGUSTUS' and Helixer's predictions. Categories 
@@
 \section{Discussion}
-With Helixer, we introduce a novel, deep-learning based tool for gene annotation
-prediction. Helixer outperforms AUGUSTUS on base pair wise metrics and on 
-consistency with independent RNASeq data.
+With Helixer, we introduce a novel, deep-learning based framework for the development 
+of more capable tools for gene annotation prediction. 
+Helixer outperforms AUGUSTUS on base pair wise metrics and on 
+consistency with independent RNASeq data while also predicting cross-species for a wide range of  
+genomes with one model.
 
 We include trained models for plants and animals which achieve 
-high prediction accuracy on cross-species gene annotation for broad 
+high prediction accuracy on gene annotation for broad 
 phylogenetic groups (land plants and vertebrates, respectively).
@@
 \section*{Acknowledgements}
-
-We are very grateful for the support of...
-
 HHU HPC-team, deNBI, Ebenh{\"o}h lab for computational (particularly GPU) resources.

### 2020-06-02 paper/main.tex
very minor

@@
 average Genic F1 (in red) and basepair wise accuracy (in purple) with respect to 
-the position in the 20000 basepair long input sequences. The dashed line shows the 
+the position in the 20,000 basepair long input sequences. The dashed line shows the 
 same without overlapping. Each value is the average performance of a 200 basepair 
@@
 \label{fig:average_vs_augustus}
-\includegraphics[width=8.6cm]{images/cov_examples/average_vs_augustus}  % should be 8.6 when floats are sorted out right
+\includegraphics[width=8.6cm]{images/cov_examples/average_vs_augustus} 
 \caption{Fraction of bp with color-indicated {\bf a)} coverage and {\bf b)} spliced coverage of genomic positions 
@@
 Coverage and Spliced coverage were broken down by the confusion matrix of both
-Helixer vs AUGUSTUS as well as Helixer vs the reference for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
+Helixer vs AUGUSTUS as well as Helixer vs the reference for all genomes (Figure \ref{fig:average_vs_augustus}). 
 Where both tools agreee on far left in the figure, coverage and spliced coverage closely matched expectations. 
@@
 With Helixer, we introduce a novel, deep-learning based framework for the development 
-of more capable tools for gene annotation prediction. 
+of more effective tools for gene annotation prediction. 
 Helixer outperforms AUGUSTUS on base pair wise metrics and on 
@@
 
+\section*{Funding}
+This work has been supported by the... \vspace*{-12pt}
+
 \section*{Acknowledgements}
-HHU HPC-team, deNBI, Ebenh{\"o}h lab for computational (particularly GPU) resources.
-
-\section*{Funding}
-
-This work has been supported by the... \vspace*{-12pt}
+We are grateful for the computational suppport and GPU access provided by
+the de.NBI cloud, and by both the Centre for Information and Media Technology (ZIM)
+and the institute of Quantitative and Theoretical Biology at Heinrich Heine University.
 

### 2020-06-04 paper/main.tex
some polishing and author rearrangements

@@
 \title[short Title]{Helixer: Cross-species Gene Annotation Of Large Eucaryotic Genomes Using Deep Learning}
-\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1,}}$$^\dagger$, Alisandra Denton\,$^{\text{\sfb 1,}}$$^\dagger$\, Marvin Steinborn\,$^{\text{\sfb 1}}$, Daniela Dey, Stephan Scholz, Andreas Weber\,$^{\text{\sfb 1,}*}$}
+\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1}}$, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Daniela Dey, Andreas P.M.
+Weber\,$^{\text{\sfb 1,}*}$, Alisandra K.
+Denton\,$^{\text{\sfb 1}}$}
 \address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
 
-\corresp{$^\dagger$Authors contributed equally. \newline $^\ast$To whom correspondence should be addressed.}
+\corresp{$^\ast$To whom correspondence should be addressed.}
 
@@
 \section*{Funding}
-This work has been supported by the... \vspace*{-12pt}
+This work has been supported by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) 
+under Germany's Excellence Strategy--EXC-2048/1--Project ID: 390686111.\vspace*{-12pt}
 
@@
 the de.NBI cloud, and by both the Centre for Information and Media Technology (ZIM)
-and the institute of Quantitative and Theoretical Biology at Heinrich Heine University.
+and the Institute of Quantitative and Theoretical Biology at Heinrich Heine University.
 

### 2020-06-04 paper/supplements.tex
minor things
animals first in coverage, authors to match main

@@
 \author[1]{Felix Stiehler}
-\author[1]{Alisandra Denton}
 \author[1]{Marvin Steinborn}
+\author[ \hspace{-1ex}]{Stephan Scholz}
 \author[ \hspace{-1ex}]{Daniela Dey}
-\author[ \hspace{-1ex}]{Stephan Scholz}
-\author[1]{Andreas Weber}
+\author[1]{Andreas P.M. Weber}
+\author[1]{Alisandra K. Denton}
 \affil[1]{Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
@@
 All of the raw metadata and the four derived features mentioned above can be found in 
-Dataset S\ref{supdata:metadata}).
+Table S\ref{supdata:metadata}.
 
@@
 are only displayed if they can be meaningfully compared by examining coverage or spliced
-coverage. The rows show the individual species: a, b) {\it M. esculenta}, c, d) {\it M. truncatula}, 
-e, f) {\it T. cacao}, g, h) {\it P. anubis}, i, j) {\it E. caballus} and k, l) {\it P. marinus} . 
+coverage. The rows show the individual species: a, b) {\it P. anubis}, c, d) {\it E. caballus}, e, f) {\it P. marinus}, 
+g, h) {\it M. esculenta}, i, j) {\it M. truncatula}, and k, l) {\it T. cacao}.
 The left-most bars show cases where the two tools agree,
@@
 are only displayed if they can be meaningfully compared by examining coverage or spliced
-coverage. The rows show the individual species: a, b) {\it M. esculenta}, c, d) {\it M. truncatula}, 
-e, f) {\it T. cacao}, g, h) {\it P. anubis}, i, j) {\it E. caballus} and k, l) {\it P. marinus} . 
+coverage. The rows show the individual species: a, b) {\it P. anubis}, c, d) {\it E. caballus}, e, f) {\it P. marinus}, 
+g, h) {\it M. esculenta}, i, j) {\it M. truncatula}, and k, l) {\it T. cacao}. 
 The left-most bars show cases where the two tools agree,

### 2020-06-04 paper/main.tex
hope it's actually ready to send around

@@
 
-\title[short Title]{Helixer: Cross-species Gene Annotation Of Large Eucaryotic Genomes Using Deep Learning}
+\title[short Title]{Helixer: Cross-species Gene Annotation Of Large Eukaryotic Genomes Using Deep Learning}
 \author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1}}$, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Daniela Dey, Andreas P.M.
-Weber\,$^{\text{\sfb 1,}*}$, Alisandra K.
-Denton\,$^{\text{\sfb 1}}$}
+Weber\,$^{\text{\sfb 1}}$, Alisandra K.
+Denton\,$^{\text{\sfb 1,}*}$}
 \address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
@@
 Public License v3.0.\\
-\textbf{Contact:} \href{andreas.weber@hhu.de}{andreas.weber@hhu.de}\\  
+\textbf{Contact:} \href{alisandra.denton@hhu.de}{alisandra.denton@hhu.de}\\  
 \textbf{Supplementary information:} Supplementary data are available at 
@@
 the use of a BLSTM after a convolutional neural net (CNN) to find detailed motifs in 
-the human genome. DeePromotor \citep{oubounyt2019deepromoter} trains a similar 
-architecture for the recognition of promotor regions. Recently, several groups 
+the human genome. DeePromoter \citep{oubounyt2019deepromoter} trains a similar 
+architecture for the recognition of promoter regions. Recently, several groups 
 \citep{jaganathan2019predicting, wang2019splicefinder} successfully used CNNs to 
@@
 greatly surpasses all other work in this field. The source code and all input data 
-is publically available.
+is publicly available.
 
@@
 The foundation of our work are 192 animal and 60 plant genomes. The data of each 
-genome consists of the latest publically available genomic assembly in form of a 
+genome consists of the latest publicly available genomic assembly in form of a 
 FASTA file and the latest annotation in the GFF format. One genome for each species of animal 
@@
 (\cite{goodstein2012phytozome}; Table S\ref{suptab:downloads_plants}); The exact 
-genomes are listed in the aformentioned tables.
+genomes are listed in the aforementioned tables.
 Data was downloaded on the 15/10/2019 and 29/03/2019 for animals and plants respectively. 
-Both data groups were used seperately during training as well as during evaluation.
+Both data groups were used separately during training and evaluation.
 
@@
 For the selection of training genomes we had to balance multiple conflicting 
-tradeoffs. On the one hand, we want to train with as much data as possible, but 
+trade-offs. On the one hand, we want to train with as much data as possible, but 
 on the other
@@
 
-We call the genomes that were ultimatily not used during training from the original 
+We call the genomes that were ultimately not used during training from the original 
 split the {\it validation genomes}. While we did split off a group of test genomes 
@@
 performance on the combined set of validation and test genomes as the validation 
-genomes are by far the largest set. We also did not find a noticible difference in 
+genomes are by far the largest set. We also did not find a noticeable difference in 
 performance between those two groups.
@@
 by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
-of masked bases lie in the intergenic region as the most prevelant error is a 
+of masked bases lie in the intergenic region as the most prevalent error is a 
 missing UTR and GeenuFF marks a potentially large intergenic region for it. Table 
@@
 
-For the training itself, we divide each continous genomic sequence into 20,000 bp 
+For the training itself, we divide each continuous genomic sequence into 20,000 bp 
 long subsequences, for which one-hot vectors of the base pairs and annotations 
@@
 FASTA sequences without a gene and any 20,000 bp subsequences that were masked as completely 
-erroneous. Each strand of DNA was counted seperately. The gene length was 
+erroneous. Each strand of DNA was counted separately. The gene length was 
 determined by the length of the pre-mRNA of the longest protein at each
@@
 including convolutional neural networks (CNN) and hybrid architectures. We also use 
-class weigths as the class frequencies are both unbalanced and 
+class weights as the class frequencies are both unbalanced and 
 vary greatly between animals and plants. 
-For example, the tradeoff between the number of intergenic and intronic bases is 
+For example, the trade-off between the number of intergenic and intronic bases is 
 quite different in both groups (see Table \ref{tab:statistics_all}).
@@
 %{\it Capsella grandiflora}, \cite{slotte2013capsella}). 
-The Ensemble animal dataset by and large
-used the Ensembl annotation pipeline, which includes no HMM-based tool.
+The Ensemble animal dataset by and large used the Ensembl annotation pipeline, which primarily
+uses extrinsic data but never-the-less incorporates Genscan \citep{burge1997prediction} predictions.
 
@@
 
-% todo, add versions to everything below
-For each of these six species the following search was performed on Sequence Read Archive
+For each of these six species the following search was performed on Sequence Read Archive \citep{leinonen2010sequence}
 `(("<species name>"[Organism] OR <species name>[All Fields]) AND stranded[All Fields]) 
@@
 samples were chosen for further processing (see Table S\ref{suptab:rnaseq_samples}).
-Each sample was quality controlled with FastQC (www.bioinformatics.babraham.ac.uk/projects/fastqc/), before and after trimming
-with Trimmomatic \citep{bolger2014trimmomatic}, and then mapped to the genome with Hisat2 \citep{kim2019graph}. 
-Samtools \citep{li2009sequence} was used for the sorting and indexing that was necessary both for
-RNAseq mapped metric collection with PicardTools (http://broadinstitute.github.io/picard/) and custom downstream 
-processing below. MultiQC \citep{ewels2016multiqc} was used to summarize the results. The above was 
-automated and the results visualized with the code here (... make available and link to RNAseq stuff).
-The samples were filtered based on many different metrics. Relatively high mapping rates, high mapping to
-exonic regions relative to other regions, low 3' bias, normal FastQC, Trimmomatic stats and a confirmation
-to be stranded with 2nd strand as sense strand were all taken into account. If more remained, seven
+Each sample was prepped, mapped, and quality controlled in a pipeline using Trimmomatic 
+\citep{bolger2014trimmomatic}, 
+Hisat2 \citep{kim2019graph}, Samtools \citep{li2009sequence}, PicardTools 
+(http://broadinstitute.github.io/picard/), 
+FastQC 
+(www.bioinformatics.babraham.ac.uk/projects/fastqc/),
+and MultiQC (\cite{ewels2016multiqc}; see
+Table S\ref{suptab:bioinfo_params} for details). The above was 
+automated and the results visualized with the code here (https://github.com/weberlab-hhu/RNAsleek).
+The samples were filtered to those with relatively high mapping rates, high mapping to
+exonic relative to non-exonic regions, low 3' bias, normal FastQC and Trimmomatic stats, 
+and a stranded mapping pattern (2nd read is sense strand). If more remained, seven
 of the high quality samples were selected randomly.
 
-Finally, the selected and mapped RNAseq samples were merged and quantified to
-get the coverage (number of reads matching; i.e. cigar =, M or X),
+Finally, the selected and mapped RNAseq samples (Table S\ref{suptab:rnaseq_samples}) were merged with Samtools 
+and quantified to get the coverage (number of reads matching; i.e. cigar =, M or X),
 and spliced coverage (number of reads with gap or splice; i.e. cigar N or D)
-for every base pair in the genomes as implemented in the "training\_rnaseq.py"
+for every base pair in the genomes as implemented in Helixer's "training\_rnaseq.py"
 script.
@@
 We do not, for example, consistently predict very well on the 
-algae nor on the non-avian reptiles. While one of the the algae were included 
+algae nor on the non-avian reptiles. While one of the algae were included 
 in our training genomes it accounted for a tiny proportion of the total training 
@@
 and the reference was supported by the RNAseq data, and {\bf d)} helixer had low
-accuracy but the reference was not supported by RNAseq data. Plotability was
+accuracy but the reference was not supported by RNAseq data. Feasibility for visualization was
 a major secondary consideration. Each subplot shows from top to
@@
 
-As an independent evaluation and to get a better qualitative understanding of
+To make an independent qualitative and quantitative evaluation of
 performance we compared Helixer predictions, AUGUSTUS predictions and the
-reference to RNAseq coverage data. RNAseq was processed for three each of plants
-and animals chosen because they had varied performance quality relative to AUGUSTUS. 
-RNAseq coverage provides support for an 
+reference to RNAseq coverage data. RNAseq evaluation was performed for
+three each animals and plants. RNAseq coverage provides support for an 
 exonic annotation (UTR or CDS), spliced coverages provides support for an 
@@
 Helixer models do not yet have postprocessing
-to make finalized single predictions, but instead output basewise probabilities.
-We see that the model exibits higher uncertainty around transitions between annotation 
+to make finalized single predictions, but instead output base-wise probabilities.
+We see that the model exhibits higher uncertainty around transitions between annotation 
 classes, for instance between UTR and CDS, or more dramatically between UTR and 
@@
 exon/intron pattern (Figures \ref{fig:cov_example_main}b-c, S\ref{supfig:cov_example_02}c). However, in the extreme,
-there are cases where the Helixer model exibits substantial indecision or confusion 
+there are cases where the Helixer model exhibits substantial indecision or confusion 
 and shifts gradually between classes with no single class receiving a high probability 
@@
 Helixer vs AUGUSTUS as well as Helixer vs the reference for all genomes (Figure \ref{fig:average_vs_augustus}). 
-Where both tools agreee on far left in the figure, coverage and spliced coverage closely matched expectations. 
+Where both tools agree on far left in the figure, coverage and spliced coverage closely matched expectations. 
 Specifically for CDS:CDS and UTR:UTR most base pairs had some, and many
@@
 These patterns were consistent in direction but varied in magnitude in the individual species,
-with the expception of {\it P. marinus} where Helixer and AUGUSTUS performed comparably
+with the exception of {\it P. marinus} where Helixer and AUGUSTUS performed comparably
 at differentiating between CDS and intron regions (Figure S\ref{supfig:each_sp_vs_augustus}).
@@
 in every conflict, to {\it T. cacao}, where Helixer models receive equivalent
-or more support in every conflict (Figure S\ref{supfig:each_sp_vs_ref}). Interestingly, these are the two species that
-were selected because they had (relative to the usual difference) low Subgenic F1 compared to 
-AUGUSTUS; for the former the RNAseq confirms relatively weak performance for Helixer, while the 
+or more support in every conflict (Figure S\ref{supfig:each_sp_vs_ref}). Interestingly, these two
+species were selected as examples where Helixer had poor Subgenic F1; 
+for the former the RNAseq confirms relatively weak performance for Helixer, while the 
 for the latter RNAseq rather indicates a sub-par reference.
@@
 
-Finally, developent of a post-processing method
+Finally, development of a post-processing method
 to go from base pair wise predictions to integrated predictions for whole 
@@
 greatly increase real world applications.
-
 \section*{Funding}
@@
 under Germany's Excellence Strategy--EXC-2048/1--Project ID: 390686111.\vspace*{-12pt}
-
 \section*{Acknowledgements}
-We are grateful for the computational suppport and GPU access provided by
+We are grateful for the computational support and GPU access provided by
 the de.NBI cloud, and by both the Centre for Information and Media Technology (ZIM)
@@
 \begin{suptab}\label{suptab:rnaseq_samples}\end{suptab}
+\begin{suptab}\label{suptab:bioinfo_params}\end{suptab}
 \begin{suptab}\label{suptab:prediction_lengths}\end{suptab}

### 2020-06-10 paper/main.tex

@@
 We first pre-process and store the raw genomic information by using GeenuFF 
-(https://github.com/weberlab-hhu/GeenuFF). 
-GeenuFF is a tool for checking and exploring genomic data, that 
+\citep{denton2019}. GeenuFF is a tool for checking and exploring genomic data, that 
 stores all information inside a SQL database. Training and evaluation-ready data is 
@@
 \end{table}
+
+During data generation, we query for the longest transcript of each gene and 
+disregard FASTA sequences that have no functional annotation. GeenuFF checks the 
+genomic annotations for potential errors during import and is able to mark those 
+areas. We use this information to effectively blank out those bases during training 
+by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
+of bases masked lie in the intergenic region as the most prevelant error is a 
+missing UTR and GeenuFF marks a potentially large intergenic region for it. Table 
+\ref{tab:statistics_all} shows statistics about the masking.
+
+For the training itself, we divide each continous genomic sequence into 20.000 bases 
+long subsequences, which are then used as input or label for the neural network 
+together with the sample weights. We append zero padding if the subsequences are 
+not long enough. If a subsequence is fully marked as erroneous, we do not include 
+it anywhere.
 
@@
 used in the animal case. The best model was chosen out of the eight models of the 
-ensemble for having the best performance on the validation set of the training genomes.}
+ensemble for having the best performance in the reported setup.}
 \end{table}

### 2020-06-10 paper/supplements.tex

@@
 \usepackage{hyperref}
+\usepackage{makecell}
 \usepackage{amssymb}
@@
 
-\title{Supplemental Materials for \\ Helixer: Cross-species Gene Annotation of Large Eucaryotic Genomes Using Deep Learning}
+\title{Supplemental Materials for \\ Helixer: Cross-species Gene Annotation of Large Eukaryotic Genomes Using Deep Learning}
 
@@
 %
+\newpage
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\centering
+\begin{tabular}{@{}llll@{}}
+\hline
+Purpose&Tool&Version&Customized Parameters\\
+\hline
+Read quality control&FastQC&v0.11.5&\\ \\
+Adapter, Quality trimming&Trimmomatic&0.36&\makecell[l]{ILLUMINACLIP:TruSeq3-PE-2.fa:3:30:10:1:true\\MAXINFO:36:0.7 MINLEN:36}\\ \\
+Mapping&Hisat2&2.1.0&\makecell[l]{Both: --max-seeds 8 --dta\\Plants only: --pen-canintronlen G,-8,1.5 --pen-noncanintronlen G,-8,1.5} \\ \\
+Bam conversion, sorting and indexing&Samtools&1.6&\\ \\
+Mapping quality control&PicardTools&52.0&\makecell[l]{CollectRnaSeqMetrics\\STRAND=SECOND\_READ\_TRANSCRIPTION\_STRAND}\\ \\
+Summarizing and organizing&MultiQC&1.8\\
+\hline
+\end{tabular}}
+\caption{List of bioinformatics software, versions, and parameters used in processing
+public RNAseq data.}
+\label{suptab:bioinfo_params}
+\end{table}
+\clearpage
 
@@
 and the reference was supported by the RNAseq data, and d) helixer had low
-accuracy but the reference was not supported by RNAseq data. Plotability was
+accuracy but the reference was not supported by RNAseq data. Feasibility for visualization was
 a major secondary consideration. Each subplot shows from top to

### 2020-07-20 paper/main.tex
co-author suggestions and just little adjustments from re-reading

--- 2020-06-04 paper/main.tex
@@
 \title[short Title]{Helixer: Cross-species Gene Annotation Of Large Eukaryotic Genomes Using Deep Learning}
-\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1}}$, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Daniela Dey, Andreas P.M.
+\author[Sample \textit{et~al}.]{Felix Stiehler\,$^{\text{\sfb 1}}$, Marvin Steinborn\,$^{\text{\sfb 1}}$, Stephan Scholz, Daniela Dey\,$^{\text{\sfb 2}}$, Andreas P.M.
 Weber\,$^{\text{\sfb 1}}$, Alisandra K.
 Denton\,$^{\text{\sfb 1,}*}$}
-\address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}
+\address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany\\
+$^{\text{\sf 2}}$Institute of Human Genetics, Medical Faculty, RWTH Aachen University, Aachen, 52062, Germany}
 
@@
 \section{Introduction}
-The annotation of genes in DNA is fundamental for many downstream tasks, which 
-depend on high quality annotations. Gene annotation can be performed at different levels 
+Annotating genes is an integral part of
+genomic DNA sequence analysis, with many downstream taks
+dependent on annotation quality.
+Gene annotation can be performed at different levels 
 of precision, from simple coding -- non-coding classification to detailed structural 
@@
 designers of HMMs assume structure in the sequence that may limit its predictive 
-power. In practice HMMs have trouble generalizing across species and any actual learning 
-related to sequence motifs is limited to very short sequences that indicate state transitions.
+power. In practice HMMs have trouble generalizing across species and the actual learning 
+of sequence motifs is limited to very short sequences that indicate state transitions.
 
@@
 architectures. LSTM units can also be used to process sequential input starting 
-from both ends, forming a bi-directional LSTM (BLSTM). It has also been shown that 
+from both ends, forming a bidirectional LSTM (BLSTM). It has also been shown that 
 HMMs can be successfully combined with DNNs \citep{liu2016novo, liu2016pedla}.
@@
 test it later on two more species. DeepAnnotator \citep{amin2018deepannotator} uses 
-bidirectional LSTMs (BLSTM) for gene finding in prokaryotes. Gene prediction in 
+BLSTMs for gene finding in prokaryotes. Gene prediction in 
 prokaryotes is considered more amenable than in eukaryotes, as genes in prokaryotes 
@@
 Both the ability to generalize across genomes as well as the scope of the evaluation 
-greatly surpasses all other work in this field. The source code and all input data 
+represent cutting edge progress in this field. The source code and all input data 
 is publicly available.
@@
 well beyond the borders given by the training data. However, it also may be difficult 
-for the model learn and generalize if the genomic patterns inside the data are 
+for the model to learn and generalize if the genomic patterns inside the data are 
 too different from each other. Practically, it was important for us to limit the size of
-our training genomes set to be small enough that enough that we could get experimental 
+our training genomes set to be small enough that we could get experimental 
 results within a couple of days and thereby be able to test many different data and 
@@
 We employed an iterative approach to effectively select a proper set of training 
-genomes. We first started with 3-4 genomes that were expected to be of the 
+genomes, starting from 3-4 genomes that were expected to be of the 
 highest quality. We then evaluated the performance of the resulting model on all 
@@
 
-We call the genomes that were ultimately not used during training from the original 
-split the {\it validation genomes}. While we did split off a group of test genomes 
-at the beginning of our work, we ended up only reporting on the generalization 
+The genomes that were ultimately not used during training from the original 
+split are hereafter refered to as the {\it validation genomes}. 
+We report only the generalization 
 performance on the combined set of validation and test genomes as the validation 
-genomes are by far the largest set. We also did not find a noticeable difference in 
+genomes are by far the largest set, and there was not a noticeable difference in 
 performance between those two groups.
@@
 \subsection{Data Pre-processing}
-We first pre-process and store the raw genomic information by using GeenuFF 
+We first pre-processed and stored the raw genomic information by using GeenuFF 
 (https://github.com/weberlab-hhu/GeenuFF). 
-GeenuFF is a tool for checking and exploring genomic data, that 
-stores all information inside a SQL database. Training and evaluation-ready data is 
+GeenuFF is a tool for checking and exploring genomic data and annotations, that 
+stores all information inside a SQL database. Training and evaluation-ready data was
 generated by querying this database and then transforming the returned data into a 
 numerical format suitable for machine learning. The encoding of the genomic sequence 
-is done in line with the IUPAC nucleic acid notation and the functional annotation 
+was done in line with the IUPAC nucleic acid notation and the functional annotation 
 used as labels during training is transformed to a one hot encoding with the four 
-classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}. See also Table 
+classes {\it Intergenic}, {\it UTR}, {\it CDS} and {\it Intron}. See Table 
 \ref{tab:arrays} for a more detailed description of the generated data types.
@@
 
-During data generation, we query for the transcript with the longest protein of each gene and 
+During data generation, we queried for the transcript with the longest protein of each gene and 
 disregard FASTA sequences that have no functional annotation as it is ambiguous
-whether such sequences contain no genes or were simply not annotated in the reference. GeenuFF checks the 
+whether such sequences contain no genes or 
+were simply not annotated in the reference. GeenuFF checks the 
 genomic annotations for potential errors during import and is able to mark those 
-areas. We use this information to effectively blank out those bases during training 
+areas. We used this information to effectively mask those bases during training 
 by using the sample weights described in Table \ref{tab:arrays}. The vast majority 
@@
 
-For the training itself, we divide each continuous genomic sequence into 20,000 bp 
+For the training itself, we divided each continuous genomic sequence into 20,000 bp 
 long subsequences, for which one-hot vectors of the base pairs and annotations 
 are generated and respectively used as input and label for the neural network 
-together with the sample weights. We append zero padding if the subsequences  
+together with the sample weights. We appended zero padding if the subsequences  
 are shorter than 20,000 bp. 
-If a subsequence is fully marked as erroneous, we exclude it from all analyses.
+If a subsequence is fully marked as erroneous, we excluded it from all analyses.
 
@@
 \label{sec:metrics}
-We mainly use two metrics to judge model performance. The {\it Genic F1} is our 
-primary metric that provides the most comprehensive picture of annotation quality 
+We mainly used two metrics to judge model performance. The {\it Genic F1} was selected as our 
+primary metric and provides the most comprehensive picture of annotation quality 
 in one number. The {\it Subgenic F1} is similar to the Genic F1, except that it does 
-not take UTR predictions into account. This is done for comparability with AUGUSTUS 
+not take UTR predictions into account. This was calculated for comparability with AUGUSTUS 
 as further explained in Section \ref{sec:augustus}. Both metrics work by summing up 
@@
 considered class before calculating precision and recall from these values and 
-combining the two into a F1 score. Neither metric takes the values of the 
-intergenic class into account, as this class is very abundant and appears to be by 
+combining the two into a F1 score. Neither metric includes the
+intergenic class, as this class is very abundant and appears to be by 
 far the easiest to predict. (See Table \ref{tab:statistics_all} for the class 
@@
 including the intergenic class). In the case of the Genic F1, this means that 
-effectively only the TP of the intergenic class are disregarded. The metrics 
+effectively only the TP of the intergenic class were disregarded. The metrics 
 essentially provide a weighted mean of the performances in the considered classes,
@@
 \label{sec:model}
-We use a 4-layer deep stacked BLSTM network with 256 units per layer and layer 
+We used a 4-layer deep stacked BLSTM network with 256 units per layer and layer 
 normalization \citep{ba2016layer} between each BLSTM layer to produce the 
@@
 tested multiple different architectures before arriving at this model configuration, 
-including convolutional neural networks (CNN) and hybrid architectures. We also use 
+including convolutional neural networks (CNN) and hybrid architectures. We also used 
 class weights as the class frequencies are both unbalanced and 
@@
 A class wise evaluation including the calculation of the Genic 
-F1 score is performed on the validation set after every epoch. The best model
+F1 score was performed on the validation set after every epoch. The best model
 from any training run was selected as the model with the highest Genic F1 after either
@@
 algorithm \citep{bergstra2011algorithms}, random search or grid search depending on 
-the situation. We used NNI (https://github.com/microsoft/nni) to facilitate the search. All relevant 
+the situation. We used NNI (https://github.com/microsoft/nni) to 
+facilitate the search. All relevant 
 hyperparameters can be found in Table S\ref{suptab:params} and 
@@
 
-We also compare our final models to a dilated CNN (dCNN) baseline as this kind of architecture is 
+We also compared our final models to a dilated CNN (dCNN) 
+baseline as this kind of architecture is 
 widely used when working with DNA data as input. Details of the neural architecture 
@@
 We also used multiple techniques to improve the prediction quality after the 
-training was done. One very effective way for genomes with larger genes was to input 
-longer sequences during inference time than during training. This was done for all 
+training was done. One very effective way for genomes with larger genes was inputting
+longer sequences during inference than during training. This was done for all 
 animal genomes except the invertebrates. The input sequences were up to 10 times 
@@
 
-The final predictions of a single model are constructed by overlapping predictions, 
-that were made from a sliding window and then cropped to a core sequence. This was done 
+The final predictions of a single model were constructed by overlapping predictions, 
+which were made from a sliding window and then cropped to a core sequence. This was done 
 to strongly reduce a typical drop in performance of the models towards the beginning 
 and end of each sequence (see Figure \ref{fig:main_results}c). It also improves the 
-average model performance in general by providing the model with multiple different 
+average model performance by providing the model with multiple different 
 starting points. The different overlapping sequences were combined by averaging 
-the individual softmax values of each base. The figures in Section \ref{ssec:overlapping} of the 
+the individual softmax values of each base. 
+The figures in Section \ref{ssec:overlapping} of the 
 Supplements show the effect of overlapping for each genome, ordered by N75. We found 
@@
 each species. For this, we performed 4 separate animal and plant training runs and 
-selected two checkpoints each, specifically the checkpoint from the epoch with the best precision and the  % please check / epoch model / version something
+selected two checkpoints each, specifically the checkpoint from the epoch with the best precision and the
 other from the epoch with the best recall. The model checkpoint with the best Genic F1 was always best in one of 
@@
 Due to the practical time constraints for retraining and running, we decided to compare our Helixer
-models only to one existing {\it de novo} gene caller, the popular gene caller, AUGUSTUS \citep{stanke2003gene}.
+models only to one existing {\it de novo} tool, namely the popular gene caller AUGUSTUS \citep{stanke2003gene}.
 Typical usage requires retraining AUGUSTUS for each species, with the exception of a
 few lucky cases where a model for a sufficiently close relative is already available. To scale this for 
-the large plant and animal datasets, we used protein-homology to create a training set and
+the large plant and animal datasets, we used protein-homology to create an AUGUSTUS training set and
 therefore could train and evaluate only models without UTRs.
@@
 we randomly selected only half the BUSCO-generated training set, which resulted in a training set 
-size and runtime comparable to the plants (about 400 genes and several days per species). For animals
+size and runtime comparable to the plants (about 400 genes and several CPU days per species). For animals
 the training was carried out with the subsetted training file and using the following scripts provided
@@
 files in the same manor as the reference, allowing for the direct comparison with both the reference and 
-Helixer predictions. While still feasible for some 237 species, this method allowed neither training
+Helixer predictions. While this method was feasible for some 237 species, it allowed neither training
 nor prediction of UTR regions with AUGUSTUS, so the metric Subgenic F1, which disregards the class UTR,
-was used for all comparisons between AUGUSTUS and Helixer (see Section \ref{sec:metrics} for more details) .
+was used for all comparisons between AUGUSTUS and Helixer (see Section \ref{sec:metrics} for more details).
 
@@
 `(("<species name>"[Organism] OR <species name>[All Fields]) AND stranded[All Fields]) 
-AND ("biomol rna"[Properties] AND "library layout paired"[Properties])'. In so far as 
-more than 50 samples were identified and every Nth sample was selected so that under 50
-samples were chosen for further processing (see Table S\ref{suptab:rnaseq_samples}).
+AND ("biomol rna"[Properties] AND "library layout paired"[Properties])'.
+If more than 50 samples were identified, every Nth samples was selected 
+so that in total under 50 samples were chosen for further processing
+(see Table S\ref{suptab:rnaseq_samples}).
 Each sample was prepped, mapped, and quality controlled in a pipeline using Trimmomatic 
@@
 and MultiQC (\cite{ewels2016multiqc}; see
-Table S\ref{suptab:bioinfo_params} for details). The above was 
-automated and the results visualized with the code here (https://github.com/weberlab-hhu/RNAsleek).
+Table S\ref{suptab:bioinfo_params} for details). This pipeline was 
+automated and the results visualized with the code available here (https://github.com/weberlab-hhu/RNAsleek).
 The samples were filtered to those with relatively high mapping rates, high mapping to
@@
 \centerline{\includegraphics[width=\textwidth]{images/main_results}}
-\caption{Main overall results and other investigations. The top row shows the animal
- data; the bottom row is concerned with the plants. Orange dots are for AUGUSTUS and 
+\caption{Main overall results and other investigations. The top and bottom rows
+show the animal and plant data, respectively. Orange dots are for AUGUSTUS and 
 the blue ones are for Helixer. {\bf a)} Swarm - and boxplots showing the Subgenic F1 
-scores for all genomes that were not used during training of Helixer, which is 
-equivalent to all validation and test genomes. Each dot represents the prediction 
+scores for all validation and test genomes. Each dot represents the prediction 
 performace for one genome. The median prediction scores are shown by the red line. 
@@
 AUGUSTUS and a dilated CNN baseline in Table \ref{tab:results}. The results show a 
-clear improvement over the AUGUSTUS both in median performance and spread. We also 
+clear improvement over the AUGUSTUS both in higher median performance and reduced spread. We also 
 significantly outperform a baseline constructed of a dilated CNN architecture.
@@
 genomes or species that are phylogenetically the furthest away from our training 
-genomes. This is the case for both animals and plants and is visualized by the 
+genomes. This is the case for both animals and plants and is visualized in
 Figures S\ref{supfig:phylogeny_f1_animals} and S\ref{supfig:phylogeny_f1_plants}. 
@@
 have longer genes and was enabled by our model architecture being a relatively 
-simple LSTM stack without any fully connected layers on top. We also constructed the 
+simple BLSTM stack without any fully connected layers on top. We also constructed the 
 final predictions out of a overlapping ones, which greatly helped to reduce 
@@
 performance we compared Helixer predictions, AUGUSTUS predictions and the
-reference to RNAseq coverage data. RNAseq evaluation was performed for
-three each animals and plants. RNAseq coverage provides support for an 
-exonic annotation (UTR or CDS), spliced coverages provides support for an 
+reference to RNAseq coverage data. Evaluation with RNAseq data was performed for
+three each of animal and plant species. RNAseq coverage provides support for an 
+exonic annotation (UTR or CDS), spliced coverage provides support for an 
 annotation of intron, and neither coverage nor spliced coverage is expected 
@@
 Helixer models also sometimes showed uncertainty for larger regions.
-In some cases, where Helixer did not receive RNAseq support for its highest probability
+In some cases where Helixer did not receive RNAseq support for its highest probability
 annotation, it assigned a low but non-trivial probability to the RNAseq-supported 
@@
 the other intergenic or intron, there was more coverage when 
-Helixer predicted CDS (Figure \ref{fig:average_vs_augustus}a). Similarly, for base pairs where one tool predicted intron and the other 
+Helixer predicted CDS (Figure \ref{fig:average_vs_augustus}a). 
+Similarly, for base pairs where one tool predicted intron and the other 
 CDS or intergenic, there was more spliced coverage when Helixer predicted intron (Figure \ref{fig:average_vs_augustus}b). 
@@
 RNAseq based comparison of Helixer and the reference was less clear cut.
-Averaged across species, helixer's predictions received slightly more support
+Averaged across species, Helixer's predictions received slightly more support
 than the reference when differentiating CDS, UTR and introns from intergenic;
-however helixer's predictions received slightly less support than the reference 
-when differentiating CDS and UTR from introns (Figure S\ref{supfig:average_vs_ref}). Performance varies between individual
+however Helixer's predictions received slightly less support than the reference 
+when differentiating CDS and UTR from introns (Figure S\ref{supfig:average_vs_ref}). 
+Performance varies between individual
 species, from {\it P. marinus}, where the reference receives more support

### 2020-07-31 paper/main.tex
minor, institute

@@
 Denton\,$^{\text{\sfb 1,}*}$}
-\address{$^{\text{\sf 1}}$Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany\\
+\address{$^{\text{\sf 1}}$Institue of Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, 40225, Germany\\
 $^{\text{\sf 2}}$Institute of Human Genetics, Medical Faculty, RWTH Aachen University, Aachen, 52062, Germany}
